### 2021-04-12 Manuscript/manuscript.Rmd
Started manuscript draft

---
title: "Edge effects and optimal field sizes in Alberta"
author:
  - name: Samuel V. J. Robinson
    email: samuel.robinson@ucalgary.ca
    affiliation: University of Calgary
    footnote: 1
  - name: Lan H. Nguyen
    email: hoanglan.nguyen@ucalgary.ca
    affiliation: University of Calgary
  - name: Paul Galpern
    email: paul.galpern@ucalgary.ca
    affiliation: University of Calgary
address:
  - code: University of Calgary
    address: 2500 University Drive NW, Calgary, AB
footnote:
  - code: 1
    text: "Corresponding Author"
date: "Spring 2021"
abstract: "`r paste(readLines('abstract.Rmd'), collapse = ' ')`"
keep_tex: true
keywords:
  - beetles;
  - spiders;
  - harvestmen;

bibliography: yieldRefs.bib
geometry: "left=3cm,right=3cm,top=3cm,bottom=3cm"
output: 
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
    toc: false
    fig_caption: true
csl: ecology-letters.csl    
header-includes:
  \makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother
  \usepackage{float} \floatplacement{figure}{H} 
  \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  \usepackage{setspace}
  \linenumbers
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

\newpage
\doublespacing

# Introduction

- Preserving SNL around fields is important for agriculture and conservation
  - Ecosystem services
  - Act as reservoirs for beneficial insects
  - Microclimate zones
  - Poorly studied in North America
  - Studies tend to be limited in scope (e.g. only certain organisms/crops) and applicability

- Ecosystem services can influence both the mean and variability of yield in agroecosystems
  - Typically only averages (means) are considered, but higher stability (lower variance) in yield is also valuable

- Field size and field boundaries are directly related to the conservation of SNL in agroecosystems
  - Size examples: wheat study from the UK
  - Boundary examples: flower strip studies, hedgerows
  - In North America, large financial incentives to make fields large and homogeneous, especially with large harvesting and planting equipment
  - Crop edge effects cause low yields at the margins of fields because of late emergence, poor microclimate, and competition with weeds
  - Ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
  - Therefore, there should be a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services

- Precision yield data holds enormous promise for agronomy
  - Limited because of:
  - Lack of standardized formats
  - Yearly calibration data
  - Clear statistical protocols

- In this paper we ask:
  1. How does crop yield mean and variance change with distance from the edge of field?
  2. Does this depend on type of field edge?
  3. Given this relationship, are there upper and lower bounds for field size if a farmer wanted to maximize average yield, or minimize variance?

# Methods

## Data collection

- Precision yield data were collected directly from farmers across Alberta
  - Farmers were solicited for yield data through local agronomists, and we received data from a total of X growers across a total of X years
  - Most fields represented only a single year of data, but we did sometimes receive multiple years of data from the same field
  - File formats vary depending on the brand of combine, so we converted data to a standard format (csv) using Ag Leader SMS
  - In total, we analyzed yield data from X field-years of data, containing a total of X million data points

- Yield data was collected in discrete rectangles of the same length as the data interval (distance = speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m)
  - We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
  - Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location
  - Seeding and application rates were constant with each field, so we did not consider inputs in our analysis

- Field boundaries were automatically digitized, then manually checked using satellite imagery and classified land cover data 
  - Crop boundaries are flexible, and often change from year to year depending on planting and emergence conditions (e.g. flooding during some years)
  - Additionally, seminatural features often change from year to year
    - Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years
    - Grass boundaries can change if fields are used for as haying or pasture during crop rotation
  - This makes accurate and consistent classification of field boundaries very difficult
  - Because of this, we used the following general categories for field boundaries:
    1. Standard: thin ($<10$ m wide) grassy field boundary, often grassy road right-of-ways 
    2. Wetland: permanent wetland, whose borders are largely unchanged from year-to-year
    3. Forest/shrub: permanent windbreaks or remnant forests
    4. Grass: larger grassy area (pasture or permanent seminatural grassland)
    5. Other crop: different crop with little or no visible boundary between planted areas 
    6. Bare: unplanted (fallow), flooded area, or unpaved roads

## Analysis

- At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor
  - Crop yield varies within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices such as strip farming
  - While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields), this does not address sensor drift over time that occurs within fields
  - The yield monitors may record lower yields as sensors accumulate debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
  - Additionally, ground speed is known to be extremely important to yield monitor accuracy [@arslan2002] 
  - To address this, we fit the following model:

  \begin{equation}
  log(yield) ~ Normal (\mu, \sigma)
  \mu = Intercept + log(Polygon Size) + f(Distance from Edge) + f(Easting, Northing) + f(Harvest Sequence)
  log(\sigma) = Intercept + log(Polygon Size) + f(Distance from Edge) + f(Easting, Northing) + f(Harvest Sequence)
  \end{equation}

# Results

Results here

# Discussion

Discussion here

# Authors' contributions

Author's contribution

# Acknowledgements

Funding for this research was provided by Ducks Unlimited Canada's Institute for Wetland and Waterfowl Research, the Alberta Canola Producers Commission, Manitoba Canola Growers Association, SaskCanola, the Alberta Biodiversity Monitoring Institute, and the Alberta Conservation Association.

# References {-}

<div id="refs"></div>

```{r child = 'supplemental.Rmd'}
```

### 2021-04-12 Manuscript/manuscript.tex
Started manuscript draft

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Edge effects and optimal field sizes in Alberta},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% Pandoc header
\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} \usepackage{setspace} \linenumbers

\begin{document}
\begin{frontmatter}

  \title{Edge effects and optimal field sizes in Alberta}
    \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
   \ead{samuel.robinson@ucalgary.ca} 
    \author[University of Calgary]{Lan H. Nguyen}
   \ead{hoanglan.nguyen@ucalgary.ca} 
    \author[University of Calgary]{Paul Galpern}
   \ead{paul.galpern@ucalgary.ca} 
      \address[University of Calgary]{2500 University Drive NW, Calgary, AB}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  Abstract goes here
  \end{abstract}
   \begin{keyword} beetles; spiders; harvestmen;\end{keyword}
 \end{frontmatter}

\newpage
\doublespacing

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\begin{itemize}
\tightlist
\item
  Preserving SNL around fields is important for agriculture and conservation

  \begin{itemize}
  \tightlist
  \item
    Ecosystem services
  \item
    Act as reservoirs for beneficial insects
  \item
    Microclimate zones
  \item
    Poorly studied in North America
  \item
    Studies tend to be limited in scope (e.g.~only certain organisms/crops) and applicability
  \end{itemize}
\item
  Ecosystem services can influence both the mean and variability of yield in agroecosystems

  \begin{itemize}
  \tightlist
  \item
    Typically only averages (means) are considered, but higher stability (lower variance) in yield is also valuable
  \end{itemize}
\item
  Field size and field boundaries are directly related to the conservation of SNL in agroecosystems

  \begin{itemize}
  \tightlist
  \item
    Size examples: wheat study from the UK
  \item
    Boundary examples: flower strip studies, hedgerows
  \item
    In North America, large financial incentives to make fields large and homogeneous, especially with large harvesting and planting equipment
  \item
    Crop edge effects cause low yields at the margins of fields because of late emergence, poor microclimate, and competition with weeds
  \item
    Ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
  \item
    Therefore, there should be a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
  \end{itemize}
\item
  Precision yield data holds enormous promise for agronomy

  \begin{itemize}
  \tightlist
  \item
    Limited because of:
  \item
    Lack of standardized formats
  \item
    Yearly calibration data
  \item
    Clear statistical protocols
  \end{itemize}
\end{itemize}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{data-collection}{%
\subsection{Data collection}\label{data-collection}}

\begin{itemize}
\tightlist
\item
  Precision yield data were collected directly from farmers across Alberta

  \begin{itemize}
  \tightlist
  \item
    Farmers were solicited for yield data through local agronomists, and we received data from a total of X growers across a total of X years
  \item
    Most fields represented only a single year of data, but we did sometimes receive multiple years of data from the same field
  \item
    File formats vary depending on the brand of combine, so we converted data to a standard format (csv) using Ag Leader SMS
  \item
    In total, we analyzed yield data from X field-years of data, containing a total of X million data points
  \end{itemize}
\item
  Yield data was collected in discrete rectangles of the same length as the data interval (distance = speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m)

  \begin{itemize}
  \tightlist
  \item
    We extracted the size of each polygon (m\(^2\)), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
  \item
    Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location
  \item
    Seeding and application rates were constant with each field, so we did not consider inputs in our analysis
  \end{itemize}
\item
  Field boundaries were automatically digitized, then manually checked using satellite imagery and classified land cover data

  \begin{itemize}
  \tightlist
  \item
    Crop boundaries are flexible, and often change from year to year depending on planting and emergence conditions (e.g.~flooding during some years)
  \item
    Additionally, seminatural features often change from year to year

    \begin{itemize}
    \tightlist
    \item
      Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years
    \item
      Grass boundaries can change if fields are used for as haying or pasture during crop rotation
    \end{itemize}
  \item
    This makes accurate and consistent classification of field boundaries very difficult
  \item
    Because of this, we used the following general categories for field boundaries:

    \begin{enumerate}
    \def\labelenumi{\arabic{enumi}.}
    \tightlist
    \item
      Standard: thin (\(<10\) m wide) grassy field boundary, often grassy road right-of-ways
    \item
      Wetland: permanent wetland, whose borders are largely unchanged from year-to-year
    \item
      Forest/shrub: permanent windbreaks or remnant forests
    \item
      Grass: larger grassy area (pasture or permanent seminatural grassland)
    \item
      Other crop: different crop with little or no visible boundary between planted areas
    \item
      Bare: unplanted (fallow), flooded area, or unpaved roads
    \end{enumerate}
  \end{itemize}
\end{itemize}

\hypertarget{analysis}{%
\subsection{Analysis}\label{analysis}}

\begin{itemize}
\tightlist
\item
  At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor

  \begin{itemize}
  \tightlist
  \item
    Crop yield varies within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices such as strip farming
  \item
    While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields), this does not address sensor drift over time that occurs within fields
  \item
    The yield monitors may record lower yields as sensors accumulate debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
  \item
    Additionally, ground speed is known to be extremely important to yield monitor accuracy (Arslan \& Colvin 2002)
  \item
    To address this, we fit the following model:
  \end{itemize}

  \begin{equation}
  ln(yield) 
  \end{equation}
\end{itemize}

\hypertarget{results}{%
\section{Results}\label{results}}

Results here

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Discussion here

\hypertarget{authors-contributions}{%
\section{Authors' contributions}\label{authors-contributions}}

Author's contribution

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

Funding for this research was provided by Ducks Unlimited Canada's Institute for Wetland and Waterfowl Research, the Alberta Canola Producers Commission, Manitoba Canola Growers Association, SaskCanola, the Alberta Biodiversity Monitoring Institute, and the Alberta Conservation Association.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-arslan2002}{}%
Arslan, S. \& Colvin, T.S. (2002). An evaluation of the response of yield monitors and combines to varying yields. \emph{Precision Agriculture}, 3, 107--122.

\end{CSLReferences}

\newpage

\hypertarget{appendix-a-supplementary-material}{%
\section*{Appendix A: Supplementary Material}\label{appendix-a-supplementary-material}}
\addcontentsline{toc}{section}{Appendix A: Supplementary Material}

Supplemental materials here

\end{document}

### 2021-04-12 Manuscript/supplemental.Rmd
Started manuscript draft

--- 2021-04-12 Manuscript/manuscript.tex
@@
-\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
-%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
-\usepackage[hyphens]{url}
-
-\usepackage{lineno} % add
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-\usepackage{graphicx}
-\usepackage{booktabs} % book-quality tables
-%%%%%%%%%%%%%%%% end my additions to header
-
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
-\bibliographystyle{elsarticle-harv}
-\usepackage{longtable}
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={Edge effects and optimal field sizes in Alberta},
-            colorlinks=false,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
-
-\setcounter{secnumdepth}{5}
-% Pandoc toggle for numbering sections (defaults to be off)
-
-% Pandoc citation processing
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc} % for calculating minipage widths
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-% Pandoc header
-\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} \usepackage{setspace} \linenumbers
-
-\begin{document}
-\begin{frontmatter}
-
-  \title{Edge effects and optimal field sizes in Alberta}
-    \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
-   \ead{samuel.robinson@ucalgary.ca} 
-    \author[University of Calgary]{Lan H. Nguyen}
-   \ead{hoanglan.nguyen@ucalgary.ca} 
-    \author[University of Calgary]{Paul Galpern}
-   \ead{paul.galpern@ucalgary.ca} 
-      \address[University of Calgary]{2500 University Drive NW, Calgary, AB}
-      \cortext[1]{Corresponding Author}
-
-  \begin{abstract}
-  Abstract goes here
-  \end{abstract}
-   \begin{keyword} beetles; spiders; harvestmen;\end{keyword}
- \end{frontmatter}
-
-\newpage
-\doublespacing
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-\begin{itemize}
-\tightlist
-\item
-  Preserving SNL around fields is important for agriculture and conservation
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Ecosystem services
-  \item
-    Act as reservoirs for beneficial insects
-  \item
-    Microclimate zones
-  \item
-    Poorly studied in North America
-  \item
-    Studies tend to be limited in scope (e.g.~only certain organisms/crops) and applicability
-  \end{itemize}
-\item
-  Ecosystem services can influence both the mean and variability of yield in agroecosystems
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Typically only averages (means) are considered, but higher stability (lower variance) in yield is also valuable
-  \end{itemize}
-\item
-  Field size and field boundaries are directly related to the conservation of SNL in agroecosystems
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Size examples: wheat study from the UK
-  \item
-    Boundary examples: flower strip studies, hedgerows
-  \item
-    In North America, large financial incentives to make fields large and homogeneous, especially with large harvesting and planting equipment
-  \item
-    Crop edge effects cause low yields at the margins of fields because of late emergence, poor microclimate, and competition with weeds
-  \item
-    Ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
-  \item
-    Therefore, there should be a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
-  \end{itemize}
-\item
-  Precision yield data holds enormous promise for agronomy
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Limited because of:
-  \item
-    Lack of standardized formats
-  \item
-    Yearly calibration data
-  \item
-    Clear statistical protocols
-  \end{itemize}
-\end{itemize}
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{data-collection}{%
-\subsection{Data collection}\label{data-collection}}
-
-\begin{itemize}
-\tightlist
-\item
-  Precision yield data were collected directly from farmers across Alberta
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Farmers were solicited for yield data through local agronomists, and we received data from a total of X growers across a total of X years
-  \item
-    Most fields represented only a single year of data, but we did sometimes receive multiple years of data from the same field
-  \item
-    File formats vary depending on the brand of combine, so we converted data to a standard format (csv) using Ag Leader SMS
-  \item
-    In total, we analyzed yield data from X field-years of data, containing a total of X million data points
-  \end{itemize}
-\item
-  Yield data was collected in discrete rectangles of the same length as the data interval (distance = speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m)
-
-  \begin{itemize}
-  \tightlist
-  \item
-    We extracted the size of each polygon (m\(^2\)), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
-  \item
-    Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location
-  \item
-    Seeding and application rates were constant with each field, so we did not consider inputs in our analysis
-  \end{itemize}
-\item
-  Field boundaries were automatically digitized, then manually checked using satellite imagery and classified land cover data
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Crop boundaries are flexible, and often change from year to year depending on planting and emergence conditions (e.g.~flooding during some years)
-  \item
-    Additionally, seminatural features often change from year to year
-
-    \begin{itemize}
-    \tightlist
-    \item
-      Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years
-    \item
-      Grass boundaries can change if fields are used for as haying or pasture during crop rotation
-    \end{itemize}
-  \item
-    This makes accurate and consistent classification of field boundaries very difficult
-  \item
-    Because of this, we used the following general categories for field boundaries:
-
-    \begin{enumerate}
-    \def\labelenumi{\arabic{enumi}.}
-    \tightlist
-    \item
-      Standard: thin (\(<10\) m wide) grassy field boundary, often grassy road right-of-ways
-    \item
-      Wetland: permanent wetland, whose borders are largely unchanged from year-to-year
-    \item
-      Forest/shrub: permanent windbreaks or remnant forests
-    \item
-      Grass: larger grassy area (pasture or permanent seminatural grassland)
-    \item
-      Other crop: different crop with little or no visible boundary between planted areas
-    \item
-      Bare: unplanted (fallow), flooded area, or unpaved roads
-    \end{enumerate}
-  \end{itemize}
-\end{itemize}
-
-\hypertarget{analysis}{%
-\subsection{Analysis}\label{analysis}}
-
-\begin{itemize}
-\tightlist
-\item
-  At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Crop yield varies within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices such as strip farming
-  \item
-    While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields), this does not address sensor drift over time that occurs within fields
-  \item
-    The yield monitors may record lower yields as sensors accumulate debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
-  \item
-    Additionally, ground speed is known to be extremely important to yield monitor accuracy (Arslan \& Colvin 2002)
-  \item
-    To address this, we fit the following model:
-  \end{itemize}
-
-  \begin{equation}
-  ln(yield) 
-  \end{equation}
-\end{itemize}
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-Results here
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-Discussion here
-
-\hypertarget{authors-contributions}{%
-\section{Authors' contributions}\label{authors-contributions}}
-
-Author's contribution
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-Funding for this research was provided by Ducks Unlimited Canada's Institute for Wetland and Waterfowl Research, the Alberta Canola Producers Commission, Manitoba Canola Growers Association, SaskCanola, the Alberta Biodiversity Monitoring Institute, and the Alberta Conservation Association.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-arslan2002}{}%
-Arslan, S. \& Colvin, T.S. (2002). An evaluation of the response of yield monitors and combines to varying yields. \emph{Precision Agriculture}, 3, 107--122.
-
-\end{CSLReferences}
-
 \newpage
 
-\hypertarget{appendix-a-supplementary-material}{%
-\section*{Appendix A: Supplementary Material}\label{appendix-a-supplementary-material}}
-\addcontentsline{toc}{section}{Appendix A: Supplementary Material}
+# Appendix A: Supplementary Material {-}
 
 Supplemental materials here
-
-\end{document}

### 2021-04-12 createBoundaries.R

@@ -96,3 +96,24 @@ cluster <- makeCluster(8) #8 procs max
 parLapply(cl=cluster,1:nrow(datSource),makeBoundary,dS=datSource,rP=rootPath,outerOnly=TRUE) 
 beep(1)
 stopCluster(cluster)
+
+# Turn polygons into linestrings ------------------------------------------
+library(tidyverse)
+library(sf)
+
+shpFiles <- paste0('./Figures/FieldBoundaries/',dir('./Figures/FieldBoundaries',pattern=".shp",recursive=TRUE))
+
+makeLinestring <- function(shp){
+  # shp <-shpFiles[1] 
+  fieldName <- strsplit(shp,'/')[[1]][length(strsplit(shp,'/')[[1]])]
+  
+  #Polygon
+  p <- read_sf(shp) %>% st_cast('MULTILINESTRING')
+  st_write(p,paste0('./Figures/FieldBoundariesLine/',fieldName),append=FALSE,driver='ESRI Shapefile')
+}
+
+sapply(shpFiles[1:5],makeLinestring)
+
+
+
+

### 2021-04-12 createBoundaries.R
Created boundary notes

@@ -88,7 +88,7 @@ makeBoundary <- function(i,dS,rP,outerOnly=TRUE,overwrite=FALSE,nSubSamp=50000){
   rm(dat,fieldEdge,fieldEdge2); gc() #Garbage collection
 }
 
-# makeBoundary(6,datSource,rootPath,overwrite=TRUE) #Test
+makeBoundary(17,datSource,rootPath,overwrite=TRUE) #Test
 
 library(parallel)
 library(beepr)

### 2021-04-21 helperFunctions.R
Worked on meta models

@@ -103,3 +103,115 @@ convertYield <- function(x,inUnits=NULL,outUnits=NULL){
   
   return(xOut)
 } 
+
+#Function to extract smooth info from modList.Rdata at specified path
+# l = number of replicates in each smooth prediction (length.out)
+# margInt = marginalize across intercepts
+# samp = sample from posterior distribution of coefficients rather than using the mean
+getSmooths <- function(path,l=c(30,100,500),margInt=c(FALSE,FALSE,FALSE),samp=FALSE){
+  require(mgcv)
+  # path <- paste0('./Figures/ModelCheck/',datSource$filename[1],' modList.Rdata')
+  # l <- c(20,100,500)
+  
+  load(path) #Load data
+  
+  smoothLabs <- sapply(modList$smooth,function(x) x$label) #Labels for smoothers
+  
+  #Polygon area (basically speed) regression
+  
+  logPArea <- seq(log(modList$pAreaRange[1]),log(modList$pAreaRange[2]),length.out=l[1])
+  pArea <- exp(logPArea) 
+  
+  meanVars <- which(grepl('(^\\(Intercept\\)$|^log\\(pArea\\)$)',names(modList$coefs)))
+  sdVars <- which(grepl('(^\\(Intercept\\)\\.1$|^log\\(pArea\\)\\.1$)',names(modList$coefs)))
+  
+  #Coefficients (intercept and slope) for mean and logSD relationship
+  if(samp){ #Sample from posterior
+    meanCoefs <- rnorm(rep(1,length(meanVars)),modList$coefs[meanVars],sqrt(diag(modList$vcv)[meanVars]))
+    sdCoefs <- rnorm(rep(1,length(meanVars)),modList$coefs[sdVars],sqrt(diag(modList$vcv)[sdVars]))
+  } else {
+    meanCoefs <- modList$coefs[meanVars] 
+    sdCoefs <- modList$coefs[sdVars]
+  }
+  
+  if(margInt[1]){ #Marginalize across intercept (set intercept coef to 0)
+    meanCoefs[1] <- 0 
+    sdCoefs[1] <- 0  
+  }
+  
+  pAreaDat <- data.frame(pArea, 
+                         mean=cbind(rep(1,length(logPArea)),logPArea) %*% meanCoefs,
+                         logSD=cbind(rep(1,length(logPArea)),logPArea) %*% sdCoefs)
+  
[…]

### 2021-04-21 runModels.R

@@ -8,6 +8,7 @@ library(ggpubr)
 library(mgcv)
 library(sf)
 library(beepr)
+source('helperFunctions.R')
 
 # Get paths to data -------------------------------------------------------
 
@@ -246,85 +247,26 @@ beep(1)
 stopCluster(cluster)
 Sys.time() #Takes ~ 7 hrs
 
-#Need to identify oundary types at each field
+#Need to identify boundary types at each field
 
 # Get smoother info from models -------------------------------------------
 
-#Function to extract smooth info from modList.Rdata at specified path
-# l = number of replicates in each smooth prediction (length.out)
-getSmooths <- function(path,l=c(30,100,500)){
-  
-  # path <- paste0('./Figures/ModelCheck/',datSource$filename[1],' modList.Rdata')
-  # l <- c(20,100,500)
-  
-  load(path) #Load data
-  
-  smoothLabs <- sapply(modList$smooth,function(x) x$label) #Labels for smoothers
-  
-  #Polygon area (basically speed) regression
-  
-  logPArea <- seq(log(modList$pAreaRange[1]),log(modList$pAreaRange[2]),length.out=l[1])
-  pArea <- exp(logPArea) 
-  
-  meanVars <- which(grepl('(^\\(Intercept\\)$|^log\\(pArea\\)$)',names(modList$coefs)))
-  sdVars <- which(grepl('(^\\(Intercept\\)\\.1$|^log\\(pArea\\)\\.1$)',names(modList$coefs)))
-  
-  pAreaDat <- data.frame(pArea, #Set intercept to 1 
-                         predMean=cbind(rep(1,length(logPArea)),logPArea) %*% modList$coefs[meanVars],
-                         predSD=cbind(rep(1,length(logPArea)),logPArea) %*% modList$coefs[sdVars])
-  
-  #Field boundary distance smoothers
-  
-  meanDistSmooth <- which(smoothLabs=='s(dist)')
-  sdDistSmooth <- which(smoothLabs=='s.1(dist)')
-  
-  meanSmoothList <- modList$smooth[[meanDistSmooth]]
-  sdSmoothList <- modList$smooth[[sdDistSmooth]]
[…]

### 2021-04-22 helperFunctions.R
Got uncertainty estimation working better

@@ -118,7 +118,6 @@ getSmooths <- function(path,l=c(30,100,500),margInt=c(FALSE,FALSE,FALSE),samp=FA
   smoothLabs <- sapply(modList$smooth,function(x) x$label) #Labels for smoothers
   
   #Polygon area (basically speed) regression
-  
   logPArea <- seq(log(modList$pAreaRange[1]),log(modList$pAreaRange[2]),length.out=l[1])
   pArea <- exp(logPArea) 
   
@@ -203,8 +202,8 @@ getSmooths <- function(path,l=c(30,100,500),margInt=c(FALSE,FALSE,FALSE),samp=FA
   }
   
   rDat <- data.frame(r=r,
-                     mean=cbind(rep(1,length(d)),PredictMat(meanSmoothList,data=data.frame(r=r))) %*% modList$coef[meanVars],
-                     logSD=cbind(rep(1,length(d)),PredictMat(sdSmoothList,data=data.frame(r=r))) %*% modList$coef[sdVars])  
+                     mean=cbind(rep(1,length(r)),PredictMat(meanSmoothList,data=data.frame(r=r))) %*% meanCoefs,
+                     logSD=cbind(rep(1,length(r)),PredictMat(sdSmoothList,data=data.frame(r=r))) %*% sdCoefs)  
   
   #Assemble into list
   datList <- list(pAreaDat=pAreaDat,distDat=distDat,rDat=rDat)
@@ -213,5 +212,17 @@ getSmooths <- function(path,l=c(30,100,500),margInt=c(FALSE,FALSE,FALSE),samp=FA
 }
 
 # debugonce(getSmooths)
-# getSmooths(paste0('./Figures/ModelCheck/',datSource$filename[1],' modList.Rdata'),margInt=c(FALSE,TRUE,TRUE),samp=FALSE) #Test
+# temp <- getSmooths(paste0('./Figures/ModelCheck/',datSource$filename[1],' modList.Rdata'),margInt=c(FALSE,TRUE,TRUE),samp=FALSE) #Test
+# temp2 <- getSmooths(paste0('./Figures/ModelCheck/',datSource$filename[1],' modList.Rdata'),margInt=c(FALSE,TRUE,TRUE),samp=TRUE) #Test
+# 
+# ggplot()+geom_line(data=temp$pAreaDat,aes(x=pArea,y=mean))+ #Works for this (I think)
[…]

### 2021-04-22 runModels.R

@@ -280,7 +280,7 @@ allEff <- lapply(names(allSmooths[[1]]),function(y){
 #Plot with individual field-level smoothers
 
 tHa2buAc <- 17.0340 #tonnes per hectare to bushels per acre (https://www.agrimoney.com/calculators/calculators)
-ylimVals <- list(mean=c(-1.2,1.2)*tHa2buAc,sd=c(-2.5,2.5)*tHa2buAc) #Y limits
+ylimVals <- list(mean=c(-1.2,1.2)*tHa2buAc,sd=c(0,5)) #Y limits
 ylabMean <- 'Mean Yield (bushels/acre)'
 ylabSD <- 'log(SD Yield)'
 
@@ -308,7 +308,7 @@ p5 <- allEff[[3]] %>% ggplot(aes(x=r,y=mean*tHa2buAc))+geom_line(aes(group=field
   # geom_hline(yintercept = 0,linetype='dashed',col='red')+
   geom_smooth(method='gam',formula=y~s(x),col='blue',se=FALSE)+
   coord_cartesian(ylim=ylimVals$mean)
-p6 <- allEff[[3]] %>% ggplot(aes(x=r,y=logSD*tHa2buAc))+geom_line(aes(group=field),alpha=0.3)+
+p6 <- allEff[[3]] %>% ggplot(aes(x=r,y=log(exp(logSD)*tHa2buAc)))+geom_line(aes(group=field),alpha=0.3)+
   labs(x='Point Order',y=ylabSD)+
   # geom_hline(yintercept = 0,linetype='dashed',col='red')+
   geom_smooth(method='gam',formula=y~s(x),col='blue',se=FALSE)+

### 2021-04-22 runModels.R
Parallel runs

@@ -318,7 +318,6 @@ p6 <- allEff[[3]] %>% ggplot(aes(x=r,y=log(exp(logSD)*tHa2buAc)))+geom_line(aes(
 ggsave(paste0('./Figures/ModelSummary.png'),p,height=6,width=12,dpi=300)  
 
 #Get smoother from each field
-#Takes about 10 seconds
 
 getFiles <- datSource$filename[datSource$use] #Field names
 files <- paste0('./Figures/ModelCheck/',getFiles,' modList.Rdata') #File paths
@@ -370,20 +369,20 @@ sampleSmooth <- function(f,s,...){
 
 temp <- sampleSmooth(f=files,s=FALSE) #Mean smoother
 
-Nrep <- 3
-# temp2 <- replicate(Nrep,sampleSmooth(f=files,s=TRUE),simplify=FALSE)
-debugonce(sampleSmooth)
-debugonce(getSmooths)
-temp2 <- sampleSmooth(f=files,s=TRUE)
+# Nrep <- 3
+# # temp2 <- replicate(Nrep,sampleSmooth(f=files,s=TRUE),simplify=FALSE)
+# debugonce(sampleSmooth)
+# debugonce(getSmooths)
+# temp2 <- sampleSmooth(f=files,s=TRUE)
+# 
+# lapply(temp2,function(x) x$r) %>% set_names(paste0('s',1:Nrep)) %>% 
+#   do.call('rbind',.) %>%  rownames_to_column('rep') %>%   mutate(rep=gsub('\\.\\d{1,3}','',rep)) %>% 
+#   group_by(rep) %>% slice(1:3)
 
-
-lapply(temp2,function(x) x$r) %>% set_names(paste0('s',1:Nrep)) %>% 
-  do.call('rbind',.) %>%  rownames_to_column('rep') %>%   mutate(rep=gsub('\\.\\d{1,3}','',rep)) %>% 
-  group_by(rep) %>% slice(1:3)
-
-library(parallel)
-Nrep <- 30
-cluster <- makeCluster(8) 
+library(parallel) #Parallel version
+Nrep <- 100
+detectCores()
+cluster <- makeCluster(14) 
 clusterExport(cl = cluster,varlist='getSmooths', envir = .GlobalEnv) #Export function to clusters
 tempSamp <- parLapply(cl=cluster,1:Nrep,fun=sampleSmooth,f=files,s=TRUE)
[…]

### 2021-04-22 runModels.R
Updated figure

@@ -325,12 +325,13 @@ files <- paste0('./Figures/ModelCheck/',getFiles,' modList.Rdata') #File paths
 #Function to get predicted smoother values from list of functions, fit models to these, and return results
 # Used for getting 
 # f = files to draw from
-# s = sample from posterior?
-sampleSmooth <- function(f,s,...){
+# s = sample from posterior
+# m = marginalize across intercept
+sampleSmooth <- function(f,s,m,...){
   require(mgcv)
   require(tidyverse)
   
-  allSmooths <- lapply(f,getSmooths,margInt=c(FALSE,TRUE,TRUE),samp=s)
+  allSmooths <- lapply(f,getSmooths,margInt=m,samp=s)
   names(allSmooths) <- gsub(' ','-',f)
   names(allSmooths[[1]]) #Variables to get from allSmooths
   
@@ -367,7 +368,7 @@ sampleSmooth <- function(f,s,...){
   return(datList)
 }
 
-temp <- sampleSmooth(f=files,s=FALSE) #Mean smoother
+temp <- sampleSmooth(f=files,s=FALSE,m=c(TRUE,TRUE,TRUE)) #Mean smoother
 
 # Nrep <- 3
 # # temp2 <- replicate(Nrep,sampleSmooth(f=files,s=TRUE),simplify=FALSE)
@@ -380,50 +381,51 @@ temp <- sampleSmooth(f=files,s=FALSE) #Mean smoother
 #   group_by(rep) %>% slice(1:3)
 
 library(parallel) #Parallel version
-Nrep <- 100
+Nrep <- 200
 detectCores()
-cluster <- makeCluster(14) 
+cluster <- makeCluster(15) 
 clusterExport(cl = cluster,varlist='getSmooths', envir = .GlobalEnv) #Export function to clusters
-tempSamp <- parLapply(cl=cluster,1:Nrep,fun=sampleSmooth,f=files,s=TRUE)
+tempSamp <- parLapply(cl=cluster,1:Nrep,fun=sampleSmooth,f=files,s=TRUE,m=c(TRUE,TRUE,TRUE))
 beep(1)
 stopCluster(cluster)
 
 #Get range of variability for models 
 
+ylabMean <- 'Mean Yield (bushels/acre)'
+ylabSD <- 'log(SD Yield)'
+
[…]

### 2021-04-30 Manuscript/manuscript.Rmd
Updated figure
Prep for second model run

@@
   - This makes accurate and consistent classification of field boundaries very difficult
-  - Because of this, we used the following general categories for field boundaries:
-    1. Standard: thin ($<10$ m wide) grassy field boundary, often grassy road right-of-ways 
+  - Because of this, we define the following general categories for field boundaries:
+    1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide) 
     2. Wetland: permanent wetland, whose borders are largely unchanged from year-to-year
-    3. Forest/shrub: permanent windbreaks or remnant forests
-    4. Grass: larger grassy area (pasture or permanent seminatural grassland)
-    5. Other crop: different crop with little or no visible boundary between planted areas 
-    6. Bare: unplanted (fallow), flooded area, or unpaved roads
+    3. Shelterbelt: permanent windbreaks, shelterbelts, or remnant forests and shrublands
+    4. Other crop: annual crop or pasture with little or no visible boundary between planted areas 
+    5. Bare: unplanted (fallow), flooded area, staging yard or road with no planted boundary
+    6. Grassland: permanent seminatural grassland
+
+    STANDARD
+    WETLAND
+    SHELTERBELT
+    OTHERCROP
+    BARE
+    GRASSLAND
 

### 2021-04-30 createBoundaries.R

@@ -105,19 +105,26 @@ stopCluster(cluster)
 library(tidyverse)
 library(sf)
 
-shpFiles <- paste0('./Figures/FieldBoundaries/',dir('./Figures/FieldBoundaries',pattern=".shp",recursive=TRUE))
+setwd("~/Documents/yield-analysis-2021")
+
+rootPath <- "/media/rsamuel/Storage/geoData/Rasters/yieldData/csv files"
+# datSource <- data.frame(path=dir(rootPath,pattern=".csv",recursive=TRUE)) %>% 
+#   separate(path,c('grower','year','field'),sep="/",remove=FALSE) %>% 
+#   mutate(field=gsub('\\.csv','',field)) %>% unite(filename,c(grower:field),sep=' ',remove=FALSE) %>% 
+#   mutate(completed=filename %in% gsub(' boundary.shp','',dir('./Figures/FieldBoundaries',pattern=".shp",recursive=TRUE)))
+
+datSource <- read.csv('./Data/datSource.csv') #Read previous datsource file
+datSource <- datSource %>% mutate(boundaryComplete=file.exists(boundaryPath)) %>% #Has boundary been made already?
+  mutate(modelComplete=file.exists(modelPath)) #Has model already been run?
+# shpFiles <- paste0('./Figures/FieldBoundaries/',dir('./Figures/FieldBoundaries',pattern=".shp",recursive=TRUE)) #Paths to shapefiles
+shpFiles <- unique(datSource$boundaryPath[datSource$use]) #Only unique shapefiles used (some fields use a single year's boundary)
 
 makeLinestring <- function(shp){
   # shp <-shpFiles[1] 
   fieldName <- strsplit(shp,'/')[[1]][length(strsplit(shp,'/')[[1]])]
-  
   #Polygon
-  p <- read_sf(shp) %>% st_cast('MULTILINESTRING')
-  st_write(p,paste0('./Figures/FieldBoundariesLine/',fieldName),append=FALSE,driver='ESRI Shapefile')
+  p <- read_sf(shp) %>% st_cast('MULTILINESTRING') %>% mutate(type='STANDARD') %>% st_simplify(dTolerance=0.5)
+  st_write(p,paste0('./Figures/FieldBoundaryLines/',fieldName),append=FALSE,driver='ESRI Shapefile')
 }
-
-sapply(shpFiles[1:5],makeLinestring)
-
-
-
-
+makeLinestring(shpFiles[1])
+sapply(shpFiles,makeLinestring)

### 2021-04-30 runModels.R

@@ -23,11 +23,13 @@ source('helperFunctions.R')
 #   mutate(modelPath=paste0('./Figures/ModelCheck/',filename,' modList.Rdata'))  #Paths to saved model files
 # write.csv(datSource,'./Data/datSource_new.csv',row.names = FALSE)
 
-datSource <- read.csv('./Data/datSource.csv') #Read previous datasource file
-datSource <- datSource %>% mutate(boundaryComplete=file.exists(boundaryPath)) %>% #Has boundary been made already?
-  mutate(modelComplete=file.exists(modelPath)) #Has model already been run?
+datSource <- read.csv('./Data/datSource.csv') %>% #Read previous datasource file
+  mutate(boundaryComplete=file.exists(boundaryPath)) %>% #Has boundary been made already?
+  mutate(modelComplete=file.exists(modelPath)) %>% #Has model1 already been run?
+  mutate(modelComplete2=file.exists(modelPath2)) #Has model2 already been run?
 
-datSource %>% count(grower,year) 
+datSource %>% count(grower,year)
+datSource %>% group_by(grower) %>% summarize(nLoc=length(unique(field)),nYears=length(unique(year)),n=n()) 
 
 # set.seed(1) # Run subset of models
 # datSource <- datSource %>%
@@ -38,9 +40,9 @@ datSource %>% count(grower,year)
 
 #Function to run the ith model
 runModI <- function(i,dS,nSubSamp=50000){ 
-  # i <- 1 #Debugging
-  # dS <- datSource
-  # nSubSamp <- 50000
+  i <- 7 #Debugging
+  dS <- datSource
+  nSubSamp <- 50000
 
   if(dS$modelComplete[i]) return('Already completed')
   if(!dS$use[i]) return('Not used')
@@ -59,6 +61,8 @@ runModI <- function(i,dS,nSubSamp=50000){
   
   print('Reading in data')
   dat <- read.csv(csvPath,stringsAsFactors=TRUE,fileEncoding='latin1') 
+  
+  cropType <- unique(dat$Product) #Crop types
   
   #Takes 40 seconds with subsamp of 50000 using full 800000 samples from Alvin French's Al Jr Field
   
@@ -205,7 +209,8 @@ runModI <- function(i,dS,nSubSamp=50000){
[…]

### 2021-05-26 createBoundaries.R
Worked on new boundaries

@@ -139,14 +139,14 @@ makeLinestring <- function(shp,overwrite=FALSE){
   #Polygon
   p <- read_sf(shp) %>% st_cast('MULTILINESTRING') %>% mutate(type='STANDARD') %>% st_simplify(dTolerance=0.5)
   
-  if(!file.exists(filename)){
+  if(!file.exists(filename)){ #If file doesn't exist
     st_write(p,filename,append=FALSE,driver='ESRI Shapefile')
-  } else if((file.exists(filename)&overwrite)){
+  } else if((file.exists(filename)&overwrite)){ #If file exists and overwrite = TRUE
     print('Overwriting existing file')
     st_write(p,filename,append=FALSE,driver='ESRI Shapefile')
-  } else {
-    print('File already exists.')
+  } else { #Otherwise
+    print('File already exists. Skipping.')
   }
 }
 makeLinestring(shpFiles[1])
-sapply(shpFiles,makeLinestring)
+sapply(shpFiles,makeLinestring) #All of them

### 2021-05-26 runModels.R

@@ -31,27 +31,26 @@ datSource <- read.csv('./Data/datSource.csv') %>% #Read previous datasource file
   mutate(modelComplete2=file.exists(modelPath2)) #Has model2 already been run?
 
 #Get summary results from models
-
 datSource %>% count(grower,year)
 datSource %>% group_by(grower) %>% summarize(nLoc=length(unique(field)),nYears=length(unique(year)),n=n()) 
 
-#First set of models - no boundary types
-resultPaths <- gsub(' modList.Rdata',' results.txt',datSource$modelPath)[datSource$use]
-modInfo <- lapply(resultPaths,getModelInfo)
-sapply(modInfo,function(x) as.numeric(x$timeTaken,units='hours')) %>% summary() #Hours taken
-sapply(modInfo,function(x) x$percDevianceExplained) %>% summary() #Explained deviance
+# #First set of models - no boundary types
+# resultPaths <- gsub(' modList.Rdata',' results.txt',datSource$modelPath)[datSource$use]
+# modInfo <- lapply(resultPaths,getModelInfo)
+# sapply(modInfo,function(x) as.numeric(x$timeTaken,units='hours')) %>% summary() #Hours taken
+# sapply(modInfo,function(x) x$percDevianceExplained) %>% summary() #Explained deviance
 
-#Second set of models - boundary types included
-resultPaths <- gsub(' modList.Rdata',' results.txt',datSource$modelPath2)[datSource$use]
-modInfo <- lapply(resultPaths,getModelInfo)
-sum(!file.exists(datSource$modelPath2[datSource$use])) #24 models not completed
-sapply(modInfo,function(x) as.numeric(x$timeTaken,units='hours')) %>% summary() #Hours taken
-sapply(modInfo,function(x) x$percDevianceExplained) %>% summary() #Explained deviance
+# #Second set of models - boundary types included
+# resultPaths <- gsub(' modList.Rdata',' results.txt',datSource$modelPath2)[datSource$use]
+# modInfo <- lapply(resultPaths,getModelInfo)
+# sum(!file.exists(datSource$modelPath2[datSource$use])) #24 models not completed
+# sapply(modInfo,function(x) as.numeric(x$timeTaken,units='hours')) %>% summary() #Hours taken
+# sapply(modInfo,function(x) x$percDevianceExplained) %>% summary() #Explained deviance
 
[…]

### 2021-05-28 Manuscript/manuscript.Rmd
Updated field boundaries

@@
 
-- Field boundaries were automatically digitized, then manually checked using satellite imagery and classified land cover data 
+- Field boundaries were automatically digitized, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC 
   - Crop boundaries are flexible, and often change from year to year depending on planting and emergence conditions (e.g. flooding during some years)
@@
     1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide) 
-    2. Wetland: permanent wetland, whose borders are largely unchanged from year-to-year
-    3. Shelterbelt: permanent windbreaks, shelterbelts, or remnant forests and shrublands
+    2. Wetland: permanent wetland; borders are largely unchanged from year-to-year
+    3. Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
     4. Other crop: annual crop or pasture with little or no visible boundary between planted areas
-    5. Bare: unplanted (fallow), flooded area, staging yard or road with no planted boundary
-    6. Grassland: permanent seminatural grassland
+    5. Bare: unplanted, fallow, flooded area, temporary wetland, staging yard, oil and gas equipment, or road with no planted boundary
+    6. Grassland: permanent seminatural grassland or pasture (not in rotation)
 
-    STANDARD
-    WETLAND
-    SHELTERBELT
-    OTHERCROP
-    BARE
-    GRASSLAND
+    <!-- STANDARD -->
+    <!-- WETLAND -->
+    <!-- SHELTERBELT -->
+    <!-- OTHERCROP -->
+    <!-- BARE -->
+    <!-- GRASSLAND -->
 

### 2021-05-28 createBoundaries.R

@@ -113,7 +113,6 @@ for(i in 1:ceiling(length(useRows)/nproc)){ #Make files in batches of 2 before k
 # Turn polygons into linestrings ------------------------------------------
 
 #NOTE: do this after cleaning up initial polygons
-
 library(tidyverse)
 library(sf)
 

### 2021-05-28 helperFunctions.R

@@ -465,9 +465,11 @@ runModI <- function(i,dS,nSubSamp=50000,kPar=c(12,60,60,12,60,60),modelCheckDir=
 #Function to run the ith model, but with boundary type
 #useClosest = use only closest boundary? Otherwise, uses distance from all boundaries
 runModII <- function(i,dS,nSubSamp=50000,kPar=c(12,60,60,12,60,60),useClosest=TRUE,modelCheckDir='./Figures/ModelCheck2',resultsDir='./Figures/YieldMaps2'){ 
-  # i <- 102 #Debugging
+  # i <- 295 #Debugging
   # dS <- datSource
   # nSubSamp <- 50000
+  # useClosest <- TRUE
+  # kPar <- c(12,60,60,12,60,60)
   
   if(dS$modelComplete2[i]) return('Already completed')
   if(!dS$use[i]) return('Not used')
@@ -490,7 +492,6 @@ runModII <- function(i,dS,nSubSamp=50000,kPar=c(12,60,60,12,60,60),useClosest=TR
   cropType <- as.character(unique(dat$Product)) #Crop types
   
   #Takes 40 seconds with subsamp of 50000 using full 800000 samples from Alvin French's Al Jr Field
-  
   if(nrow(dat)>nSubSamp){
     #Limit to nSubSamp sequential samples
     dat <- dat %>% slice(round(seq(1,nrow(dat),length.out=nSubSamp)))
@@ -525,7 +526,7 @@ runModII <- function(i,dS,nSubSamp=50000,kPar=c(12,60,60,12,60,60),useClosest=TR
   if(useClosest){
     #Get distance and type of closest boundary
     dat <- dat %>% 
-      bind_cols(.,data.frame(t(apply(st_distance(dat,fieldEdge),1,function(x) c(dist=min(x),boundaryType=fieldEdge$type[which.min(x)]))))) %>% 
+      bind_cols(.,data.frame(t(apply(st_distance(dat,fieldEdge),1,function(x) c(dist=min(x,na.rm=TRUE),boundaryType=fieldEdge$type[which.min(x)]))))) %>% 
       mutate(dist=as.numeric(dist),boundaryType=factor(boundaryType))
   } else {
     #Get distance columns for each boundary type

### 2021-05-28 runModels.R

@@ -262,9 +262,9 @@ ggsave(paste0('./Figures/ModelSummary1a.png'),p,height=6,width=12,dpi=300)
 # Run second set of models - boundary type included ---------------------------
 
 # a <- Sys.time()
-# runModII(102,dS=datSource) #Trent Clark Swanny's 2019 (small field)
+# runModII(295,dS=datSource) 
 # Sys.time()-a
-# beep(1)
+# # beep(1)
 # debugonce(runModII)
 
 library(parallel)

### 2021-06-01 runModels.R
Ran model2 and extracted results

@@ -269,12 +269,13 @@ ggsave(paste0('./Figures/ModelSummary1a.png'),p,height=6,width=12,dpi=300)
 
 library(parallel)
 detectCores()
-cluster <- makeCluster(15) #10 procs uses about 30% of memory - could probably max it out 
+cluster <- makeCluster(15) #10 procs uses about 30% of memory - could probably max it out
 a <- Sys.time()
 parLapply(cl=cluster,1:nrow(datSource),runModII,dS=datSource,useClosest=TRUE) #All models
 beep(1)
 stopCluster(cluster)
 Sys.time()-a #Takes ~ 10 hrs for 50 models at 15 procs
+#Takes 2.364893 days to do all fields at 15 procs
 
 # Get smoother info from second set of models ---------------------------
 
@@ -286,13 +287,22 @@ est <- getPreds(paste0('./Figures/ModelCheck2/',datSource$filename[2],' modList.
 
 #Draw posterior samples at field 1
 Nsamp <- 100 #Number of samples
-samp <- lapply(1:Nsamp,
-          function(x){
-            getPreds(paste0('./Figures/ModelCheck2/',datSource$filename[2],' modList.Rdata'),margInt=c(FALSE,TRUE,TRUE),samp=TRUE)$distDat %>%
-              bind_rows(.id='dist_type')
-          }) %>% bind_rows(.id='N')
+# samp <- lapply(1:Nsamp,
+#           function(x){
+#             getPreds(paste0('./Figures/ModelCheck2/',datSource$filename[2],' modList.Rdata'),margInt=c(FALSE,TRUE,TRUE),samp=TRUE)$distDat %>%
+#               bind_rows(.id='dist_type')
+#           }) %>% bind_rows(.id='N')
+library(parallel)
+cluster <- makeCluster(15) #10 procs uses about 30% of memory - could probably max it out
+clusterExport(cluster,c('datSource'))
+samp <- parLapply(cl=cluster,1:Nsamp,fun=function(x){
+  require(tidyverse)
+  source('helperFunctions.R')
[…]

### 2021-06-01 runModels.R
Got overall smoothers

@@ -1,6 +1,6 @@
 #Run non-isotropic GAMs on multiple field years
 
-# Load required libraries ---------------------------------------------------------
+# Load everything ---------------------------------------------------------
 
 library(tidyverse)
 theme_set(theme_bw())
@@ -10,19 +10,34 @@ library(sf)
 library(beepr)
 source('helperFunctions.R')
 
-# Get paths to data -------------------------------------------------------
+# Get paths to data 
 
-# #Generate new dataSource dataframe - NOTE: NAMING SCHEME HAS CHANGED, SO THIS SHOULD BE UPDATED - MAY 13 2021
+# #Generate new dataSource dataframe
 # rootPath <- "/media/rsamuel/Storage/geoData/Rasters/yieldData/csv files" #Path to csv files
 # datSource <- data.frame(dataPath=dir(rootPath,pattern=".csv",recursive=TRUE,full.names = TRUE)) %>%
 #   mutate(path=gsub('/media/rsamuel/Storage/geoData/Rasters/yieldData/csv files/','',dataPath)) %>%
-#   separate(path,c('grower','year','field'),sep="/",remove=FALSE) %>% select(-path) %>%
-#   mutate(field=gsub('\\.csv','',field)) %>%
-#   unite(filename,c(grower:field),sep=' ',remove=FALSE) %>%
+#   separate(path,c('grower','field','year'),sep=" ",remove=FALSE) %>% select(-path) %>%
+#   mutate(year=gsub('\\.csv','',year)) %>%
+#   unite(filename,c(grower:year),sep=' ',remove=FALSE) %>%
 #   mutate(boundaryPath=paste0('./Figures/FieldBoundaries/',filename,' boundary.shp')) %>% #Paths to boundary shapefiles
 #   mutate(boundaryPath2=paste0('./Figures/FieldBoundarySegments/',filename,' boundary.shp')) %>%
 #   mutate(modelPath=paste0('./Figures/ModelCheck/',filename,' modList.Rdata')) %>% #Paths to saved model files
 #   mutate(modelPath2=paste0('./Figures/ModelCheck2/',filename,' modList.Rdata'))
+# 
+# #Read 10000 lines from each csv, and get counts of products - takes about 20 seconds
[…]

### 2021-06-02 helperFunctions.R
Created zero-th model runs

@@ -263,7 +263,7 @@ getPreds <- function(path,l=c(30,100,500),margInt=c(FALSE,FALSE,FALSE),samp=FALS
 
 #Function to run the ith model, using datSource ds, and sampling below nSubSamp points if necessary
 #kPar = basis dimensions for distance, E/N, and time smoothers + basis dimensions for logSD smoothers
-runModI <- function(i,dS,nSubSamp=50000,kPar=c(12,60,60,12,60,60),modelCheckDir='./Figures/ModelCheck',resultsDir='./Figures/YieldMaps'){ 
+runModI <- function(i,dS,nSubSamp=50000,kPar=c(12,60,60,12,60,60),modelCheckDir='./Figures/ModelCheck1',resultsDir='./Figures/YieldMaps'){ 
   if(dS$modelComplete[i]) return('Already completed')
   if(!dS$use[i]) return('Not used')
   
@@ -648,13 +648,183 @@ runModII <- function(i,dS,nSubSamp=50000,kPar=c(12,60,60,12,60,60),useClosest=TR
   gc()
 }
 
+#Function to run the ith model, but with no boundary included ("null" model)
+runMod0 <- function(i,dS,nSubSamp=50000,kPar=c(60,60,60,60),useClosest=TRUE,modelCheckDir='./Figures/ModelCheck0',resultsDir='./Figures/YieldMaps0',startCoefs=NULL){ 
+  i <- 1 #Debugging
+  dS <- datSource
+  nSubSamp <- 50000
+  useClosest <- TRUE
+  kPar <- c(60,60,60,60)
+  startCoefs <- NULL
+  startCoefs <- modList$coefs
+  
+  if(dS$modelComplete0[i]) return('Already completed')
+  if(!dS$use[i]) return('Not used')
+  
+  require(tidyverse)
+  theme_set(theme_bw())
+  require(mgcv)
+  require(sf)
+  require(ggpubr)
+  
+  source('helperFunctions.R')
+  
+  csvPath <- dS$dataPath[i]
+  fieldName <- dS$filename[i]
+  boundaryPath <- dS$boundaryPath[i]
+  fieldEdge <- read_sf(boundaryPath) %>% st_cast('MULTILINESTRING') #Read in boundary file
+  
+  print('Reading in data')
+  dat <- read.csv(csvPath,stringsAsFactors=TRUE,fileEncoding='latin1') 
+  
+  cropType <- as.character(unique(dat$Product)) #Crop types
+  
[…]

### 2021-06-02 runModels.R

@@ -21,8 +21,9 @@ source('helperFunctions.R')
 #   unite(filename,c(grower:year),sep=' ',remove=FALSE) %>%
 #   mutate(boundaryPath=paste0('./Figures/FieldBoundaries/',filename,' boundary.shp')) %>% #Paths to boundary shapefiles
 #   mutate(boundaryPath2=paste0('./Figures/FieldBoundarySegments/',filename,' boundary.shp')) %>%
-#   mutate(modelPath=paste0('./Figures/ModelCheck/',filename,' modList.Rdata')) %>% #Paths to saved model files
-#   mutate(modelPath2=paste0('./Figures/ModelCheck2/',filename,' modList.Rdata'))
+#   mutate(modelPath1=paste0('./Figures/ModelCheck1/',filename,' modList.Rdata')) %>% #Paths to saved model files
+#   mutate(modelPath2=paste0('./Figures/ModelCheck2/',filename,' modList.Rdata')) %>% 
+#   mutate(modelPath0=paste0('./Figures/ModelCheck0/',filename,' modList.Rdata')) 
 # 
 # #Read 10000 lines from each csv, and get counts of products - takes about 20 seconds
 # cropType <- sapply(datSource$dataPath,function(path){
@@ -42,8 +43,9 @@ source('helperFunctions.R')
 
 datSource <- read.csv('./Data/datSource.csv') %>% #Read previous datasource file
   mutate(boundaryComplete=file.exists(boundaryPath)) %>% #Has boundary been made already?
-  mutate(modelComplete=file.exists(modelPath)) %>% #Has model1 already been run?
-  mutate(modelComplete2=file.exists(modelPath2)) #Has model2 already been run?
+  mutate(modelComplete1=file.exists(modelPath1)) %>% #Has model1 already been run?
+  mutate(modelComplete2=file.exists(modelPath2)) %>% #Has model2 already been run?
+  mutate(modelComplete0=file.exists(modelPath0)) #Has model0 already been run?
 
 #Get summary results from models
 datSource %>% pull(use) %>% sum()
@@ -52,7 +54,7 @@ datSource %>% group_by(grower) %>% summarize(nLoc=length(unique(field)),nYears=l
[…]

### 2021-06-02 Manuscript/manuscript.Rmd
Worked on manuscript

@@
   - Boundary examples: flower strip studies, hedgerows
-  - In North America, large financial incentives to make fields large and homogeneous, especially with large harvesting and planting equipment
+  - In North America, there are financial incentives to make fields large and homogeneous, especially with large agricultural equipment
   - Crop edge effects cause low yields at the margins of fields because of late emergence, poor microclimate, and competition with weeds
-  - Ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
-  - Therefore, there should be a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
+  - At the same time, ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
+  - This suggests that yield may be maximized at distances between the edge and centre of the field, where ecosystem services cancel out negative edge effects
+
+  <!-- - This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services -->
 
@@
 - In this paper we ask:
-  1. How does crop yield mean and variance change with distance from the edge of field?
+  1. How does crop yield change with distance from the edge of field?
   2. Does this depend on type of field edge? 
-  3. Given this relationship, are there upper and lower bounds for field size if a farmer wanted to maximize average yield, or minimize variance?
+  3. Is there an intermediate distance where yield is maximized or variance is minimized?
 
@@
 
-- Yield data was collected in discrete rectangles of the same length as the data interval (distance = speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m)
+- Yield data was collected in rectangles of the same length as the data interval (distance = speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m)
   - We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
-  - Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location
-  - Seeding and application rates were constant with each field, so we did not consider inputs in our analysis
+  - Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
+  - Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
+  - We used dry yield (tonnes of seed/hectare after accounting for crop moisture) as our measure of crop yield
+  - Due to the high number of samples at each field, we sub-sampled to 50,000 data points per field to reduce computation time
 
-- Field boundaries were automatically digitized, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC 
-  - Crop boundaries are flexible, and often change from year to year depending on planting and emergence conditions (e.g. flooding during some years)
-  - Additionally, seminatural features often change from year to year
+- Field boundaries were automatically digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC 
+  - Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g. flooding during some years)
+  - Additionally, seminatural features often change yearly
     - Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years
     - Grass boundaries can change if fields are used for as haying or pasture during crop rotation
-  - This makes accurate and consistent classification of field boundaries very difficult
-  - Because of this, we define the following general categories for field boundaries:
+  - This makes accurate and consistent classification of field boundaries difficult
+  - We defined the following general categories for field boundaries:
     1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide) 
@@
     4. Other crop: annual crop or pasture with little or no visible boundary between planted areas
-    5. Bare: unplanted, fallow, flooded area, temporary wetland, staging yard, oil and gas equipment, or road with no planted boundary
+    5. Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
     6. Grassland: permanent seminatural grassland or pasture (not in rotation)
@@
 - At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor
-  - Crop yield varies within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices such as strip farming
-  - While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields), this does not address sensor drift over time that occurs within fields
-  - The yield monitors may record lower yields as sensors accumulate debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
-  - Additionally, ground speed is known to be extremely important to yield monitor accuracy [@arslan2002] 
-  - To address this, we fit the following model:
+  - Ground speed is known to be extremely important to yield monitor accuracy [@arslan2002], with low ground speed registering higher yields 
+  - Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices 
+    <!-- - such as strip farming -->
+  - While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time  within fields
+  - This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
+
+  - To model this, we fit the following model:
 
   \begin{equation}
-  log(yield) ~ Normal (\mu, \sigma)
-  \mu = Intercept + log(Polygon Size) + f(Distance from Edge) + f(Easting, Northing) + f(Harvest Sequence)
-  log(\sigma) = Intercept + log(Polygon Size) + f(Distance from Edge) + f(Easting, Northing) + f(Harvest Sequence)
+  sqrt(yield) ~ Normal (\mu, \sigma)
+  \mu = Intercept + log(Polygon Size) + f(Distance from Edge_i, b=12)_i + f(Easting, Northing, b=60) + f(Harvest Sequence, b=60)
+  log(\sigma) = Intercept + log(Polygon Size) + f(Distance from Edge, b=12) + f(Easting, Northing, b=60) + f(Harvest Sequence, b=60)
   \end{equation}
+
+  - where:
+  1. Polygon Size = distance traveled x width of header bar ($m^2$)
+  2. Distance from Edge = distance from field edge type _i_ (m)
+  3. Easting, Northing = distance from centroid of field (m)
+  4. Harvest Sequence = order of harvest within field (1--N points)
+  5. f(x,b) = penalized thin-plate regression spline, where x is the predictor and b is the number of basis dimensions
+
+- This model accounts for a) combine speed differences, b) edge effects, c) within-field spatial variation not related to edges, and d) temporal differences in harvest 
+  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (e.g. Kriging), but this was computationally infeasible with 50,000 data points per field
+  - Penalized regression splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with substantially reduced computation time
+  - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
+  - A smoother was not used for polygon size because the relationship was easily approximated with a log-linear relationship using a single slope term
+  - All models were fit in _R_ using the _mgcv_ library, and figures were created with _ggplot2_ and _ggpubr_
+
+- Most models still had a large amount of spatio-temporal autocorrelation in residuals
+  - We basically just ignored this
+
+- To combine results from all field-level models while also accounting for field-level uncertainty, we used an approach similar to bootstrapping of hierarchical mixed effects models
+  - The approach was as follows
+  1. Extract posterior samples of smoother parameters from each field-level model
+  2. Fit meta-model of all smoothers, and save parameters
+  3. Repeat 1000 times, and calculate credible intervals (10-90th percentiles of draws)
+
+- We compared this
 

### 2021-06-02 gam_TMB/gamTests.R
Trying out GAMs in TMB

@@ -1,0 +1,28 @@
+#GAM testing using TMB
+
+library(TMB)
+library(mgcv)
+library(tidyverse)
+library(gamair)
+data(engine)
+
+#Cubic splines, as in example ------------------------
+
+s1 <- smoothCon(object=s(wear,bs='cr'),data=engine)
+
+s1[[1]]$X #Model matrix
+s1[[1]]$S #Penalty matrix
+
+# Thin-plate splines ------------------------------------------------------
+
+s1 <- smoothCon(object=s(wear,bs='cs'),data=engine)
+
+s1[[1]]$X #Model matrix
+s1[[1]]$S #Penalty matrix #Two zero columns. Not sure this is going to work
+
+# Thin-plate splines ------------------------------------------------------
+
+s1 <- smoothCon(object=s(wear,bs='tp'),data=engine)
+
+s1[[1]]$X #Model matrix
+s1[[1]]$S #Penalty matrix #Two zero columns. Not sure this is going to work

### 2021-06-02 gam_TMB/pSplines/pSplines.Rmd
Added TMB experiments

```{r echo=FALSE, eval=TRUE, results='hide',include=FALSE}
library(TMB)
library(mgcv)
library(Matrix)
source("../utils.R")
source("pSplines.R")
```

## **GAM model based on P-splines**
In this example you learn:

* How to estimate a model with P-splines (penalized splines) in TMB.
* How to set up the spline part of the model in the `mgcv` package. 
* How to extract standard errors of a spline in TMB.
* How to input "complex datastructures" from R to C++ via _sparse matrices_.

***
#### Description of the practical situation
Let $Y$ be the biodiversity, as given by the variable `Richness` in the data. We assume that $Y \sim N(\mu,\sigma^2)$, were 
\begin{equation}
\mu = \beta_0 +  f_1(rock) +f_2(litter) + f_3(soil) + f_4(fallPrec) + f_5(sprTmax).\qquad (1)
\label{def:mu}
\end{equation}
Here, $f_1,\ldots,f_5$ are splines, modelled as function of covariates `rock`, `litter`,... (see data excerpt below).
See page 63 in Zuur (2009) for a similar example with use of the same data.
<details> <summary> "Penalizing the splines": how to get an appropriate level of smoothness?</summary>
We let the coefficients ($\beta$) of the $i$th spline have a multivariate normal distribution with precision matrix $\lambda_i \pmb{S}_i$, where $\lambda_i$ is a penalization parameter and $\pmb{S}_i$ is a penalization matrix extracted from the R-package `mgcv`. The matrices $\pmb{S}_i$ are singular, and hence the multivariate normal distributions are degenerate. However, when bringing in data through the likelihood, the (posterior) precision matrix becomes non-singular. See [Wood (2017)](https://books.google.no/books?id=JTkkDwAAQBAJ&printsec=frontcover&dq=generalized+additive+models+wood+2017&hl=no&sa=X&ved=0ahUKEwjp4o2sicHbAhWEFZoKHaHYBQ0Q6AEIJzAA#v=onepage&q=generalized%20additive%20models%20wood%202017&f=false) for details.
Implementation details are given in under the section "C++ code step-by-step" below.

```{r,echo=FALSE, comment="",fig.cap="Figure: Illustration of estimated splines with (red) and without (black) the penalization.", out.width = "300px"}
knitr::include_graphics("withAndWithout.png")
```

</details>

<details> <summary>Data excerpt</summary>
```{r,echo=FALSE, comment=""}
head(Vegetation,5)
```
</details>

***

#### Full code for example
<details> <summary> R code</summary>
```{r,echo=FALSE, comment=""}
include_source("pSplines.R", linesToInclud=c(1:3, 10:999))
```
</details>
<details> <summary> C++ code</summary>
```{r,echo=FALSE, comment=""}
include_source("pSplines.cpp", linesToInclud=c(1:999))
```
</details>

***

#### C++ code step-by-step
<details>  
  <summary>Read data from R</summary>
```c++
    DATA_VECTOR(Y);       //The response
    DATA_MATRIX(X);       //Design matrix for splines
    DATA_SPARSE_MATRIX(S);//Penalization matrix diag(S1,S2,S3,S4,S5) without
                          //storing off-diagonal zeros.
    DATA_VECTOR(Sdims);   //Dimensions of S1,S2,S3,S4 and S5
    DATA_INTEGER(flag);     //Used for calculating the normalizing constant of the prior for spline regression coefficients on the R side

```

The `X` matrix combines the 5 design matrices for the 5 splines.
$$
X =\left[X_1,X_2,X_3,X_4,X_5\right]
= \underbrace{
                \left.\left( 
                      \begin{array}{cc|cc|cc|cc}
                             1.2&\cdots &4.3&\cdots &5.0&\cdots &\cdots&0.2\\
                             -.2&\cdots &1.3&\cdots &2.0&\cdots &\cdots&9.1\\
                             3.1&\cdots &0.0&\cdots &1.2&\cdots &\cdots&-20.1\\
                             \vdots&&&&&&\vdots&\\
                             4.2&\cdots &9.1&\cdots &2.3&\cdots &\cdots&1.3
                      \end{array}
                \right)\right\}
              }_{\text{Design matrices for 5 splines stacked columnwise}} 
              \,\text{ observations }
$$
</details>
<details>
  <summary>Read parameters from R</summary>
```c++
    PARAMETER(beta0);       //Intercept
    PARAMETER_VECTOR(beta); //Spline regression parameters (5 splines stacked)
    PARAMETER_VECTOR(log_lambda);// Log penalization parameters (for 5 splines)
    PARAMETER(log_sigma);   // log(SD) of measurement error
```
</details>
<details>

<summary>Transform the standard deviation, $\sigma$, and the penalize parameter, $\lambda$, from log scale</summary>
```c++
    Type sigma = exp(log_sigma);
    vector<Type> lambda = exp(log_lambda);
```
</details>
<details>

<summary>Define the objective function</summary>
```c++
    Type nll = 0.0;
```
</details>
<details>

  <summary>Calculate the prior contribution from each of the 5 spline</summary>
  This is where the "complicated stuff" is going on. The following code does two things:

  * Retrieve the five $S$ matrices and $\beta$ vectors corresponding to each spline. 
  * Calculate the (multivariate) normal prior on each spline coeff vector $\beta_i$ (with precission matrix $\lambda_i S_i$).
```c++
  int k=0;  // Counter
  for(int i=0;i<Sdims.size();i++){
    int m_i = Sdims(i);
    vector<Type> beta_i = beta.segment(k,m_i);       // Retrieve beta_i
    SparseMatrix<Type> S_i = S.block(k,k,m_i,m_i);   // Retrieve S_i
    nll -= Type(0.5)*m_i*log_lambda(i) - 0.5*lambda(i)*GMRF(S_i).Quadform(beta_i);
    k += m_i;
  }
```

* `beta.segment(k,m_i)` returns elements $\beta_k,\beta_{k+1},\ldots,\beta_{k+m_i}$.  
* `S.block(k,k,m_i,m_i)` returns submatrix of $S$ with upper left corner $(k,k)$ and lower
  right corner $(k+m_i,k+m_i)$.
* The contribution from the $i$th spline is $0.5m_i\log(\lambda_i) - 0.5\lambda_i\beta_i^T S_i \beta_i$. A bit of explanation
is needed here. This prior on $\beta$ is a zero mean Gaussian distribution with precission (inverse covariance) matrix
$\lambda_iS_i$. Note that $S_i$  is not full rank, so we are dealing with a degenerate mutivariate distribution.
The part $0.5m_i\log(\lambda_i)$ is comming from the normalizing constant of the multivariate mean Gaussian distribution. A reference is needed here........
* `GMRF(S_i).Quadform(beta_i)` is the quadratic form $\beta_i^T S_i \beta_i$ (it is hard
to get the compiler to accept the  code `beta_i*(S_i*beta_i)  `).
</details>
<details>

  <summary>Return the prior for the spline regression coefficents on request </summary>
```c++
  if (flag == 0) return nll;
```
</details>
<details>

  <summary>Calculate the linear predictor, $\mu$, defined in eqn. (1).   </summary>
```c++
    vector<Type> mu(Y.size());
    mu = beta0 + X*beta;
```

* $X\beta=\sum_{j=1}^5X_j\beta_j$ gives the sum of all 5 splines (see above).
</details>
<details>

<summary>Calculate the negative log likelihood $-\sum_{i=1}^{n}\log P(Y_i|\beta_0,\beta,\sigma)$</summary>
```c++
    for(int i=0; i<Y.size(); i++){
      nll -= dnorm(Y(i), mu(i), sigma, true);
    }
```
</details>
<details>

<summary>Return the objective function </summary>
```c++
  return nll;
```
</details>

***

#### R code step-by-step
<details>  
  <summary>Include libraries</summary>
```r
  library(TMB)
  library(mgcv) #Use gam
  library(Matrix) #Use sparse matrices
```
</details>
<details>  
  <summary>Compile and load the C++ code</summary>
```r
  compile("pSplines.cpp")
  dyn.load(dynlib("pSplines"))
```
</details>
<details>  
  <summary>Read data</summary>
```r
  Vegetation <- read.table(file = "Vegetation.txt", header = TRUE, dec = ".")
  Vegetation = Vegetation[!is.na(Vegetation$Richness),]
```
</details>
<details>  
  <summary>Set up spline structure by using `mgcv`</summary>
```r
  gam_setup = gam(Richness ~ s(ROCK, bs = "cs") +
        s(LITTER, bs = "cs") + s(BARESOIL, bs = "cs") +
        s(FallPrec, bs = "cs") + s(SprTmax, bs = "cs"),
      data = Vegetation,fit=FALSE)
```
* `fit=FALSE` is because you only want `mgcv` to set up the spline bases,
   not fit the model.
</details>
<details>  
  <summary>Extract penalization matrices and combine them in a sparse block diagonal matrix</summary>
```r
  S_ROCK = gam_setup$smooth[[1]]$S[[1]]
  S_LITTER = gam_setup$smooth[[2]]$S[[1]]
  S_BARESOIL = gam_setup$smooth[[3]]$S[[1]]
  S_FallPrec = gam_setup$smooth[[4]]$S[[1]]
  S_SprTmax = gam_setup$smooth[[5]]$S[[1]]
  S_list = list(S_ROCK,S_LITTER,S_BARESOIL,S_FallPrec,S_SprTmax)
  S_combined = .bdiag(S_list)         # join S's in sparse matrix
  Sdims = unlist(lapply(S_list,nrow)) # Find dimension of each S
```
</details>
<details>  
  <summary>Construct the data list which is given to TMB (C++)</summary>
```r
data = list(Y = Vegetation$Richness, # Response
            X = gam_setup$X[,-1],  # Design matrix, without intercept
            S = S_combined,      # Combined penalty matrix
            Sdims = Sdims,
            flag = 1)
```
</details>
<details>  
  <summary>Construct the parameter list with initial values which is given to TMB (C++)</summary>
```r
  par = list(
    beta0 = 0,  # Intercept
    beta = rep(0,sum(Sdims)),  # Spline coefficients
    log_lambda = rep(rep(0,length(Sdims))), #Log spline penalization coefficients
    log_sigma = 0
  )
```
</details>
<details>  
  <summary>Fit the model</summary>
```r
  obj = MakeADFun(data = data, parameters = par,random="beta",DLL = "pSplines")
  obj <- normalize(obj, flag="flag")  #Inform obj that we calculate the prior normalizing constant in R. 
  opt = nlminb(obj$par,obj$fn,obj$gr)
  rep = sdreport(obj)
```
</details>

***
#### Exercise

1. There are 5 penalty matrices (S-matrices) in this example that are sent
to the .cpp code jointly as a blockdiagonal sparse matrix `S_combined`.
Understand this mechanism by looking at the following miniature example:
<details>  
  <summary>R code to be run line by line</summary>
```r
  library(Matrix)
  M1 = matrix(1:4,2,2)               # ordinary 2x2 matrix
  M2 = matrix(5:8,2,2)               # ordinary 2x2 matrix
  M_block_diag = .bdiag(list(M1,M2)) # Block diagonal (sparse) matrix
  data.class(M_block_diag)           # Check data.class
  print(M_block_diag)                # dots means 0 value
```
</details>

2. Plot the estimated splines with 95% confidence intervalls.
<details>
  <summary>Solution</summary>
    First we define a design matrix for the splines we want to plot:
    ```r
      #R code
      ROCK=min(Vegetation$ROCK):max(Vegetation$ROCK)
      LITTER=min(Vegetation$LITTER):max(Vegetation$LITTER)
      BARESOIL=min(Vegetation$BARESOIL):max(Vegetation$BARESOIL)
      FallPrec=min(Vegetation$FallPrec):max(Vegetation$FallPrec)
      SprTmax=min(Vegetation$SprTmax):max(Vegetation$SprTmax)

      rockReport = PredictMat(gam_setup$smooth[[1]],data = data.frame(ROCK))
      litterReport = PredictMat(gam_setup$smooth[[2]],data = data.frame(LITTER)) 
      soilReport = PredictMat(gam_setup$smooth[[3]],data = data.frame(BARESOIL)) 
      fallReport = PredictMat(gam_setup$smooth[[4]],data = data.frame(FallPrec)) 
      sprReport = PredictMat(gam_setup$smooth[[5]],data = data.frame(SprTmax))

      designMatrixForReport = list(rockReport,litterReport,soilReport,fallReport,sprReport)
      data$designMatrixForReport = .bdiag(designMatrixForReport) 
    ```

    Then we estimate those splines in TMB when we estimate the model:
    ```c++
        //c++ code
        DATA_SPARSE_MATRIX(designMatrixForReport); //place were the data is included

        vector<Type> splineForReport = designMatrixForReport*beta; //place just before return nll
        ADREPORT(splineForReport);

    ```
    Then we extract the estimated splines with standard errors, and plot the splines:

    ```{r, fig.height=7, fig.width=7}
      #R code
      muSpline = rep$value[names(rep$value)=="splineForReport"]
      sdSpline<-rep$sd[names(rep$value)=="splineForReport"]

      par(mfrow=c(2,3))
      start = 1
      stop = start + length(ROCK) -1
      plot(ROCK, muSpline[start:stop], lty=1,type = 'l',ylim = c(-6,5),ylab = "f(rock)",main = "Spline for ROCK")
      lines(ROCK, muSpline[start:stop]- 1.96*sdSpline[start:stop], lty=2)
      lines(ROCK, muSpline[start:stop]+ 1.96*sdSpline[start:stop], lty=2)
      abline(h = 0)

      start = stop +1
      stop = start+ length(LITTER)-1
      plot(LITTER, muSpline[start:stop], lty=1,type = 'l',ylim = c(-6,5),ylab = "f(litter)",main = "Spline for LITTER")
      lines(LITTER, muSpline[start:stop]- 1.96*sdSpline[start:stop], lty=2)
      lines(LITTER, muSpline[start:stop]+ 1.96*sdSpline[start:stop], lty=2)
      abline(h = 0)

      start = stop +1
      stop = start+ length(BARESOIL)-1
      plot(BARESOIL, muSpline[start:stop], lty=1,type = 'l',ylim = c(-6,5),ylab = "f(soil)",main = "Spline for BARESOIL")
      lines(BARESOIL, muSpline[start:stop]- 1.96*sdSpline[start:stop], lty=2)
      lines(BARESOIL, muSpline[start:stop]+ 1.96*sdSpline[start:stop], lty=2)
      abline(h = 0)

      start = stop +1
      stop = start+ length(FallPrec)-1
      plot(FallPrec, muSpline[start:stop], lty=1,type = 'l',ylim = c(-6,5),ylab = "f(fallPrec)",main = "Spline for FallPrec")
      lines(FallPrec, muSpline[start:stop]- 1.96*sdSpline[start:stop], lty=2)
      lines(FallPrec, muSpline[start:stop]+ 1.96*sdSpline[start:stop], lty=2)
      abline(h = 0)

      start = stop +1
      stop = start+ length(SprTmax)-1
      plot(SprTmax, muSpline[start:stop], lty=1,type = 'l',ylim = c(-6,5),ylab = "f(sprTMax)",main = "Spline for SprTmax")
      lines(SprTmax, muSpline[start:stop]- 1.96*sdSpline[start:stop], lty=2)
      lines(SprTmax, muSpline[start:stop]+ 1.96*sdSpline[start:stop], lty=2)
      abline(h = 0)
    ```
</details>

</details>

3. Verify you get the same results using the mgcv package.
  <details>
    <summary>Solution</summary>
    ```{r, fig.height=7, fig.width=7}
        gam_setup = gam(Richness ~ s(ROCK, bs = "cs") +
                          s(LITTER, bs = "cs") + s(BARESOIL, bs = "cs") +
                          s(FallPrec, bs = "cs") + s(SprTmax, bs = "cs"),
                        data = Vegetation,method = "ML",fit=TRUE) #Note that we use maximum likelihood method in the gam function.
        par(mfrow=c(2,3))
        plot(gam_setup)
    ```
</details>

4. Include `fallPrec` as a linear effect.
  <details>
    <summary>Solution</summary>
    On the R side we change the gam_setup:
    ```r
      gam_setup = gam(Richness ~ FallPrec+s(ROCK, bs = "cs") +
                        s(LITTER, bs = "cs") + s(BARESOIL, bs = "cs") +
                         s(SprTmax, bs = "cs"),
                      data = Vegetation,fit=FALSE)
    ```
    and provide TMB with separate design matrices for spline and linear covariates:
    ```r
    data = list(Y = Vegetation$Richness, # Response
            #X = gam_setup$X[,-1],
            Xlinear = gam_setup$X[,c(1,2)],  # Design matrix for intercept and linear effect
            Xsplines = gam_setup$X[,-c(1,2)],  # Design matrix, without intercept and linear effect
            S = S_combined,      # Combined penalty matrix
            Sdims = Sdims,
            designMatrixForReport = .bdiag(designMatrixForReport),
            flag = 1)
    ```
    and modify the parameters:
    ```r
    par = list(
          #beta0 = 0,  
          #beta = rep(0,sum(Sdims)),
          betaLinear = c(0,0),  # Intercept and linear effect
          betaSplines = rep(0,sum(Sdims)),  # Spline coefficients
          log_lambda = rep(rep(0,length(Sdims))), #Log spline penalization coefficients
          log_sigma = 0 
    )
    ```

    On the C++ side we include the two design matrices:
    ```c++
      //DATA_MATRIX(X);
      DATA_MATRIX(Xsplines);       //Design matrix for splines
      DATA_MATRIX(Xlinear);       //Design matrix for linear effects and intercept
    ```
    and modify how the linear predictor is calculated:
    ```c++
      //mu = beta0 + X*beta;
      mu = Xlinear*betaLinear;
      mu += Xsplines*betaSplines;
    ```
</details>

### 2021-06-02 gam_TMB/simpleExample/simpleExample.Rmd
Added TMB experiments

```{r echo=FALSE, eval=TRUE, results='hide',include=FALSE}
library(TMB)
library(mgcv)
library(Matrix)
source("../utils.R")
source("simpleExample.R")
```

## **Simple example ** 
In this example you learn:

* How to set up and fit a linear regression model in TMB.
* Basis functionality of TMB
* Some useful C++ details if you have little experience with C++

***
#### Description of the practical situation
Nitrogen ratios (NL) in whale theets are assumed to be a linear function of the age of the individual. The model is $\text{NL}_{age} \sim N(\mu_{age},\sigma^2)$, where $\mu_{age} = \beta_0 + \beta_{age}age$.
This example is found on page 27 in Zuur (2006).
</details>

<details> <summary> Overview of data</summary>
The five first records:
```{r,echo=FALSE, comment=""}
TN = read.table("TeethNitrogen.txt",header = T)
head(TN,5)
```
</details>

***

#### Full code for example
<details> <summary> R code</summary>
```{r,echo=FALSE, comment=""}
include_source("simpleExample.R", linesToInclud=c(1:21))
```
</details>
<details> <summary> C++ code</summary>
```{r,echo=FALSE, comment=""}
include_source("simpleExample.cpp", linesToInclud=c(1:999))
```
</details>

***
#### C++ code step-by-step
<details>
<summary>Read data from R</summary>
```c++
DATA_VECTOR(NL);  // Response
DATA_VECTOR(x);   // Covariate

```
</details>

<details>
<summary>Read parameters from R</summary>
```c++
PARAMETER(beta0);
PARAMETER(beta1);
PARAMETER(logSigma);
```
</details>

<details>
<summary>Transform some of the parameters. </summary>
```c++
Type sigma = exp(logSigma);
```
* For numerical and statistical reasons we let the underlying parameter be
   `logSigma`.
* In a C++ every variable must explicitely be given in a type. In TMB 
  all variables should as a general rule have type `Type` (for a long and    
  compicated reason). Integers of `int` are also commonly used as counting
  variables and indices.
</details>

<details>
<summary>Define the objective function (where we will later store) 
$\text{nll}=-\log P(\text{NL}|\pmb{\beta}, \sigma)$. </summary>
```c++
Type nll = 0; //negative log-likelihood
```
* Initialize to 0.
</details>

<details>
<summary>Calculates the regression line </summary>
```c++
vector<Type> mu = beta0 + x*beta1;
```
* This is a vectorized calculation. 
* `vector<Type>` is the vector type in TMB.
  (Similary you can have `matrix<Type>` and `array<Type>`)
</details>

<details>
<summary>Calculates the likelihood contribution from the observations </summary>
```c++
nll -= sum(dnorm(NL,mu,sigma,true));
```
* `dnorm(NL,mu,sigma,true)` is normal density (vectorized) in R-style notation.
* Argument `true` returns log-density (rather than density).
* `sum()` is adding the vector as an alternative to using a loop.
* `-=` subtracts the right hand side from `nll` (which we have initialized to    0).

</details>

<details>
<summary>Return the objective function </summary>
```c++
return nll;
```
* At this point we return the objective function value to R.
</details>

***

#### R code step-by-step
<details>  
<summary>Include the TMB library</summary>
```r
library(TMB)
```
</details>

<details>  
<summary>Compile and load the C++ code</summary>
```r
compile("simpleExample.cpp")  # Compile the C++ file
dyn.load("simpleExample")     # Link C++ file into R session
```
</details>

<details>  
<summary>Read data from file</summary>
```r
TN = read.table("TeethNitrogen.txt",header = T)
```
</details>

<details>  
<summary> Construct the data list which is given to TMB (C++)</summary>
```r
data = list()
data$NL = TN$X15N[TN$Tooth=="Moby"]
data$x = TN$Age[TN$Tooth=="Moby"]
```
* Elements of this list are picked up on the C++ side as for instance
  `DATA_VECTOR(NL)`.
* Names and order in the list must match exactly the C++ side.
</details>

<details>  
<summary> Construct the parameter list of initial values for parameters (C++)</summary>
```r
parameters = list(
beta0 = 0,
beta1 = 0,
logSigma = 0
)
```
* Serves the purpose of defining (on the R side) actually what the parameters in the model are.
* Names and order in the list must match exactly the C++ side (`PARAMETER` objects).
* The numerical values given are for practical purposes the initial values given to the function minimizer (eg. `nlminb`).

</details>

<details>  
<summary> Sets up the model (key step)</summary>
```r
obj = MakeADFun(data,parameters,DLL = "simpleExample")  # Link data and model
```
* Pulls together data, parameters, and C++ code (via `DLL`)
* Sets up data structures internally, and may take time for large models (not this one).
* Runs the objective function, and may crash R if your objective function for instance tries to access an array outside the legal range.
* Read the help page `help(MakeADFun)` to learn more features. 
</details>

<details>  
<summary> Fit the model</summary>
```r
opt = nlminb(obj$par,obj$fn, obj$gr)   # Estimate parameters
rep = sdreport(obj)                    # Evaluate standard deviations
```
</details>

***
#### Exercise

1. 
<details>  
<summary> Go through the code and confirm that you understand what is happening. Plot the fitted regression line together with the data.
</summary>
    ```{r, fig.height=4, fig.width=4}
        rep$par.fixed
        plot(data$x,data$NL,xla = "Age", ylab ="NL")
        abline(a = rep$par.fixed[1], b = rep$par.fixed[2])
    ```
</details>

2. 
<details>  
<summary> Evaluate the objective function value 
at $\beta_0=0.3$, $\beta_{age}=0.4$ and $\log(\sigma)=0.2$.
</summary>
```r
obj$fn(c(0.3,0.4,0.2))
```
</details>

3. 
<details>  
<summary> Evaluate the gradient (derivative with respect to
the parameters) of the objective function at $\beta_0=0.3$, $\beta_{age}=0.4$ and $\log(\sigma)=0.2$.
</summary>
```r
obj$gr(c(0.3,0.4,0.2))
```
</details>

### 2021-06-02 gam_TMB/gamTests.R

@@ -6,6 +6,8 @@ library(tidyverse)
 library(gamair)
 data(engine)
 
+plot(wear~size,data=engine,)
+
 #Cubic splines, as in example ------------------------
 
 s1 <- smoothCon(object=s(wear,bs='cr'),data=engine)
@@ -13,9 +15,11 @@ s1 <- smoothCon(object=s(wear,bs='cr'),data=engine)
 s1[[1]]$X #Model matrix
 s1[[1]]$S #Penalty matrix
 
-# Thin-plate splines ------------------------------------------------------
 
-s1 <- smoothCon(object=s(wear,bs='cs'),data=engine)
+
+# Cubic splines ------------------------------------------------------
+
+s1 <- smoothCon(object=s(wear,bs='tr'),data=engine)
 
 s1[[1]]$X #Model matrix
 s1[[1]]$S #Penalty matrix #Two zero columns. Not sure this is going to work

### 2021-06-02 gam_TMB/pSplines/pSplines.R

@@ -1,0 +1,121 @@
+setwd("~/Projects/UofC postdoc/TMB Meeting Group/tmb-case-studies/pSplines")
+
+library(TMB)
+library(mgcv) #Use gam
+library(Matrix) #Use sparse matrices
+
+# Tutorial: sparse matrices in R and TMB
+M = matrix(1:4,2,2)   # ordinary 2x2 matrix
+M_block_diag = .bdiag(list(M,M)) # Block diagonal (sparse) matrix
+data.class(M_block_diag)     # Check data.class
+print(M_block_diag)    # dots means 0 value
+
+#Load the data and compile c++ code------
+Vegetation <- read.table(file = "Vegetation.txt", header = TRUE, dec = ".")
+Vegetation = Vegetation[!is.na(Vegetation$Richness),]
+
+compile("pSplines.cpp")
+dyn.load(dynlib("pSplines"))
+#----------------------------------------
+
+#Set up spline structure by using mgcv---
+gam_setup = gam(Richness ~ s(ROCK, bs = "ts") +
+      s(LITTER, bs = "ts") + s(BARESOIL, bs = "ts") +
+      s(FallPrec, bs = "ts") + s(SprTmax, bs = "ts"),
+    data = Vegetation,fit=FALSE)
+
+#Extrtact penalization matrices
+S_ROCK = gam_setup$smooth[[1]]$S[[1]]
+S_LITTER = gam_setup$smooth[[2]]$S[[1]]
+S_BARESOIL = gam_setup$smooth[[3]]$S[[1]]
+S_FallPrec = gam_setup$smooth[[4]]$S[[1]]
+S_SprTmax = gam_setup$smooth[[5]]$S[[1]]
+
+S_list = list(S_ROCK,S_LITTER,S_BARESOIL,S_FallPrec,S_SprTmax)
+S_combined = .bdiag(S_list)         # join S's in sparse matrix
+Sdims = unlist(lapply(S_list,nrow)) # Find dimension of each S
+#----------------------------------------
+
+
+#For report, used for constructing plots----
+ROCK=seq(min(Vegetation$ROCK),max(Vegetation$ROCK),by = 1)
+LITTER=seq(min(Vegetation$LITTER),max(Vegetation$LITTER),by = 1)
+BARESOIL=seq(min(Vegetation$BARESOIL),max(Vegetation$BARESOIL),by = 1)
+FallPrec=seq(min(Vegetation$FallPrec),max(Vegetation$FallPrec),by = 0.2)
[…]

### 2021-06-02 gam_TMB/pSplines/pSplines.cpp

@@ -1,0 +1,60 @@
+#include <TMB.hpp>   //Links in the TMB libraries
+template<class Type>
+Type objective_function<Type>::operator() (){
+  using namespace density;
+  using namespace Eigen; //Needed for utilisation of sparse structures
+  
+  
+  //Read data from R------------------
+  DATA_VECTOR(Y);       //The response
+  DATA_MATRIX(X);       //Design matrix for splines
+  DATA_SPARSE_MATRIX(S);//Penalization matrix diag(S1,S2,S3,S4,S5) without storing off-diagonal zeros.
+  DATA_IVECTOR(Sdims);   //Dimensions of S1,S2,S3,S4 and S5
+  DATA_SPARSE_MATRIX(designMatrixForReport);//Design matrix for report of splines
+  DATA_INTEGER(flag);     //Used for calculating the normalizing constant of the prior for beta on R side
+  //----------------------------------
+  
+  //Read parameters from R------------
+  PARAMETER(beta0);       //Intercept
+  PARAMETER_VECTOR(beta); //Spline regression parameters
+  PARAMETER_VECTOR(log_lambda);//Penalization parameters
+  PARAMETER(log_sigma);   
+  //----------------------------------
+  
+  //Transform some of the parameters--
+  Type sigma = exp(log_sigma);
+  vector<Type> lambda = exp(log_lambda);
+  //----------------------------------
+  
+  //Calculate the objective function--
+  Type nll=0;
+
+  int k=0;  // Counter
+  for(int i=0;i<Sdims.size();i++){
+    int m_i = Sdims(i); //Dimensions of sub-matrix i
+    vector<Type> beta_i = beta.segment(k,m_i);       // Get coefficients for i-th smoother
+    SparseMatrix<Type> S_i = S.block(k,k,m_i,m_i);  // Get penalty (sub)matrix for i-th smoother
+    
+    //This bit is the logLik contribution from smoother i
+    //Formula comes from: beta_i ~ mvNorm(0,lambda_i * S_i), where lambda = penalty param, S = penalty matrix
+    //GMRF bit = t(beta_i) %*% S_i %*% beta_i
+    nll -= Type(0.5)*m_i*log_lambda(i) - 0.5*lambda(i)*GMRF(S_i).Quadform(beta_i); //Decrement logLik for each set of coefs
+    k += m_i; //Increment to next sub-matrix
+  }
+  
[…]

### 2021-06-02 gam_TMB/simpleExample/simpleExample.R

@@ -1,0 +1,25 @@
+library(TMB)
+
+compile("simpleExample.cpp")
+dyn.load(dynlib("simpleExample"))
+
+TN = read.table("TeethNitrogen.txt",header = T)
+
+data = list()
+data$NL = TN$X15N[TN$Tooth=="Moby"]
+data$x = TN$Age[TN$Tooth=="Moby"]
+data
+
+parameters = list( #Starting parameters
+  beta0 = 0,
+  beta1 = 0,
+  logSigma = 0
+)
+
+obj = MakeADFun(data,parameters,DLL = "simpleExample")
+opt = nlminb(obj$par,obj$fn, obj$gr) #Fit model
+rep = sdreport(obj)
+
+plot(data$x,data$NL)
+abline(a = rep$par.fixed[1], b = rep$par.fixed[2])
+

### 2021-06-02 gam_TMB/simpleExample/simpleExample.cpp

@@ -1,0 +1,19 @@
+#include <TMB.hpp>
+template<class Type>
+  Type objective_function<Type>::operator() ()
+{
+  DATA_VECTOR(NL); //Tooth nitrogen
+  DATA_VECTOR(x); //Age
+  
+  PARAMETER(beta0); //Intercept
+  PARAMETER(beta1); //Slope
+  PARAMETER(logSigma); //log SD
+  
+  Type sigma= exp(logSigma); //Get SD
+  Type nll = 0; //Set negative log-likelihood to 0
+  
+  vector<Type> mu = beta0 + x*beta1; //Expected value
+  nll -= sum(dnorm(NL,mu,sigma,true)); //Subtract sum of log likelihood from 0
+
+  return(nll);
+}

### 2021-06-02 gam_TMB/gamTests.R
Moved files to different directory

@@ -6,7 +6,7 @@ library(tidyverse)
 library(gamair)
 data(engine)
 
-plot(wear~size,data=engine,)
+plot(wear~size,data=engine)
 
 #Cubic splines, as in example ------------------------
 

### 2021-06-02 gam_TMB/pSplines/pSplines.R

@@ -1,121 +1,0 @@
-setwd("~/Projects/UofC postdoc/TMB Meeting Group/tmb-case-studies/pSplines")
-
-library(TMB)
-library(mgcv) #Use gam
-library(Matrix) #Use sparse matrices
-
-# Tutorial: sparse matrices in R and TMB
-M = matrix(1:4,2,2)   # ordinary 2x2 matrix
-M_block_diag = .bdiag(list(M,M)) # Block diagonal (sparse) matrix
-data.class(M_block_diag)     # Check data.class
-print(M_block_diag)    # dots means 0 value
-
-#Load the data and compile c++ code------
-Vegetation <- read.table(file = "Vegetation.txt", header = TRUE, dec = ".")
-Vegetation = Vegetation[!is.na(Vegetation$Richness),]
-
-compile("pSplines.cpp")
-dyn.load(dynlib("pSplines"))
-#----------------------------------------
-
-#Set up spline structure by using mgcv---
-gam_setup = gam(Richness ~ s(ROCK, bs = "ts") +
-      s(LITTER, bs = "ts") + s(BARESOIL, bs = "ts") +
-      s(FallPrec, bs = "ts") + s(SprTmax, bs = "ts"),
-    data = Vegetation,fit=FALSE)
-
-#Extrtact penalization matrices
-S_ROCK = gam_setup$smooth[[1]]$S[[1]]
-S_LITTER = gam_setup$smooth[[2]]$S[[1]]
-S_BARESOIL = gam_setup$smooth[[3]]$S[[1]]
-S_FallPrec = gam_setup$smooth[[4]]$S[[1]]
-S_SprTmax = gam_setup$smooth[[5]]$S[[1]]
-
-S_list = list(S_ROCK,S_LITTER,S_BARESOIL,S_FallPrec,S_SprTmax)
-S_combined = .bdiag(S_list)         # join S's in sparse matrix
-Sdims = unlist(lapply(S_list,nrow)) # Find dimension of each S
-#----------------------------------------
-
-
-#For report, used for constructing plots----
-ROCK=seq(min(Vegetation$ROCK),max(Vegetation$ROCK),by = 1)
-LITTER=seq(min(Vegetation$LITTER),max(Vegetation$LITTER),by = 1)
-BARESOIL=seq(min(Vegetation$BARESOIL),max(Vegetation$BARESOIL),by = 1)
-FallPrec=seq(min(Vegetation$FallPrec),max(Vegetation$FallPrec),by = 0.2)
[…]

### 2021-06-02 gam_TMB/pSplines/pSplines.cpp

@@ -1,60 +1,0 @@
-#include <TMB.hpp>   //Links in the TMB libraries
-template<class Type>
-Type objective_function<Type>::operator() (){
-  using namespace density;
-  using namespace Eigen; //Needed for utilisation of sparse structures
-  
-  
-  //Read data from R------------------
-  DATA_VECTOR(Y);       //The response
-  DATA_MATRIX(X);       //Design matrix for splines
-  DATA_SPARSE_MATRIX(S);//Penalization matrix diag(S1,S2,S3,S4,S5) without storing off-diagonal zeros.
-  DATA_IVECTOR(Sdims);   //Dimensions of S1,S2,S3,S4 and S5
-  DATA_SPARSE_MATRIX(designMatrixForReport);//Design matrix for report of splines
-  DATA_INTEGER(flag);     //Used for calculating the normalizing constant of the prior for beta on R side
-  //----------------------------------
-  
-  //Read parameters from R------------
-  PARAMETER(beta0);       //Intercept
-  PARAMETER_VECTOR(beta); //Spline regression parameters
-  PARAMETER_VECTOR(log_lambda);//Penalization parameters
-  PARAMETER(log_sigma);   
-  //----------------------------------
-  
-  //Transform some of the parameters--
-  Type sigma = exp(log_sigma);
-  vector<Type> lambda = exp(log_lambda);
-  //----------------------------------
-  
-  //Calculate the objective function--
-  Type nll=0;
-
-  int k=0;  // Counter
-  for(int i=0;i<Sdims.size();i++){
-    int m_i = Sdims(i); //Dimensions of sub-matrix i
-    vector<Type> beta_i = beta.segment(k,m_i);       // Get coefficients for i-th smoother
-    SparseMatrix<Type> S_i = S.block(k,k,m_i,m_i);  // Get penalty (sub)matrix for i-th smoother
-    
-    //This bit is the logLik contribution from smoother i
-    //Formula comes from: beta_i ~ mvNorm(0,lambda_i * S_i), where lambda = penalty param, S = penalty matrix
-    //GMRF bit = t(beta_i) %*% S_i %*% beta_i
-    nll -= Type(0.5)*m_i*log_lambda(i) - 0.5*lambda(i)*GMRF(S_i).Quadform(beta_i); //Decrement logLik for each set of coefs
-    k += m_i; //Increment to next sub-matrix
-  }
-  
[…]

### 2021-06-02 gam_TMB/simpleExample/simpleExample.R

@@ -1,25 +1,0 @@
-library(TMB)
-
-compile("simpleExample.cpp")
-dyn.load(dynlib("simpleExample"))
-
-TN = read.table("TeethNitrogen.txt",header = T)
-
-data = list()
-data$NL = TN$X15N[TN$Tooth=="Moby"]
-data$x = TN$Age[TN$Tooth=="Moby"]
-data
-
-parameters = list( #Starting parameters
-  beta0 = 0,
-  beta1 = 0,
-  logSigma = 0
-)
-
-obj = MakeADFun(data,parameters,DLL = "simpleExample")
-opt = nlminb(obj$par,obj$fn, obj$gr) #Fit model
-rep = sdreport(obj)
-
-plot(data$x,data$NL)
-abline(a = rep$par.fixed[1], b = rep$par.fixed[2])
-

### 2021-06-02 gam_TMB/simpleExample/simpleExample.cpp

@@ -1,19 +1,0 @@
-#include <TMB.hpp>
-template<class Type>
-  Type objective_function<Type>::operator() ()
-{
-  DATA_VECTOR(NL); //Tooth nitrogen
-  DATA_VECTOR(x); //Age
-  
-  PARAMETER(beta0); //Intercept
-  PARAMETER(beta1); //Slope
-  PARAMETER(logSigma); //log SD
-  
-  Type sigma= exp(logSigma); //Get SD
-  Type nll = 0; //Set negative log-likelihood to 0
-  
-  vector<Type> mu = beta0 + x*beta1; //Expected value
-  nll -= sum(dnorm(NL,mu,sigma,true)); //Subtract sum of log likelihood from 0
-
-  return(nll);
-}

### 2021-06-04 gam_TMB/gamTests.R
Got GAMs working in TMB
Next steps
- Multiple smoothers
- 2D smoothers
- location-scale smoothers
- Spatiotemporal correlation
- All of the above

@@ -1,32 +1,82 @@
 #GAM testing using TMB
 
+setwd("~/Documents/yield-analysis-2021/gam_TMB")
+
+library(Matrix) #Use sparse matrices
 library(TMB)
 library(mgcv)
 library(tidyverse)
+theme_set(theme_classic())
 library(gamair)
+
 data(engine)
 
-plot(wear~size,data=engine)
+plot(wear~size,data=engine,pch=19)
 
-#Cubic splines, as in example ------------------------
+#Thin plate spline -------------------
 
-s1 <- smoothCon(object=s(wear,bs='cr'),data=engine)
+s0 <- gam(wear~s(size,k=7,bs='tp'),data=engine,fit=FALSE) #Problems start to occur if k<=6; Hessian non-positive definite because lambda gradient = 0
 
-s1[[1]]$X #Model matrix
-s1[[1]]$S #Penalty matrix
+#Input
+modMat <- s0$X[,-1] #Model matrix (no intercept)
+penMat <- s0$smooth[[1]]$S[[1]] #Penalty matrix (sparse)
+penMat <- as(penMat,'sparseMatrix')
+penDim <- nrow(penMat) #Dimensions of penalty matrix
+
+#For making predictions:
+predSize <- with(engine,seq(min(size),max(size),length.out=50)) #Size of engines to predict for
+predModMat <- PredictMat(s0$smooth[[1]],data = data.frame(size=predSize)) #New model matrix to predict with (no intercept)
+
+#Input for model
+datList <- list(Y=engine$wear, X=modMat, S1=penMat, S1dim=penDim, newX=predModMat) #Data
+parList <- list(b0=0, smoothCoefs=rep(0,datList$S1dim), log_lambda=0, log_sigma=0) #Parameters
+
+compile("gam1.cpp")
+dyn.unload(dynlib("gam1"))
+dyn.load(dynlib("gam1"))
+obj <- MakeADFun(data = datList, parameters = parList, random=c('b0','smoothCoefs'), DLL = "gam1")
+
+obj$par
+obj$fn() #Works
+opt <- nlminb(obj$par,obj$fn,obj$gr)
+rep <- sdreport(obj) #Estimates and SEs
+rep #Looks OK
+names(rep$value)
+
+data.frame(name=names(obj$par), Est=opt$par, final_gradient=as.vector(obj$gr(opt$par)))
+
[…]

### 2021-06-04 gam_TMB/gam1.cpp
Added C code

@@ -1,0 +1,52 @@
+#include <TMB.hpp>   //Links in the TMB libraries
+template<class Type>
+Type objective_function<Type>::operator() (){
+  using namespace density;
+  using namespace Eigen; //Needed for sparse structures
+  // using namespace atomic; //Needed for logdet
+  
+  //Simple 1-D gam
+  
+  //Read data from R------------------
+  DATA_VECTOR(Y);               //Response
+  DATA_MATRIX(X);               //Design matrix (no intercept)
+  DATA_SPARSE_MATRIX(S1);       //Penalty matrix - must be sparse for GMRF to work
+  DATA_INTEGER(S1dim);          //Dimensions of S1
+  DATA_MATRIX(newX);            //New design matrix for predictions
+  
+  //Read parameters from R------------
+  PARAMETER(b0);       //Intercept
+  PARAMETER_VECTOR(smoothCoefs); //Smooth coefficients
+  PARAMETER(log_lambda);// Log penalization parameter
+  PARAMETER(log_sigma);   // log(SD) of measurement error 
+  
+  
+  //Transform SD and penalization parameters 
+  Type sigma = exp(log_sigma);
+  Type lambda = exp(log_lambda);
+  
+  //Define objective function
+  Type nll=0;
+  
+  // Penalization part ------------------------------------
+  SparseMatrix<Type> S = lambda*S1; //Penalty term * Penalty matrix
+  
+  //logdet requires a matrix type, and GMRF needs a sparse matrix
+  // nll -= 0.5*(logdet(matrix<Type>(S))) - 0.5*(GMRF(S).Quadform(smoothCoefs));
+  nll -= Type(0.5)*S1dim*log_lambda - 0.5*lambda*GMRF(S).Quadform(smoothCoefs); //This approach avoids logdet calculation - results are identical as far as I can tell
+  
+  vector<Type> mu = b0 + X*smoothCoefs; //Expected value: mu = Intercept + model matrix %*% coefs
+  
+  nll -= sum(dnorm(Y, mu, sigma, true)); //Decrement logLik
+  // for(int i=0; i<Y.size(); i++){ //Other approach that uses a loop
+  //   nll -= dnorm(Y(i), mu(i), sigma, true); 
+  // }
+
+  vector<Type> splineForReport = b0 + newX*smoothCoefs; //Generate predictions 
+  
+  // ADREPORT(b0); //Intercept
[…]

### 2021-06-04 gam_TMB/thingo.cpp

@@ -1,0 +1,47 @@
+// code auto-generated by TMBam
+#include <TMB.hpp>
+
+template<class Type>
+Type objective_function<Type>::operator() ()
+{
+  using namespace density;
+  using namespace Eigen;
+  using namespace R_inla;
+
+  DATA_MATRIX(X); // design matrix
+  DATA_VECTOR(y); // response
+  // penalty matrices
+  DATA_SPARSE_MATRIX(S1);
+
+  PARAMETER(mu); // intercept
+  PARAMETER_VECTOR(beta);
+  PARAMETER(log_lambda); // smoopar
+  PARAMETER(log_sigma);
+
+  // transform smoopars and scale
+  Type lambda = exp(log_lambda);
+  Type sigma = exp(log_sigma);
+
+  // split up the random effects coefs
+  vector<Type> beta1 = beta.segment(0,9);
+
+  // define the real penalties
+  Eigen::SparseMatrix<Type> K1 = lambda(0)*S1;
+
+  // initialize log-likelihood
+  Type nll=0;
+
+  using namespace atomic; // for logdet
+
+  // calculate REML penalty
+  nll = -0.5*(logdet(matrix<Type>(K1))) + 0.5*(GMRF(K1).Quadform(beta1));
+
+  // linear predictor
+  vector<Type> eta = mu + X*beta;
+
+  for(int i=0; i<y.size(); i++)
+      nll -= dnorm(y(i), eta(i), sigma, true);
+
+  return nll;
+
+}

### 2021-06-04 gam_TMB/thingo2.cpp

@@ -1,0 +1,51 @@
+// code auto-generated by TMBam
+#include <TMB.hpp>
+
+template<class Type>
+Type objective_function<Type>::operator() ()
+{
+  using namespace density;
+  using namespace Eigen;
+  using namespace R_inla;
+
+  DATA_MATRIX(X); // design matrix
+  DATA_VECTOR(y); // response
+  // penalty matrices
+  DATA_SPARSE_MATRIX(S1);
+  DATA_SPARSE_MATRIX(S2);
+
+  PARAMETER(mu); // intercept
+  PARAMETER_VECTOR(beta);
+  PARAMETER_VECTOR(log_lambda); // smoopar
+  PARAMETER(log_sigma);
+
+  // transform smoopars and scale
+  vector<Type> lambda = exp(log_lambda);
+  Type sigma = exp(log_sigma);
+
+  // split up the random effects coefs
+  vector<Type> beta1 = beta.segment(0,9);
+  vector<Type> beta2 = beta.segment(9,9);
+
+  // define the real penalties
+  Eigen::SparseMatrix<Type> K1 = lambda(0)*S1;
+  Eigen::SparseMatrix<Type> K2 = lambda(1)*S2;
+
+  // initialize log-likelihood
+  Type nll=0;
+
+  using namespace atomic; // for logdet
+
+  // calculate REML penalty
+  nll = -0.5*(logdet(matrix<Type>(K1)) + logdet(matrix<Type>(K2))) +
+    0.5*(GMRF(K1).Quadform(beta1) + GMRF(K2).Quadform(beta2));
+
+  // linear predictor
+  vector<Type> eta = mu + X*beta;
+
+  for(int i=0; i<y.size(); i++)
+      nll -= dnorm(y(i), eta(i), sigma, true);
+
+  return nll;
+
+}

### 2021-06-04 gam_TMB/gam1.cpp
Added 2-spline method, but something went wrong

@@ -20,7 +20,6 @@ Type objective_function<Type>::operator() (){
   PARAMETER(log_lambda);// Log penalization parameter
   PARAMETER(log_sigma);   // log(SD) of measurement error 
   
-  
   //Transform SD and penalization parameters 
   Type sigma = exp(log_sigma);
   Type lambda = exp(log_lambda);
@@ -29,11 +28,13 @@ Type objective_function<Type>::operator() (){
   Type nll=0;
   
   // Penalization part ------------------------------------
-  SparseMatrix<Type> S = lambda*S1; //Penalty term * Penalty matrix
   
   //logdet requires a matrix type, and GMRF needs a sparse matrix
-  // nll -= 0.5*(logdet(matrix<Type>(S))) - 0.5*(GMRF(S).Quadform(smoothCoefs));
-  nll -= Type(0.5)*S1dim*log_lambda - 0.5*lambda*GMRF(S).Quadform(smoothCoefs); //This approach avoids logdet calculation - results are identical as far as I can tell
+  // SparseMatrix<Type> S = lambda*S1; //Penalty term * Penalty matrix
+  // nll -= 0.5*(logdet(matrix<Type>(S))) - 0.5*(GMRF(S).Quadform(smoothCoefs)); //Returns NaNs. I think this is because logDet(S) is infinite (S = rank deficient?)
+  
+  //This approach avoids logdet calculation - results are identical as far as I can tell
+  nll -= Type(0.5)*S1dim*log_lambda - 0.5*lambda*GMRF(S1).Quadform(smoothCoefs); 
   
   vector<Type> mu = b0 + X*smoothCoefs; //Expected value: mu = Intercept + model matrix %*% coefs
   
@@ -43,10 +44,10 @@ Type objective_function<Type>::operator() (){
   // }
 
   vector<Type> splineForReport = b0 + newX*smoothCoefs; //Generate predictions 
+  ADREPORT(splineForReport); //Predicted smoother values
   
   // ADREPORT(b0); //Intercept
   // ADREPORT(smoothCoefs); //Smoothing coefficients - these seem to be left in the model report no matter what
-  // ADREPORT(splineForReport); //Predicted smoother values
   
   return nll;
 }

### 2021-06-04 gam_TMB/gam2.cpp

@@ -1,0 +1,51 @@
+#include <TMB.hpp>   //Links in the TMB libraries
+template<class Type>
+Type objective_function<Type>::operator() (){
+  using namespace density;
+  using namespace Eigen; //Needed for sparse structures
+  
+  //1-D gam with 2 predictors
+  
+  //Read data from R------------------
+  DATA_VECTOR(Y);               //Response
+  DATA_MATRIX(X);               //Design matrix (no intercept)
+  DATA_SPARSE_MATRIX(S1);       //Penalty matrix for smoother 1
+  DATA_SPARSE_MATRIX(S2);       //Penalty matrix for smoother 2
+  DATA_INTEGER(S1dim);          //Dimensions of S1
+  DATA_INTEGER(S2dim);          //Dimensions of S2
+  DATA_MATRIX(newX);            //New design matrix for predictions
+  
+  //Read parameters from R------------
+  PARAMETER(b0);       //Intercept
+  PARAMETER_VECTOR(smoothCoefs); //Smooth coefficients
+  PARAMETER(log_lambda1);// Log penalization parameter
+  PARAMETER(log_lambda2);
+  PARAMETER(log_sigma);   // log(SD) of measurement error 
+  
+  
+  //Transform SD and penalization parameters 
+  Type sigma = exp(log_sigma);
+  Type lambda1 = exp(log_lambda1);
+  Type lambda2 = exp(log_lambda2);
+  
+  // Get smoothing coefficients
+  vector<Type> smoothCoefs1 = smoothCoefs.segment(0,S1dim);
+  vector<Type> smoothCoefs2 = smoothCoefs.segment(S1dim-1,S2dim);
+  
+  //Define objective function
+  Type nll=0;
+  
+  // Penalization 
+  nll -= Type(0.5)*(S1dim*log_lambda1) - 0.5*(lambda1*GMRF(S1).Quadform(smoothCoefs1));
+  nll -= Type(0.5)*(S2dim*log_lambda2) - 0.5*(lambda2*GMRF(S2).Quadform(smoothCoefs2)); 
+
+  //Expected value: mu = Intercept + model matrix %*% coefs
+  vector<Type> mu = b0 + X*smoothCoefs; 
+  
+  nll -= sum(dnorm(Y, mu, sigma, true)); //Decrement logLik
+  
+  // vector<Type> splineForReport = b0 + newX*smoothCoefs; //Generate predictions 
+  // ADREPORT(splineForReport); //Predicted smoother values
+  
+  return nll;
+}

### 2021-06-04 gam_TMB/gamTests1.R

@@ -1,4 +1,5 @@
 #GAM testing using TMB
+#Thin plate spline  - single smoother 
 
 setwd("~/Documents/yield-analysis-2021/gam_TMB")
 
@@ -7,15 +8,14 @@ library(TMB)
 library(mgcv)
 library(tidyverse)
 theme_set(theme_classic())
-library(gamair)
 
+library(gamair)
 data(engine)
 
-plot(wear~size,data=engine,pch=19)
+ggplot(data=engine,aes(x=size,y=wear))+geom_point()+
+  geom_smooth(method='gam',formula=y~s(x,k=7))
 
-#Thin plate spline -------------------
-
-s0 <- gam(wear~s(size,k=7,bs='tp'),data=engine,fit=FALSE) #Problems start to occur if k<=6; Hessian non-positive definite because lambda gradient = 0
+s0 <- gam(wear~s(size,k=8,bs='tp'),data=engine,fit=FALSE) #Problems start to occur if k<=6; Hessian non-positive definite because lambda gradient = 0
 
 #Input
 modMat <- s0$X[,-1] #Model matrix (no intercept)
@@ -32,7 +32,6 @@ datList <- list(Y=engine$wear, X=modMat, S1=penMat, S1dim=penDim, newX=predModMa
 parList <- list(b0=0, smoothCoefs=rep(0,datList$S1dim), log_lambda=0, log_sigma=0) #Parameters
 
 compile("gam1.cpp")
-dyn.unload(dynlib("gam1"))
 dyn.load(dynlib("gam1"))
 obj <- MakeADFun(data = datList, parameters = parList, random=c('b0','smoothCoefs'), DLL = "gam1")
 
@@ -41,18 +40,13 @@ obj$fn() #Works
 opt <- nlminb(obj$par,obj$fn,obj$gr)
 rep <- sdreport(obj) #Estimates and SEs
 rep #Looks OK
-names(rep$value)
+rep$value
 
 data.frame(name=names(obj$par), Est=opt$par, final_gradient=as.vector(obj$gr(opt$par)))
 
 #Get coefs from model
 m1Tmb <- rbind(with(rep,data.frame(coef=names(par.fixed),est=par.fixed,sd=sqrt(diag(cov.fixed)),row.names=NULL)),
-      with(rep,data.frame(coef=names(value),est=value,sd=sd)))
-
[…]

### 2021-06-04 gam_TMB/gamTests2.R

@@ -1,0 +1,44 @@
+#GAM testing using TMB
+# Thin plate spline - 2 smoothers 
+
+setwd("~/Documents/yield-analysis-2021/gam_TMB")
+
+library(Matrix) #Use sparse matrices
+library(TMB)
+library(mgcv)
+library(tidyverse)
+theme_set(theme_classic())
+
+s0 <- gam(Volume~s(Girth,k=6,bs='tp')+s(Height,k=6,bs='tp'),data=trees) 
+
+#Input
+modMat <- model.matrix(s0)[,-1] #Model matrix (no intercept)
+penMat1 <- s0$smooth[[1]]$S[[1]] #Penalty matrices (sparse)
+penMat1 <- as(penMat1,'sparseMatrix')
+penMat2 <- s0$smooth[[2]]$S[[1]] 
+penMat2 <- as(penMat2,'sparseMatrix')
+penDim1 <- nrow(penMat1) #Dimensions of penalty matrices
+penDim2 <- nrow(penMat2) 
+
+#For making predictions:
+predSize <- with(trees,rbind(expand.grid(Girth=mean(Girth),Height=seq(min(Height),max(Height),1)),
+                 expand.grid(Girth=seq(min(Girth),max(Girth),0.5),Height=mean(Height))))
+predMat1 <- PredictMat(s0$smooth[[1]],data = predSize) 
+predMat2 <- PredictMat(s0$smooth[[2]],data = predSize) 
+predModMat <- cbind(predMat1,predMat2) #New model matrix to predict with (no intercept)
+
+#Input for model
+datList <- list(Y=trees$Volume, X=modMat, 
+                S1=penMat1, S2=penMat2, S1dim=penDim1, S2dim=penDim2, newX=predModMat) #Data
+parList <- list(b0=coef(s0)[1], smoothCoefs=coef(s0)[-1], log_lambda1=1, log_lambda2=1, log_sigma=1) #Parameters
+
+compile("gam2.cpp")
+dyn.load(dynlib("gam2"))
+obj <- MakeADFun(data = datList, parameters = parList, random=c('b0','smoothCoefs'), DLL = "gam2")
+
+obj$par
+obj$fn() #Works
+opt <- nlminb(obj$par,obj$fn,obj$gr)
+rep <- sdreport(obj) #Estimates and SEs
+rep #Looks OK
+rep$value

### 2021-06-15 Manuscript/manuscript.Rmd
Worked on Intro and Methods

@@
 
-- Preserving SNL around fields is important for agriculture and conservation
-  - Ecosystem services
-  - Act as reservoirs for beneficial insects
-  - Microclimate zones
-  - Poorly studied in North America
-  - Studies tend to be limited in scope (e.g. only certain organisms/crops) and applicability
+- Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth
+  - This has allows increases in human population and increased (global) stability in production
+  - However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms
+  - Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios
+  - Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems
 
+- SNL in and around crops is important for both agricultural production and conservation
+  - They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control
+  - They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed
+  - Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
+  - In particular, North American agroecosystems have larger fields, and different varieties and agronomic practices, all of which could negate effects of SNL
+
+<!-- - SNL has conflicting effects on crop yield because of ecosystem service provision as well as negative edge effects -->
+- SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate
+  - Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds
+  - At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges
+  - For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance
+  - Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects 
+  - This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
+
+<!-- - Less-considered are ecosystem ``disservices", where weeds and pests  -->
+
+<!-- - In North America, there are financial incentives to make fields large, homogeneous, and with simple geometry -->
+
+- Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality (but see wheat study)
+  - For this reason, large-scale precision yield data holds enormous promise for agronomy, as it allows 
+  - However, its use is limited for several reasons: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) unfamiliarity with spatial statistics
   - Ecosystem services can influence both the mean and variability of yield in agroecosystems
-  - Typically only averages (means) are considered, but higher stability (lower variance) in yield is also valuable
-
-- Field size and field boundaries are directly related to the conservation of SNL in agroecosystems
+  - Typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable
   - Size examples: wheat study from the UK
-  - Boundary examples: flower strip studies, hedgerows
-  - In North America, there are financial incentives to make fields large and homogeneous, especially with large agricultural equipment
-  - Crop edge effects cause low yields at the margins of fields because of late emergence, poor microclimate, and competition with weeds
-  - At the same time, ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
-  - This suggests that yield may be maximized at distances between the edge and centre of the field, where ecosystem services cancel out negative edge effects
-
-  <!-- - This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services -->
-
-- Precision yield data holds enormous promise for agronomy
-  - Limited because of:
-  - Lack of standardized formats
-  - Yearly calibration data
-  - Clear statistical protocols
+  - There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield
 
@@
   - where:
-  1. Polygon Size = distance traveled x width of header bar ($m^2$)
+  1. Polygon Size = distance traveled $\times$ width of header bar ($m^2$)
   2. Distance from Edge = distance from field edge type _i_ (m)
@@
 
-- This model accounts for a) combine speed differences, b) edge effects, c) within-field spatial variation not related to edges, and d) temporal differences in harvest 
-  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (e.g. Kriging), but this was computationally infeasible with 50,000 data points per field
-  - Penalized regression splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with substantially reduced computation time
+- In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest
+  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (e.g.
+Kriging) or Stochastic Partial Differential Equations (e.g.
+INLA), but this was computationally infeasible with 50,000 data points per field
+  - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with substantially reduced computation time
   - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
@@
 - Most models still had a large amount of spatio-temporal autocorrelation in residuals
-  - We basically just ignored this
+  - We have 50000 data points per field, so we ignored it
 
-- To combine results from all field-level models while also accounting for field-level uncertainty, we used an approach similar to bootstrapping of hierarchical mixed effects models
+- To consider results from all field-level models while also accounting for field-level uncertainty, we used an approach similar to bootstrapping of hierarchical mixed effects models
   - The approach was as follows
-  1. Extract posterior samples of smoother parameters from each field-level model
-  2. Fit meta-model of all smoothers, and save parameters
-  3. Repeat 1000 times, and calculate credible intervals (10-90th percentiles of draws)
-
-- We compared this
+  1. Extract single posterior sample (_rnorm_ in R) of smoother parameters from each field-level model using coefficient estimates and standard error
+  2. Use posterior sample to create new simulated smoother from each field
+  3. Fit model of all simulated smoothers, and save this meta-smoother
+  4. Repeat 1000 times, and calculate coverage intervals (5-95% percentiles) on saved meta-smoothers
+  - This gives credible intervals for the "average" smoother while accounting for field-level variability
 

### 2021-06-15 gam_TMB/gamTests1.R

@@ -1,7 +1,5 @@
 #GAM testing using TMB
 
-
-
 setwd("~/Documents/yield-analysis-2021/gam_TMB")
 
 library(Matrix) #Use sparse matrices

### 2021-06-22 Manuscript/manuscript.Rmd
Worked on manuscript and started TMB tests

@@
 
-- Yield data was collected in rectangles of the same length as the data interval (distance = speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m)
+- Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m)
   - We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
@@
   - We used dry yield (tonnes of seed/hectare after accounting for crop moisture) as our measure of crop yield
-  - Due to the high number of samples at each field, we sub-sampled to 50,000 data points per field to reduce computation time
+  - Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time
 

### 2021-06-22 gam_TMB/mod0A.cpp

@@ -1,0 +1,51 @@
+#include <TMB.hpp>   //Links in the TMB libraries
+template<class Type>
+Type objective_function<Type>::operator() (){
+  using namespace density;
+  using namespace Eigen; //Needed for sparse structures
+  
+  //Simple 1-D gam
+  
+  //Read data from R------------------
+  DATA_VECTOR(yield);                 //Response
+  DATA_MATRIX(smoothMat);             //Design matrix (no intercept)
+  DATA_SPARSE_MATRIX(penaltyMat);     //Penalty matrix - must be sparse for GMRF to work
+  DATA_INTEGER(penaltyDim);           //Dimensions of penaltyMat
+  DATA_MATRIX(newSmoothMat);          //New smoothing matrix for predictions
+
+  //Read parameters from R------------
+  PARAMETER(b0);       //Intercept
+  PARAMETER_VECTOR(smoothCoefs); //Smooth coefficients
+  PARAMETER(log_lambda);// Log penalization parameter
+  PARAMETER(log_sigma);   // log(SD) of measurement error 
+  
+  //Transform SD and penalization parameters 
+  Type sigma = exp(log_sigma);
+  Type lambda = exp(log_lambda);
+  
+  //Initialize objective function
+  Type nll=0;
+  
+  // Penalization part ------------------------------------
+  
+  //GMRF needs a sparse matrix - This approach avoids logdet calculation
+  SparseMatrix<Type> S = lambda*penaltyMat; //Penalty term * Penalty matrix
+  // Penalty that works with bs = 'ts'/'cs': extra shrinkage
+  nll -= 0.5*penaltyDim*log_lambda - 0.5*GMRF(S).Quadform(smoothCoefs);  
+
+  // Main model -------------------------------------------
+  vector<Type> mu = b0 + smoothMat*smoothCoefs; //Expected value: mu = Intercept + model matrix %*% coefs
+  
+  nll -= sum(dnorm(yield, mu, sigma, true)); //Decrement logLik
+  
+  
+  // Other things to report -------------------------------
+  
+  vector<Type> splineForReport = b0 + newSmoothMat*smoothCoefs; //Generate predictions
+  vector<Type> residuals = yield - mu; //Residuals 
+  
+  REPORT(residuals); //Residuals
+  ADREPORT(splineForReport); //Predicted smoother values
+  
+  return nll; //Return likelihood
+}

### 2021-06-22 gam_TMB/mod0B.cpp

@@ -1,0 +1,48 @@
+#include <TMB.hpp>   //Links in the TMB libraries
+template<class Type>
+Type objective_function<Type>::operator() (){
+  using namespace R_inla; //includes SPDE-specific functions, e.g. Q_spde()
+  using namespace density;
+  using namespace Eigen; //Needed for sparse structures
+  
+  //Read data from R------------------
+  DATA_VECTOR(yield);                 //Response
+  DATA_STRUCT(spdeMatrices,spde_t);   //Three matrices needed for representing the GMRF, see p. 8 in Lindgren et al. (2011)
+  DATA_SPARSE_MATRIX(A);              //Matrix for interpolating points within triangles   
+  
+  //Read parameters from R------------
+  PARAMETER(b0);                  //Intercept
+  PARAMETER_VECTOR(spdeCoefs);    //SPDE coefficients (basically random effects coefs)
+  PARAMETER(log_tau);             // Tau (Precision parameter)
+  PARAMETER(log_kappa);           // Kappa (Spatial decay parameter)
+  PARAMETER(log_sigma);           // Sigma (Data variance)
+  
+  Type tau = exp(log_tau);      // Precision
+  Type kappa = exp(log_kappa);  // Spatial decay range
+  Type sigma = exp(log_sigma);  // Data variance
+
+  // Spatial interpolation
+  vector<Type> delta = (A*spdeCoefs)/tau; //Match location to random intercept, divide by tau (precision parameter) to get spatial random effect
+  
+  //Construct sparse precision matrix for latent field
+  SparseMatrix<Type> Q = Q_spde(spdeMatrices,kappa);
+
+  //Initialize objective function
+  Type nll = 0.0;
+  
+  nll = GMRF(Q)(spdeCoefs); //Basically: ~MVNorm(0, Matern Covariance Matrix) but using SPDE approach
+  
+  // Main model -------------------------------------------
+  vector<Type> mu = b0 + delta; //Expected value: mu = Intercept + spatial random effect
+  
+  nll -= sum(dnorm(yield, mu, sigma, true)); //Decrement log-likelihood
+  
+  // Other things to report -------------------------------
+  vector<Type> residuals = yield - mu; //Residuals 
+  Type range = sqrt(8)/kappa;   //Distance at which correlation has dropped to 0.1, see p. 4 in Lindgren et al. (2011)
+  
+  REPORT(residuals); //Residuals
+  ADREPORT(range);
+  
+  return nll; //Return likelihood
+}

### 2021-06-22 gam_TMB/yieldTMBtest.R

@@ -1,0 +1,177 @@
+#Tests of R-INLA models for yield, using SPDE approx of Matern covariance
+
+#Models to try:
+# 0A. Yield ~ s(dist) - done
+# 0B. Yield ~ SPDE(E,N)
+# 0C. Yield ~ SPDE(Time)
+# 1A Yield ~ s(dist) + SPDE(E,N) + SPDE(Time)
+# 1B Yield ~ s(dist) + SPDE(E,N) + SPDE(Time), Var ~ s(dist) + SPDE(E,N) + SPDE(Time)
+# 1C. Yield ~ s(dist,by=Type) + SPDE(E,N) + SPDE(Time), Var ~ s(dist,by=Type) + SPDE(E,N) + SPDE(Time)
+
+#Load everything --------------------
+
+library(tidyverse)
+theme_set(theme_classic())
+library(sf)
+library(mgcv)
+library(INLA)
+library(TMB)
+setwd("~/Documents/yield-analysis-2021/gam_TMB")
+source('../helperFunctions.R')
+
+nSubSamp <- 20000 #Number of points to limit to
+csvPath <- "/media/rsamuel/Storage/geoData/Rasters/yieldData/csv files/Alvin_French C41 2020.csv"
+boundaryPath <- "/home/rsamuel/Documents/yield-analysis-2021/Figures/FieldBoundaries/Alvin_French C41 2020 boundary.shp"
+
+dat <- read.csv(csvPath,stringsAsFactors=TRUE,fileEncoding='latin1') 
+
+if(nrow(dat)>nSubSamp){
+  dat <- dat %>% slice(round(seq(1,nrow(dat),length.out=nSubSamp)))
+}
+
+dat <- dat %>% rename_with(.fn = ~gsub('..L.ha.$','_lHa',.x)) %>%
+  rename_with(.fn = ~gsub('..tonne.ha.$','_tHa',.x)) %>%
+  rename_with(.fn = ~gsub('..m.s.$','_ms',.x)) %>%
+  rename_with(.fn = ~gsub('.km.h.$','_kmh',.x)) %>%
+  rename_with(.fn = ~gsub('..tonne.h.$','_th',.x)) %>%
+  rename_with(.fn = ~gsub('.ha.h.','_hah',.x)) %>%
+  rename_with(.fn = ~gsub('.m.','_m',.x)) %>%
+  rename_with(.fn = ~gsub('.deg.','Angle',.x)) %>%
+  rename_with(.fn = ~gsub('\\.','',.x)) %>%
[…]

### 2021-06-22 getPreds.R

@@ -1,0 +1,131 @@
+#Script to get parameter estimates from saved GAM models, and fit meta-model that incorporates uncertainty at the lower levels
+# Written by Sam Robinson for Lan Nguyen, Summer 2021
+
+# Load libraries and functions --------------------------
+
+library(mgcv)
+library(tidyverse)
+theme_set(theme_classic())
+
+# Convenience function to get parameters from models
+# d = data frame of new parameters to use. Useful if you want predictions of 1 thing at a time (e.g. Distance at specific Lat/Lon values)
+#   Can also be: 'original' uses original data, 'distance' uses unique distances and median row/col values.
+getPars <- function(path,d='distance'){
+  load(path) #Load file into environment
+  beta <- coefficients(mod) #Coefficients
+  sigma <- sqrt(diag(vcov(mod))) #Standard errors
+  
+  #Range of predictors in original model
+  predRange <- cbind(summary(with(mod$smooth[[1]],data.frame(Dist=as.vector(Xu)+as.vector(shift)))), 
+                     summary(with(mod$smooth[[2]],data.frame(Row=Xu[,1]+shift[1],Col=Xu[,2]+shift[2]))))
+  predNames <- gsub('\\s*(:|:\\s).*','',predRange[,1])
+  predRange <- apply(predRange,2,function(x) as.numeric(gsub('.+(:|:\\s)','',x)))
+  
+  nUnique <- c(with(mod$smooth[[1]],length(unique(Xu))), #Number of unique values
+               with(mod$smooth[[2]],apply(Xu,2,function(x) length(unique(x)))))
+  predRange <- rbind(predRange,nUnique)
+  rownames(predRange) <- c(predNames,'Nunique')
+  
+  if(is.null(d)){
+    mm <- model.matrix(mod)
+    d <- NULL
+    } else if(is.data.frame(d)) {
+      mm <- predict(mod,type='lpmatrix',newdata=d) #Model matrix from new dataframe
+    } else if(d=='distance'){
+      d <- data.frame(Dist=with(mod$smooth[[1]],sort(unique(as.vector(Xu)+as.vector(shift)))), #Unique distances
+                      Row=with(mod$smooth[[2]],median(Xu[,1]+shift[1])), #Median row
+                      Col=with(mod$smooth[[2]],median(Xu[,2]+shift[2]))) #Median column
[…]

### 2021-06-22 helperFunctions.R

@@ -1,5 +1,8 @@
 # Helper functions
 
+#Scales x between 0 and 1
+range01 <- function(x) (x-min(x))/(max(x)-min(x))
+
 seqGroup <- function(x,singles=TRUE){ #Turns sequential IDs into numbered groups.
   #Singles = TRUE, sequential values must be exactly 1 greater (gaps matter)
   #Singles = FALSE, sequential values must be greater (gaps don't matter)
@@ -930,9 +933,9 @@ getModelInfo <- function(path){
   
   return(retList)
 }
-getModelInfo("/home/rsamuel/Documents/yield-analysis-2021/Figures/ModelCheck0/Alvin_French Al_Jr 2020 results.txt")
-getModelInfo("/home/rsamuel/Documents/yield-analysis-2021/Figures/ModelCheck1/Alvin_French Al_Jr 2020 results.txt")$REML
-getModelInfo("/home/rsamuel/Documents/yield-analysis-2021/Figures/ModelCheck2/Alvin_French Al_Jr 2020 results.txt")$REML
-
-debugonce(getModelInfo)
+# getModelInfo("/home/rsamuel/Documents/yield-analysis-2021/Figures/ModelCheck0/Alvin_French Al_Jr 2020 results.txt")
+# getModelInfo("/home/rsamuel/Documents/yield-analysis-2021/Figures/ModelCheck1/Alvin_French Al_Jr 2020 results.txt")$REML
+# getModelInfo("/home/rsamuel/Documents/yield-analysis-2021/Figures/ModelCheck2/Alvin_French Al_Jr 2020 results.txt")$REML
+# 
+# debugonce(getModelInfo)
 # getModelInfo("./Figures/ModelCheck/Dean Hubbard 2018 All_27_11_25 results.txt")

### 2021-06-22 Manuscript/manuscript.Rmd
Tried out AR1 models
- Model fits, but runs into problems when running adreport() - takes an incredibly long time

@@
 - At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor
-  - Ground speed is known to be extremely important to yield monitor accuracy [@arslan2002], with low ground speed registering higher yields 
   - Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices 
     <!-- - such as strip farming -->
+  - Ground speed is extremely important to yield monitor accuracy [@arslan2002], with low ground speed registering higher yields 
   - While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time  within fields

### 2021-06-22 gam_TMB/yieldTMBtest.R

@@ -3,7 +3,7 @@
 #Models to try:
 # 0A. Yield ~ s(dist) - done
 # 0B. Yield ~ SPDE(E,N)
-# 0C. Yield ~ SPDE(Time)
+# 0C. Yield ~ AR1(Time)
 # 1A Yield ~ s(dist) + SPDE(E,N) + SPDE(Time)
 # 1B Yield ~ s(dist) + SPDE(E,N) + SPDE(Time), Var ~ s(dist) + SPDE(E,N) + SPDE(Time)
 # 1C. Yield ~ s(dist,by=Type) + SPDE(E,N) + SPDE(Time), Var ~ s(dist,by=Type) + SPDE(E,N) + SPDE(Time)
@@ -144,14 +144,18 @@ parList <- list(b0 = 1.9, spdeCoefs = rep(0.0, nrow(datList$spdeMatrices$M0)), l
 dyn.unload(dynlib("mod0B"))
 compile("mod0B.cpp")
 dyn.load(dynlib("mod0B"))
-obj <- MakeADFun(data = datList, parameters = parList, random=c('spdeCoefs'), DLL = "mod0B") #Takes about 5 mins to run if nSubSamp == 10000, crashes if 20000
 
-obj$par
-obj$fn() #Works
-a <- Sys.time()
-opt <- nlminb(obj$par,obj$fn,obj$gr) 
-Sys.time()-a #Takes 13 mins with 20000 samples
-report <- sdreport(obj) #Estimates and SEs - takes 3 mins
+# obj <- MakeADFun(data = datList, parameters = parList, random=c('spdeCoefs'), DLL = "mod0B") #Takes about 5 mins to run if nSubSamp == 10000, crashes if 20000
+# obj$par
+# obj$fn() #Works
+# opt <- nlminb(obj$par,obj$fn,obj$gr)  #Takes 13 mins with 20000 samples
+# report <- sdreport(obj) #Estimates and SEs - takes 3 mins
+# save(obj,opt,report,file = 'mod0B.Rdata')
+
+load('mod0B.Rdata')
+names(modList) <- c('obj','opt','report')
+attach(modList)
+dyn.load(dynlib("mod0B"))
 
 report #Looks OK
 report$value #Correlation drops below 0.1 at ~50 meters
@@ -160,18 +164,53 @@ report$par.random
 
 #Residuals are still horrible
 qqnorm(obj$report()$residuals); abline(0,1)
[…]

### 2021-06-24 gam_TMB/mod0A.cpp
Added SPDE models

@@ -8,6 +8,8 @@ Type objective_function<Type>::operator() (){
   
   //Read data from R------------------
   DATA_VECTOR(yield);                 //Response
+  DATA_VECTOR(logArea);               //log-Area 
+  DATA_SCALAR(meanLogArea);           //Mean of log(area)
   DATA_MATRIX(smoothMat);             //Design matrix (no intercept)
   DATA_SPARSE_MATRIX(penaltyMat);     //Penalty matrix - must be sparse for GMRF to work
   DATA_INTEGER(penaltyDim);           //Dimensions of penaltyMat
@@ -15,6 +17,7 @@ Type objective_function<Type>::operator() (){
 
   //Read parameters from R------------
   PARAMETER(b0);       //Intercept
+  PARAMETER(b_area);       //Area (combine speed)
   PARAMETER_VECTOR(smoothCoefs); //Smooth coefficients
   PARAMETER(log_lambda);// Log penalization parameter
   PARAMETER(log_sigma);   // log(SD) of measurement error 
@@ -34,14 +37,13 @@ Type objective_function<Type>::operator() (){
   nll -= 0.5*penaltyDim*log_lambda - 0.5*GMRF(S).Quadform(smoothCoefs);  
 
   // Main model -------------------------------------------
-  vector<Type> mu = b0 + smoothMat*smoothCoefs; //Expected value: mu = Intercept + model matrix %*% coefs
+  vector<Type> mu = b0 + logArea*b_area + smoothMat*smoothCoefs; //Expected value: mu = Intercept + model matrix %*% coefs
   
   nll -= sum(dnorm(yield, mu, sigma, true)); //Decrement logLik
   
-  
   // Other things to report -------------------------------
   
-  vector<Type> splineForReport = b0 + newSmoothMat*smoothCoefs; //Generate predictions
+  vector<Type> splineForReport = b0 + meanLogArea*b_area + newSmoothMat*smoothCoefs; //Generate predictions at mean combine speed
   vector<Type> residuals = yield - mu; //Residuals 
   
   REPORT(residuals); //Residuals

### 2021-06-24 gam_TMB/yieldTMBtest.R

@@ -3,7 +3,7 @@
 #Models to try:
 # 0A. Yield ~ s(dist) - done
 # 0B. Yield ~ SPDE(E,N)
-# 0C. Yield ~ AR1(Time)
+# 0C. Yield ~ SPDE(Time)
 # 1A Yield ~ s(dist) + SPDE(E,N) + SPDE(Time)
 # 1B Yield ~ s(dist) + SPDE(E,N) + SPDE(Time), Var ~ s(dist) + SPDE(E,N) + SPDE(Time)
 # 1C. Yield ~ s(dist,by=Type) + SPDE(E,N) + SPDE(Time), Var ~ s(dist,by=Type) + SPDE(E,N) + SPDE(Time)
@@ -22,6 +22,7 @@ source('../helperFunctions.R')
 nSubSamp <- 20000 #Number of points to limit to
 csvPath <- "/media/rsamuel/Storage/geoData/Rasters/yieldData/csv files/Alvin_French C41 2020.csv"
 boundaryPath <- "/home/rsamuel/Documents/yield-analysis-2021/Figures/FieldBoundaries/Alvin_French C41 2020 boundary.shp"
+boundaryPath2 <- "/home/rsamuel/Documents/yield-analysis-2021/Figures/FieldBoundarySegments/Alvin_French C41 2020 boundary.shp"
 
 dat <- read.csv(csvPath,stringsAsFactors=TRUE,fileEncoding='latin1') 
 
@@ -51,12 +52,15 @@ dat <- dat %>% rename_with(.fn = ~gsub('..L.ha.$','_lHa',.x)) %>%
   mutate(E=(E-mean(E)),N=(N-mean(N))) #Center coordinates
 # ggplot(dat)+geom_sf()
 
-fieldEdge <- read_sf(boundaryPath) %>% st_cast('MULTILINESTRING') #Read in boundary file
+fieldEdge <- read_sf(boundaryPath) #Read in boundary polygon
+fieldEdgeType <- read_sf(boundaryPath2) #Read in boundary type linestrings
 
-#Distance from boundary
-dat <- dat %>% mutate(dist=as.numeric(st_distance(.,fieldEdge))[1:nrow(.)]) 
+#Get distance and type of closest boundary
+dat <- dat %>% 
[…]

### 2021-07-07 gam_TMB/mod1B.cpp
Added TMB files
Some observations:
- Smoothers are OK at low dimensions, but steal tons of memory at higher dimensions
- e.g. s(E,N,k=60) using TMB takes about 16GB just to run. Bigger ones are even worse, and just crash when running for the first time
- SPDE approach is pretty good, but using the same basis for mean and SD is probably not going to work
- For nonstationary models, SD sub-model needs to vary "more slowly" than mean model

@@ -1,0 +1,150 @@
+#include <TMB.hpp>   //Links in the TMB libraries
+template<class Type>
+Type objective_function<Type>::operator() (){
+  using namespace density; //includes GMRF
+  // using namespace R_inla; //includes SPDE-specific functions, e.g. Q_spde()
+  using namespace Eigen; //Needed for sparse structures
+                      
+  //Read data from R------------------
+  DATA_VECTOR(yield);                        // Response
+  DATA_VECTOR(logArea);                      // log-Area of harvested polygons 
+  DATA_SCALAR(meanLogArea);                  // Mean of log(area)
+  DATA_MATRIX(smoothMat);                    // Design matrix (no intercept)
+  DATA_SPARSE_MATRIX(penaltyMat);            // Penalty matrix - must be sparse for GMRF to work
+  DATA_IVECTOR(penaltyDim);                  // Dimensions of penalty sub-matrices
+  // DATA_MATRIX(predModMat_dist);              // New smoothing matrices for predictions
+  // DATA_MATRIX(predModMat_spatial);
+  // DATA_MATRIX(predModMat_temporal);
+  // DATA_STRUCT(spatialSPDEMatrices,spde_t);   // Matrices needed for representing the spatial GMRF
+  // DATA_SPARSE_MATRIX(spatialA);              // Matrix for interpolating points within spatial mesh
+  // DATA_STRUCT(temporalSPDEMatrices,spde_t);  // Matrices needed for representing the temporal GMRF
+  // DATA_SPARSE_MATRIX(temporalA);             // Matrix for interpolating points within temporal mesh  
+
+  //Read parameters from R------------
+  
+  //Mean parameters
+  PARAMETER(b0);                              // Intercept
+  PARAMETER(b_areaMean);                      // Area (combine speed)
+  PARAMETER_VECTOR(smoothCoefsMean);          // Smooth coefficients
+  PARAMETER_VECTOR(log_lambdasMean);          // Log penalization parameter
+  //Variance parameters
+  PARAMETER(b0SD);                            // Intercept
+  // PARAMETER(b_areaSD);                        // Area (combine speed)
+  // PARAMETER_VECTOR(smoothCoefsSD);            // Smooth coefficients
[…]

### 2021-07-10 Manuscript/manuscript.Rmd
Gave up on TMB
Doesn't work well enough for large numbers of samples

@@
 ---
-title: "Edge effects and optimal field sizes in Alberta"
+title: "Edge effects in Alberta"
 author:
@@
 knitr::opts_chunk$set(echo = TRUE)
+
+library(tidyverse)
+
+datSource <- read.csv('../Data/datSource.csv') #Get data source
+datSourceFinal <- datSource %>% filter(use) %>% filter(grepl('Wheat|Canola',crop))
+yearsPerField <- datSourceFinal %>% count(field) %>% pull(n) #How many years of data at each unique location?
 
@@
 - Precision yield data were collected directly from farmers across Alberta
-  - Farmers were solicited for yield data through local agronomists, and we received data from a total of X growers across a total of X years
-  - Most fields represented only a single year of data, but we did sometimes receive multiple years of data from the same field
-  - File formats vary depending on the brand of combine, so we converted data to a standard format (csv) using Ag Leader SMS
-  - In total, we analyzed yield data from X field-years of data, containing a total of X million data points
+  - Farmers were solicited for yield data through local agronomists, and we received `r nrow(datSource)` field-years of data from `r length(unique(datSource$grower))` growers across a total of `r length(unique(datSource$year))` years (`r min(unique(datSource$year))`-`r max(unique(datSource$year))`)
+  - We converted data to a standard csv format using Ag Leader SMS
+  - `r round(100*sum(grepl('(Wheat|Canola)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_) or canola (_Brassica napus_), two of the most common crops in rotation in Alberta
+  - The remaining crop types were poorly replicated in our sample, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`) or canola (`r sum(datSourceFinal$crop=='Canola')`)
+  - Individual fields contained between `r min(yearsPerField)` to `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`)
+  - Containing a total of `r round(sum(datSource$npoints)/1e6,1)` million data points
 
@@
 
-  - To model this, we fit the following model:
+  - To model this, we fit the following model to each field-year of data:
 
   \begin{equation}
-  sqrt(yield) ~ Normal (\mu, \sigma)
-  \mu = Intercept + log(Polygon Size) + f(Distance from Edge_i, b=12)_i + f(Easting, Northing, b=60) + f(Harvest Sequence, b=60)
-  log(\sigma) = Intercept + log(Polygon Size) + f(Distance from Edge, b=12) + f(Easting, Northing, b=60) + f(Harvest Sequence, b=60)
+  \begin{split}
+  sqrt(yield) \sim & Normal (\mu, \sigma)\\
+  \mu = & Intercept + log(Polygon Size) + f(Distance from Edge_i, b=12)_i + \\
+   & f(Easting, Northing, b=60) + f(Sequence, b=60) \\
+  log(\sigma) = & Intercept + log(Polygon Size) + f(Distance from Edge, b=12) + \\
+   & f(Easting, Northing, b=60) + f(Sequence, b=60)
+  \end{split}
   \end{equation}
@@
   2. Distance from Edge = distance from field edge type _i_ (m)
-  3. Easting, Northing = distance from centroid of field (m)
-  4. Harvest Sequence = order of harvest within field (1--N points)
+  3. Easting, Northing = distance from centre of field (m)
+  4. Sequence = order of harvest within field (1--N points)
   5. f(x,b) = penalized thin-plate regression spline, where x is the predictor and b is the number of basis dimensions
@@
 - In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest
-  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (e.g.
-Kriging) or Stochastic Partial Differential Equations (e.g.
+  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (Kriging) or approximations such as Stochastic Partial Differential Equations (e.g.
 INLA), but this was computationally infeasible with 50,000 data points per field
-  - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with substantially reduced computation time
+  - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with reduced dimensionality and computation time
   - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
-  - A smoother was not used for polygon size because the relationship was easily approximated with a log-linear relationship using a single slope term
-  - All models were fit in _R_ using the _mgcv_ library, and figures were created with _ggplot2_ and _ggpubr_
+  - The relationship between polygon size (i.e. ground speed) and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions
+  - All models were fit in _R_ using the _mgcv_ library [version `r packageVersion('mgcv')`, @wood2017], and figures were created with _ggplot2_ and _ggpubr_ [versions `r packageVersion('ggplot2')`, @wickham2016; and `r packageVersion('ggpubr')`, @kassambara2020].
 
-- Most models still had a large amount of spatio-temporal autocorrelation in residuals
-  - We have 50000 data points per field, so we ignored it
-
-- To consider results from all field-level models while also accounting for field-level uncertainty, we used an approach similar to bootstrapping of hierarchical mixed effects models
-  - The approach was as follows
+- To consider results from all field-level models, we fit models independent of each other, and "overall" smoothers were taken as averages of the field-level smoothers
+  - However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models
   1. Extract single posterior sample (_rnorm_ in R) of smoother parameters from each field-level model using coefficient estimates and standard error
   2. Use posterior sample to create new simulated smoother from each field
-  3. Fit model of all simulated smoothers, and save this meta-smoother
+  3. Fit new model of simulated smoothers from all fields, and save this "meta-smoother"
   4. Repeat 1000 times, and calculate coverage intervals (5-95% percentiles) on saved meta-smoothers
-  - This gives credible intervals for the "average" smoother while accounting for field-level variability
+  - This gives coverage intervals (CIs) for the "average" smoother while accounting for field-level variability
 
@@
 
-Results here
+```{r , echo=FALSE, fig.cap='Fig1', out.width='100%'}  
+  include_graphics('../figures/frExample.png',dpi=NA)
+```
 

### 2021-07-10 Manuscript/manuscript.tex
Gave up on TMB
Doesn't work well enough for large numbers of samples

@@
             pdfauthor={},
-            pdftitle={Edge effects and optimal field sizes in Alberta},
+            pdftitle={Edge effects in Alberta},
             colorlinks=false,
@@
 
-  \title{Edge effects and optimal field sizes in Alberta}
+  \title{Edge effects in Alberta}
     \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
@@
 \item
-  Preserving SNL around fields is important for agriculture and conservation
+  Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth
 
@@
   \item
-    Ecosystem services
+    This has allows increases in human population and increased (global) stability in production
   \item
-    Act as reservoirs for beneficial insects
+    However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms
   \item
-    Microclimate zones
+    Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios
   \item
-    Poorly studied in North America
-  \item
-    Studies tend to be limited in scope (e.g.~only certain organisms/crops) and applicability
+    Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems
   \end{itemize}
 \item
-  Ecosystem services can influence both the mean and variability of yield in agroecosystems
+  SNL in and around crops is important for both agricultural production and conservation
 
@@
   \item
-    Typically only averages (means) are considered, but higher stability (lower variance) in yield is also valuable
+    They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control
+  \item
+    They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed
+  \item
+    Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
+  \item
+    In particular, North American agroecosystems have larger fields, and different varieties and agronomic practices, all of which could negate effects of SNL
   \end{itemize}
-\item
-  Field size and field boundaries are directly related to the conservation of SNL in agroecosystems
+\end{itemize}
 
@@
 \item
-    Size examples: wheat study from the UK
-  \item
-    Boundary examples: flower strip studies, hedgerows
-  \item
-    In North America, large financial incentives to make fields large and homogeneous, especially with large harvesting and planting equipment
-  \item
-    Crop edge effects cause low yields at the margins of fields because of late emergence, poor microclimate, and competition with weeds
-  \item
-    Ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
-  \item
-    Therefore, there should be a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
-  \end{itemize}
-\item
-  Precision yield data holds enormous promise for agronomy
+  SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate
 
@@
   \item
-    Limited because of:
+    Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds
   \item
-    Lack of standardized formats
+    At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges
   \item
-    Yearly calibration data
+    For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance
   \item
-    Clear statistical protocols
+    Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects
+  \item
+    This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
   \end{itemize}
+\end{itemize}
+
+\begin{itemize}
+\tightlist
+\item
+  Studies of SNL effects on crop yield also suffer from limited scope (e.g.~few crop types) and small sample sizes, limiting inference and reducing generality (but see wheat study)
+
+  \begin{itemize}
+  \tightlist
+  \item
+    For this reason, large-scale precision yield data holds enormous promise for agronomy, as it allows
+  \item
+    However, its use is limited for several reasons: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) unfamiliarity with spatial statistics
+  \item
+    Ecosystem services can influence both the mean and variability of yield in agroecosystems
+  \item
+    Typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable
+  \item
+    Size examples: wheat study from the UK
+  \item
+    There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield
+  \end{itemize}
+\item
+  In this paper we ask:
+
+  \begin{enumerate}
+  \def\labelenumi{\arabic{enumi}.}
+  \tightlist
+  \item
+    How does crop yield change with distance from the edge of field?
+  \item
+    Does this depend on type of field edge?
+  \item
+    Is there an intermediate distance where yield is maximized or variance is minimized?
+  \end{enumerate}
 \end{itemize}
@@
   \item
-    Farmers were solicited for yield data through local agronomists, and we received data from a total of X growers across a total of X years
+    Farmers were solicited for yield data through local agronomists, and we received 298 field-years of data from 5 growers across a total of 7 years (2014-2020)
   \item
-    Most fields represented only a single year of data, but we did sometimes receive multiple years of data from the same field
+    We converted data to a standard csv format using Ag Leader SMS
   \item
-    File formats vary depending on the brand of combine, so we converted data to a standard format (csv) using Ag Leader SMS
+    72\% of the crop types where either wheat (\emph{Triticum aestivum}) or canola (\emph{Brassica napus}), two of the most common crops in rotation in Alberta
   \item
-    In total, we analyzed yield data from X field-years of data, containing a total of X million data points
+    The remaining crop types were poorly replicated in our sample, so we constrained our analysis to only field-years containing wheat (94) or canola (119)
+  \item
+    Individual fields contained between 1 to 5 years of data (mean: 2.7)
+  \item
+    Containing a total of 18.4 million data points
   \end{itemize}
 \item
-  Yield data was collected in discrete rectangles of the same length as the data interval (distance = speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m)
+  Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m)
 
@@
   \item
-    Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location
+    Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
   \item
-    Seeding and application rates were constant with each field, so we did not consider inputs in our analysis
+    Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
+  \item
+    We used dry yield (tonnes of seed/hectare after accounting for crop moisture) as our measure of crop yield
+  \item
+    Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time
   \end{itemize}
 \item
-  Field boundaries were automatically digitized, then manually checked using satellite imagery and classified land cover data
+  Field boundaries were automatically digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC
 
@@
   \item
-    Crop boundaries are flexible, and often change from year to year depending on planting and emergence conditions (e.g.~flooding during some years)
+    Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g.~flooding during some years)
   \item
-    Additionally, seminatural features often change from year to year
+    Additionally, seminatural features often change yearly
 
@@
   \item
-    This makes accurate and consistent classification of field boundaries very difficult
+    This makes accurate and consistent classification of field boundaries difficult
   \item
-    Because of this, we used the following general categories for field boundaries:
+    We defined the following general categories for field boundaries:
 
@@
     \item
-      Standard: thin (\(<10\) m wide) grassy field boundary, often grassy road right-of-ways
+      Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
     \item
-      Wetland: permanent wetland, whose borders are largely unchanged from year-to-year
+      Wetland: permanent wetland; borders are largely unchanged from year-to-year
     \item
-      Forest/shrub: permanent windbreaks or remnant forests
+      Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
     \item
-      Grass: larger grassy area (pasture or permanent seminatural grassland)
+      Other crop: annual crop or pasture with little or no visible boundary between planted areas
     \item
-      Other crop: different crop with little or no visible boundary between planted areas
+      Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
     \item
-      Bare: unplanted (fallow), flooded area, or unpaved roads
+      Grassland: permanent seminatural grassland or pasture (not in rotation)
     \end{enumerate}
@@
 \begin{itemize}
-\tightlist
 \item
@@
   \begin{itemize}
-  \tightlist
   \item
-    Crop yield varies within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices such as strip farming
+    Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices
   \item
-    While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields), this does not address sensor drift over time that occurs within fields
+    Ground speed is extremely important to yield monitor accuracy (Arslan \& Colvin 2002), with low ground speed registering higher yields
   \item
-    The yield monitors may record lower yields as sensors accumulate debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
+    While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time within fields
   \item
-    Additionally, ground speed is known to be extremely important to yield monitor accuracy (Arslan \& Colvin 2002)
+    This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
   \item
-    To address this, we fit the following model:
+    To model this, we fit the following model to each field-year of data:
   \end{itemize}
@@
   \begin{equation}
-  ln(yield) 
+  \begin{split}
+  sqrt(yield) \sim & Normal (\mu, \sigma)\\
+  \mu = & Intercept + log(Polygon Size) + f(Distance from Edge_i, b=12)_i + \\
+   & f(Easting, Northing, b=60) + f(Sequence, b=60) \\
+  log(\sigma) = & Intercept + log(Polygon Size) + f(Distance from Edge, b=12) + \\
+   & f(Easting, Northing, b=60) + f(Sequence, b=60)
+  \end{split}
   \end{equation}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    where:
+  \end{itemize}
+
+  \begin{enumerate}
+  \def\labelenumi{\arabic{enumi}.}
+  \tightlist
+  \item
+    Polygon Size = distance traveled \(\times\) width of header bar (\(m^2\))
+  \item
+    Distance from Edge = distance from field edge type \emph{i} (m)
+  \item
+    Easting, Northing = distance from centre of field (m)
+  \item
+    Sequence = order of harvest within field (1--N points)
+  \item
+    f(x,b) = penalized thin-plate regression spline, where x is the predictor and b is the number of basis dimensions
+  \end{enumerate}
+\item
+  In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest
+
+  \begin{itemize}
+  \tightlist
+  \item
+    Spatial or temporal variation is typically modelled using a Gaussian Process Model (Kriging) or approximations such as Stochastic Partial Differential Equations (e.g.~INLA), but this was computationally infeasible with 50,000 data points per field
+  \item
+    Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with substantially reduced computation time
+  \item
+    The number of basis dimensions was checked with the \emph{gam.check} function from \emph{mgcv}
+  \item
+    The relationship between polygon size (i.e.~ground speed) and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions
+  \item
+    All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).
+  \end{itemize}
+\item
+  To consider results from all field-level models, we fit models independent of each other, and ``overall'' smoothers were taken as averages of the field-level smoothers
+
+  \begin{itemize}
+  \tightlist
+  \item
+    However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models
+  \end{itemize}
+
+  \begin{enumerate}
+  \def\labelenumi{\arabic{enumi}.}
+  \tightlist
+  \item
+    Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard error
+  \item
+    Use posterior sample to create new simulated smoother from each field
+  \item
+    Fit new model of simulated smoothers from all fields, and save this ``meta-smoother''
+  \item
+    Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers
+  \end{enumerate}
+
+  \begin{itemize}
+  \tightlist
+  \item
+    This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability
+  \end{itemize}
 \end{itemize}
@@
 
+\leavevmode\hypertarget{ref-kassambara2020}{}%
+Kassambara, A. (2020). \emph{{ggpubr}: 'ggplot2' based publication ready plots}.
+
+\leavevmode\hypertarget{ref-wickham2016}{}%
+Wickham, H. (2016). \emph{{ggplot2}: Elegant graphics for data analysis}. Springer-Verlag New York.
+
+\leavevmode\hypertarget{ref-wood2017}{}%
+Wood, S.N. (2017). \emph{Generalized additive models: An introduction with {R}}. CRC press.
+
 \end{CSLReferences}

### 2021-07-10 gam_TMB/gam_tmb_tp.R

@@ -1,0 +1,48 @@
+library(mgcv)
+library(TMB)
+library(ggplot2)
+
+set.seed(1)
+
+#Make 20000 sampling locations
+N <- 20000
+E <- runif(N,-10,10)
+N <- runif(N,-10,10)
+dat <- data.frame(y=0,E,N) #Fill in y with zeros for now
+
+s0 <- gam(y~s(E,N,bs='ts',k=60),data=dat,fit=FALSE) #Create model matrix
+modMat <- s0$X[,-1] #Model matrix (no intercept)
+penMat <- s0$smooth[[1]]$S[[1]] #Penalty matrix
+penMat <- as(penMat,'sparseMatrix') #Change to sparse matrix
+penDim <- nrow(penMat) #Dimensions of penalty matrix
+
+b0 <- 1 #Intercept
+(smoothCoefs <- rnorm(ncol(modMat))) #Smoothing coefficients
+sigma <- 2 #Residual SD
+yhat <- b0 + modMat %*% smoothCoefs #Mean values at each location
+dat$y <- rnorm(N,yhat,sigma) #Generate data
+dat$field <- modMat %*% smoothCoefs #Spatial field
+
+ggplot(dat)+geom_point(aes(x=E,y=N,col=field)) #Plot spatial field (no noise)
+ggplot(dat)+geom_point(aes(x=E,y=N,col=y)) #Plot data
+
+#Input for model
+datList <- list(y=dat$y, smoothMat=modMat, penaltyMat=penMat, penaltyDim=penDim) #Data
+parList <- list(b0=0, smoothCoefs=rep(0,ncol(modMat)), log_lambda=0, log_sigma=0) #Parameters
+
+compile("gam_tmb_tp.cpp")
+dyn.load(dynlib("gam_tmb_tp"))
+obj <- MakeADFun(data = datList, parameters = parList, random=c('smoothCoefs'), DLL = "gam_tmb_tp")
+
+#Memory usage spikes from 4.5GB to 16GB during Optimizing Tape stage
+opt <- nlminb(obj$par,obj$fn,obj$gr) #Run model
+report <- sdreport(obj)
+
+opt$convergence==0 #Did model converge?
+report #Looks OK. b0 and sigma are very close to actual values (1 and 2)
+
+#Estimates smoothing coefficients fairly well
+ggplot(data.frame(predicted=report$par.random,actual=smoothCoefs))+
[…]

### 2021-07-10 gam_TMB/gam_tmb_tp.cpp

@@ -1,0 +1,39 @@
+#include <TMB.hpp>  
+template<class Type>
+Type objective_function<Type>::operator() (){
+  using namespace density;
+  using namespace Eigen; 
+  
+  //2D thin-plate regression spline from mgcv
+  
+  //Read data from R------------------
+  DATA_VECTOR(y);                     // Response
+  DATA_MATRIX(smoothMat);             // Design matrix (no intercept)
+  DATA_SPARSE_MATRIX(penaltyMat);     // Penalty matrix
+  DATA_INTEGER(penaltyDim);           // Dimensions of penalty matrix
+  
+  //Read parameters from R------------
+  PARAMETER(b0);                      // Intercept
+  PARAMETER_VECTOR(smoothCoefs);      // Smooth coefficients
+  PARAMETER(log_lambda);              // Log penalization parameter
+  PARAMETER(log_sigma);               // Log(SD) of measurement error 
+  
+  //Transform SD and penalization parameters 
+  Type sigma = exp(log_sigma);
+  Type lambda = exp(log_lambda);
+  
+  //Initialize objective function
+  Type nll=0;
+  
+  // Penalization ------------------------------------
+  
+  SparseMatrix<Type> S = lambda*penaltyMat; //Penalty term * Penalty matrix
+  nll -= 0.5*penaltyDim*log_lambda - 0.5*GMRF(S).Quadform(smoothCoefs);
+  
+  // Main model -------------------------------------------
+  vector<Type> mu = b0 + smoothMat*smoothCoefs; //Expected value
+  
+  nll -= sum(dnorm(y, mu, sigma, true)); //Decrement logLik
+  
+  return nll; //Return likelihood
+}

### 2021-07-10 gam_TMB/mod0A.cpp

@@ -34,7 +34,16 @@ Type objective_function<Type>::operator() (){
   //GMRF needs a sparse matrix - This approach avoids logdet calculation
   SparseMatrix<Type> S = lambda*penaltyMat; //Penalty term * Penalty matrix
   // Penalty that works with bs = 'ts'/'cs': extra shrinkage
-  nll -= 0.5*penaltyDim*log_lambda - 0.5*GMRF(S).Quadform(smoothCoefs);  
+  nll -= 0.5*penaltyDim*log_lambda - 0.5*GMRF(S).Quadform(smoothCoefs);
+  
+  // //Second approach using 
+  // //2nd approach using precision matrix and GMFR - taken from spatialALK example
+  // // SparseMatrix<Type> sparseLambdaDiag = lambdaDiag.sparseView(); //Turn into sparse diagonal matrix of lambdas
+  // SparseMatrix<Type> lambdaS = penaltyMat*lambda; //Penalty matrices %*% lambda matrix = precision matrix: lambdaS
+  // 
+  // //NOTE: GMRF returns negative log-likelihood, so must be added to objective function
+  // nll += GMRF(lambdaS,false)(smoothCoefs); //Likelihood of beta, given precision matrix lambdaS
+  // 
 
   // Main model -------------------------------------------
   vector<Type> mu = b0 + logArea*b_area + smoothMat*smoothCoefs; //Expected value: mu = Intercept + model matrix %*% coefs

### 2021-07-10 gam_TMB/temp.R

@@ -1,0 +1,2 @@
+library(TMB)
+compile("mod1B.cpp")

### 2021-07-10 gam_TMB/yieldTMBtest.R

@@ -1,5 +1,7 @@
 #Tests of R-INLA models for yield, using SPDE approx of Matern covariance
 
+#Jul 8, 2021: Tried getting thin-plate splines to run, but memory usage spikes during Optimizing Tape routine. Only real option is SPDE at this point, and I'm not sure how to get "slowly changing" variance estimates to work using SPDE
+
 #Models to try:
 # 0A. Yield ~ s(dist) - done
 # 0B. Yield ~ SPDE(E,N)
@@ -156,7 +158,6 @@ data.frame(dist=predDist,fit=report$value,se.fit=report$sd) %>%
 # ggarrange(p1,p2)
 
 #Try TMB model with s(E,N) instead of s(Dist)
-
 #Get smoother specifications
 s0 <- gam(DryYield ~ s(E,N,bs='ts',k=60), data = dat, fit = FALSE)
 
@@ -200,23 +201,61 @@ dat %>% mutate(resid=obj$report()$residuals) %>% #Weird variogram
   ggplot(aes(x=dist,y=gamma))+geom_point()+geom_line()+
   labs(title='Residual variogram',x='Distance',y='Semivariance')
 
-dat %>% mutate(yield=sqrt(DryYield)-mean(sqrt(DryYield))) %>% 
-  select(.,ID,yield) %>% as_Spatial(.) %>% 
-  gstat::variogram(yield~1, data=.,width=10,cutoff=150) %>% #Variogram
+#Predictions
+data.frame(dist=predDF_spatial,fit=report$value) %>% 
+  mutate(across(c(fit),~.x^2)) %>% 
+  ggplot()+geom_sf(aes(geometry=dist.geometry,col=fit),size=3)
+
+#Try TMB model with s(r) instead of s(Dist)
+#Get smoother specifications
+s0 <- gam(DryYield ~ s(r,bs='ts',k=60), data = dat, fit = FALSE)
+
+#Input
+modMat <- s0$X[,-1] #Model matrix (no intercept)
+penMat <- s0$smooth[[1]]$S[[1]] #Penalty matrix (sparse)
+penMat <- as(penMat,'sparseMatrix')
+penDim <- nrow(penMat) #Dimensions of penalty matrix
+
+predDF_temporal <- with(dat,data.frame(r=seq(min(r),max(r),length.out=1000)))
[…]

### 2021-07-10 helperFunctions.R

@@ -156,6 +156,7 @@ getSmooths <- function(smoothLabel,modList,xvals,postSamp=FALSE,noIntercept=TRUE
 # l = number of replicates in each smooth prediction (length.out)
 # margInt = marginalize across intercepts
 # samp = sample from posterior distribution of coefficients rather than using the mean
+
 getPreds <- function(path,l=c(30,100,500),margInt=c(FALSE,FALSE,FALSE),samp=FALSE){
   
   # #Debugging
@@ -164,7 +165,6 @@ getPreds <- function(path,l=c(30,100,500),margInt=c(FALSE,FALSE,FALSE),samp=FALS
   # path <- paste0('./Figures/ModelCheck2/',datSource$filename[2],' modList.Rdata') #Model type 2
   # margInt <- c(FALSE,TRUE,TRUE)
   # samp <- FALSE
-  # 
   require(mgcv)
   
   load(path) #Load data
@@ -261,9 +261,58 @@ getPreds <- function(path,l=c(30,100,500),margInt=c(FALSE,FALSE,FALSE),samp=FALS
 # 
 # ggplot()+geom_line(data=temp$r,aes(x=r,y=mean))+ #Works for this
 #   geom_line(data=temp2$r,aes(x=r,y=mean),linetype='dashed',col='red')
-
 # (temp <- getPreds(paste0('./Figures/ModelCheck2/',datSource$filename[2],' modList.Rdata'),margInt=c(FALSE,TRUE,TRUE),samp=FALSE)) #Test with second type of model
 
+
+#Shorter version that only gets estimates of log(pArea) relationship and turns them into speed measurements
+# Combine ground speed is talked about a lot, so this is more appropriate for growers/agronomists
+
+getPredsSpeed <- function(path,l=30,margInt=FALSE,samp=FALSE, minSpeed = NA, maxSpeed = NA, width = NA, speed2Distance = NA){
+  
+  # #Debugging
+  # l <- 30
+  # # path <- paste0('./Figures/ModelCheck1/',datSource$filename[2],' modList.Rdata') #Model type 1
+  # path <- paste0('./Figures/ModelCheck2/',datSource$filename[2],' modList.Rdata') #Model type 2
+  # minSpeed <- 0.1
+  # maxSpeed <- 15 
+  # margInt <- FALSE
+  # samp <- FALSE
+  # width <- 10.21
+  # speed2Distance <- 1/18.043
+  
[…]

### 2021-07-10 runModels.R

@@ -1,4 +1,4 @@
-#Run non-isotropic GAMs on multiple field years
+#Run non-stationary GAMs on multiple field years
 
 # Load everything ---------------------------------------------------------
 
@@ -31,6 +31,8 @@ source('helperFunctions.R')
 #   set_names(datSource$filename) %>% 
 #   sapply(.,function(x) names(x)[1])
 # 
+# 
+# 
 # datSource$crop <- case_when(grepl('(CANOLA|InVigor|Dekalb 74-54|Dekalb 75-42|Brett Young 6056|Dupont D3152|VT 9562)',cropType) ~ 'Canola',
 #                             grepl('(WHEAT|Wheat|CPS AC Foremost|CWRS|CWSW)',cropType) ~ 'Wheat',
 #                             grepl('Peas',cropType) ~ 'Peas',
@@ -39,6 +41,7 @@ source('helperFunctions.R')
 #                             grepl('Oats',cropType) ~ 'Oats',
 #                             grepl('Rye',cropType) ~ 'Rye',
 #                             TRUE ~ cropType) 
+# datSource$npoints <- sapply(datSource$dataPath,function(x) length(readLines(x)))
 # write.csv(datSource,'./Data/datSource_new.csv',row.names = FALSE)
 
 datSource <- read.csv('./Data/datSource.csv') %>% #Read previous datasource file
@@ -49,9 +52,9 @@ datSource <- read.csv('./Data/datSource.csv') %>% #Read previous datasource file
 
 #Get summary results from models
 datSource %>% pull(use) %>% sum()
-datSource %>% count(grower,year) #How many fields per grower per year
+datSource %>% filter(use) %>% count(grower,year) #How many fields per grower per year
 datSource %>% group_by(grower) %>% summarize(nLoc=length(unique(field)),nYears=length(unique(year)),n=n()) #Number of distinct fields 
-datSource %>% count(crop) #Enough to do separate analyses on canola and wheat, maybe peas
+datSource %>% filter(use) %>% count(crop) #Enough to do separate analyses on canola and wheat, maybe peas
 
[…]

### 2021-07-26 gam_TMB/data_update.R
Added TMB examples
- Conclusion: TMB can't really handle large datasets :(

@@ -1,0 +1,54 @@
+## Demonstrate how to change data without re-taping
+library(TMB)
+setwd("~/Documents/yield-analysis-2021/gam_TMB")
+## Compile and load example
+compile("data_update.cpp")
+dyn.load(dynlib("data_update"))
+
+## Generate random objects
+set.seed(123)
+a <- rep(0, 2)
+b <- array(0, c(2,3,4) )
+c <- matrix(0,2,2)
+d <- 1
+x <- rnorm(1)
+data <- list(a=a,b=b,c=c,d=d)
+parameters <- list(x=x)
+
+## Check that data objects can be changed from R
+obj <- MakeADFun(data=data,parameters=parameters,DLL="data_update")
+
+## Function to check value
+check <- function(x) {
+  res.true <- -dnorm(x, 0, sum(unlist(obj$env$data)), TRUE)
+  res.obs  <- obj$fn(x)
+  abs.diff <- abs(res.true - res.obs)
+  stopifnot( abs.diff < 1e-10 )
+  data.frame(res.true, res.obs, abs.diff)
+}
+
+## No change made yet:
+check(x=1)
+check(x=2)
+
+## Change 'a' without re-taping
+obj$env$data$a <- obj$env$data$a + .1
+check(x=1) #Estimates are essentially identical
+check(x=2)
+
+## Change 'b' without re-taping
+obj$env$data$b <- obj$env$data$b + .01
+check(x=1) #Estimates are essentially identical
+check(x=2)
+
+## Change 'c' without re-taping
+obj$env$data$c <- obj$env$data$c + .05
+check(x=1) #Estimates are essentially identical
+check(x=2)
+
+## Change 'd' without re-taping
+obj$env$data$d <- obj$env$data$d + .1
+check(x=1) #Estimates are essentially identical
+check(x=2)
+
+

### 2021-07-26 gam_TMB/data_update.cpp

@@ -1,0 +1,15 @@
+// Demonstrate how to change data without re-taping
+#include <TMB.hpp>
+template<class Type>
+Type objective_function<Type>::operator() ()
+{
+  DATA_VECTOR(a); DATA_UPDATE(a);
+  DATA_ARRAY (b); DATA_UPDATE(b);
+  DATA_MATRIX(c); DATA_UPDATE(c);
+  DATA_SCALAR(d); DATA_UPDATE(d);
+  PARAMETER  (x);
+  
+  Type sd = a.sum() + b.sum() + c.sum() + d;
+  Type nll = -dnorm(x, Type(0), sd, true);
+  return nll;
+}

### 2021-07-26 gam_TMB/gam_tmb_batches.R

@@ -2,7 +2,6 @@ library(mgcv)
 library(TMB)
 library(ggplot2)
 library(ggpubr)
-
 set.seed(1)
 
 #Make 20000 sampling locations
@@ -31,19 +30,25 @@ ggplot(dat)+geom_point(aes(x=E,y=N,col=field)) #Plot spatial field (no noise)
 ggplot(dat)+geom_point(aes(x=E,y=N,col=y)) #Plot data
 
 #Input for model
-datList <- list(y=dat$y, smoothMat=modMat, penaltyMat=penMat, penaltyDim=penDim,flag=1) #Data
+
+#Do matrix multiplication in 5 "batches"
+nbatches <- 5
+batches <- rep(1:nbatches,each=nrow(dat)/nbatches)
+bStart <- sapply(1:nbatches,function(x) min(which(batches==x)))-1 #Minus one because of c++ indexing
+bLength <- nrow(dat)/nbatches #How long is each batch?
+datList <- list(y=dat$y, smoothMat=modMat, penaltyMat=penMat, penaltyDim=penDim,bStart=bStart,bLength=bLength,nbatches=nbatches) #Data
 parList <- list(b0=0, smoothCoefs=rep(0,ncol(modMat)), log_lambda=0, log_sigma=1) #Parameters
 
-compile("gam_tmb_tp.cpp")
-dyn.load(dynlib("gam_tmb_tp"))
-obj <- MakeADFun(data = datList, parameters = parList, random=c('smoothCoefs'), DLL = "gam_tmb_tp")
+compile("gam_tmb_batches.cpp")
+dyn.load(dynlib("gam_tmb_batches"))
+obj <- MakeADFun(data = datList, parameters = parList, random=c('smoothCoefs'), DLL = "gam_tmb_batches")
 
 #Memory usage spikes from 4.5GB to 16GB during Optimizing Tape stage
 opt <- nlminb(obj$par,obj$fn,obj$gr) #Run model
 report <- sdreport(obj)
 
 opt$convergence==0 #Did model converge?
-report #Looks OK. b0 and sigma are very close to actual values (1 and 2)
+report #Looks OK. b0 and logsigma are very close to actual values (1 and 0.693)
 
 #Estimates smoothing coefficients fairly well
 
@@ -57,5 +62,3 @@ p2 <- ggplot(data.frame(predicted=report$par.random,actual=smoothCoefs))+
[…]

### 2021-07-26 gam_TMB/gam_tmb_batches.cpp

@@ -11,14 +11,15 @@ Type objective_function<Type>::operator() (){
   DATA_MATRIX(smoothMat);             // Design matrix (no intercept)
   DATA_SPARSE_MATRIX(penaltyMat);     // Penalty matrix
   DATA_INTEGER(penaltyDim);           // Dimensions of penalty matrix
-  DATA_INTEGER(flag);                 // Which penalty type should be used?
+  DATA_VECTOR(bStart);               // Start of batches
+  DATA_INTEGER(bLength);              // Length of batches
+  DATA_INTEGER(nbatches);             // Number of batches
   
   //Read parameters from R------------
   PARAMETER(b0);                      // Intercept
   PARAMETER_VECTOR(smoothCoefs);      // Smooth coefficients
   PARAMETER(log_lambda);              // Log penalization parameter
   PARAMETER(log_sigma);               // Log(SD) of measurement error 
-  
   
   //Transform SD and penalization parameters 
   Type sigma = exp(log_sigma);
@@ -30,16 +31,21 @@ Type objective_function<Type>::operator() (){
   // Penalization ------------------------------------
   
   SparseMatrix<Type> S = lambda*penaltyMat; //Penalty term * Penalty matrix
-  if(flag==1){
-    nll -= 0.5*penaltyDim*log_lambda - 0.5*GMRF(S).Quadform(smoothCoefs);
-  } else {
-    nll += 0.5*(smoothCoefs * (S * smoothCoefs.matrix()).array()).sum();
-  }
+  nll -= 0.5*penaltyDim*log_lambda - 0.5*GMRF(S).Quadform(smoothCoefs);
   
   // Main model -------------------------------------------
-  vector<Type> mu = b0 + smoothMat*smoothCoefs; //Expected value
+  int k = 0; //Counter
+  int nCoefs = smoothCoefs.size(); //Number of smoothing coefficients
   
-  nll -= sum(dnorm(y, mu, sigma, true)); //Decrement logLik
+  for(int i=0;i<nbatches;i++){ //For each batch
+    matrix<Type> sMatBlock = smoothMat.block(k,0,bLength,nCoefs); //Pull out a block of model matrix starting at 0,k going for bLength rows, nCoefs columns
+    vector<Type> ySeg = y.segment(k,bLength); //Pull out data for that block
+    vector<Type> muSeg = b0 + sMatBlock*smoothCoefs; //Expected value for that block
+    
[…]

### 2021-07-26 gam_TMB/gam_tmb_batches2.R

@@ -1,0 +1,124 @@
+library(mgcv)
+library(TMB)
+library(ggplot2)
+library(ggpubr)
+set.seed(1)
+setwd("~/Documents/yield-analysis-2021/gam_TMB")
+
+#Make 20000 sampling locations
+N <- 20000
+E <- runif(N,-10,10)
+N <- runif(N,-10,10)
+dat <- data.frame(y=0,E,N) #Fill in y with zeros for now
+
+s0 <- gam(y~s(E,N,bs='ts',k=60),data=dat,fit=FALSE) #Create model matrix
+modMat <- s0$X[,-1] #Model matrix (no intercept)
+penMat <- s0$smooth[[1]]$S[[1]] #Penalty matrix
+penMat <- as(penMat,'sparseMatrix') #Change to sparse matrix
+penDim <- nrow(penMat) #Dimensions of penalty matrix
+
+b0 <- 1 #Intercept
+smoothCoefs <- rnorm(ncol(modMat)) #Smoothing coefficients
+
+sigma <- 2 #Residual SD
+yhat <- b0 + modMat %*% smoothCoefs #Mean values at each location
+dat$y <- rnorm(N,yhat,sigma) #Generate data
+dat$field <- modMat %*% smoothCoefs #Spatial field
+
+ggplot(dat)+geom_point(aes(x=E,y=N,col=field)) #Plot spatial field (no noise)
+ggplot(dat)+geom_point(aes(x=E,y=N,col=y)) #Plot data
+
+#Input for model
+
+#Fit model using 5 "batches" of data
+nbatches <- 5
+batches <- rep(1:nbatches,each=nrow(dat)/nbatches)
+
+#Data, only using first batch
+datList <- list(y=dat$y[batches==1], smoothMat=modMat[batches==1,], penaltyMat=penMat, penaltyDim=penDim,usePenalty=1.0) 
+parList <- list(b0=0, smoothCoefs=rep(0,ncol(modMat)), log_lambda=0, log_sigma=1) #Parameters
+
+compile("gam_tmb_batches2.cpp")
+dyn.load(dynlib("gam_tmb_batches2"))
+obj <- MakeADFun(data = datList, parameters = parList, random=c('smoothCoefs'), DLL = "gam_tmb_batches2")
+
+obj$fn() #Objective function works. Spike in memory during Optimizing Tape much reduced from before
+
+#Make new objective function that accumulates values from all batches
+oFun <- function(params){ 
[…]

### 2021-07-26 gam_TMB/gam_tmb_batches2.cpp

@@ -7,18 +7,17 @@ Type objective_function<Type>::operator() (){
   //2D thin-plate regression spline from mgcv
   
   //Read data from R------------------
-  DATA_VECTOR(y);                     // Response
-  DATA_MATRIX(smoothMat);             // Design matrix (no intercept)
-  DATA_SPARSE_MATRIX(penaltyMat);     // Penalty matrix
-  DATA_INTEGER(penaltyDim);           // Dimensions of penalty matrix
-  DATA_INTEGER(flag);                 // Which penalty type should be used?
+  DATA_VECTOR(y); DATA_UPDATE(y);                     // Response
+  DATA_MATRIX(smoothMat); DATA_UPDATE(smoothMat);     // Design matrix (no intercept)
+  DATA_SPARSE_MATRIX(penaltyMat);                     // Penalty matrix
+  DATA_INTEGER(penaltyDim);                           // Dimensions of penalty matrix (ie number of smoothing coefficients)
+  DATA_SCALAR(usePenalty); DATA_UPDATE(usePenalty);  // Should penalty be calculated?
   
   //Read parameters from R------------
   PARAMETER(b0);                      // Intercept
   PARAMETER_VECTOR(smoothCoefs);      // Smooth coefficients
   PARAMETER(log_lambda);              // Log penalization parameter
   PARAMETER(log_sigma);               // Log(SD) of measurement error 
-  
   
   //Transform SD and penalization parameters 
   Type sigma = exp(log_sigma);
@@ -28,17 +27,12 @@ Type objective_function<Type>::operator() (){
   Type nll=0;
   
   // Penalization ------------------------------------
-  
   SparseMatrix<Type> S = lambda*penaltyMat; //Penalty term * Penalty matrix
-  if(flag==1){
-    nll -= 0.5*penaltyDim*log_lambda - 0.5*GMRF(S).Quadform(smoothCoefs);
-  } else {
-    nll += 0.5*(smoothCoefs * (S * smoothCoefs.matrix()).array()).sum();
-  }
+  nll -= (0.5*penaltyDim*log_lambda - 0.5*GMRF(S).Quadform(smoothCoefs))*usePenalty;  
   
   // Main model -------------------------------------------
-  vector<Type> mu = b0 + smoothMat*smoothCoefs; //Expected value
   
+  vector<Type> mu = b0 + smoothMat*smoothCoefs; //Expected value
   nll -= sum(dnorm(y, mu, sigma, true)); //Decrement logLik
   
   return nll; //Return likelihood

### 2021-07-26 gam_TMB/mod1B.cpp

@@ -2,23 +2,24 @@
 template<class Type>
 Type objective_function<Type>::operator() (){
   using namespace density; //includes GMRF
-  // using namespace R_inla; //includes SPDE-specific functions, e.g. Q_spde()
+  using namespace R_inla; //includes SPDE-specific functions, e.g. Q_spde()
   using namespace Eigen; //Needed for sparse structures
                       
   //Read data from R------------------
-  DATA_VECTOR(yield);                        // Response
-  DATA_VECTOR(logArea);                      // log-Area of harvested polygons 
-  DATA_SCALAR(meanLogArea);                  // Mean of log(area)
-  DATA_MATRIX(smoothMat);                    // Design matrix (no intercept)
-  DATA_SPARSE_MATRIX(penaltyMat);            // Penalty matrix - must be sparse for GMRF to work
-  DATA_IVECTOR(penaltyDim);                  // Dimensions of penalty sub-matrices
-  // DATA_MATRIX(predModMat_dist);              // New smoothing matrices for predictions
-  // DATA_MATRIX(predModMat_spatial);
-  // DATA_MATRIX(predModMat_temporal);
-  // DATA_STRUCT(spatialSPDEMatrices,spde_t);   // Matrices needed for representing the spatial GMRF
-  // DATA_SPARSE_MATRIX(spatialA);              // Matrix for interpolating points within spatial mesh
-  // DATA_STRUCT(temporalSPDEMatrices,spde_t);  // Matrices needed for representing the temporal GMRF
-  // DATA_SPARSE_MATRIX(temporalA);             // Matrix for interpolating points within temporal mesh  
+  DATA_VECTOR(yield); DATA_UPDATE(yield);           // Response
+  DATA_VECTOR(logArea); DATA_UPDATE(logArea);       // log-Area of harvested polygons 
+  DATA_SCALAR(meanLogArea);                         // Mean of log(area)
+  DATA_MATRIX(smoothMat); DATA_UPDATE(smoothMat);   // Design matrix (no intercept)
+  DATA_SPARSE_MATRIX(penaltyMat);                   // Penalty matrix - must be sparse for GMRF to work
+  DATA_IVECTOR(penaltyDim);                         // Dimensions of penalty sub-matrices
[…]

### 2021-07-26 gam_TMB/yieldTMBtest.R

@@ -639,36 +639,43 @@ temporalSPDEMatrices <- temporalSPDE$param.inla[c("M0","M1","M2")] #Matrices nee
 dim(temporalA) #4999 vertices
 
 #Input for model:
+
+#Divide data into 10 batches
+nbatches <- 10
+batches <- rep(1:nbatches,each=nrow(dat)/nbatches)
+use <- batches==1
+
 #Data
-datList <- list(yield=sqrt(dat$DryYield), #Transformed Data
-                logArea=log(dat$pArea), meanLogArea = mean(log(dat$pArea)), #log(area) and mean(log(area))
-                smoothMat=modMat, penaltyMat=penMat, penaltyDim=penDim) #Smoother input
-                # predModMat_dist=predModMat_dist, #New smoother matrices to predict on
-                # predModMat_spatial=predModMat_spatial,
-                # predModMat_temporal=predModMat_temporal,
-                # spatialSPDEMatrices=spatialSPDEMatrices, spatialA = spatialA, #Spatial SPDE input
-                # temporalSPDEMatrices = temporalSPDEMatrices, temporalA = temporalA) #Temporal SPDE input
+datList <- list(yield=sqrt(dat$DryYield)[use], #Transformed Data
+                logArea=log(dat$pArea)[use], meanLogArea = mean(log(dat$pArea)), #log(area) and mean(log(area))
+                smoothMat=modMat[use,], penaltyMat=penMat, penaltyDim=penDim, #Smoother input
+                predModMat_dist=predModMat_dist, #New smoother matrices to predict on
+                predModMat_spatial=predModMat_spatial,
+                predModMat_temporal=predModMat_temporal,
+                spatialSPDEMatrices=spatialSPDEMatrices, spatialA = as.matrix(spatialA[use,]), #Spatial SPDE input
+                temporalSPDEMatrices = temporalSPDEMatrices, temporalA = as.matrix(temporalA[use,]),
+                usePenalty=1.0) #Temporal SPDE input
 #Parameters
 parList <- list(b0=0, #Intercept
                 b_areaMean=0,
                 smoothCoefsMean=rep(0,ncol(modMat)), log_lambdasMean=rep(0,length(penDim)), #Smooth coefficients and penalization parameter for means
-                b0SD=0 #log-SD intercept
-                # b_areaSD=0
[…]

### 2021-07-27 Manuscript/manuscript.Rmd
Worked on manuscript and data filtering

@@
 - SNL in and around crops is important for both agricultural production and conservation
-  - They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control
-  - They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed
+  - They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control (albrecht2020)
+  - They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (kort1988)
   - Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
-  - In particular, North American agroecosystems have larger fields, and different varieties and agronomic practices, all of which could negate effects of SNL
+  - In particular, North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL
 
@@
   - Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds
-  - At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges
+  - At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (kowalchuk1995)
   - For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance
-  - Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects 
+  - Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (e.g. vanVooren2017)
   - This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
@@
 - Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality (but see wheat study)
-  - For this reason, large-scale precision yield data holds enormous promise for agronomy, as it allows 
+  - Meta-analyses have been the standard solution to this problem, but precision yield data holds enormous promise
   - However, its use is limited for several reasons: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) unfamiliarity with spatial statistics
-  - Ecosystem services can influence both the mean and variability of yield in agroecosystems
+  - Ecosystem services can influence both the mean and variability of yield in agroecosystems (redhead2020)
   - Typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable
-  - Size examples: wheat study from the UK
   - There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield
@@
   - Farmers were solicited for yield data through local agronomists, and we received `r nrow(datSource)` field-years of data from `r length(unique(datSource$grower))` growers across a total of `r length(unique(datSource$year))` years (`r min(unique(datSource$year))`-`r max(unique(datSource$year))`)
-  - We converted data to a standard csv format using Ag Leader SMS
+  - We converted raw data to a standard csv format using Ag Leader SMS
   - `r round(100*sum(grepl('(Wheat|Canola)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_) or canola (_Brassica napus_), two of the most common crops in rotation in Alberta
-  - The remaining crop types were poorly replicated in our sample, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`) or canola (`r sum(datSourceFinal$crop=='Canola')`)
+  - The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`) or canola (`r sum(datSourceFinal$crop=='Canola')`)
   - Individual fields contained between `r min(yearsPerField)` to `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`)
@@
   - Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
-  - We used dry yield (tonnes of seed/hectare after accounting for crop moisture) as our measure of crop yield
+  - We used dry yield (tonnes of seed/hectare minus recorded crop moisture) as our measure of crop yield
+  - Precision yield data can be highly variable and is prone to extreme outliers (both positive and negative), especially when combine ground speeds are low and at the beginning of rows
+  - Therefore, we filtered data that lay outside the 95th percentiles of yield within each field, values less than or equal to zero, and values outside of the 95th percentiles of ground speed (high: 
   - Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time
@@
 
-- At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor
+- Yield data from each field contains both random and systematic errors that must be modeled in order to reveal underlying changes in yield
+  - We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the data
   - Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices 
@@
   - Ground speed is extremely important to yield monitor accuracy [@arslan2002], with low ground speed registering higher yields 
-  - While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time  within fields
+  - Sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time
   - This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
@@
 - In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest
-  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (Kriging) or approximations such as Stochastic Partial Differential Equations (e.g.
-INLA), but this was computationally infeasible with 50,000 data points per field
-  - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with reduced dimensionality and computation time
+  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (e.g. Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field
+  - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but using a reduced number of dimensions (as well as computation time)
   - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
@@
 
-- To consider results from all field-level models, we fit models independent of each other, and "overall" smoothers were taken as averages of the field-level smoothers
+- To consider results from all field-level models together, we fit models independent of each other and fit an "overall" smoother as averages of the field-level smoothers
   - However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models
   1. Extract single posterior sample (_rnorm_ in R) of smoother parameters from each field-level model using coefficient estimates and standard error
-  2. Use posterior sample to create new simulated smoother from each field
+  2. Use posterior sample to simulated a new smoother from each field
   3. Fit new model of simulated smoothers from all fields, and save this "meta-smoother"

### 2021-07-27 filtering.R

@@ -1,0 +1,193 @@
+#Seems to be a consensus that filter/cleaning is necessary for yield data. Trying to see how much of a difference this makes for our models
+
+library(tidyverse)
+theme_set(theme_classic())
+library(sf)
+library(mgcv)
+library(beepr)
+library(ggpubr)
+setwd("~/Documents/yield-analysis-2021/gam_TMB")
+source('../helperFunctions.R')
+
+nSubSamp <- 50000 #Number of points to start with
+csvPath <- "/media/rsamuel/Storage/geoData/Rasters/yieldData/csv files/Alvin_French C41 2020.csv"
+boundaryPath <- "/home/rsamuel/Documents/yield-analysis-2021/Figures/FieldBoundaries/Alvin_French C41 2020 boundary.shp"
+boundaryPath2 <- "/home/rsamuel/Documents/yield-analysis-2021/Figures/FieldBoundarySegments/Alvin_French C41 2020 boundary.shp"
+
+#Also try: Gibbons Bill Visser 2017 (bad residual structure)
+
+dat <- read.csv(csvPath,stringsAsFactors=TRUE,fileEncoding='latin1') 
+
+if(nrow(dat)>nSubSamp){
+  dat <- dat %>% slice(round(seq(1,nrow(dat),length.out=nSubSamp)))
+}
+
+dat <- dat %>% rename_with(.fn = ~gsub('..L.ha.$','_lHa',.x)) %>%
+  rename_with(.fn = ~gsub('..tonne.ha.$','_tHa',.x)) %>%
+  rename_with(.fn = ~gsub('..m.s.$','_ms',.x)) %>%
+  rename_with(.fn = ~gsub('.km.h.$','_kmh',.x)) %>%
+  rename_with(.fn = ~gsub('..tonne.h.$','_th',.x)) %>%
+  rename_with(.fn = ~gsub('.ha.h.','_hah',.x)) %>%
+  rename_with(.fn = ~gsub('.m.','_m',.x)) %>%
+  rename_with(.fn = ~gsub('.deg.','Angle',.x)) %>%
+  rename_with(.fn = ~gsub('\\.','',.x)) %>%
+  rename('ID'='ObjId','DryYield'='YldMassDry_tHa','Lon'='Longitude','Lat'='Latitude','Pass'='PassNum','Speed'='Speed__m') %>%
+  st_as_sf(coords=c('Lon','Lat')) %>% #Add spatial feature info
[…]

### 2021-07-30 Manuscript/manuscript.Rmd
Updated model to filter extreme data

@@
 
-- Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m)
+- Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted to account for overlap with adjacent swaths)
   - We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
@@
 - In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest
-  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (e.g. Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field
+  - Spatial or temporal variation can be modelled using a Gaussian Process Model (e.g. Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field
   - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but using a reduced number of dimensions (as well as computation time)
@@
   - All models were fit in _R_ using the _mgcv_ library [version `r packageVersion('mgcv')`, @wood2017], and figures were created with _ggplot2_ and _ggpubr_ [versions `r packageVersion('ggplot2')`, @wickham2016; and `r packageVersion('ggpubr')`, @kassambara2020].
+
+- Despite the use of spatial and temporal smoothers, the residuals in most of the models displayed a large amount of spatial and temporal autocorrelation
+   - This is extremely common in precision yield data, but
 
@@
 # Results
+
+## Data
+
+## Boundary effects
 

### 2021-07-30 filtering.R

@@ -1,20 +1,32 @@
 #Seems to be a consensus that filter/cleaning is necessary for yield data. Trying to see how much of a difference this makes for our models
 
+# Load everything -----------------------------------------------
+
 library(tidyverse)
 theme_set(theme_classic())
 library(sf)
 library(mgcv)
 library(beepr)
 library(ggpubr)
-setwd("~/Documents/yield-analysis-2021/gam_TMB")
-source('../helperFunctions.R')
+setwd("~/Documents/yield-analysis-2021")
+source('./helperFunctions.R')
+
+datSource <- read.csv('./Data/datSource.csv') %>% #Read previous datasource file
+  mutate(boundaryComplete=file.exists(boundaryPath)) %>% #Has boundary been made already?
+  mutate(modelComplete1=file.exists(modelPath1)) %>% #Has model1 already been run?
+  mutate(modelComplete2=file.exists(modelPath2)) %>% #Has model2 already been run?
+  mutate(modelComplete0=file.exists(modelPath0)) #Has model0 already been run?
 
 nSubSamp <- 50000 #Number of points to start with
-csvPath <- "/media/rsamuel/Storage/geoData/Rasters/yieldData/csv files/Alvin_French C41 2020.csv"
-boundaryPath <- "/home/rsamuel/Documents/yield-analysis-2021/Figures/FieldBoundaries/Alvin_French C41 2020 boundary.shp"
-boundaryPath2 <- "/home/rsamuel/Documents/yield-analysis-2021/Figures/FieldBoundarySegments/Alvin_French C41 2020 boundary.shp"
 
-#Also try: Gibbons Bill Visser 2017 (bad residual structure)
+# fieldName <- 'Alvin_French C41 2020'
+fieldName <- 'Gibbons Bill Visser 2017'
+# fieldName <- 'Trent_Clark E 06 2018'
+
+rownum <- which(datSource$filename==fieldName)
+csvPath <- datSource$dataPath[rownum] #Get file paths
+boundaryPath <- datSource$boundaryPath[rownum]
+boundaryPath2 <- datSource$boundaryPath2[rownum]
 
[…]

### 2021-07-30 helperFunctions.R

@@ -264,58 +264,9 @@ getPreds <- function(path,l=c(30,100,500),margInt=c(FALSE,FALSE,FALSE),samp=FALS
 # (temp <- getPreds(paste0('./Figures/ModelCheck2/',datSource$filename[2],' modList.Rdata'),margInt=c(FALSE,TRUE,TRUE),samp=FALSE)) #Test with second type of model
 
 
-#Shorter version that only gets estimates of log(pArea) relationship and turns them into speed measurements
-# Combine ground speed is talked about a lot, so this is more appropriate for growers/agronomists
-
-getPredsSpeed <- function(path,l=30,margInt=FALSE,samp=FALSE, minSpeed = NA, maxSpeed = NA, width = NA, speed2Distance = NA){
-  
-  # #Debugging
-  # l <- 30
-  # # path <- paste0('./Figures/ModelCheck1/',datSource$filename[2],' modList.Rdata') #Model type 1
-  # path <- paste0('./Figures/ModelCheck2/',datSource$filename[2],' modList.Rdata') #Model type 2
-  # minSpeed <- 0.1
-  # maxSpeed <- 15 
-  # margInt <- FALSE
-  # samp <- FALSE
-  # width <- 10.21
-  # speed2Distance <- 1/18.043
-  
-  if(any(is.na(c(minSpeed,maxSpeed,width,speed2Distance)))) stop('Missing term')
-  load(path) #Load data
-  
-  #Converts speed to area term for use with original coefficients (holding width at max value)
-  logSpeed <- seq(log(minSpeed),log(maxSpeed),length.out=l)
-  speed <- exp(logSpeed)
-  WAS <- width*speed2Distance*speed #Width * alpha * speed = proxy for pArea
-  logWAS <- log(WAS) #proxy for log(pArea)
-  
-  meanVars <- which(grepl('(^\\(Intercept\\)$|^log\\(pArea\\)$)',names(modList$coefs)))
-  sdVars <- which(grepl('(^\\(Intercept\\)\\.1$|^log\\(pArea\\)\\.1$)',names(modList$coefs)))
-  
-  #Coefficients (intercept and slope) for mean and logSD relationship
-  if(samp){ #Sample from posterior
[…]

### 2021-07-30 runModels.R

@@ -281,8 +281,9 @@ ggsave(paste0('./Figures/ModelSummary1a.png'),p,height=6,width=12,dpi=300)
 
 # Run second set of models - boundary type included ---------------------------
 
+# #Test with single model
 # a <- Sys.time()
-# runModII(295,dS=datSource) 
+# runModII(295,dS=datSource)
 # Sys.time()-a
 # # beep(1)
 # debugonce(runModII)
@@ -395,15 +396,15 @@ ggsave(paste0('./Figures/ModelSummary2a.png'),p,height=6,width=12,dpi=300)
 
 # Function to sample from posterior of each prediction, amalgamate, fit meta-model, return results of meta-model
 # My way of getting around the problem of random effects
-# a = does nothing (argument for parLapply), useThese = rows of ds to be used, ds = datSource csv, nX = number of samples along range of X
-samplePreds <- function(a=NULL,useThese=NULL,ds=datSource,nX=c(100,100,100)){ 
-  if(is.null(useThese)){
+# a = does nothing (argument for parLapply), useRows = rows of ds to be used, ds = datSource csv, nX = number of samples along range of X
+samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(100,100,100)){ 
+  if(is.null(useRows)){
     warning('Rows not specified. Using entire dataset.')
-    useThese <- 1:nrow(datSource)
+    useRows <- 1:nrow(datSource)
   } 
   require(tidyverse); require(mgcv)
   source('helperFunctions.R')
-  getFiles <- ds %>% slice(useThese) %>% filter(use,modelComplete2) #Completed models
+  getFiles <- ds %>% slice(useRows) %>% filter(use,modelComplete2) #Completed models
   
   #Sample from posterior, get new lines for each field
   allSmooths <- lapply(getFiles$modelPath2,getPreds,margInt=c(FALSE,TRUE,TRUE),samp=TRUE) %>% set_names(getFiles$filename)
@@ -586,7 +587,7 @@ ggsave(paste0('./Figures/ModelSummary3a.png'),p,height=6,width=12,dpi=300)
 # cluster <- makeCluster(10)
[…]

### 2021-08-24 Manuscript/manuscript.Rmd
Worked on manuscript
- Also updated figures: marginalize across intercepts for yield:pArea figures

--- 2021-07-10 Manuscript/manuscript.Rmd
@@
 ---
-title: "Edge effects in Alberta"
+title: "Living on the edge: crop boundary effects on canola and wheat yields in Alberta"
 author:
@@
 library(tidyverse)
+library(knitr)
 
@@
 
-- Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth
-  - This has allows increases in human population and increased (global) stability in production
-  - However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms
-  - Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios
-  - Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems
+Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth
+This has allowed increases in human population and increased (global) stability in production
+However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms
+Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios
+Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems
 
-- SNL in and around crops is important for both agricultural production and conservation
-  - They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control
-  - They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed
-  - Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
-  - In particular, North American agroecosystems have larger fields, and different varieties and agronomic practices, all of which could negate effects of SNL
+SNL in and around crops is important for both agricultural production and conservation
+They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control (@albrecht2020; @gardner2021)
+They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (@kort1988)
+Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
+In particular, North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL
 
 <!-- - SNL has conflicting effects on crop yield because of ecosystem service provision as well as negative edge effects -->
-- SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate
-  - Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds
-  - At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges
-  - For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance
-  - Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects 
-  - This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
+SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate
+Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds
+At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (@kowalchuk1995)
+For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance
+Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (e.g. @vanVooren2017)
+This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
 
@@
 
-- Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality (but see wheat study)
-  - For this reason, large-scale precision yield data holds enormous promise for agronomy, as it allows 
-  - However, its use is limited for several reasons: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) unfamiliarity with spatial statistics
-  - Ecosystem services can influence both the mean and variability of yield in agroecosystems
-  - Typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable
-  - Size examples: wheat study from the UK
-  - There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield
+Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality (but see wheat study)
+Meta-analyses have been the standard solution to this problem, but precision yield data holds enormous promise.
+However, its use is limited for several reasons: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) unfamiliarity with spatial statistics.
+Ecosystem services can influence both the mean and variability of yield in agroecosystems (@redhead2020).
+Typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable.
+There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
 
-- In this paper we ask:
+In this paper we ask:
   1. How does crop yield change with distance from the edge of field?
@@
 
-- Precision yield data were collected directly from farmers across Alberta
-  - Farmers were solicited for yield data through local agronomists, and we received `r nrow(datSource)` field-years of data from `r length(unique(datSource$grower))` growers across a total of `r length(unique(datSource$year))` years (`r min(unique(datSource$year))`-`r max(unique(datSource$year))`)
-  - We converted data to a standard csv format using Ag Leader SMS
-  - `r round(100*sum(grepl('(Wheat|Canola)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_) or canola (_Brassica napus_), two of the most common crops in rotation in Alberta
-  - The remaining crop types were poorly replicated in our sample, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`) or canola (`r sum(datSourceFinal$crop=='Canola')`)
-  - Individual fields contained between `r min(yearsPerField)` to `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`)
-  - Containing a total of `r round(sum(datSource$npoints)/1e6,1)` million data points
+Precision yield data were collected directly from farmers across Alberta.
+Farmers were solicited for yield data through local agronomists, and we received `r nrow(datSource)` field-years of data from `r length(unique(datSource$grower))` growers across a total of `r length(unique(datSource$year))` years (`r min(unique(datSource$year))`-`r max(unique(datSource$year))`).
+We converted raw data to a standard csv format using Ag Leader SMS.
+`r round(100*sum(grepl('(Wheat|Canola)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_) or canola (_Brassica napus_), two of the most common crops in rotation in Alberta.
+The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`) or canola (`r sum(datSourceFinal$crop=='Canola')`).
+Individual fields contained between `r min(yearsPerField)` to `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`), containing a total of `r round(sum(datSource$npoints)/1e6,1)` million data points.
 
-- Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m)
-  - We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
-  - Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
-  - Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
-  - We used dry yield (tonnes of seed/hectare after accounting for crop moisture) as our measure of crop yield
-  - Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time
+Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted to account for overlap with adjacent swaths).
+We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
+Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
+Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
+We used dry yield (tonnes of seed/hectare minus recorded crop moisture) as our measure of crop yield
+Precision yield data can be highly variable and is prone to extreme outliers (both positive and negative), especially when combine ground speeds are low and at the beginning of rows
+Therefore, we filtered data that lay outside the 95th percentiles of yield within each field, values less than or equal to zero, and values outside of the 95th percentiles of ground speed.
+Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time.
 
-- Field boundaries were automatically digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC 
-  - Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g. flooding during some years)
-  - Additionally, seminatural features often change yearly
-    - Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years
-    - Grass boundaries can change if fields are used for as haying or pasture during crop rotation
-  - This makes accurate and consistent classification of field boundaries difficult
-  - We defined the following general categories for field boundaries:
+Field boundaries were automatically digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC. 
+Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g. flooding during some years)
+Additionally, seminatural features often change yearly. 
+Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
+This makes accurate and consistent classification of field boundaries difficult.
+We defined the following general categories for field boundaries:
     1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide) 
@@
 
-- At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor
-  - Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices 
-    <!-- - such as strip farming -->
-  - Ground speed is extremely important to yield monitor accuracy [@arslan2002], with low ground speed registering higher yields 
-  - While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time  within fields
-  - This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
-
-  - To model this, we fit the following model to each field-year of data:
+Yield data from each field contains both random and systematic errors that must be modeled in order to reveal underlying changes in yield.
+We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the data.
+Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices.      <!-- - such as strip farming -->
+Ground speed is extremely important to yield monitor accuracy [@arslan2002], with low ground speed registering higher yields.
+Sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time.
+This may be caused by sensors accumulating debris during harvest (especially in canola, pers. comm. Trent Clark), leading to changes in accuracy and bias over time.
+To model this, we fit the following model to each field-year of data:
 
@@
   \begin{split}
-  sqrt(yield) \sim & Normal (\mu, \sigma)\\
-  \mu = & Intercept + log(Polygon Size) + f(Distance from Edge_i, b=12)_i + \\
-   & f(Easting, Northing, b=60) + f(Sequence, b=60) \\
-  log(\sigma) = & Intercept + log(Polygon Size) + f(Distance from Edge, b=12) + \\
-   & f(Easting, Northing, b=60) + f(Sequence, b=60)
+  \sqrt{yield} \sim & Normal (\mu, \sigma)\\
+  \mu = & Intercept + log(\text{Polygon Size}) + f(\text{Edge Distance}, b=12)_i + \\
+   & f(\text{Easting}, \text{Northing}, b=60) + f(\text{Sequence}, b=60) \\
+  log(\sigma) =  & Intercept + log(\text{Polygon Size}) + f(\text{Edge Distance}, b=12)_i + \\
+   & f(\text{Easting}, \text{Northing}, b=60) + f(\text{Sequence}, b=60) \\
   \end{split}
@@
   1. Polygon Size = distance traveled $\times$ width of header bar ($m^2$)
-  2. Distance from Edge = distance from field edge type _i_ (m)
+  2. Edge Distance = distance from field edge type _i_ (m)
   3. Easting, Northing = distance from centre of field (m)
   4. Sequence = order of harvest within field (1--N points)
-  5. f(x,b) = penalized thin-plate regression spline, where x is the predictor and b is the number of basis dimensions
+  5. f(x,b) = penalized thin-plate regression spline, where _x_ is the predictor and _b_ is the number of basis dimensions
 
-- In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest
-  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (Kriging) or approximations such as Stochastic Partial Differential Equations (e.g.
-INLA), but this was computationally infeasible with 50,000 data points per field
-  - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with reduced dimensionality and computation time
-  - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
-  - The relationship between polygon size (i.e. ground speed) and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions
-  - All models were fit in _R_ using the _mgcv_ library [version `r packageVersion('mgcv')`, @wood2017], and figures were created with _ggplot2_ and _ggpubr_ [versions `r packageVersion('ggplot2')`, @wickham2016; and `r packageVersion('ggpubr')`, @kassambara2020].
+In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest.
+Spatial or temporal variation can be modelled using a Gaussian Process Model (e.g. Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
+Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but using a reduced number of dimensions and computation time.
+The number of basis dimensions was checked with the _gam.check_ function from _mgcv_.
+The relationship between polygon size (i.e. ground speed) and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions.
+All models were fit in _R_ using the _mgcv_ library [version `r packageVersion('mgcv')`, @wood2017], and figures were created with _ggplot2_ and _ggpubr_ [versions `r packageVersion('ggplot2')`, @wickham2016; and `r packageVersion('ggpubr')`, @kassambara2020].
 
-- To consider results from all field-level models, we fit models independent of each other, and "overall" smoothers were taken as averages of the field-level smoothers
-  - However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models
+Despite the use of spatial and temporal smoothers, the residuals in most of the models displayed a large amount of spatial and temporal autocorrelation
+   <!-- - This is extremely common in precision yield data, but  -->
+
+To consider results from all field-level models together, we fit models independent of each other and fit an "overall" smoother as averages of the field-level smoothers.
+However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models:
   1. Extract single posterior sample (_rnorm_ in R) of smoother parameters from each field-level model using coefficient estimates and standard error
-  2. Use posterior sample to create new simulated smoother from each field
-  3. Fit new model of simulated smoothers from all fields, and save this "meta-smoother"
+  2. Use posterior sample to create a new smoother from each field
+  3. Fit new model of new smoothers from all fields, and save this "meta-smoother"
   4. Repeat 1000 times, and calculate coverage intervals (5-95% percentiles) on saved meta-smoothers
-  - This gives coverage intervals (CIs) for the "average" smoother while accounting for field-level variability
+This gives coverage intervals (CIs) for the "average" smoother while accounting for field-level variability.
 
@@
 
-```{r , echo=FALSE, fig.cap='Fig1', out.width='100%'}  
-  include_graphics('../figures/frExample.png',dpi=NA)
+## Canola
+
+```{r , echo=FALSE, fig.cap='Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
+Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
+"patchiness").', out.width='100%'}
+  include_graphics('../Figures/ModelSummary3a_canola.png',dpi=NA)
+```
+
+## Wheat
+
+```{r , echo=FALSE, fig.cap='Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
+Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
+"patchiness").', out.width='100%'}
+  include_graphics('../Figures/ModelSummary3a_wheat.png',dpi=NA)
 ```
@@
 
-# Authors' contributions
+# Author Contributions
 
-Author's contribution
+SVJR and PG conceived of the project.
+SVJR collected data, conducted analysis, and wrote the manuscript.
 
@@
 
+We thank Trent Clark, Dean Hubbard, Alvin French, and Kristina Polziehn for providing yield data for us, and providing insight on their yield data. 
+We also thank Autumn Barnes and Keith Gabert of the Canola Council of Canada, and Laurel Thompson and JP Pettyjohn of Lakeland College for connecting us with farmers and agronomists. 
 Funding for this research was provided by Ducks Unlimited Canada's Institute for Wetland and Waterfowl Research, the Alberta Canola Producers Commission, Manitoba Canola Growers Association, SaskCanola, the Alberta Biodiversity Monitoring Institute, and the Alberta Conservation Association.

### 2021-08-24 Manuscript/manuscript.tex
Worked on manuscript
- Also updated figures: marginalize across intercepts for yield:pArea figures

--- 2021-04-12 Manuscript/manuscript.tex
@@
             pdfauthor={},
-            pdftitle={Edge effects and optimal field sizes in Alberta},
+            pdftitle={Living on the edge: crop boundary effects on canola and wheat yields in Alberta},
             colorlinks=false,
@@
 
-  \title{Edge effects and optimal field sizes in Alberta}
+  \title{Living on the edge: crop boundary effects on canola and wheat yields in Alberta}
     \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Preserving SNL around fields is important for agriculture and conservation
+Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth
+This has allowed increases in human population and increased (global) stability in production
+However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms
+Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios
+Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems
 
-  \begin{itemize}
-  \tightlist
-  \item
-    Ecosystem services
-  \item
-    Act as reservoirs for beneficial insects
-  \item
-    Microclimate zones
-  \item
-    Poorly studied in North America
-  \item
-    Studies tend to be limited in scope (e.g.~only certain organisms/crops) and applicability
-  \end{itemize}
-\item
-  Ecosystem services can influence both the mean and variability of yield in agroecosystems
+SNL in and around crops is important for both agricultural production and conservation
+They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control (Albrecht \emph{et al.} (2020); Gardner \emph{et al.} (2021))
+They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (Kort (1988))
+Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
+In particular, North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL
 
-  \begin{itemize}
-  \tightlist
-  \item
-    Typically only averages (means) are considered, but higher stability (lower variance) in yield is also valuable
-  \end{itemize}
-\item
-  Field size and field boundaries are directly related to the conservation of SNL in agroecosystems
+SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate
+Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds
+At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (Kowalchuk \& Jong (1995))
+For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance
+Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (e.g. Van Vooren \emph{et al.} (2017))
+This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
 
-  \begin{itemize}
-  \tightlist
-  \item
-    Size examples: wheat study from the UK
-  \item
-    Boundary examples: flower strip studies, hedgerows
-  \item
-    In North America, large financial incentives to make fields large and homogeneous, especially with large harvesting and planting equipment
-  \item
-    Crop edge effects cause low yields at the margins of fields because of late emergence, poor microclimate, and competition with weeds
-  \item
-    Ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
-  \item
-    Therefore, there should be a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
-  \end{itemize}
-\item
-  Precision yield data holds enormous promise for agronomy
+Studies of SNL effects on crop yield also suffer from limited scope (e.g.~few crop types) and small sample sizes, limiting inference and reducing generality (but see wheat study)
+Meta-analyses have been the standard solution to this problem, but precision yield data holds enormous promise.
+However, its use is limited for several reasons: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) unfamiliarity with spatial statistics.
+Ecosystem services can influence both the mean and variability of yield in agroecosystems (Redhead \emph{et al.} (2020)).
+Typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable.
+There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
 
-  \begin{itemize}
-  \tightlist
-  \item
-    Limited because of:
-  \item
-    Lack of standardized formats
-  \item
-    Yearly calibration data
-  \item
-    Clear statistical protocols
-  \end{itemize}
-\end{itemize}
+In this paper we ask:
+1. How does crop yield change with distance from the edge of field?
+2. Does this depend on type of field edge?
+3. Is there an intermediate distance where yield is maximized or variance is minimized?
 
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Precision yield data were collected directly from farmers across Alberta
+Precision yield data were collected directly from farmers across Alberta.
+Farmers were solicited for yield data through local agronomists, and we received 298 field-years of data from 5 growers across a total of 7 years (2014-2020).
+We converted raw data to a standard csv format using Ag Leader SMS.
+72\% of the crop types where either wheat (\emph{Triticum aestivum}) or canola (\emph{Brassica napus}), two of the most common crops in rotation in Alberta.
+The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94) or canola (119).
+Individual fields contained between 1 to 5 years of data (mean: 2.7), containing a total of 18.4 million data points.
+
+Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted to account for overlap with adjacent swaths).
+We extracted the size of each polygon (m\(^2\)), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
+Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
+Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
+We used dry yield (tonnes of seed/hectare minus recorded crop moisture) as our measure of crop yield
+Precision yield data can be highly variable and is prone to extreme outliers (both positive and negative), especially when combine ground speeds are low and at the beginning of rows
+Therefore, we filtered data that lay outside the 95th percentiles of yield within each field, values less than or equal to zero, and values outside of the 95th percentiles of ground speed.
+Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time.
+
+Field boundaries were automatically digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC.
+Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g.~flooding during some years)
+Additionally, seminatural features often change yearly.
+Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
+This makes accurate and consistent classification of field boundaries difficult.
+We defined the following general categories for field boundaries:
+1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
+2. Wetland: permanent wetland; borders are largely unchanged from year-to-year
+3. Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
+4. Other crop: annual crop or pasture with little or no visible boundary between planted areas
+5. Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
+6. Grassland: permanent seminatural grassland or pasture (not in rotation)
+
+\begin{verbatim}
+<!-- STANDARD -->
+<!-- WETLAND -->
+<!-- SHELTERBELT -->
+<!-- OTHERCROP -->
+<!-- BARE -->
+<!-- GRASSLAND -->
+\end{verbatim}
+
+\hypertarget{analysis}{%
+\subsection{Analysis}\label{analysis}}
+
+Yield data from each field contains both random and systematic errors that must be modeled in order to reveal underlying changes in yield.
+We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the data.
+Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices.
+Ground speed is extremely important to yield monitor accuracy (Arslan \& Colvin 2002), with low ground speed registering higher yields.
+Sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time.
+This may be caused by sensors accumulating debris during harvest (especially in canola, pers. comm. Trent Clark), leading to changes in accuracy and bias over time.
+To model this, we fit the following model to each field-year of data:
+
+\begin{equation}
+  \begin{split}
+  \sqrt{yield} \sim & Normal (\mu, \sigma)\\
+  \mu = & Intercept + log(\text{Polygon Size}) + f(\text{Edge Distance}, b=12)_i + \\
+   & f(\text{Easting}, \text{Northing}, b=60) + f(\text{Sequence}, b=60) \\
+  log(\sigma) =  & Intercept + log(\text{Polygon Size}) + f(\text{Edge Distance}, b=12)_i + \\
+   & f(\text{Easting}, \text{Northing}, b=60) + f(\text{Sequence}, b=60) \\
+  \end{split}
+  \end{equation}
 
@@
 \item
-    Farmers were solicited for yield data through local agronomists, and we received data from a total of X growers across a total of X years
-  \item
-    Most fields represented only a single year of data, but we did sometimes receive multiple years of data from the same field
-  \item
-    File formats vary depending on the brand of combine, so we converted data to a standard format (csv) using Ag Leader SMS
-  \item
-    In total, we analyzed yield data from X field-years of data, containing a total of X million data points
+  where:
 \end{itemize}
-\item
-  Yield data was collected in discrete rectangles of the same length as the data interval (distance = speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m)
-
-  \begin{itemize}
-  \tightlist
-  \item
-    We extracted the size of each polygon (m\(^2\)), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
-  \item
-    Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location
-  \item
-    Seeding and application rates were constant with each field, so we did not consider inputs in our analysis
-  \end{itemize}
-\item
-  Field boundaries were automatically digitized, then manually checked using satellite imagery and classified land cover data
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Crop boundaries are flexible, and often change from year to year depending on planting and emergence conditions (e.g.~flooding during some years)
-  \item
-    Additionally, seminatural features often change from year to year
-
-    \begin{itemize}
-    \tightlist
-    \item
-      Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years
-    \item
-      Grass boundaries can change if fields are used for as haying or pasture during crop rotation
-    \end{itemize}
-  \item
-    This makes accurate and consistent classification of field boundaries very difficult
-  \item
-    Because of this, we used the following general categories for field boundaries:
 
@@
 \item
-      Standard: thin (\(<10\) m wide) grassy field boundary, often grassy road right-of-ways
+  Polygon Size = distance traveled \(\times\) width of header bar (\(m^2\))
 \item
-      Wetland: permanent wetland, whose borders are largely unchanged from year-to-year
+  Edge Distance = distance from field edge type \emph{i} (m)
 \item
-      Forest/shrub: permanent windbreaks or remnant forests
+  Easting, Northing = distance from centre of field (m)
 \item
-      Grass: larger grassy area (pasture or permanent seminatural grassland)
+  Sequence = order of harvest within field (1--N points)
 \item
-      Other crop: different crop with little or no visible boundary between planted areas
-    \item
-      Bare: unplanted (fallow), flooded area, or unpaved roads
+  f(x,b) = penalized thin-plate regression spline, where \emph{x} is the predictor and \emph{b} is the number of basis dimensions
 \end{enumerate}
-  \end{itemize}
-\end{itemize}
 
-\hypertarget{analysis}{%
-\subsection{Analysis}\label{analysis}}
+In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest.
+Spatial or temporal variation can be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
+Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but using a reduced number of dimensions and computation time.
+The number of basis dimensions was checked with the \emph{gam.check} function from \emph{mgcv}.
+The relationship between polygon size (i.e.~ground speed) and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions.
+All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).
 
-\begin{itemize}
-\tightlist
-\item
-  At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor
+Despite the use of spatial and temporal smoothers, the residuals in most of the models displayed a large amount of spatial and temporal autocorrelation
 
-  \begin{itemize}
-  \tightlist
-  \item
-    Crop yield varies within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices such as strip farming
-  \item
-    While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields), this does not address sensor drift over time that occurs within fields
-  \item
-    The yield monitors may record lower yields as sensors accumulate debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
-  \item
-    Additionally, ground speed is known to be extremely important to yield monitor accuracy (Arslan \& Colvin 2002)
-  \item
-    To address this, we fit the following model:
-  \end{itemize}
-
-  \begin{equation}
-  ln(yield) 
-  \end{equation}
-\end{itemize}
+To consider results from all field-level models together, we fit models independent of each other and fit an ``overall'' smoother as averages of the field-level smoothers.
+However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models:
+1. Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard error
+2. Use posterior sample to create a new smoother from each field
+3. Fit new model of new smoothers from all fields, and save this ``meta-smoother''
+4. Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers
+This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability.
 
@@
 
-Results here
+\hypertarget{canola}{%
+\subsection{Canola}\label{canola}}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_canola} \caption{Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
+Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
+"patchiness").}\label{fig:unnamed-chunk-1}
+\end{figure}
+
+\hypertarget{wheat}{%
+\subsection{Wheat}\label{wheat}}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_wheat} \caption{Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
+Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
+"patchiness").}\label{fig:unnamed-chunk-2}
+\end{figure}
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-albrecht2020}{}%
+Albrecht, M., Kleijn, D., Williams, N.M., Tschumi, M., Blaauw, B.R., Bommarco, R., \emph{et al.} (2020).
+The effectiveness of flower strips and hedgerows on pest control, pollination services and crop yield: A quantitative synthesis.
+\emph{Ecology Letters}.
+
 \leavevmode\hypertarget{ref-arslan2002}{}%
 Arslan, S. \& Colvin, T.S. (2002). An evaluation of the response of yield monitors and combines to varying yields. \emph{Precision Agriculture}, 3, 107--122.
+
+\leavevmode\hypertarget{ref-gardner2021}{}%
+Gardner, E., Breeze, T.D., Clough, Y., Smith, H.G., Baldock, K.C.R., Campbell, A., \emph{et al.} (2021).
+Field boundary features can stabilise bee populations and the pollination of mass-flowering crops in rotational systems.
+\emph{Journal of Applied Ecology}, In press.
+
+\leavevmode\hypertarget{ref-kassambara2020}{}%
+Kassambara, A. (2020). \emph{{ggpubr}: 'ggplot2' based publication ready plots}.
+
+\leavevmode\hypertarget{ref-kort1988}{}%
+Kort, J. (1988). 9. Benefits of windbreaks to field and forage crops. \emph{Agriculture, Ecosystems {\&} Environment}, 22-23, 165--190.
+
+\leavevmode\hypertarget{ref-kowalchuk1995}{}%
+Kowalchuk, T.E. \& Jong, E. de. (1995). Shelterbelts and their effect on crop yield. \emph{Canadian Journal of Soil Science}, 75, 543--550.
+
+\leavevmode\hypertarget{ref-redhead2020}{}%
+Redhead, J.W., Oliver, T.H., Woodcock, B.A. \& Pywell, R.F. (2020). The influence of landscape composition and configuration on crop yield resilience. \emph{Journal of Applied Ecology}.
+
+\leavevmode\hypertarget{ref-vanVooren2017}{}%
+Van Vooren, L., Bert, R., Steven, B., Pieter, D.F., Victoria, N., Paul, P., \emph{et al.} (2017).
+Ecosystem service delivery of agri-environment measures: A synthesis for hedgerows and grass strips on arable land.
+\emph{Agriculture, Ecosystems {\&} Environment}, 244, 32--51.
+
+\leavevmode\hypertarget{ref-wickham2016}{}%
+Wickham, H. (2016). \emph{{ggplot2}: Elegant graphics for data analysis}. Springer-Verlag New York.
+
+\leavevmode\hypertarget{ref-wood2017}{}%
+Wood, S.N. (2017). \emph{Generalized additive models: An introduction with {R}}. CRC press.
 

### 2021-08-24 Manuscript/supplemental.Rmd
Worked on manuscript
- Also updated figures: marginalize across intercepts for yield:pArea figures

@@
 
-Supplemental materials here
+```{r , echo=FALSE, fig.cap='Effect of polygon area and harvest sequence on canola yield. Polygon area is equivalent to ground ', out.width='100%'}  
+  include_graphics('../Figures/ModelSummary3b_canola.png',dpi=NA)
+```
+
+```{r , echo=FALSE, fig.cap='Temp', out.width='100%'}  
+  include_graphics('../Figures/ModelSummary3b_wheat.png',dpi=NA)
+```

### 2021-08-24 runModels.R

@@ -398,7 +398,7 @@ ggsave(paste0('./Figures/ModelSummary2a.png'),p,height=6,width=12,dpi=300)
 # My way of getting around the problem of random effects
 # a = does nothing (mandatory argument for parLapply), useRows = rows of ds to be used, ds = datSource csv, nX = number of samples along range of X
 # rCutoff = restrict point number (some fields have hundreds of thousands of points, but all data uses approximately the same size yield rectangles, so cuts off meta-smoother at "about" the maximum field size - i.e. the most common number of points per field)
-samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutoff=50000){ 
+samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutoff=50000,margInt=c(FALSE,TRUE,TRUE)){ 
   if(is.null(useRows)){
     warning('Rows not specified. Using entire dataset.')
     useRows <- 1:nrow(datSource)
@@ -408,7 +408,7 @@ samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutof
   getFiles <- ds %>% slice(useRows) %>% filter(use,modelComplete2) #Completed models
   
   #Sample from posterior, get new lines for each field
-  allSmooths <- lapply(getFiles$modelPath2,getPreds,l=nX,margInt=c(FALSE,TRUE,TRUE),samp=TRUE) %>% set_names(getFiles$filename)
+  allSmooths <- lapply(getFiles$modelPath2,getPreds,l=nX,margInt=margInt,samp=TRUE) %>% set_names(getFiles$filename)
   
   #log(Area) meta-models
   pAreaDf <- lapply(allSmooths,function(x) x$pAreaDat) %>% bind_rows(.id = 'field')
@@ -583,35 +583,32 @@ samplePreds() #Test
 
 # Get crop-specific smoother info from second set of models 
 
-# # Add to current samples - canola
+## Add to current samples - canola
 # isCanola <- which(datSource$crop=='Canola') #Canola crops only
-# Nsamp <- 985 #Number of samples
[…]

### 2021-08-27 Manuscript/manuscript.Rmd
Worked on marginal plots
- Because of intercept/slope correlation in pArea term, need to fit lmer models
- Entire model is on sqrt scale, so need to consider this in marginal plots

@@
     text: "Corresponding Author"
-date: "Spring 2021"
+date: "Fall 2021"
 abstract: "`r paste(readLines('abstract.Rmd'), collapse = ' ')`"
@@
 keywords:
-  - beetles;
-  - spiders;
-  - harvestmen;
+  - precision yield;
+  - spillover;
+  - landscape;
 
@@
 library(knitr)
+library(kableExtra)
 
@@
 yearsPerField <- datSourceFinal %>% count(field) %>% pull(n) #How many years of data at each unique location?
+
+#Boundary lengths
+bLengths <- read.csv('../Data/boundaryLengths.csv')
+bNames <- c('Bare','Grassland','Other Crop','Shelterbelt','Standard','Wetland')
+
+bLengthSummary <- bLengths %>% select(filename,contains('prop_')) %>% pivot_longer(contains('prop_')) %>% 
+  mutate(name=gsub('prop_','',name)) %>% na.omit() %>% 
+  group_by(name) %>% summarize(mean=mean(value),sd=sd(value),Nfields=sum(value>0)) %>% 
+  mutate(name=bNames)
+
+#Model smoother summaries
+use <- with(datSource,use&modelComplete2&crop=='Canola')
+paths <- paste0('.',gsub('modList.Rdata','results.txt',datSource$modelPath2))
+
+canolaMods <- lapply(paths[use],function(x) getModelInfo(x)$smooths) %>% 
+  set_names(nm=datSource$filename[use]) %>% 
+  bind_rows(.id='field') %>% 
+  mutate(pval=as.numeric(pval)) %>% 
+  group_by(Smoother) %>% 
+  summarize(meanP=mean(pval),maxP=max(pval),propSig=sum(pval<0.05)/n(),meanEDF=mean(edf),meanRef=mean(Ref.df))
+
+use <- with(datSource,use&modelComplete2&crop=='Wheat')
+wheatMods <- lapply(paths[use],function(x) getModelInfo(x)$smooths) %>% 
+  set_names(nm=datSource$filename[use]) %>% 
+  bind_rows(.id='field') %>% 
+  mutate(pval=as.numeric(pval)) %>% 
+  group_by(Smoother) %>% 
+  summarize(meanP=mean(pval),maxP=max(pval),propSig=sum(pval<0.05)/n(),meanEDF=mean(edf),meanRef=mean(Ref.df))
+rm(use,paths)
+
+# #Average model parameters (mean/sd slopes with pArea)
+# lapply(paths[use],function(x) getModelInfo(x)$params) %>% 
+#   set_names(nm=datSource$filename[use]) %>% 
+#   bind_rows(.id='field') %>% group_by(Parameter) %>% 
+#   summarize(meanEst=mean(Estimate),Z=mean(Estimate/SE))
 
@@
 Field boundaries were automatically digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC. 
-Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g. flooding during some years)
-Additionally, seminatural features often change yearly. 
+Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g. flooding during some years).
 Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
@@
 
-In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest.
+In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in harvested area, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest.
 Spatial or temporal variation can be modelled using a Gaussian Process Model (e.g. Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
@@
 The number of basis dimensions was checked with the _gam.check_ function from _mgcv_.
-The relationship between polygon size (i.e. ground speed) and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions.
+The relationship between polygon size and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions.
+Because polygon size is not intuitive, we converted it into ground speed by regressing speed on distance traveled per polygon, then using this to back-calculate the ground speed for a given polygon size (given a fixed width of header).
 All models were fit in _R_ using the _mgcv_ library [version `r packageVersion('mgcv')`, @wood2017], and figures were created with _ggplot2_ and _ggpubr_ [versions `r packageVersion('ggplot2')`, @wickham2016; and `r packageVersion('ggpubr')`, @kassambara2020].
 
-Despite the use of spatial and temporal smoothers, the residuals in most of the models displayed a large amount of spatial and temporal autocorrelation
+Despite the use of spatial and temporal smoothers, the residuals in most of the models displayed a large amount of spatial and temporal autocorrelation.
+_Not really sure how we can get around this.
+In practice, this means that SEs for all coefficients are too small (i.e.
+we gathered 1000 points, but because they're spatially related, it's only equivalent to 100 points)._
 <!-- - This is extremely common in precision yield data, but  -->
@@
   4. Repeat 1000 times, and calculate coverage intervals (5-95% percentiles) on saved meta-smoothers
-This gives coverage intervals (CIs) for the "average" smoother while accounting for field-level variability.
+This gives coverage intervals (CIs) for the "average" smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
+Spatial variability was not averaged between fields because differences in field geometry make it impossible to compare fields.
 
@@
 
+## Field boundaries
+
+"Standard" (grassy border) was the most common boundary type, followed by "Shelterbelt" and "Other Crop" (Table \@ref(tab:boundaryTypes)).
+
+```{r boundaryTypes, echo=FALSE, out.width='100%'}  
+bLengthSummary %>% 
+  kable(.,format='latex',caption='Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.',col.names = c('Type','Mean','SD','Number of Fields'),digits = c(1,4,4,1),escape = FALSE)
+```
+
 ## Canola
 
-```{r , echo=FALSE, fig.cap='Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
+There was a strong negative effect combine ground speed on both the average and variability of canola yield.
+At the field level, both yield average and yield variability (SD of yield) of yield were extremely spatially (average p-values: `r canolaMods$meanP[canolaMods$Smoother=='s(E,N)']`, `r canolaMods$meanP[canolaMods$Smoother=='s.1(E,N)']`) and temporally dependent (`r canolaMods$meanP[canolaMods$Smoother=='s(r)']`, `r canolaMods$meanP[canolaMods$Smoother=='s.1(r)']`), meaning that average and variablity changed with location in the field, as well as sequentially during the harvest.
+However, there was no difference in average or variability with harvest sequence when all fields were considered, meaning that combines did not appear to distort yield data from start to finish in any systematic way.
+
+Average canola yield tended to increase with distance from Standard and Other Crop boundaries, leveling off at about 25 m (Figure \@ref(fig:canolaPlot) top row, Other crop and Standard panels); yield was approximately 0.1 T/ha (1.5 bu/ac) lower at the edge for both of these boundary types.
+Interestingly, Shelterbelts were well-represented in the data (Table \@ref(tab:boundaryTypes)) but the negative edge effect was not strong (Figure \@ref(fig:canolaPlot) top row, Shelterbelt panel), indicating that they may provide better micro-climates for crops located near them.
+Other boundary types were less common in the data and had correspondingly weaker effects on mean yield, but all showed a small negative edge effect.
+Yield variability tended to decrease with distance from the edge of the field, indicating that edge effects contribute to yield "patchiness" as well as averages.
+At Standard, Bare, and Wetland field boundaries, variability decreased until about 100 m, after which variablity was constant, while the pattern was less consistent in Grassland, Other Crop, and Shelterbelt.
+
+```{r canolaPlot, echo=FALSE, fig.cap='Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
 Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
@@
 
-```{r , echo=FALSE, fig.cap='Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
+Similar to canola, there was a strong negative effect of ground speed on both yield and variability of wheat yield.
+
+Average wheat yield (Figure \@ref(fig:wheatPlot)
+
+```{r wheatPlot, echo=FALSE, fig.cap='Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
 Upper panel represents mean yield, while the lower panel represents yield variation (i.e.

### 2021-08-27 Manuscript/manuscript.tex
Worked on marginal plots
- Because of intercept/slope correlation in pArea term, need to fit lmer models
- Entire model is on sqrt scale, so need to consider this in marginal plots

@@
 \makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} \usepackage{setspace} \linenumbers
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 Field boundaries were automatically digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC.
-Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g.~flooding during some years)
-Additionally, seminatural features often change yearly.
+Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g.~flooding during some years).
 Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
@@
 
-Despite the use of spatial and temporal smoothers, the residuals in most of the models displayed a large amount of spatial and temporal autocorrelation
+Despite the use of spatial and temporal smoothers, the residuals in most of the models displayed a large amount of spatial and temporal autocorrelation.
+\emph{Not really sure how we can get around this.
+In practice, this means that SEs for all coefficients are too small (i.e.~we gathered 1000 points, but because they're spatially related, it's only equivalent to 100 points).}
 
@@
 4. Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers
-This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability.
+This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
 
@@
 
+\hypertarget{field-boundaries}{%
+\subsection{Field boundaries}\label{field-boundaries}}
+
+``Standard'' (grassy border) was the most common boundary type, followed by ``Shelterbelt'' and ``Other Crop'' (Table \ref{tab:boundaryTypes}).
+
+\begin{table}
+
+\caption{\label{tab:boundaryTypes}Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.}
+\centering
+\begin{tabular}[t]{l|r|r|r}
+\hline
+Type & Mean & SD & Number of Fields\\
+\hline
+Bare & 0.0419 & 0.0618 & 126\\
+\hline
+Grassland & 0.0506 & 0.1254 & 84\\
+\hline
+Other Crop & 0.1515 & 0.2187 & 143\\
+\hline
+Shelterbelt & 0.3180 & 0.2236 & 274\\
+\hline
+Standard & 0.3870 & 0.2041 & 283\\
+\hline
+Wetland & 0.0509 & 0.0836 & 120\\
+\hline
+\end{tabular}
+\end{table}
+
 \hypertarget{canola}{%
 \subsection{Canola}\label{canola}}
+
+Average canola yield tended to increase with distance from Standard and Other Crop boundaries, leveling off at about 25 m (Figure \ref{fig:canolaPlot} top row, Other crop and Standard panels); yield was approximately 0.1 T/ha (1.5 bu/ac) lower at the edge for both of these boundary types.
+Interestingly, Shelterbelts were well-represented in the data (Table \ref{tab:boundaryTypes}) but the negative edge effect was not strong (Figure \ref{fig:canolaPlot} top row, Shelterbelt panel), indicating that they may provide better micro-climates for crops located near them.
+Other boundary types were less common in the data and had correspondingly weaker effects on mean yield, but all showed a small negative edge effect.
+
+Yield variability (SD of yield) also varied with distance from the edge of the field, but this
 
@@
 Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
-"patchiness").}\label{fig:unnamed-chunk-1}
+"patchiness").}\label{fig:canolaPlot}
 \end{figure}
@@
 
+Average wheat yield (Figure \ref{fig:wheatPlot}
+
 \begin{figure}
@@
 Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
-"patchiness").}\label{fig:unnamed-chunk-2}
+"patchiness").}\label{fig:wheatPlot}
 \end{figure}
@@
 
-\hypertarget{authors-contributions}{%
-\section{Authors' contributions}\label{authors-contributions}}
+\hypertarget{author-contributions}{%
+\section{Author Contributions}\label{author-contributions}}
 
-Author's contribution
+SVJR and PG conceived of the project.
+SVJR collected data, conducted analysis, and wrote the manuscript.
 
@@
 
+We thank Trent Clark, Dean Hubbard, Alvin French, and Kristina Polziehn for providing yield data for us, and providing insight on their yield data.
+We also thank Autumn Barnes and Keith Gabert of the Canola Council of Canada, and Laurel Thompson and JP Pettyjohn of Lakeland College for connecting us with farmers and agronomists.
 Funding for this research was provided by Ducks Unlimited Canada's Institute for Wetland and Waterfowl Research, the Alberta Canola Producers Commission, Manitoba Canola Growers Association, SaskCanola, the Alberta Biodiversity Monitoring Institute, and the Alberta Conservation Association.
@@
 
-Supplemental materials here
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/ModelSummary3b_canola} \caption{Effect of polygon area and harvest sequence on canola yield.
+Polygon area is equivalent to ground }\label{fig:unnamed-chunk-2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/ModelSummary3b_wheat} \caption{Temp}\label{fig:unnamed-chunk-3}
+\end{figure}
 

### 2021-08-27 compareFilteredData.R

@@ -21,7 +21,7 @@ samp_unfilt <- samp
 rm(samp); gc()
 
 ylabMean <- 'Mean Yield (T/ha)'
-ylabSD <- 'SD Yield'
+ylabSD <- 'log SD Yield'
 alphaVal <- 0.1 #Transparancy 
 qs <- c(0.1,0.5,0.9) #Quantiles
 scalecols <- c('blue','red')
@@ -29,7 +29,7 @@ scalecols <- c('blue','red')
 pAreaSamp <- lapply(list(samp_filt,samp_unfilt),function(samp){
   lapply(samp,function(x) x$pArea) %>% bind_rows(.id = 'sample') %>% 
     group_by(pArea) %>% 
-    summarise(predMean = quantile(predMean, qs), predLogSD = exp(quantile(predLogSD, qs)), q = qs) %>% 
+    summarise(predMean = quantile(predMean, qs), predLogSD = quantile(predLogSD, qs), q = qs) %>% 
     ungroup() %>% 
     mutate(q=factor(q,labels=c('lwr','med','upr'))) %>% 
     pivot_wider(names_from=q,values_from=c(predMean,predLogSD)) %>% 
@@ -47,8 +47,9 @@ p2 <- ggplot(pAreaSamp,aes(x=pArea))+
   geom_line(aes(y=predLogSD_med,col=Data),size=1)+
   labs(x='Polygon Area',y=ylabSD)+
   scale_color_manual(values=scalecols)+
-  scale_fill_manual(values=scalecols)+
-  coord_cartesian(ylim=c(0,1))
+  scale_fill_manual(values=scalecols)
+# +
+#   coord_cartesian(ylim=c(0,1))
 
 coverSamp <- lapply(list(samp_filt,samp_unfilt),function(samp){
   lapply(samp,function(x) x$coverDist %>% bind_rows(.id = 'dist_type')) %>% 

### 2021-08-27 runModels.R

@@ -32,7 +32,7 @@ source('helperFunctions.R')
 #   sapply(.,function(x) names(x)[1])
 # 
 # 
-# 
+# #Get crop types
 # datSource$crop <- case_when(grepl('(CANOLA|InVigor|Dekalb 74-54|Dekalb 75-42|Brett Young 6056|Dupont D3152|VT 9562)',cropType) ~ 'Canola',
 #                             grepl('(WHEAT|Wheat|CPS AC Foremost|CWRS|CWSW)',cropType) ~ 'Wheat',
 #                             grepl('Peas',cropType) ~ 'Peas',
@@ -43,6 +43,24 @@ source('helperFunctions.R')
 #                             TRUE ~ cropType) 
 # datSource$npoints <- sapply(datSource$dataPath,function(x) length(readLines(x)))
 # write.csv(datSource,'./Data/datSource_new.csv',row.names = FALSE)
+
+# #Get length of boundary features at each field
+# bLengths <- lapply(datSource$boundaryPath2,function(x){
+#   bTypes <- c('BARE','GRASSLAND','OTHERCROP','SHELTERBELT','STANDARD','WETLAND')
+#   if(!file.exists(x)){
+#     return(data.frame(type=bTypes,len=NA))
+#   }
+#   st_read(x,quiet=TRUE) %>%
+#     mutate(len=as.numeric(st_length(.))) %>%
+#     filter(len>0) %>%
+#     mutate(type=factor(type,levels=bTypes)) %>%
+#     st_drop_geometry() %>% group_by(type,.drop=FALSE) %>%
+#     summarize(len=sum(len))}) %>%
+#   set_names(nm = datSource$filename) %>%
+#   bind_rows(.id='filename') %>% group_by(filename) %>% mutate(prop=len/sum(len)) %>%
+#   pivot_wider(names_from=type,values_from=c(len,prop)) %>%
+#   data.frame()
+# write.csv(bLengths,file = './Data/boundaryLengths.csv')
 
 datSource <- read.csv('./Data/datSource.csv') %>% #Read previous datasource file
   mutate(boundaryComplete=file.exists(boundaryPath)) %>% #Has boundary been made already?
@@ -468,8 +486,8 @@ samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutof
   return(retList) 
 }
 
[…]

### 2021-08-31 runModels.R
Changed meta-model approach
- uses lmer for polygon area sub-model
- adds bs='re' smoother for distance/time smoothers
- got rid of generic crop approach

@@ -421,7 +421,7 @@ samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutof
     warning('Rows not specified. Using entire dataset.')
     useRows <- 1:nrow(datSource)
   } 
-  require(tidyverse); require(mgcv)
+  require(tidyverse); require(mgcv); require(lme4)
   source('helperFunctions.R')
   getFiles <- ds %>% slice(useRows) %>% filter(use,modelComplete2) #Completed models
   
@@ -431,8 +431,12 @@ samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutof
   #log(Area) meta-models
   pAreaDf <- lapply(allSmooths,function(x) x$pAreaDat) %>% bind_rows(.id = 'field')
   
-  m1 <- lm(mean~log(pArea),data=pAreaDf) 
-  m2 <- lm(logSD~log(pArea),data=pAreaDf)
+  #Adds a small amount of noise so model fit isn't singular
+  m1 <- pAreaDf %>% 
+    mutate(mean=mean+rnorm(n(),0,(max(mean)-min(mean))*0.005)) %>% 
+    lmer(mean~log(pArea)+(log(pArea)|field),data=.) 
+  m2 <- pAreaDf %>% mutate(logSD=logSD+rnorm(n(),0,(max(logSD)-min(logSD))*0.005)) %>% 
+    lmer(logSD~log(pArea)+(log(pArea)|field),data=.)
   
   #cover class meta-models
   
@@ -448,36 +452,50 @@ samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutof
   for(i in 1:ncol(containsCover)){
     useThese <- unname(which(containsCover[,i]))
     cname <- colnames(containsCover)[i]
-    coverDF <- lapply(allSmooths[useThese],function(x) x$distDat[[cname]]) %>% bind_rows(.id = 'field')
-    meanModList[[i]] <- gam(mean~s(dist,k=20),data=coverDF)
-    logSDModList[[i]] <- gam(logSD~s(dist,k=20),data=coverDF)
+    coverDF <- lapply(allSmooths[useThese],function(x) x$distDat[[cname]]) %>% 
+      bind_rows(.id = 'field') %>% mutate(field=factor(field))
[…]

### 2021-09-01 Manuscript/manuscript.Rmd
Added example section with single field

@@
 
-datSource <- read.csv('../Data/datSource.csv') #Get data source
+source('../helperFunctions.R')
+
+datSource <- read.csv('../Data/datSource.csv') %>% #Get data source
+    mutate(modelComplete2=file.exists(paste0('.',modelPath2)))  #Has model2 already been run?
+
 datSourceFinal <- datSource %>% filter(use) %>% filter(grepl('Wheat|Canola',crop)) 
@@
 
-## Field boundaries
+## Field-level exanoke
+
+The filtered data from each field still contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
+Figure \@ref(fig:examplePlot1)b shows the results from a single field to illustrate this, where areas of higher yield are clearly visible within the field.
+Additionally, Figure \@ref(fig:examplePlot1)c shows that yield "patchiness" (standard deviation, SD) also varies within fields, with areas of higher and lower patchiness also clearly visible within fields.
+
+```{r examplePlot1, echo=FALSE, fig.cap='Example output from a single field.', out.width='100%'}  
+  include_graphics('../Figures/ExamplePlots/Trent_Clark JOHNSON 2014.png',dpi=NA)
+```
+```{r examplePlot1, echo=FALSE, fig.cap='Distance smoothers from a single field.', out.width='100%'}  
+  include_graphics('../Figures/ExamplePlots/Trent_Clark JOHNSON 2014 Summary.png',dpi=NA)
+```
+
+## All canola fields
 
@@
 
-## Canola
-
 There was a strong negative effect combine ground speed on both the average and variability of canola yield.
-At the field level, both yield average and yield variability (SD of yield) of yield were extremely spatially (average p-values: `r canolaMods$meanP[canolaMods$Smoother=='s(E,N)']`, `r canolaMods$meanP[canolaMods$Smoother=='s.1(E,N)']`) and temporally dependent (`r canolaMods$meanP[canolaMods$Smoother=='s(r)']`, `r canolaMods$meanP[canolaMods$Smoother=='s.1(r)']`), meaning that average and variablity changed with location in the field, as well as sequentially during the harvest.
+At the field level, both yield average and yield variability (SD of yield) of yield were extremely spatially (average p-values: `r with(canolaMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(canolaMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`) and temporally dependent (`r with(canolaMods,formatC(meanP[Smoother=='s(r)'],3))`, `r with(canolaMods,formatC(meanP[Smoother=='s.1(r)'],3))`), meaning that average and variability changed with location in the field, as well as sequentially during the harvest.
 However, there was no difference in average or variability with harvest sequence when all fields were considered, meaning that combines did not appear to distort yield data from start to finish in any systematic way.
@@
 
-## Wheat
+## All wheat fields
 

### 2021-09-01 Manuscript/supplemental.Rmd
Added example section with single field

--- 2021-04-12 Manuscript/supplemental.Rmd
@@
 
-Supplemental materials here
+```{r , echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on canola yield.', out.width='100%'}  
+  include_graphics('../Figures/groundSpeed_canola.png',dpi=NA)
+```
+
+```{r , echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on wheat yield.', out.width='100%'}  
+  include_graphics('../Figures/groundSpeed_wheat.png',dpi=NA)
+```

### 2021-09-01 helperFunctions.R

@@ -115,7 +115,7 @@ convertYield <- function(x,inUnits=NULL,outUnits=NULL){
 #postSamp = sample posterior?
 #noIntercept = set intercept to 0?
 #returnDF = return DF? Otherwise returns vector of predictions
-getSmooths <- function(smoothLabel,modList,xvals,postSamp=FALSE,noIntercept=TRUE){
+getSmooths <- function(smoothLabel,modList,xvals,postSamp=FALSE,noIntercept=TRUE,returnSE=FALSE){
   
   # #Debugging
   # smoothLabel <- 's(dist)'
@@ -123,11 +123,14 @@ getSmooths <- function(smoothLabel,modList,xvals,postSamp=FALSE,noIntercept=TRUE
   # noIntercept <- TRUE
   # xvals <- seq(predRange[1],predRange[2],length.out=lengthOut) #x-values to use for smoother predictions
   
+  require(mgcv)
+  
   smoothLabs <- sapply(modList$smooth,function(x) x$label) #Labels for smoothers
   meanSmoothList <- modList$smooth[[which(smoothLabs==smoothLabel)]] #Get smoother info
   
   #Get locations of coefficients
-  meanVars <- c(which(grepl('Intercept',names(modList$coefs)))[1],meanSmoothList$first.para:meanSmoothList$last.para)
+  whichIntercept <- ifelse(grepl('s.1',smoothLabel),2,1) #Is intercept for mean or logSD?
+  meanVars <- c(which(grepl('Intercept',names(modList$coefs)))[whichIntercept],meanSmoothList$first.para:meanSmoothList$last.para)
   
   if(postSamp){ #Sample from posterior
     meanCoefs <- rnorm(rep(1,length(meanVars)),modList$coefs[meanVars],sqrt(diag(modList$vcv)[meanVars]))
@@ -138,7 +141,11 @@ getSmooths <- function(smoothLabel,modList,xvals,postSamp=FALSE,noIntercept=TRUE
   #Marginalize across intercept (set intercept coef to 0)
   if(noIntercept) meanCoefs[1] <- 0
   
-  predDF <- data.frame(x=xvals) #Dataframe for holding predictions
+  if('matrix' %in% class(xvals)){ #If xvals is a matrix (2D smooth)
+    predDF <- data.frame(x=xvals[,1],y=xvals[,2]) #Dataframe for holding predictions
+  } else { #If not (1D smooth)
[…]

### 2021-09-01 runModels.R

@@ -416,7 +416,7 @@ ggsave(paste0('./Figures/ModelSummary2a.png'),p,height=6,width=12,dpi=300)
 # My way of getting around the problem of random effects
 # a = does nothing (mandatory argument for parLapply), useRows = rows of ds to be used, ds = datSource csv, nX = number of samples along range of X
 # rCutoff = restrict point number (some fields have hundreds of thousands of points, but all data uses approximately the same size yield rectangles, so cuts off meta-smoother at "about" the maximum field size - i.e. the most common number of points per field)
-samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutoff=50000,margInt=c(FALSE,TRUE,TRUE)){ 
+samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutoff=50000,margInt=c(FALSE,FALSE,FALSE)){ 
   if(is.null(useRows)){
     warning('Rows not specified. Using entire dataset.')
     useRows <- 1:nrow(datSource)
@@ -469,16 +469,6 @@ samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutof
   #Assemble predictions
   pAreaPred <- data.frame(pArea=exp(seq(log(min(pAreaDf$pArea)),log(max(pAreaDf$pArea)),length.out=nX[1]))) %>% 
     mutate(predMean=predict(m1,newdata=.,re.form=~0),predLogSD=predict(m2,newdata=.,re.form=~0))
-  
-  # coverPred <- lapply(1:length(coverNames),function(i){
-  #   useThese <- unname(which(containsCover[,i]))
-  #   cname <- colnames(containsCover)[i]
-  #   coverDF <- lapply(allSmooths[useThese],function(x) x$distDat[[cname]]) %>% 
-  #     bind_rows(.id = 'field') %>% mutate(field=factor(field))
-  #   
-  #   data.frame(dist=seq(min(coverDF$dist),max(coverDF$dist),length.out=nX[2]),field=coverDF$field[1]) %>% 
[…]

### 2021-09-03 Manuscript/manuscript.Rmd
Added plots

@@
 library(kableExtra)
+theme_set(theme_bw())
 
@@
 <!-- - SNL has conflicting effects on crop yield because of ecosystem service provision as well as negative edge effects -->
-SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate
-Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds
-At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (@kowalchuk1995)
-For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance
-Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (e.g. @vanVooren2017)
-This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
+SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate.
+Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds.
+At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (@kowalchuk1995).
+For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance.
+Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (e.g. @vanVooren2017).
+This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
+<!-- - Less-considered are ecosystem ``disservices", where weeds and pests  -->
+<!-- - In North America, there are financial incentives to make fields large, homogeneous, and with simple geometry -->
 
-<!-- - Less-considered are ecosystem ``disservices", where weeds and pests  -->
-
-<!-- - In North America, there are financial incentives to make fields large, homogeneous, and with simple geometry -->
+```{r rawData, echo=FALSE, fig.cap='Potential yield patterns, depending on ecosystem service effects with distance.
+In both figures, negative environmental effects (edge effects) are shown in red, and decrease with distance from the edge of the crop, creating an increase in yield with distance.
+The left figure shows a decline in ecosystem services with distance, leading to an intermediate peak in yield, while the right figure shows an increase in services with distance, leading to a peak at the centre of the field (note: this also occurs if ecosystem services do not depend on distance).', out.width='75%'}
+  include_graphics('../Figures/ExamplePlots/hypotheses.png',dpi=NA)
+```
 
@@
 we gathered 1000 points, but because they're spatially related, it's only equivalent to 100 points)._
+
 <!-- - This is extremely common in precision yield data, but  -->
@@
 
-## Field-level exanoke
+## Field-level example
 
 The filtered data from each field still contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
-Figure \@ref(fig:examplePlot1)b shows the results from a single field to illustrate this, where areas of higher yield are clearly visible within the field.
-Additionally, Figure \@ref(fig:examplePlot1)c shows that yield "patchiness" (standard deviation, SD) also varies within fields, with areas of higher and lower patchiness also clearly visible within fields.
+To illustrate this, we show the results from a single wheat field, where we show the relative impact of ground speed, harvest sequence, boundary distance, and spatial variation on both yield mean and yield SD ("patchiness").
+Figure \@ref(fig:rawData) shows the raw data collected from a single field, demonstrating a large amount of variation even among the filtered data.
+Combine ground speed had a large effect on both yield mean and SD, and there was also a strong effect of harvest sequence on yield mean and SD (Figure \@ref(fig:orderSmooths).
+Yield mean and SD tended to decrease with distance from field boundaries, but there were no consistent differences between boundary types (Figure \@ref(fig:distSmooths).
+Finally, the random spatial smoothers in Figure \@ref(fig:spatialSmooths) show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions (Figure \@ref(fig:spatialSmooths).
 
-```{r examplePlot1, echo=FALSE, fig.cap='Example output from a single field.', out.width='100%'}  
-  include_graphics('../Figures/ExamplePlots/Trent_Clark JOHNSON 2014.png',dpi=NA)
-```
-```{r examplePlot1, echo=FALSE, fig.cap='Distance smoothers from a single field.', out.width='100%'}  
-  include_graphics('../Figures/ExamplePlots/Trent_Clark JOHNSON 2014 Summary.png',dpi=NA)
+```{r rawData, echo=FALSE, fig.cap='Example output from a single field.', out.width='75%'}  
+  include_graphics('../Figures/ExamplePlots/rawData.png',dpi=NA)
 ```
 
-## All canola fields
-
-"Standard" (grassy border) was the most common boundary type, followed by "Shelterbelt" and "Other Crop" (Table \@ref(tab:boundaryTypes)).
-
-```{r boundaryTypes, echo=FALSE, out.width='100%'}  
-bLengthSummary %>% 
-  kable(.,format='latex',caption='Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.',col.names = c('Type','Mean','SD','Number of Fields'),digits = c(1,4,4,1),escape = FALSE)
+```{r orderSmooths, echo=FALSE, fig.cap='Ground speed and harvest sequence (random temporal variation) smoothers from a single field.', out.width='100%'}  
+  include_graphics('../Figures/ExamplePlots/orderSmooths.png',dpi=NA)
 ```
 
-There was a strong negative effect combine ground speed on both the average and variability of canola yield.
+```{r distSmooths, echo=FALSE, fig.cap='Distance smoothers from a single field.', out.width='100%'}  
+  include_graphics('../Figures/ExamplePlots/distSmooths.png',dpi=NA)
+```
+
+```{r spatialSmooths, echo=FALSE, fig.cap='Spatial (random spatial variation) smoothers from a single field.', out.width='100%'}  
+  include_graphics('../Figures/ExamplePlots/spatialSmooths.png',dpi=NA)
+```
+
+## Overall results
+
+Across all fields, Standard (grassy border) field boundaries were the most common boundary type, followed by Shelterbelts and Other Crops.
+<!-- Not sure if we should talk about this. Not very many fields from southern regions, so it's hard to compare. Perhaps mention in Discussion -->
+
+```{r boundaryTypeTable, echo=FALSE, fig.cap='Percentage of field boundaries across sampled fields.', out.width='100%'}
+bLengthSummary %>% #Table version
+  kable(.,format='latex',caption='Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.',col.names = c('Type','Mean','SD','Number of Fields'),digits = c(1,4,4,1),escape = FALSE)
+
+```
+
+### Canola fields
+
+Overall, there was a strong negative effect of combine ground speed on both the average and variability of canola yield.
 At the field level, both yield average and yield variability (SD of yield) of yield were extremely spatially (average p-values: `r with(canolaMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(canolaMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`) and temporally dependent (`r with(canolaMods,formatC(meanP[Smoother=='s(r)'],3))`, `r with(canolaMods,formatC(meanP[Smoother=='s.1(r)'],3))`), meaning that average and variability changed with location in the field, as well as sequentially during the harvest.
@@
 Average canola yield tended to increase with distance from Standard and Other Crop boundaries, leveling off at about 25 m (Figure \@ref(fig:canolaPlot) top row, Other crop and Standard panels); yield was approximately 0.1 T/ha (1.5 bu/ac) lower at the edge for both of these boundary types.
-Interestingly, Shelterbelts were well-represented in the data (Table \@ref(tab:boundaryTypes)) but the negative edge effect was not strong (Figure \@ref(fig:canolaPlot) top row, Shelterbelt panel), indicating that they may provide better micro-climates for crops located near them.
+Interestingly, Shelterbelts were well-represented in the data (Table \@ref(tab:boundaryTypeTable)) but the negative edge effect was not strong (Figure \@ref(fig:canolaPlot) top row, Shelterbelt panel), indicating that they may provide better micro-climates for crops located near them.
 Other boundary types were less common in the data and had correspondingly weaker effects on mean yield, but all showed a small negative edge effect.
 Yield variability tended to decrease with distance from the edge of the field, indicating that edge effects contribute to yield "patchiness" as well as averages.
-At Standard, Bare, and Wetland field boundaries, variability decreased until about 100 m, after which variablity was constant, while the pattern was less consistent in Grassland, Other Crop, and Shelterbelt.
+At Standard, Bare, and Wetland field boundaries, variability decreased until about 100 m, after which variability was constant, while the pattern was less consistent in Grassland, Other Crop, and Shelterbelt.
 
@@
 
-## All wheat fields
+### Wheat fields
 
-Similar to canola, there was a strong negative effect of ground speed on both yield and variability of wheat yield.
+Similar to canola, there was a strong negative effect of ground speed on both yield and variability of wheat yield (Figure \@ref(fig:wheatPlotSpeed), left panels).
+At the field level, both yield average and yield variability (SD of yield) of yield were spatially (average p-values: `r with(wheatMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(wheatMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`) and temporally dependent (`r with(wheatMods,formatC(meanP[Smoother=='s(r)'],3))`, `r with(wheatMods,formatC(meanP[Smoother=='s.1(r)'],3))`), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \@ref(fig:wheatPlotSpeed), right panel).
 
-Average wheat yield (Figure \@ref(fig:wheatPlot)
+Average wheat yield increased slightly with distance from the field boundary, but this was much more variable than the patterns seen in canola yield (Figure \@ref(fig:wheatPlot)).
+There was also a slight increase in variability away from the field boundary, at least for Standard and Other Crop boundaries, but this was inconsistent among other boundary types.
 
@@
 
-Discussion here
+- Summary:
+  - At the level of individual fields, there was substantial variation in yield with distance from the field boundary, as well as with harvest sequence and spatially within each field
+  - Overall, however, the effect of field boundary on yield was inconsistent and the change in mean yield tended to be small ($\sim$0.5 T/ha for canola)
 

### 2021-09-03 Manuscript/manuscript.tex
Added plots

@@
   \end{abstract}
-   \begin{keyword} beetles; spiders; harvestmen;\end{keyword}
+   \begin{keyword} precision yield; spillover; landscape;\end{keyword}
  \end{frontmatter}
@@
 
-In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest.
+In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in harvested area, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest.
 Spatial or temporal variation can be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
@@
 The number of basis dimensions was checked with the \emph{gam.check} function from \emph{mgcv}.
-The relationship between polygon size (i.e.~ground speed) and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions.
+The relationship between polygon size and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions.
+Because polygon size is not intuitive, we converted it into ground speed by regressing speed on distance traveled per polygon, then using this to back-calculate the ground speed for a given polygon size (given a fixed width of header).
 All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).
@@
 This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
+Spatial variability was not averaged between fields because differences in field geometry make it impossible to compare fields.
 
@@
 
-\hypertarget{field-boundaries}{%
-\subsection{Field boundaries}\label{field-boundaries}}
+\hypertarget{field-level-example}{%
+\subsection{Field-level example}\label{field-level-example}}
 
-``Standard'' (grassy border) was the most common boundary type, followed by ``Shelterbelt'' and ``Other Crop'' (Table \ref{tab:boundaryTypes}).
+The filtered data from each field still contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
+To illustrate this, we show the results from a single wheat field, where we show the relative impact of ground speed, harvest sequence, boundary distance, and spatial variation on both yield mean and yield SD (``patchiness'').
+Figure \ref{fig:rawData} shows the raw data collected from a single field, demonstrating a large amount of variation even among the filtered data.
+Combine ground speed had a large effect on both yield mean and SD, and there was also a strong effect of harvest sequence on yield mean and SD (Figure \ref{fig:orderSmooths}.
+Yield mean and SD tended to decrease with distance from field boundaries, but there were no consistent differences between boundary types (Figure \ref{fig:distSmooths}.
+Finally, the random spatial smoothers in Figure \ref{fig:spatialSmooths} show that yield mean and SD systematically vary within fields, likely due to moisture and soil conditions (Figure \ref{fig:spatialSmooths}.
 
-\begin{table}
+\begin{figure}
+\includegraphics[width=0.75\linewidth]{../Figures/ExamplePlots/rawData} \caption{Example output from a single field.}\label{fig:rawData}
+\end{figure}
 
-\caption{\label{tab:boundaryTypes}Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.}
-\centering
-\begin{tabular}[t]{l|r|r|r}
-\hline
-Type & Mean & SD & Number of Fields\\
-\hline
-Bare & 0.0419 & 0.0618 & 126\\
-\hline
-Grassland & 0.0506 & 0.1254 & 84\\
-\hline
-Other Crop & 0.1515 & 0.2187 & 143\\
-\hline
-Shelterbelt & 0.3180 & 0.2236 & 274\\
-\hline
-Standard & 0.3870 & 0.2041 & 283\\
-\hline
-Wetland & 0.0509 & 0.0836 & 120\\
-\hline
-\end{tabular}
-\end{table}
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/orderSmooths} \caption{Ground speed and harvest sequence (random temporal variation) smoothers from a single field.}\label{fig:orderSmooths}
+\end{figure}
 
-\hypertarget{canola}{%
-\subsection{Canola}\label{canola}}
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/distSmooths} \caption{Distance smoothers from a single field.}\label{fig:distSmooths}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/spatialSmooths} \caption{Spatial (random spatial variation) smoothers from a single field.}\label{fig:spatialSmooths}
+\end{figure}
+
+\hypertarget{overall-results}{%
+\subsection{Overall results}\label{overall-results}}
+
+Across all fields, ``Standard'' (grassy border) was the most common boundary type, followed by Shelterbelts and Other Crops.
+
+\hypertarget{canola-fields}{%
+\subsubsection{Canola fields}\label{canola-fields}}
+
+Overall, there was a strong negative effect of combine ground speed on both the average and variability of canola yield.
+At the field level, both yield average and yield variability (SD of yield) of yield were extremely spatially (average p-values: 2e-16, 1.77e-05) and temporally dependent (2e-16, 0.000243), meaning that average and variability changed with location in the field, as well as sequentially during the harvest.
+However, there was no difference in average or variability with harvest sequence when all fields were considered, meaning that combines did not appear to distort yield data from start to finish in any systematic way.
 
@@
 Other boundary types were less common in the data and had correspondingly weaker effects on mean yield, but all showed a small negative edge effect.
-
-Yield variability (SD of yield) also varied with distance from the edge of the field, but this
+Yield variability tended to decrease with distance from the edge of the field, indicating that edge effects contribute to yield ``patchiness'' as well as averages.
+At Standard, Bare, and Wetland field boundaries, variability decreased until about 100 m, after which variablity was constant, while the pattern was less consistent in Grassland, Other Crop, and Shelterbelt.
 
@@
 
-\hypertarget{wheat}{%
-\subsection{Wheat}\label{wheat}}
+\hypertarget{wheat-fields}{%
+\subsubsection{Wheat fields}\label{wheat-fields}}
 
-Average wheat yield (Figure \ref{fig:wheatPlot}
+Similar to canola, there was a strong negative effect of ground speed on both yield and variability of wheat yield (Figure \ref{fig:wheatPlotSpeed}, left panels).
+At the field level, both yield average and yield variability (SD of yield) of yield were spatially (average p-values: 2e-16, 2e-16) and temporally dependent (2e-16, 2e-16), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \ref{fig:wheatPlotSpeed}, right panel).
+
+Average wheat yield increased slightly with distance from the field boundary, but this was much more variable than the patterns seen in canola yield (Figure \ref{fig:wheatPlot}).
+There was also a slight increase in variability away from the field boundary, at least for Standard and Other Crop boundaries, but this was inconsistent among other boundary types.
 
@@
 
-Discussion here
+\begin{itemize}
+\tightlist
+\item
+  Summary:
+
+  \begin{itemize}
+  \tightlist
+  \item
+    At the level of an individual field, there were
+  \end{itemize}
+\end{itemize}
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3b_canola} \caption{Effect of polygon area and harvest sequence on canola yield.
-Polygon area is equivalent to ground }\label{fig:unnamed-chunk-2}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/boundaryTypes-1} \caption{Percentage of field boundaries across sampled fields.}\label{fig:boundaryTypes}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3b_wheat} \caption{Temp}\label{fig:unnamed-chunk-3}
+\includegraphics[width=1\linewidth]{../Figures/groundSpeed_canola} \caption{Effect of ground speed and order (sequence of harvest) on canola yield.}\label{fig:canolaPlotSpeed}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/groundSpeed_wheat} \caption{Effect of ground speed and order (sequence of harvest) on wheat yield.}\label{fig:wheatPlotSpeed}
 \end{figure}

### 2021-09-03 Manuscript/supplemental.Rmd
Added plots

--- 2021-04-12 Manuscript/supplemental.Rmd
@@
 
-Supplemental materials here
+```{r boundaryTypes, echo=FALSE, fig.cap='Percentage of field boundaries across sampled fields.', out.width='100%'}
+bLengths %>% 
+  mutate(croptype=datSource$crop) %>% filter(croptype=='Canola'|croptype=='Wheat') %>% 
+  # mutate(location=ifelse(grepl('(Alvin|Dean)',filename),'Southern','Central')) %>% 
+  select(filename,croptype,contains('prop_')) %>% 
+  pivot_longer(contains('prop_')) %>% 
+  mutate(name=gsub('prop_','',name),value=value*100) %>% na.omit() %>%
+  mutate(name=factor(name,labels=c('Bare','Grassland','Other\nCrop','Shelterbelt','Standard','Wetland'))) %>% 
+  ggplot(aes(x=name,y=value,fill=croptype))+geom_boxplot()+
+  labs(y='Percentage of field boundary',fill='Crop')+
+  scale_fill_manual(values=c('grey50','white'))+
+  theme(axis.title.x = element_blank(),legend.position = 'bottom')
+
+# bLengthSummary %>% #Table version
+#   kable(.,format='latex',caption='Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.',col.names = c('Type','Mean','SD','Number of Fields'),digits = c(1,4,4,1),escape = FALSE)
+
+```
+
+```{r canolaPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on canola yield.', out.width='100%'}  
+  include_graphics('../Figures/groundSpeed_canola.png',dpi=NA)
+```
+
+```{r wheatPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on wheat yield.', out.width='100%'}  
+  include_graphics('../Figures/groundSpeed_wheat.png',dpi=NA)
+```

### 2021-09-03 helperFunctions.R

@@ -494,6 +494,7 @@ runModII <- function(i,dS,nSubSamp=50000,kPar=c(12,60,60,12,60,60),useClosest=TR
   # nSubSamp <- 50000
   # useClosest <- TRUE
   # kPar <- c(12,60,60,12,60,60)
+  # filterData <- TRUE
   
   if(dS$modelComplete2[i]) return('Already completed')
   if(!dS$use[i]) return('Not used')

### 2021-09-03 runModels.R

@@ -66,7 +66,8 @@ datSource <- read.csv('./Data/datSource.csv') %>% #Read previous datasource file
   mutate(boundaryComplete=file.exists(boundaryPath)) %>% #Has boundary been made already?
   mutate(modelComplete1=file.exists(modelPath1)) %>% #Has model1 already been run?
   mutate(modelComplete2=file.exists(modelPath2)) %>% #Has model2 already been run?
-  mutate(modelComplete0=file.exists(modelPath0)) #Has model0 already been run?
+  mutate(modelComplete0=file.exists(modelPath0)) %>% #Has model0 already been run?
+  mutate(location=ifelse(grepl('(Alvin|Dean)',filename),'Southern','Central')) #Field locations (southern or central AB)
 
 #Get summary results from models
 datSource %>% pull(use) %>% sum()
@@ -523,8 +524,8 @@ samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutof
 # stopCluster(cluster)
 # save(samp,file='./Data/postSamples_wheat.Rdata')
 
-# croptype <- 'canola'
-croptype <- 'wheat'
+croptype <- 'canola'
+# croptype <- 'wheat'
 
 load(paste0('./Data/postSamples_',croptype,'.Rdata'))
 
@@ -895,13 +896,14 @@ ggsave(paste0('./Figures/groundSpeed_',croptype,'.png'),p,height=6,width=8,dpi=3
 
 # Example figures (Trent Clark Johnson 2014) ---------------------------------------------------------
 
-use <- which(datSource$filename == 'Trent_Clark JOHNSON 2014')
+use <- which(datSource$filename == 'Trent_Clark W 34 2014')
 load(datSource$modelPath2[use]) #Load model
 
 #Spatial smoothers
 
 library(sf)
 fieldBoundary <- st_read(datSource$boundaryPath[use]) #Get boundary
+fieldBoundaryType <- st_read(datSource$boundaryPath2[use]) %>% mutate(type=factor(type)) #Get boundary
 crs <- st_crs(fieldBoundary) #Save CRS
[…]

### 2021-09-03 yield simulations.R

@@ -53,22 +53,27 @@ curve(leq(x,2,-0.1),0,100,ylim=c(0,1)) #Insects
 curve(leq(x,1,exp(-2))*leq(x,1,-0.1),0,100,ylim=c(0,1)) #Both
 curve(leq2(x,1,exp(-2),1,-0.1),0,100,ylim=c(0,1)) #Both
 
-par(mfrow=c(2,1))
+
+png(paste0('./Figures/ExamplePlots/hypotheses.png'),width=12,height=6,units='in',res=200)
+
+par(mfrow=c(1,2))
 #Scenario 1 - beneficial bugs decrease with distance into field
 a1 <- -2; b1 <- 0.2; a2 <- 3; b2 <- -0.03
 curve(leq2(x,a1,b1,a2,b2),0,100,ylim=c(0,1),
-      lwd=4,ylab='Yield',xlab='Distance from Edge',main='Scenario 1: bugs - with dist')
+      lwd=4,ylab='Yield',xlab='Distance from Edge')#,main='Scenario 1: bugs - with dist')
 curve(leq(x,a1,b1),0,100,xlim=range(dist),col='red',add=TRUE,lwd=2)
 curve(leq(x,a2,b2),0,100,xlim=range(dist),col='blue',add=TRUE,lwd=2)
-text(5,0.9,'Bugs',col='blue'); text(80,0.9,'Environment',col='red'); text(80,0.6,'Yield')
+text(5,0.85,'Ecosystem\nservices',col='blue'); text(80,0.9,'Environment',col='red'); text(80,0.6,'Yield')
 
 #Scenario 2 - beneficial bugs increase with distance into field (or pest insects at edge)
 a2 <- 0; b2 <- 0.03
 curve(leq2(x,a1,b1,a2,b2),0,100,ylim=c(0,1),
-      lwd=4,ylab='Yield',xlab='Distance from Edge',main='Scenario 2: bugs + with dist')
+      lwd=4,ylab='Yield',xlab='Distance from Edge')#,main='Scenario 2: bugs + with dist')
 curve(leq(x,a1,b1),0,100,xlim=range(dist),col='red',add=TRUE,lwd=2)
 curve(leq(x,a2,b2),0,100,xlim=range(dist),col='blue',add=TRUE,lwd=2)
 par(mfrow=c(1,1))
+
+dev.off()
 
 #Simulate data using logit-normal distribution --------------
 

### 2021-09-07 Manuscript/manuscript.Rmd
Added "raster" test
- aggregate data into hexagons, run similar analysis

@@
 
-- Summary:
-  - At the level of individual fields, there was substantial variation in yield with distance from the field boundary, as well as with harvest sequence and spatially within each field
-  - Overall, however, the effect of field boundary on yield was inconsistent and the change in mean yield tended to be small ($\sim$0.5 T/ha for canola)
+At the level of individual fields, both average and SD of yield varied with distance from the field boundary.
+Both metrics also varied with ground speed and harvest sequence, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
+However, the overall effect of field boundary on yield was inconsistent and the change tended to be small.
+For example, average canola yield was $\sim$0.5 T/ha lower close to Standard field edges, and yield SD was $\sim$0.3 T/ha SD higher.
+
+- Why no consistent effect for other boundary types? 
+  - Our analysis only looks at boundary types, without considering larger landscape context (e.g. Lan's paper)
+  - Boundary type itself is not well defined, so perhaps other metrics would be more helpful (height of veg, width of shelterbelt - I think this is important for hedgerows)
+  - Ecosystem disservices (e.g. weeds, pest insects) could also be coming from field boundaries, which may negate any effect of ecosystem services; this is still poorly studied (but see Woodcock 2015)
+  - Could vary by location; most fields were from central AB, but effects from southern AB may be confounding the signal
+  - Additional confounding variables (crop history, differences in tillage)
 

### 2021-09-07 rasterTest.R

@@ -1,0 +1,87 @@
+#Test analysis using discrete (rasterized) data
+# Reduces processing time from ~20 mins to ~3 mins
+
+library(tidyverse)
+theme_set(theme_bw())
+library(ggpubr)
+library(mgcv)
+library(sf)
+library(beepr)
+source('helperFunctions.R')
+
+datSource <- read.csv('./Data/datSource.csv') %>% #Read previous datasource file
+  mutate(boundaryComplete=file.exists(boundaryPath)) %>% #Has boundary been made already?
+  mutate(modelComplete1=file.exists(modelPath1)) %>% #Has model1 already been run?
+  mutate(modelComplete2=file.exists(modelPath2)) %>% #Has model2 already been run?
+  mutate(modelComplete0=file.exists(modelPath0)) %>% #Has model0 already been run?
+  mutate(location=ifelse(grepl('(Alvin|Dean)',filename),'Southern','Central')) #Field locations (southern or central AB)
+
+use <- which(datSource$filename == 'Trent_Clark W 34 2014')
+load(datSource$modelPath2[use]) #Load model
+
+#Spatial smoothers
+
+library(sf)
+fieldBoundary <- st_read(datSource$boundaryPath[use]) #Get boundary
+fieldBoundaryType <- st_read(datSource$boundaryPath2[use])# %>% mutate(type=factor(type)) #Get boundary
+crs <- st_crs(fieldBoundary) #Save CRS
+fieldBoundary <- st_sfc(st_polygon(lapply(fieldBoundary$geometry,function(x) st_coordinates(x)[,c('X','Y')])),crs=crs) #Fix hole geometry
+hexGrid <- st_make_grid(fieldBoundary,square=FALSE,n=75)  #Make hexagonal grid
+hexGrid <- hexGrid[sapply(st_within(st_centroid(hexGrid),fieldBoundary),function(x) length(x)==1)] %>%  #Strip out points outside polygon
+  st_sf() #Set as sf object
+hexGrid %>% ggplot()+geom_sf()+geom_sf(data=fieldBoundary,fill=NA,col='red') #Looks OK
+
+dat <- read.csv(datSource$dataPath[use],stringsAsFactors=TRUE,fileEncoding='latin1') %>% 
[…]

### 2021-09-07 runModels.R

@@ -897,6 +897,7 @@ ggsave(paste0('./Figures/groundSpeed_',croptype,'.png'),p,height=6,width=8,dpi=3
 # Example figures (Trent Clark Johnson 2014) ---------------------------------------------------------
 
 use <- which(datSource$filename == 'Trent_Clark W 34 2014')
+#Could also use 'Trent_Clark JOHNSON 2014'
 load(datSource$modelPath2[use]) #Load model
 
 #Spatial smoothers

### 2021-09-07 Manuscript/manuscript.Rmd
Worked on discussion

@@
 
-SNL in and around crops is important for both agricultural production and conservation
-They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control (@albrecht2020; @gardner2021)
-They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (@kort1988)
+SNL in and around crops is important for both agricultural production and conservation, as it has the potential to create better biotic and abiotic conditions (Case 2019, https://doi.org/10.1101/624635, Weninger 2021 https://doi.org/10.1088/1748-9326/ac1d0d)
+They are habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control (@albrecht2020; @gardner2021), but this can vary depending on the system (e.g.
+Quinn et al 2017, https://doi.org/10.1093/jee/tow306)
+They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (@kort1988, Brandle et al 2004 https://doi.org/10.1007/978-94-017-2424-1_5)
 Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
@@
 
-Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality (but see wheat study)
-Meta-analyses have been the standard solution to this problem, but precision yield data holds enormous promise.
+Studies of SNL effects on crop yield also suffer from limited scope (e.g.
+few crop types) and small sample sizes, limiting inference and reducing generality (Baker et al 2018, https://doi.org/10.1071/CP17242) (but see also wheat study)
+Meta-analyses have been the standard solution to this problem (e.g. @albrecht2020), but precision yield data holds enormous promise.
 However, its use is limited for several reasons: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) unfamiliarity with spatial statistics.
@@
 For example, average canola yield was $\sim$0.5 T/ha lower close to Standard field edges, and yield SD was $\sim$0.3 T/ha SD higher.
+So, why did we find few consistent effects of boundary type on canola and wheat yield?
+We consider a few potential reasons for this below:
+1. Landscape influence
+2. Boundary variables
+3. Ecosystem disservices
+4. Yearly/location variation
+5. Lack of pest pressure/pollination deficit
 
-- Why no consistent effect for other boundary types? 
-  - Our analysis only looks at boundary types, without considering larger landscape context (e.g. Lan's paper)
-  - Boundary type itself is not well defined, so perhaps other metrics would be more helpful (height of veg, width of shelterbelt - I think this is important for hedgerows)
-  - Ecosystem disservices (e.g. weeds, pest insects) could also be coming from field boundaries, which may negate any effect of ecosystem services; this is still poorly studied (but see Woodcock 2015)
+Our analysis only examined boundary types, without considering the landscape context that fields were embedded in
+   - Some studies show the influence of landscape composition at larger spatial scales (e.g.
+Lan's paper) which could be particularly important when beneficial arthropods are highly mobile (e.g.
+bumblebee foraging) or disperse very broadly (Lynx spider paper from Sweden)
+   - There could also be interactions between landscape composition at different scales (e.g. our paper in AGEE), but this is not well-studied
+   - Landscape complexity (Fahrig et al 2010,  https://doi.org/10.1111/j.1461-0248.2010.01559.x) and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others)
+   - This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat (but see Redhead 2020(?) ie. SLOSS problem)
+
+Our definitions of boundary type were necessarily loose, as we used satellite imagery to manually classify them
+  - However, it may be that other boundary traits, such as size or plant species composition, are more important for spillover of ecosystems services into fields
+  - For example, larger hedgerows tend to support more biodiversity than smaller hedgerows (Lecq et al 2018, https://doi.org/10.1016/j.ufug.2018.08.002, Stasiov 2020 https://doi.org/10.1016/j.baae.2020.05.001) or more biodiverse (Bonifacio et al 2011 https://doi.org/10.1071/PC110022)
+  - Boughey et al 2011 (https://doi.org/10.1016/j.biocon.2011.02.017) showed that trees within hedgerows are important for bat diversity
+  - Increased litter cover and soil compaction increased pollen beetle predators (Sutter et al 2018 https://doi.org/10.1016/j.agee.2018.06.030)
+
+The net flow of ecosystem services may not be positive, if beneficial organisms do not disperse from the SNL, or if pests disperse along with them.
+- Ecosystem disservices such as pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang et al 2007 https://doi.org/10.1016/j.ecolecon.2007.02.024)
+  - Gray and Lewis 2014 showed that remnant forest fragments do not provide significant pest control services (https://doi.org/10.1016/j.baae.2014.09.009)
+  - However, this is still poorly studied (but see Woodcock 2015), and parsing out services from disservices is impossible using crop yield data _alone_
+
+Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield
+  - (crop history, differences in tillage)
   - Could vary by location; most fields were from central AB, but effects from southern AB may be confounding the signal
-  - Additional confounding variables (crop history, differences in tillage)
+  - Between-year effects may be important, especially when accounting for
+
+Finally, pest pressure or pollination deficit may be at low enough levels that ecosystem service provision does not significantly change yield
+  - Canola is largely self-pollinating (at least in the hybrid canola varieties commonly used in Canada) and wheat is wind-pollinated, meaning that insect pollination may not change yields significantly
+  - Pest pressures are also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty
+
+We used harvest time as smoothed variable to deal with temporal autocorrelation in the data, but this is a crude way to deal with this
+  - This is because in-field combine driving patterns _almost always_ induce correlation between spatial and temporal measurements (e.g.
+northern section of the field was harvested before the southern section), meaning that temporal and spatial random effects are difficult to separate
+  - There are few solutions to this using harvest data as-is, other than to ignore temporal random effects entirely, thereby constraining all random variation to be spatial
+  - However, it is possible to remove correlation between spatial and temporal measurements by using specially designed harvest patterns within a field (harvesting alternating strips sequentially across the field, then harvesting the remaining strips on a second pass)
+  - Future studies could use this to estimate drift in combine sensor measurements over time
 

### 2021-09-08 Manuscript/manuscript.Rmd
Worked on discussion

--- 2021-09-03 Manuscript/manuscript.Rmd
@@
 
-datSourceFinal <- datSource %>% filter(use) %>% filter(grepl('Wheat|Canola',crop)) 
+useThese <- with(datSource,use&grepl('Wheat|Canola|Peas',crop)) #Rows to use
+
+datSourceFinal <- datSource %>% filter(useThese)
 yearsPerField <- datSourceFinal %>% count(field) %>% pull(n) #How many years of data at each unique location?
@@
 #Boundary lengths
-bLengths <- read.csv('../Data/boundaryLengths.csv')
+bLengths <- read.csv('../Data/boundaryLengths.csv') %>% filter(useThese)
 bNames <- c('Bare','Grassland','Other Crop','Shelterbelt','Standard','Wetland')
@@
   summarize(meanP=mean(pval),maxP=max(pval),propSig=sum(pval<0.05)/n(),meanEDF=mean(edf),meanRef=mean(Ref.df))
+
+use <- with(datSource,use&modelComplete2&crop=='Peas')
+peaMods <- lapply(paths[use],function(x) getModelInfo(x)$smooths) %>% 
+  set_names(nm=datSource$filename[use]) %>% 
+  bind_rows(.id='field') %>% 
+  mutate(pval=as.numeric(pval)) %>% 
+  group_by(Smoother) %>% 
+  summarize(meanP=mean(pval),maxP=max(pval),propSig=sum(pval<0.05)/n(),meanEDF=mean(edf),meanRef=mean(Ref.df))
+
 rm(use,paths)
@@
 
-Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth
-This has allowed increases in human population and increased (global) stability in production
-However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms
-Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios
-Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems
+Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth.
+This has allowed increases in human population and increased global stability in production.
+However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
+Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios.
+Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems.
 
-SNL in and around crops is important for both agricultural production and conservation
-They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control (@albrecht2020; @gardner2021)
-They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (@kort1988)
+SNL in and around crops is important for both agricultural production and conservation, as it has the potential to create better biotic and abiotic conditions [@case2019; @weninger2021].
+They are habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control [@albrecht2020; @gardner2021], but this can vary depending on the system [@gray2014; @quinn2017]
+They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed [@kort1988; @brandle2004]
 Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
@@
 Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (e.g. @vanVooren2017).
-This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
+This suggests a "goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
 <!-- - Less-considered are ecosystem ``disservices", where weeds and pests  -->
@@
 
-```{r rawData, echo=FALSE, fig.cap='Potential yield patterns, depending on ecosystem service effects with distance.
+```{r hypotheses, echo=FALSE, fig.cap='Potential yield patterns, depending on ecosystem service effects with distance.
 In both figures, negative environmental effects (edge effects) are shown in red, and decrease with distance from the edge of the crop, creating an increase in yield with distance.
-The left figure shows a decline in ecosystem services with distance, leading to an intermediate peak in yield, while the right figure shows an increase in services with distance, leading to a peak at the centre of the field (note: this also occurs if ecosystem services do not depend on distance).', out.width='75%'}
+The left figure shows a decline in ecosystem services with distance, leading to an intermediate peak in yield, while the right figure shows an increase in services with distance, leading to a peak at the centre of the field (note: Scenario 2 also occurs if ecosystem services do not change with distance).', out.width='100%'}
   include_graphics('../Figures/ExamplePlots/hypotheses.png',dpi=NA)
@@
 
-Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality (but see wheat study)
-Meta-analyses have been the standard solution to this problem, but precision yield data holds enormous promise.
-However, its use is limited for several reasons: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) unfamiliarity with spatial statistics.
-Ecosystem services can influence both the mean and variability of yield in agroecosystems (@redhead2020).
-Typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable.
+Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality (@baker2018; but see also @redhead2020)
+Meta-analyses have been the standard solution to this problem (e.g. @albrecht2020), but precision yield data holds enormous promise.
+However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
+Ecosystem services can influence both the mean and variability of yield in agroecosystems (@redhead2020); typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from the perspective of a grower.
 There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
@@
   3. Is there an intermediate distance where yield is maximized or variance is minimized?
+To our knowledge, no other study has addressed the effect of field boundaries on productivity using precision yield, making this study highly novel.
 
@@
 `r round(100*sum(grepl('(Wheat|Canola)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_) or canola (_Brassica napus_), two of the most common crops in rotation in Alberta.
-The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`) or canola (`r sum(datSourceFinal$crop=='Canola')`).
-Individual fields contained between `r min(yearsPerField)` to `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`), containing a total of `r round(sum(datSource$npoints)/1e6,1)` million data points.
+The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`), canola (`r sum(datSourceFinal$crop=='Canola')`), or peas (`r sum(datSourceFinal$crop=='Peas')`).
+Individual fields contained between `r min(yearsPerField)` and `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`), containing a total of `r round(sum(datSourceFinal$npoints)/1e6,1)` million data points.
 
-Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted to account for overlap with adjacent swaths).
+Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted to account for overlap with adjacent swaths).
 We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
@@
 
-Across all fields, Standard (grassy border) field boundaries were the most common boundary type, followed by Shelterbelts and Other Crops.
+Across all fields, Standard (grassy border) field boundaries were the most common boundary type, followed by Shelterbelts and Other Crops (Figure \@ref(fig:boundaryTypes)).
 <!-- Not sure if we should talk about this. Not very many fields from southern regions, so it's hard to compare. Perhaps mention in Discussion -->
 
-```{r boundaryTypeTable, echo=FALSE, fig.cap='Percentage of field boundaries across sampled fields.', out.width='100%'}
-bLengthSummary %>% #Table version
-  kable(.,format='latex',caption='Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.',col.names = c('Type','Mean','SD','Number of Fields'),digits = c(1,4,4,1),escape = FALSE)
+```{r boundaryTypes, echo=FALSE, fig.cap='Percentage of field boundaries across sampled fields.', out.width='100%'}
+bLengths %>% 
+  mutate(croptype=datSourceFinal$crop) %>% 
+  mutate(location=ifelse(grepl('(Alvin|Dean)',filename),'Southern Alberta','Central Alberta')) %>%
+  group_by(location) %>% mutate(nFields=length(unique(filename))) %>% ungroup() %>% 
+  mutate(location=paste0(location,' (N field-years = ',nFields,')')) %>%
+  select(filename,croptype,location,contains('prop_')) %>% 
+  pivot_longer(contains('prop_')) %>% 
+  mutate(name=gsub('prop_','',name),value=value*100) %>% na.omit() %>%
+  mutate(name=factor(name,labels=c('Bare','Grassland','Other\nCrop','Shelterbelt','Standard','Wetland'))) %>% 
+  ggplot(aes(x=name,y=value,fill=croptype))+geom_boxplot()+
+  facet_wrap(~location,ncol=1) +
+  labs(y='Percentage of field boundary',fill='Crop')+
+  scale_fill_manual(values=c('grey50','grey80','white'))+
+  theme(axis.title.x = element_blank(),legend.position = 'bottom')
+
+# bLengthSummary %>% #Table version
+#   kable(.,format='latex',caption='Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.',col.names = c('Type','Mean','SD','Number of Fields'),digits = c(1,4,4,1),escape = FALSE)
 
@@
 Average canola yield tended to increase with distance from Standard and Other Crop boundaries, leveling off at about 25 m (Figure \@ref(fig:canolaPlot) top row, Other crop and Standard panels); yield was approximately 0.1 T/ha (1.5 bu/ac) lower at the edge for both of these boundary types.
-Interestingly, Shelterbelts were well-represented in the data (Table \@ref(tab:boundaryTypeTable)) but the negative edge effect was not strong (Figure \@ref(fig:canolaPlot) top row, Shelterbelt panel), indicating that they may provide better micro-climates for crops located near them.
+Interestingly, Shelterbelts were well-represented in the data (Figure \@ref(fig:boundaryTypes)) but the negative edge effect was not strong (Figure \@ref(fig:canolaPlot) top row, Shelterbelt panel), indicating that they may provide better micro-climates for crops located near them.
 Other boundary types were less common in the data and had correspondingly weaker effects on mean yield, but all showed a small negative edge effect.
@@
 
+### Pea fields
+
+At the field level, both yield average and yield variability (SD of yield) of yield were spatially (average p-values: `r with(peaMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(peaMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`) and temporally dependent (`r with(peaMods,formatC(meanP[Smoother=='s(r)'],3))`, `r with(peaMods,formatC(meanP[Smoother=='s.1(r)'],3))`),
+There was a strong negative effect of ground speed on yield and variability of peas (Figure \@ref(fig:peasPlotSpeed)), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \@ref(fig:wheatPlotSpeed), right panel).
+
+Average pea yields increased with distance from Shelterbelt, Standard, and Wetland field boundaries.
+Interestingly, the reduction at the edge was slightly higher for Wetlands, but yield increased with distance more quickly, i.e. yields were higher at distance for Wetlands.
+Variability decreased with distance from Grassland and Standard boundaries
+
+```{r peaPlot, echo=FALSE, fig.cap='Field boundary effect on pea yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
+Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
+"patchiness").', out.width='100%'}
+  include_graphics('../Figures/ModelSummary3a_peas.png',dpi=NA)
+```
+
 # Discussion
 
-- Summary:
-  - At the level of individual fields, there was substantial variation in yield with distance from the field boundary, as well as with harvest sequence and spatially within each field
-  - Overall, however, the effect of field boundary on yield was inconsistent and the change in mean yield tended to be small ($\sim$0.5 T/ha for canola)
+At the level of individual fields, both average and SD of yield varied with distance from the field boundary.
+Both metrics also varied with ground speed and harvest sequence, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
+Yield tended to increase with distance from field boundaries before plateauing, and there was little evidence of intermediate boosts in yield, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \@ref(fig:hypotheses)).
+However, the effect of field boundary types was inconsistent, and the change tended to be small, especially for wheat.
+<!-- For example, average canola yield was about 0.5 T/ha lower close to Standard field edges, and yield SD was about 0.3 T/ha SD higher. -->
+So, why did we find few consistent effects of boundary type on canola and wheat yield?
+We consider a few potential reasons for this below:
+1. Landscape influence
+2. Boundary variables
+3. Ecosystem disservices
+4. Yearly/location variation
+5. Lack of pest pressure/pollination deficit
+
+Our analysis only examined boundary types, without considering the landscape context that fields were embedded in.
+Some studies show the influence of landscape composition at larger spatial scales (e.g.
+Lan's paper) which could be particularly important when beneficial arthropods are highly mobile (e.g.
+bumblebee foraging) or disperse very broadly [@oberg2007].
+There could also be interactions between landscape composition at different scales (e.g. our paper in AGEE), but this is not well-studied.
+Landscape complexity [@fahrig2011] and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others).
+This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat (but see @fahrig2019).
+
+Our definitions of boundary type were necessarily loose, as we used satellite imagery to manually classify them.
+However, it may be that other boundary traits, such as size or plant species composition, are more important for spillover of ecosystems services into fields.
+For example, larger [@lecq2018; @stasiov2020] or more functionally diverse (@bonifacio2011; @boughey2011) hedgerows tend to support more biodiversity than smaller hedgerows.
+Increased litter cover and soil compaction increased pollen beetle predators [@sutter2018b].
+
+Finally, pest pressure or pollination deficit may be at low enough levels that ecosystem service provision does not significantly change yield.
+Canola is largely self-pollinating (at least in hybrid canola varieties, @lindstrom2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop [@woodcock2016].
+Pest pressures may also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty.
+
+The net flow of ecosystem services may not be positive, if beneficial organisms do not disperse from the SNL, or if pests disperse along with them.
+Ecosystem disservices such as pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (@zhang2007).
+@gray2014 showed that remnant forest fragments do not provide significant pest control services , but also do not depress yield.
+However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data _alone_.
+
+Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
+Most fields were located in Central Alberta (Dry Mixedwood region, @ABRegions2006), where boundaries contained a higher proportion of Shelterbelts (Figure \@ref(fig:boundaryTypes)), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
+Between-year effects may be important, especially during dry years [@kort1988], but this was not considered in our analysis.
+Future work could examine whether boundaries have differing effects during hot or cool years.
+
+We used harvest time as smoothed variable to deal with temporal autocorrelation in the data, but this is a crude way to deal with this.
+This is because in-field combine driving patterns _almost always_ induce correlation between spatial and temporal measurements (e.g.
+northern section of the field was harvested before the southern section), meaning that temporal and spatial random effects are difficult to separate.
+There are few solutions to this using harvest data as-is, other than to ignore temporal random effects entirely, thereby constraining all random variation to be spatial.
+However, it is possible to remove correlation between spatial and temporal measurements by using specially designed harvest patterns within a field (harvesting alternating strips sequentially across the field, then harvesting the remaining strips on a second pass).
+Future studies could use this to estimate drift in combine sensor measurements over time.
 

### 2021-09-08 Manuscript/manuscript.tex
Worked on discussion

@@
 
-Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth
-This has allowed increases in human population and increased (global) stability in production
-However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms
-Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios
-Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems
+Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth.
+This has allowed increases in human population and increased global stability in production.
+However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
+Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios.
+Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems.
 
-SNL in and around crops is important for both agricultural production and conservation
-They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control (Albrecht \emph{et al.} (2020); Gardner \emph{et al.} (2021))
-They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (Kort (1988))
+SNL in and around crops is important for both agricultural production and conservation, as it has the potential to create better biotic and abiotic conditions (Case \emph{et al.} 2019; Weninger \emph{et al.} 2021).
+They are habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this can vary depending on the system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017)
+They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (Kort 1988; Brandle \emph{et al.} 2004)
 Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
@@
 
-SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate
-Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds
-At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (Kowalchuk \& Jong (1995))
-For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance
-Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (e.g. Van Vooren \emph{et al.} (2017))
-This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
+SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate.
+Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds.
+At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (Kowalchuk \& Jong (1995)).
+For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance.
+Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (e.g. Van Vooren \emph{et al.} (2017)).
+This suggests a ``goldilocks'' field size, where negative edge effects are canceled out by ecosystem services.
 
-Studies of SNL effects on crop yield also suffer from limited scope (e.g.~few crop types) and small sample sizes, limiting inference and reducing generality (but see wheat study)
-Meta-analyses have been the standard solution to this problem, but precision yield data holds enormous promise.
-However, its use is limited for several reasons: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) unfamiliarity with spatial statistics.
-Ecosystem services can influence both the mean and variability of yield in agroecosystems (Redhead \emph{et al.} (2020)).
-Typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable.
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/hypotheses} \caption{Potential yield patterns, depending on ecosystem service effects with distance.
+In both figures, negative environmental effects (edge effects) are shown in red, and decrease with distance from the edge of the crop, creating an increase in yield with distance.
+The left figure shows a decline in ecosystem services with distance, leading to an intermediate peak in yield, while the right figure shows an increase in services with distance, leading to a peak at the centre of the field (note: Scenario 2 also occurs if ecosystem services do not change with distance).}\label{fig:hypotheses}
+\end{figure}
+
+Studies of SNL effects on crop yield also suffer from limited scope (e.g.~few crop types) and small sample sizes, limiting inference and reducing generality (Baker \emph{et al.} (2018); but see also Redhead \emph{et al.} (2020))
+Meta-analyses have been the standard solution to this problem (e.g. Albrecht \emph{et al.} (2020)), but precision yield data holds enormous promise.
+However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
+Ecosystem services can influence both the mean and variability of yield in agroecosystems (Redhead \emph{et al.} (2020)); typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from the perspective of a grower.
 There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
@@
 3. Is there an intermediate distance where yield is maximized or variance is minimized?
+To our knowledge, no other study has addressed the effect of field boundaries on productivity using precision yield, making this study highly novel.
 
@@
 72\% of the crop types where either wheat (\emph{Triticum aestivum}) or canola (\emph{Brassica napus}), two of the most common crops in rotation in Alberta.
-The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94) or canola (119).
-Individual fields contained between 1 to 5 years of data (mean: 2.7), containing a total of 18.4 million data points.
+The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39).
+Individual fields contained between 1 and 6 years of data (mean: 3.1), containing a total of 14.4 million data points.
 
-Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted to account for overlap with adjacent swaths).
+Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted to account for overlap with adjacent swaths).
 We extracted the size of each polygon (m\(^2\)), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
@@
 6. Grassland: permanent seminatural grassland or pasture (not in rotation)
-
-\begin{verbatim}
-<!-- STANDARD -->
-<!-- WETLAND -->
-<!-- SHELTERBELT -->
-<!-- OTHERCROP -->
-<!-- BARE -->
-<!-- GRASSLAND -->
-\end{verbatim}
 
@@
 Yield mean and SD tended to decrease with distance from field boundaries, but there were no consistent differences between boundary types (Figure \ref{fig:distSmooths}.
-Finally, the random spatial smoothers in Figure \ref{fig:spatialSmooths} show that yield mean and SD systematically vary within fields, likely due to moisture and soil conditions (Figure \ref{fig:spatialSmooths}.
+Finally, the random spatial smoothers in Figure \ref{fig:spatialSmooths} show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions (Figure \ref{fig:spatialSmooths}.
 
@@
 
-Across all fields, ``Standard'' (grassy border) was the most common boundary type, followed by Shelterbelts and Other Crops.
+Across all fields, Standard (grassy border) field boundaries were the most common boundary type, followed by Shelterbelts and Other Crops (Figure \ref{fig:boundaryTypes}).
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/boundaryTypes-1} \caption{Percentage of field boundaries across sampled fields.}\label{fig:boundaryTypes}
+\end{figure}
 
@@
 Average canola yield tended to increase with distance from Standard and Other Crop boundaries, leveling off at about 25 m (Figure \ref{fig:canolaPlot} top row, Other crop and Standard panels); yield was approximately 0.1 T/ha (1.5 bu/ac) lower at the edge for both of these boundary types.
-Interestingly, Shelterbelts were well-represented in the data (Table \ref{tab:boundaryTypes}) but the negative edge effect was not strong (Figure \ref{fig:canolaPlot} top row, Shelterbelt panel), indicating that they may provide better micro-climates for crops located near them.
+Interestingly, Shelterbelts were well-represented in the data (Figure \ref{fig:boundaryTypes}) but the negative edge effect was not strong (Figure \ref{fig:canolaPlot} top row, Shelterbelt panel), indicating that they may provide better micro-climates for crops located near them.
 Other boundary types were less common in the data and had correspondingly weaker effects on mean yield, but all showed a small negative edge effect.
 Yield variability tended to decrease with distance from the edge of the field, indicating that edge effects contribute to yield ``patchiness'' as well as averages.
-At Standard, Bare, and Wetland field boundaries, variability decreased until about 100 m, after which variablity was constant, while the pattern was less consistent in Grassland, Other Crop, and Shelterbelt.
+At Standard, Bare, and Wetland field boundaries, variability decreased until about 100 m, after which variability was constant, while the pattern was less consistent in Grassland, Other Crop, and Shelterbelt.
 
@@
 
+\hypertarget{pea-fields}{%
+\subsubsection{Pea fields}\label{pea-fields}}
+
+At the field level, both yield average and yield variability (SD of yield) of yield were spatially (average p-values: 2e-16, 2e-16) and temporally dependent (2e-16, 3.62e-07),
+There was a strong negative effect of ground speed on yield and variability of peas (Figure \ref{fig:peasPlotSpeed}), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \ref{fig:wheatPlotSpeed}, right panel).
+
+Average pea yields increased with distance from Shelterbelt, Standard, and Wetland field boundaries.
+Interestingly, the reduction at the edge was slightly higher for Wetlands, but yield increased with distance more quickly, i.e.~yields were higher at distance for Wetlands.
+Variability decreased with distance from Grassland and Standard boundaries
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_peas} \caption{Field boundary effect on pea yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
+Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
+"patchiness").}\label{fig:peaPlot}
+\end{figure}
+
 \hypertarget{discussion}{%
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Summary:
+At the level of individual fields, both average and SD of yield varied with distance from the field boundary.
+Both metrics also varied with ground speed and harvest sequence, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
+Yield tended to increase with distance from field boundaries before plateauing, and there was little evidence of intermediate boosts in yield, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \ref{fig:hypotheses}).
+However, the effect of field boundary types was inconsistent, and the change tended to be small, especially for wheat.
+So, why did we find few consistent effects of boundary type on canola and wheat yield?
+We consider a few potential reasons for this below:
+1. Landscape influence
+2. Boundary variables
+3. Ecosystem disservices
+4. Yearly/location variation
+5. Lack of pest pressure/pollination deficit
 
-  \begin{itemize}
-  \tightlist
-  \item
-    At the level of an individual field, there were
-  \end{itemize}
-\end{itemize}
+Our analysis only examined boundary types, without considering the landscape context that fields were embedded in.
+Some studies show the influence of landscape composition at larger spatial scales (e.g.~Lan's paper) which could be particularly important when beneficial arthropods are highly mobile (e.g.~bumblebee foraging) or disperse very broadly (Öberg \emph{et al.} 2007).
+There could also be interactions between landscape composition at different scales (e.g.~our paper in AGEE), but this is not well-studied.
+Landscape complexity (Fahrig \emph{et al.} 2011) and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others).
+This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat (but see Fahrig \emph{et al.} (2019)).
+
+Our definitions of boundary type were necessarily loose, as we used satellite imagery to manually classify them.
+However, it may be that other boundary traits, such as size or plant species composition, are more important for spillover of ecosystems services into fields.
+For example, larger (Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020) or more functionally diverse (Bonifacio \emph{et al.} (2011); Boughey \emph{et al.} (2011)) hedgerows tend to support more biodiversity than smaller hedgerows.
+Increased litter cover and soil compaction increased pollen beetle predators (Sutter \emph{et al.} 2018).
+
+Finally, pest pressure or pollination deficit may be at low enough levels that ecosystem service provision does not significantly change yield.
+Canola is largely self-pollinating (at least in hybrid canola varieties, Lindström \emph{et al.} (2016)) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop (Woodcock \emph{et al.} 2016).
+Pest pressures may also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty.
+
+The net flow of ecosystem services may not be positive, if beneficial organisms do not disperse from the SNL, or if pests disperse along with them.
+Ecosystem disservices such as pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang \emph{et al.} (2007)).
+Gray \& Lewis (2014) showed that remnant forest fragments do not provide significant pest control services , but also do not depress yield.
+However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data \emph{alone}.
+
+Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
+Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee (2006)), where boundaries contained a higher proportion of Shelterbelts (Figure \ref{fig:boundaryTypes}), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
+Between-year effects may be important, especially during dry years (Kort 1988), but this was not considered in our analysis.
+Future work could examine whether boundaries have differing effects during hot or cool years.
+
+We used harvest time as smoothed variable to deal with temporal autocorrelation in the data, but this is a crude way to deal with this.
+This is because in-field combine driving patterns \emph{almost always} induce correlation between spatial and temporal measurements (e.g.~northern section of the field was harvested before the southern section), meaning that temporal and spatial random effects are difficult to separate.
+There are few solutions to this using harvest data as-is, other than to ignore temporal random effects entirely, thereby constraining all random variation to be spatial.
+However, it is possible to remove correlation between spatial and temporal measurements by using specially designed harvest patterns within a field (harvesting alternating strips sequentially across the field, then harvesting the remaining strips on a second pass).
+Future studies could use this to estimate drift in combine sensor measurements over time.
 
@@
 
+\leavevmode\hypertarget{ref-baker2018}{}%
+Baker, T.P., Moroni, M.T., Mendham, D.S., Smith, R. \& Hunt, M.A. (2018). Impacts of windbreak shelter on crop and livestock production. \emph{Crop and Pasture Science}, 69, 785.
+
+\leavevmode\hypertarget{ref-bonifacio2011}{}%
+Bonifacio, R.S., Kinross, C.M., Gurr, G.M.
+\& Nicol, H.
+(2011).
+The effect of woody plant diversity and other stand and landscape factors on the diversity and abundance of birds using farm shelterbelts.
+\emph{Pacific Conservation Biology}, 17, 22.
+
+\leavevmode\hypertarget{ref-boughey2011}{}%
+Boughey, K.L., Lake, I.R., Haysom, K.A.
+\& Dolman, P.M.
+(2011).
+Improving the biodiversity benefits of hedgerows: How physical characteristics and the proximity of foraging habitat affect the use of linear features by bats.
+\emph{Biological Conservation}, 144, 1790--1798.
+
+\leavevmode\hypertarget{ref-brandle2004}{}%
+Brandle, J.R., Hodges, L.
+\& Zhou, X.H.
+(2004).
+Windbreaks in {North American} agricultural systems.
+In: \emph{New vistas in agroforestry: A compendium for 1st world congress of agroforestry, 2004} (eds.
+Nair, P.K.R., Rao, M.R.
+\& Buck, L.E.).
+Springer Netherlands, Dordrecht, pp.
+65--78.
+
+\leavevmode\hypertarget{ref-case2019}{}%
+Case, B.S., Pannell, J.L., Stanley, M.C., Norton, D.A., Brugman, A., Funaki, M., \emph{et al.} (2019).
+Non-production vegetation has a positive effect on ecological processes in agroecosystems.
+\emph{bioRxiv}, 624635.
+
+\leavevmode\hypertarget{ref-fahrig2019}{}%
+Fahrig, L., Arroyo-Rodríguez, V., Bennett, J.R., Boucher-Lalonde, V., Cazetta, E., Currie, D.J., \emph{et al.} (2019).
+Is habitat fragmentation bad for biodiversity?
+\emph{Biological Conservation}, 230, 179--186.
+
+\leavevmode\hypertarget{ref-fahrig2011}{}%
+Fahrig, L., Baudry, J., Brotons, L., Burel, F.G., Crist, T.O., Fuller, R.J., \emph{et al.} (2011).
+Functional landscape heterogeneity and animal biodiversity in agricultural landscapes.
+\emph{Ecology Letters}, 14, 101--112.
+
 \leavevmode\hypertarget{ref-gardner2021}{}%
@@
 \emph{Journal of Applied Ecology}, In press.
+
+\leavevmode\hypertarget{ref-gray2014}{}%
+Gray, C.L. \& Lewis, O.T. (2014). Do riparian forest fragments provide ecosystem services or disservices in surrounding oil palm plantations? \emph{Basic and Applied Ecology}, 15, 693--700.
 
@@
 
+\leavevmode\hypertarget{ref-lecq2018}{}%
+Lecq, S., Loisel, A., Mullin, S.J.
+\& Bonnet, X.
+(2018).
+Manipulating hedgerow quality: Embankment size influences animal biodiversity in a peri-urban context.
+\emph{Urban Forestry {\&} Urban Greening}, 35, 1--7.
+
+\leavevmode\hypertarget{ref-lindstrom2016}{}%
+Lindström, S.A.M., Herbertsson, L., Rundlöf, M., Smith, H.G.
+\& Bommarco, R.
+(2016).
+Large-scale pollination experiment demonstrates the importance of insect pollination in winter oilseed rape.
+\emph{Oecologia}, 180, 759--769.
+
+\leavevmode\hypertarget{ref-ABRegions2006}{}%
+Natural Regions Committee. (2006). \emph{Natural regions and subregions of {Alberta}}. Government of Alberta, Edmonton.
+
+\leavevmode\hypertarget{ref-oberg2007}{}%
+Öberg, S., Ekbom, B.
+\& Bommarco, R.
+(2007).
+Influence of habitat type and surrounding landscape on spider diversity in {Swedish} agroecosystems.
+\emph{Agriculture, Ecosystems {\&} Environment}, 122, 211--219.
+
+\leavevmode\hypertarget{ref-quinn2017}{}%
+Quinn, N.F., Brainard, D.C. \& Szendrei, Z. (2017). Floral strips attract beneficial insects but do not enhance yield in cucumber fields. \emph{Journal of Economic Entomology}, 110, 517--524.
+
 \leavevmode\hypertarget{ref-redhead2020}{}%
-Redhead, J.W., Oliver, T.H., Woodcock, B.A. \& Pywell, R.F. (2020). The influence of landscape composition and configuration on crop yield resilience. \emph{Journal of Applied Ecology}.
+Redhead, J.W., Oliver, T.H., Woodcock, B.A.
+\& Pywell, R.F.
+(2020).
+The influence of landscape composition and configuration on crop yield resilience.
+\emph{Journal of Applied Ecology}, 57, 2180--2190.
+
+\leavevmode\hypertarget{ref-stasiov2020}{}%
+Stašiov, S., Diviaková, A., Svitok, M., Novikmec, M.
+\& Dovciak, M.
+(2020).
+Hedgerows support rich communities of harvestmen ({Opiliones}) in upland agricultural landscape.
+\emph{Basic and Applied Ecology}, 47, 73--82.
+
+\leavevmode\hypertarget{ref-sutter2018b}{}%
+Sutter, L., Amato, M., Jeanneret, P.
+\& Albrecht, M.
+(2018).
+Overwintering of pollen beetles and their predators in oilseed rape and semi-natural habitats.
+\emph{Agriculture, Ecosystems {\&} Environment}, 265, 275--281.
 
@@
 
+\leavevmode\hypertarget{ref-weninger2021}{}%
+Weninger, T., Scheper, S., Lackóová, L., Kitzler, B., Gartner, K., King, N.W., \emph{et al.} (2021).
+Ecosystem services of tree windbreaks in rural landscapes -- a systematic review.
+\emph{Environmental Research Letters}, In press.
+
 \leavevmode\hypertarget{ref-wickham2016}{}%
@@
 Wood, S.N. (2017). \emph{Generalized additive models: An introduction with {R}}. CRC press.
+
+\leavevmode\hypertarget{ref-woodcock2016}{}%
+Woodcock, B.A., Bullock, J.M., McCracken, M., Chapman, R.E., Ball, S.L., Edwards, M.E., \emph{et al.} (2016).
+Spill-over of pest control and pollination services into arable crops.
+\emph{Agriculture, Ecosystems \& Environment}, 231, 15--23.
+
+\leavevmode\hypertarget{ref-zhang2007}{}%
+Zhang, W., Ricketts, T.H., Kremen, C., Carney, K. \& Swinton, S.M. (2007). Ecosystem services and dis-services to agriculture. \emph{Ecological Economics}, 64, 253--260.
 
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/boundaryTypes-1} \caption{Percentage of field boundaries across sampled fields.}\label{fig:boundaryTypes}
-\end{figure}
-
-\begin{figure}
 \includegraphics[width=1\linewidth]{../Figures/groundSpeed_canola} \caption{Effect of ground speed and order (sequence of harvest) on canola yield.}\label{fig:canolaPlotSpeed}
@@
 
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/groundSpeed_peas} \caption{Effect of ground speed and order (sequence of harvest) on pea yield.}\label{fig:peasPlotSpeed}
+\end{figure}
+
 \end{document}

### 2021-09-08 Manuscript/supplemental.Rmd
Worked on discussion

--- 2021-08-24 Manuscript/supplemental.Rmd
@@
 
-```{r , echo=FALSE, fig.cap='Effect of polygon area and harvest sequence on canola yield. Polygon area is equivalent to ground ', out.width='100%'}  
-  include_graphics('../Figures/ModelSummary3b_canola.png',dpi=NA)
+```{r canolaPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on canola yield.', out.width='100%'}  
+  include_graphics('../Figures/groundSpeed_canola.png',dpi=NA)
 ```
 
-```{r , echo=FALSE, fig.cap='Temp', out.width='100%'}  
-  include_graphics('../Figures/ModelSummary3b_wheat.png',dpi=NA)
+```{r wheatPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on wheat yield.', out.width='100%'}  
+  include_graphics('../Figures/groundSpeed_wheat.png',dpi=NA)
 ```
+
+```{r peasPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on pea yield.', out.width='100%'}  
+  include_graphics('../Figures/groundSpeed_peas.png',dpi=NA)
+```

### 2021-09-08 runModels.R

@@ -524,8 +524,22 @@ samplePreds <- function(a=NULL,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutof
 # stopCluster(cluster)
 # save(samp,file='./Data/postSamples_wheat.Rdata')
 
-croptype <- 'canola'
+# # Add to current samples - peas
+# isPeas <- which(datSource$crop=='Peas') #Peas
+# Nsamp <- 500 #Number of samples
+# library(parallel)
+# cluster <- makeCluster(15)
+# clusterExport(cluster,c('datSource'))
+# a <- Sys.time()
+# samp <- parLapply(cl=cluster,1:Nsamp,fun=samplePreds,useRows=isPeas,margInt=c(FALSE,FALSE,FALSE))
+# Sys.time()-a
+# stopCluster(cluster)
+# save(samp,file='./Data/postSamples_peas.Rdata')
+
+# croptype <- 'canola'
 # croptype <- 'wheat'
+croptype <- 'peas'
+
 
 load(paste0('./Data/postSamples_',croptype,'.Rdata'))
 
@@ -554,9 +568,12 @@ ylabSD <- 'Yield variablity (log(T/ha)) '
 if(croptype=='wheat'){
   ylimsMean <- c(5,13)
   ylimsSD <- c(-4,2)
-} else {
+} else if(croptype=='canola'){
   ylimsMean <- c(1,10)
   ylimsSD <- c(-4,4)
+} else if(croptype=='peas'){
+  ylimsMean <- c(5,12)
+  ylimsSD <- c(-5,5)
 }
 
 distLims <- c(0,200)
@@ -597,7 +614,6 @@ p3 <- lapply(samp,function(x) x$coverDist %>% bind_rows(.id = 'dist_type')) %>%
   # geom_line(data=allEff[[2]],aes(x=dist,y=mean,group=field),alpha=0.1) + #"Raw" smoothers
   geom_ribbon(aes(ymax=upr,ymin=lwr),alpha=0.3,fill=fillshade) + 
   geom_line(aes(y=med),col=colshade) + #Meta-model
-  # geom_hline(yintercept = 0,col='red',linetype='dashed')+
   facet_wrap(~dist_type,scales='free') + labs(x='Distance (m)',y=ylabMean)+ 
   coord_cartesian(xlim=distLims,ylim=ylimsMean)
 
[…]

### 2021-09-08 yield simulations.R

@@ -30,7 +30,7 @@ text(5,0.9,'Bugs',col='blue'); text(80,0.9,'Environment',col='red'); text(80,0.6
 
 #Scenario 2 - beneficial bugs increase with distance into field
 curve(expFun1(x,a=1,b=3)*(expFun2(x,a=0.5,b=-100)+0.5),0,100,xlim=range(dist),
-      lwd=4,ylab='Yield',main='Scenario 1')
+      lwd=4,ylab='Yield',main='Scenario 2')
 curve(expFun1(x,a=1,b=3),0,100,ylim=c(0,1),xlim=range(dist),col='red',add=TRUE,lwd=2)
 curve(expFun2(x,a=0.5,b=-100)+0.5,0,100,ylim=c(0,1),xlim=range(dist),col='blue',add=TRUE,lwd=2)
 text(5,0.9,'Bugs',col='blue'); text(80,0.9,'Environment',col='red'); text(80,0.6,'Yield')
@@ -60,15 +60,15 @@ par(mfrow=c(1,2))
 #Scenario 1 - beneficial bugs decrease with distance into field
 a1 <- -2; b1 <- 0.2; a2 <- 3; b2 <- -0.03
 curve(leq2(x,a1,b1,a2,b2),0,100,ylim=c(0,1),
-      lwd=4,ylab='Yield',xlab='Distance from Edge')#,main='Scenario 1: bugs - with dist')
+      lwd=4,ylab='Yield',xlab='Distance from Edge',main='Scenario 1')#,main='Scenario 1: bugs - with dist')
 curve(leq(x,a1,b1),0,100,xlim=range(dist),col='red',add=TRUE,lwd=2)
 curve(leq(x,a2,b2),0,100,xlim=range(dist),col='blue',add=TRUE,lwd=2)
-text(5,0.85,'Ecosystem\nservices',col='blue'); text(80,0.9,'Environment',col='red'); text(80,0.6,'Yield')
+text(5,0.85,'Ecosystem\nservices',col='blue'); text(80,0.9,'Environmental\nEffect',col='red'); text(80,0.6,'Yield')
 
 #Scenario 2 - beneficial bugs increase with distance into field (or pest insects at edge)
 a2 <- 0; b2 <- 0.03
 curve(leq2(x,a1,b1,a2,b2),0,100,ylim=c(0,1),
[…]

### 2021-09-09 Manuscript/manuscript.Rmd
Worked on discussion, added refs

@@
 
-SNL in and around crops is important for both agricultural production and conservation, as it has the potential to create better biotic and abiotic conditions [@case2019; @weninger2021].
+SNL in and around crops is important for both agricultural production and conservation [@prietoBenitez2011], as it has the potential to create better biotic and abiotic conditions for crop growth [@case2019; @weninger2021; but see @lowe2021].
 They are habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control [@albrecht2020; @gardner2021], but this can vary depending on the system [@gray2014; @quinn2017]
-They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed [@kort1988; @brandle2004]
-Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
-In particular, North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL
+They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed [@kort1988; @brandle2004].
+North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL
 
@@
 At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (@kowalchuk1995).
-For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance.
+For example, pollination services from central place foragers (bees) that nest in SNL but forage in crops drops rapidly with distance [@garibaldi2011].
 Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (e.g. @vanVooren2017).
 This suggests a "goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
-<!-- - Less-considered are ecosystem ``disservices", where weeds and pests  -->
-<!-- - In North America, there are financial incentives to make fields large, homogeneous, and with simple geometry -->
+<!-- In North America there are financial incentives to make fields large, homogeneous, and geometrically simple [@clare2021], but  -->
 
@@
 
-Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality (@baker2018; but see also @redhead2020)
-Meta-analyses have been the standard solution to this problem (e.g. @albrecht2020), but precision yield data holds enormous promise.
+Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality (@baker2018; but see also @redhead2020).
+Meta-analyses have been the standard solution to this problem (e.g. @albrecht2020; @lowe2021), but precision yield data holds enormous promise.
 However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
-Ecosystem services can influence both the mean and variability of yield in agroecosystems (@redhead2020); typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from the perspective of a grower.
+Ecosystem services can influence both the mean and variability of yield in agroecosystems (@redhead2020); typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from a grower's perspective.
 There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
@@
 Both metrics also varied with ground speed and harvest sequence, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
-Yield tended to increase with distance from field boundaries before plateauing, and there was little evidence of intermediate boosts in yield, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \@ref(fig:hypotheses)).
-However, the effect of field boundary types was inconsistent, and the change tended to be small, especially for wheat.
+Yield tended to increase with distance from field boundaries before plateauing, but there was little evidence of intermediate boosts in yield, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \@ref(fig:hypotheses)).
+Some crops displayed a larger decrease in yield next to field boundaries; for example, pea yield directly adjacent to Wetland was lower overall, but increased with distance more than other crop types, indicating that Wetlands may locally decrease pea yields but are beneficial to the crop in general.
+Canola yields next to Other Crops had a reduced edge effect compared to Standard field boundaries, indicating that Other Crops may provide a more consistent microclimate for canola than grassy field boundaries.
+However, the difference between boundary types tended to be both small and variable, especially for wheat.
 <!-- For example, average canola yield was about 0.5 T/ha lower close to Standard field edges, and yield SD was about 0.3 T/ha SD higher. -->
-So, why did we find few consistent effects of boundary type on canola and wheat yield?
+
+Why did we find few consistent effects of boundary type on crop yield?
 We consider a few potential reasons for this below:
@@
 bumblebee foraging) or disperse very broadly [@oberg2007].
-There could also be interactions between landscape composition at different scales (e.g. our paper in AGEE), but this is not well-studied.
+There could also be interactions between landscape composition at different scales [@motzke2016; @robinson2021], but this is not well-studied.
 Landscape complexity [@fahrig2011] and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others).
 This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat (but see @fahrig2019).
+Our definitions of boundary type were also (necessarily) loose, as we used satellite imagery to manually classify them.
+However, it may be that other boundary traits, such as size or plant species composition, are more important for spillover of ecosystems services into fields.
+For example, larger [@lecq2018; @stasiov2020] or more functionally diverse (@bonifacio2011; @boughey2011) hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction matters to predatory beetle communities [@sutter2018b].
+These fine-scale boundary characteristics were not measured in our study, but could be used as variables in more biologically-meaningful classifications of field boundaries.
 
-Our definitions of boundary type were necessarily loose, as we used satellite imagery to manually classify them.
-However, it may be that other boundary traits, such as size or plant species composition, are more important for spillover of ecosystems services into fields.
-For example, larger [@lecq2018; @stasiov2020] or more functionally diverse (@bonifacio2011; @boughey2011) hedgerows tend to support more biodiversity than smaller hedgerows.
-Increased litter cover and soil compaction increased pollen beetle predators [@sutter2018b].
-
-Finally, pest pressure or pollination deficit may be at low enough levels that ecosystem service provision does not significantly change yield.
-Canola is largely self-pollinating (at least in hybrid canola varieties, @lindstrom2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop [@woodcock2016].
-Pest pressures may also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty.
-
-The net flow of ecosystem services may not be positive, if beneficial organisms do not disperse from the SNL, or if pests disperse along with them.
+Our first scenario (Scenario 1, Figure \@ref(fig:hypotheses)) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
+However, field boundaries may not have a positive effect if a) pests (ecosystem dis-services) also disperse along with beneficials, b) if beneficials do not disperse from the boundary environment, or if c) the effect of beneficial organisms is small.
 Ecosystem disservices such as pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (@zhang2007).
@@
 However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data _alone_.
+Ecosystem services from field boundaries may be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales (@oberg2007; @zamorano2020, but see @robinson2021).
+Finally, pest pressure or pollination deficit may be low enough that ecosystem service provision does not significantly change yield.
+Canola is largely self-pollinating (at least in hybrid canola varieties, @lindstrom2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop [@woodcock2016].
+Pests are also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty.
 

### 2021-09-09 Manuscript/abstract.Rmd
Added abstract

Field boundaries can improve crop yields by creating better abiotic conditions for crop growth (e.g. mosture, temperature), and can act as refuges for beneficial arthropods.
This suggests that beneficial crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services (e.g. pollination or pest suppression) from the field boundary.
However, there is little large-scale evidence showing this, largely because yield is costly and time-consuming to measure.
Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is relatively cheap and commonly measured by growers.
<!-- In this study, we used 252 field-years of yield monitor data from three common crops -- wheat (_Triticum aestivum_), canola (_Brassica napus_), or peas (_Pisum sativum_) -- recorded across Alberta, Canada, and examined how yield varied with distances from common crop boundary types. -->
Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease concurrently with distance.
Differences between boundary types tended to be weak and variable, especially for wheat; there was some evidence of certain field boundaries causing a greater decrease in yield at the edge, but these differences tended to be small in magnitude (<5% of yield).
Furthermore, there was little evidence of an intermediate increase in yield, suggesting that a) ecosystem services do not strongly depend on distance or b) their effect is relatively small.
This study represents one of the first uses of precision yield data to measure ecosystem service provision at large spatial scales.

### 2021-09-09 Manuscript/manuscript.Rmd
Added abstract

@@
 
+Potential journals: Agriculture Ecosystems and Environment, Basic and Applied Ecology, Ecological Economics, Journal of Applied Ecology (probably won't like it)
+
+Potential reviewers: Teja Tscharntke, Ignasi Bartomeus, Ben Woodcock, John Redhead, some Canadian/Albertan reviewers who know the crop systems?
+
 \newpage
@@
 Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds.
-At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (@kowalchuk1995).
+At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges [@kowalchuk1995].
 For example, pollination services from central place foragers (bees) that nest in SNL but forage in crops drops rapidly with distance [@garibaldi2011].
-Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (e.g. @vanVooren2017).
+Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects [@vanVooren2017].
 This suggests a "goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
@@
 
-Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality (@baker2018; but see also @redhead2020).
-Meta-analyses have been the standard solution to this problem (e.g. @albrecht2020; @lowe2021), but precision yield data holds enormous promise.
+Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality [@baker2018; but see also @redhead2020].
+Meta-analyses have been the standard solution to this problem [e.g. @albrecht2020; @lowe2021], but precision yield data holds enormous promise.
 However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
-Ecosystem services can influence both the mean and variability of yield in agroecosystems (@redhead2020); typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from a grower's perspective.
+Ecosystem services can influence both the mean and variability of yield in agroecosystems [@redhead2020]; typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from a grower's perspective.
 There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
@@
 We converted raw data to a standard csv format using Ag Leader SMS.
-`r round(100*sum(grepl('(Wheat|Canola)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_) or canola (_Brassica napus_), two of the most common crops in rotation in Alberta.
-The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`), canola (`r sum(datSourceFinal$crop=='Canola')`), or peas (`r sum(datSourceFinal$crop=='Peas')`).
+`r round(100*sum(grepl('(Wheat|Canola|Peas)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_), canola (_Brassica napus_), or peas (_Pisum sativum_), three of the most common crops in rotation in Alberta.
+The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`), canola (`r sum(datSourceFinal$crop=='Canola')`), or peas (`r sum(datSourceFinal$crop=='Peas')`), for a total of `r nrow(datSourceFinal)` field-years of data.
 Individual fields contained between `r min(yearsPerField)` and `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`), containing a total of `r round(sum(datSourceFinal$npoints)/1e6,1)` million data points.
@@
 Landscape complexity [@fahrig2011] and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others).
-This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat (but see @fahrig2019).
+This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat [but see @fahrig2019].
 Our definitions of boundary type were also (necessarily) loose, as we used satellite imagery to manually classify them.
 However, it may be that other boundary traits, such as size or plant species composition, are more important for spillover of ecosystems services into fields.
-For example, larger [@lecq2018; @stasiov2020] or more functionally diverse (@bonifacio2011; @boughey2011) hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction matters to predatory beetle communities [@sutter2018b].
+For example, larger [@lecq2018; @stasiov2020] or more functionally diverse [@bonifacio2011; @boughey2011] hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction matters to predatory beetle communities [@sutter2018b].
 These fine-scale boundary characteristics were not measured in our study, but could be used as variables in more biologically-meaningful classifications of field boundaries.
@@
 However, field boundaries may not have a positive effect if a) pests (ecosystem dis-services) also disperse along with beneficials, b) if beneficials do not disperse from the boundary environment, or if c) the effect of beneficial organisms is small.
-Ecosystem disservices such as pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (@zhang2007).
+Ecosystem disservices such as pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services [@zhang2007].
 @gray2014 showed that remnant forest fragments do not provide significant pest control services , but also do not depress yield.
 However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data _alone_.
-Ecosystem services from field boundaries may be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales (@oberg2007; @zamorano2020, but see @robinson2021).
+Ecosystem services from field boundaries may be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales [@oberg2007; @zamorano2020, but see @robinson2021].
 Finally, pest pressure or pollination deficit may be low enough that ecosystem service provision does not significantly change yield.
-Canola is largely self-pollinating (at least in hybrid canola varieties, @lindstrom2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop [@woodcock2016].
+Canola is largely self-pollinating [at least in hybrid canola varieties, @lindstrom2016] and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop [@woodcock2016].
 Pests are also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty.
@@
 Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
-Most fields were located in Central Alberta (Dry Mixedwood region, @ABRegions2006), where boundaries contained a higher proportion of Shelterbelts (Figure \@ref(fig:boundaryTypes)), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
+Most fields were located in Central Alberta [Dry Mixedwood region, @ABRegions2006], where boundaries contained a higher proportion of Shelterbelts (Figure \@ref(fig:boundaryTypes)), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
 Between-year effects may be important, especially during dry years [@kort1988], but this was not considered in our analysis.

### 2021-09-09 Manuscript/manuscript.tex
Added abstract

@@
   \begin{abstract}
-  Abstract goes here
+  Field boundaries can improve crop yields by creating better abiotic conditions for crop growth (e.g.~mosture, temperature), and can act as refuges for beneficial arthropods.
+This suggests that beneficial crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services (e.g.~pollination or pest suppression) from the field boundary.
+However, there is little large-scale evidence showing this, largely because yield is costly and time-consuming to measure.
+Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is relatively cheap and commonly measured by growers.
+Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease concurrently with distance.
+Differences between boundary types tended to be weak and variable, especially for wheat; there was some evidence of certain field boundaries causing a greater decrease in yield at the edge, but these differences tended to be small in magnitude (\textless5\% of yield).
+Furthermore, there was little evidence of an intermediate increase in yield, suggesting that a) ecosystem services do not strongly depend on distance or b) their effect is relatively small.
+This study represents one of the first uses of precision yield data to measure ecosystem service provision at large spatial scales.
   \end{abstract}
@@
  \end{frontmatter}
+
+Potential journals: Agriculture Ecosystems and Environment, Basic and Applied Ecology, Ecological Economics, Journal of Applied Ecology (probably won't like it)
+
+Potential reviewers: Teja Tscharntke, Ignasi Bartomeus, Ben Woodcock, John Redhead, some Canadian/Albertan reviewers who know the crop systems?
 
@@
 
-SNL in and around crops is important for both agricultural production and conservation, as it has the potential to create better biotic and abiotic conditions (Case \emph{et al.} 2019; Weninger \emph{et al.} 2021).
+SNL in and around crops is important for both agricultural production and conservation (Prieto-Benítez \& Méndez 2011), as it has the potential to create better biotic and abiotic conditions for crop growth (Case \emph{et al.} 2019; Weninger \emph{et al.} 2021; but see Lowe \emph{et al.} 2021).
 They are habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this can vary depending on the system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017)
-They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (Kort 1988; Brandle \emph{et al.} 2004)
-Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
-In particular, North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL
+They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (Kort 1988; Brandle \emph{et al.} 2004).
+North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL
 
@@
 Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds.
-At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (Kowalchuk \& Jong (1995)).
-For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance.
-Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (e.g. Van Vooren \emph{et al.} (2017)).
+At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (Kowalchuk \& Jong 1995).
+For example, pollination services from central place foragers (bees) that nest in SNL but forage in crops drops rapidly with distance (Garibaldi \emph{et al.} 2011).
+Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (Van Vooren \emph{et al.} 2017).
 This suggests a ``goldilocks'' field size, where negative edge effects are canceled out by ecosystem services.
@@
 
-Studies of SNL effects on crop yield also suffer from limited scope (e.g.~few crop types) and small sample sizes, limiting inference and reducing generality (Baker \emph{et al.} (2018); but see also Redhead \emph{et al.} (2020))
-Meta-analyses have been the standard solution to this problem (e.g. Albrecht \emph{et al.} (2020)), but precision yield data holds enormous promise.
+Studies of SNL effects on crop yield also suffer from limited scope (e.g.~few crop types) and small sample sizes, limiting inference and reducing generality (Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
+Meta-analyses have been the standard solution to this problem (e.g. Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise.
 However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
-Ecosystem services can influence both the mean and variability of yield in agroecosystems (Redhead \emph{et al.} (2020)); typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from the perspective of a grower.
+Ecosystem services can influence both the mean and variability of yield in agroecosystems (Redhead \emph{et al.} 2020); typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from a grower's perspective.
 There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
@@
 We converted raw data to a standard csv format using Ag Leader SMS.
-72\% of the crop types where either wheat (\emph{Triticum aestivum}) or canola (\emph{Brassica napus}), two of the most common crops in rotation in Alberta.
-The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39).
+85\% of the crop types where either wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}), three of the most common crops in rotation in Alberta.
+The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
 Individual fields contained between 1 and 6 years of data (mean: 3.1), containing a total of 14.4 million data points.
@@
 Both metrics also varied with ground speed and harvest sequence, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
-Yield tended to increase with distance from field boundaries before plateauing, and there was little evidence of intermediate boosts in yield, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \ref{fig:hypotheses}).
-However, the effect of field boundary types was inconsistent, and the change tended to be small, especially for wheat.
-So, why did we find few consistent effects of boundary type on canola and wheat yield?
+Yield tended to increase with distance from field boundaries before plateauing, but there was little evidence of intermediate boosts in yield, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \ref{fig:hypotheses}).
+Some crops displayed a larger decrease in yield next to field boundaries; for example, pea yield directly adjacent to Wetland was lower overall, but increased with distance more than other crop types, indicating that Wetlands may locally decrease pea yields but are beneficial to the crop in general.
+Canola yields next to Other Crops had a reduced edge effect compared to Standard field boundaries, indicating that Other Crops may provide a more consistent microclimate for canola than grassy field boundaries.
+However, the difference between boundary types tended to be both small and variable, especially for wheat.
+
+Why did we find few consistent effects of boundary type on crop yield?
 We consider a few potential reasons for this below:
@@
 Some studies show the influence of landscape composition at larger spatial scales (e.g.~Lan's paper) which could be particularly important when beneficial arthropods are highly mobile (e.g.~bumblebee foraging) or disperse very broadly (Öberg \emph{et al.} 2007).
-There could also be interactions between landscape composition at different scales (e.g.~our paper in AGEE), but this is not well-studied.
+There could also be interactions between landscape composition at different scales (Motzke \emph{et al.} 2016; Robinson \emph{et al.} 2021), but this is not well-studied.
 Landscape complexity (Fahrig \emph{et al.} 2011) and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others).
-This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat (but see Fahrig \emph{et al.} (2019)).
+This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat (but see Fahrig \emph{et al.} 2019).
+Our definitions of boundary type were also (necessarily) loose, as we used satellite imagery to manually classify them.
+However, it may be that other boundary traits, such as size or plant species composition, are more important for spillover of ecosystems services into fields.
+For example, larger (Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020) or more functionally diverse (Bonifacio \emph{et al.} 2011; Boughey \emph{et al.} 2011) hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction matters to predatory beetle communities (Sutter \emph{et al.} 2018).
+These fine-scale boundary characteristics were not measured in our study, but could be used as variables in more biologically-meaningful classifications of field boundaries.
 
-Our definitions of boundary type were necessarily loose, as we used satellite imagery to manually classify them.
-However, it may be that other boundary traits, such as size or plant species composition, are more important for spillover of ecosystems services into fields.
-For example, larger (Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020) or more functionally diverse (Bonifacio \emph{et al.} (2011); Boughey \emph{et al.} (2011)) hedgerows tend to support more biodiversity than smaller hedgerows.
-Increased litter cover and soil compaction increased pollen beetle predators (Sutter \emph{et al.} 2018).
-
-Finally, pest pressure or pollination deficit may be at low enough levels that ecosystem service provision does not significantly change yield.
-Canola is largely self-pollinating (at least in hybrid canola varieties, Lindström \emph{et al.} (2016)) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop (Woodcock \emph{et al.} 2016).
-Pest pressures may also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty.
-
-The net flow of ecosystem services may not be positive, if beneficial organisms do not disperse from the SNL, or if pests disperse along with them.
-Ecosystem disservices such as pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang \emph{et al.} (2007)).
+Our first scenario (Scenario 1, Figure \ref{fig:hypotheses}) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
+However, field boundaries may not have a positive effect if a) pests (ecosystem dis-services) also disperse along with beneficials, b) if beneficials do not disperse from the boundary environment, or if c) the effect of beneficial organisms is small.
+Ecosystem disservices such as pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang \emph{et al.} 2007).
 Gray \& Lewis (2014) showed that remnant forest fragments do not provide significant pest control services , but also do not depress yield.
 However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data \emph{alone}.
+Ecosystem services from field boundaries may be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales Robinson \emph{et al.} (2021).
+Finally, pest pressure or pollination deficit may be low enough that ecosystem service provision does not significantly change yield.
+Canola is largely self-pollinating (at least in hybrid canola varieties, Lindström \emph{et al.} 2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop (Woodcock \emph{et al.} 2016).
+Pests are also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty.
 
 Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
-Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee (2006)), where boundaries contained a higher proportion of Shelterbelts (Figure \ref{fig:boundaryTypes}), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
+Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee 2006), where boundaries contained a higher proportion of Shelterbelts (Figure \ref{fig:boundaryTypes}), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
 Between-year effects may be important, especially during dry years (Kort 1988), but this was not considered in our analysis.
@@
 
+\leavevmode\hypertarget{ref-garibaldi2011}{}%
+Garibaldi, L.A., Steffan-Dewenter, I., Kremen, C., Morales, J.M., Bommarco, R., Cunningham, S.A., \emph{et al.} (2011).
+Stability of pollination services decreases with isolation from natural areas despite honey bee visits.
+\emph{Ecology Letters}, 14, 1062--1072.
+
 \leavevmode\hypertarget{ref-gray2014}{}%
@@
 
+\leavevmode\hypertarget{ref-lowe2021}{}%
+Lowe, E.B., Groves, R.
+\& Gratton, C.
+(2021).
+Impacts of field-edge flower plantings on pollinator conservation and ecosystem service delivery {{}} a meta-analysis.
+\emph{Agriculture, Ecosystems {\&} Environment}, 310, 107290.
+
+\leavevmode\hypertarget{ref-motzke2016}{}%
+Motzke, I., Klein, A.-M., Saleh, S., Wanger, T.C.
+\& Tscharntke, T.
+(2016).
+Habitat management on multiple spatial scales can enhance bee pollination and crop yield in tropical homegardens.
+\emph{Agriculture, Ecosystems \& Environment}, 223, 144--151.
+
 \leavevmode\hypertarget{ref-ABRegions2006}{}%
@@
 
+\leavevmode\hypertarget{ref-prietoBenitez2011}{}%
+Prieto-Benítez, S. \& Méndez, M. (2011). Effects of land management on the abundance and richness of spiders ({Araneae}): A meta-analysis. \emph{Biological Conservation}, 144, 683--691.
+
 \leavevmode\hypertarget{ref-quinn2017}{}%
@@
 \emph{Journal of Applied Ecology}, 57, 2180--2190.
+
+\leavevmode\hypertarget{ref-robinson2021}{}%
+Robinson, S.V.J., Edwards, D., Vickruck, J.L., Best, L.R.
+\& Galpern, P.
+(2021).
+Non-crop sources of beneficial arthropods vary within-season across a prairie agroecosystem.
+\emph{Agriculture, Ecosystems {\&} Environment}, 320, 107581.
 
@@
 
+\leavevmode\hypertarget{ref-zamorano2020}{}%
+Zamorano, J., Bartomeus, I., Grez, A.A.
+\& Garibaldi, L.A.
+(2020).
+Field margin floral enhancements increase pollinator diversity at the field edge but show no consistent spillover into the crop field: A meta-analysis.
+\emph{Insect Conservation and Diversity}, 13, 519--531.
+
 \leavevmode\hypertarget{ref-zhang2007}{}%

### 2021-09-10 Manuscript/manuscript.Rmd
Added notes from meeting

@@
 
+<!-- Things to think about: -->
+<!-- - incorporate precision yield in title -->
+<!-- - “are we looking at the right thing?” -->
+<!-- - remove y axis from hypothesis figure – make it completely theoretical -->
+<!-- - setup: SNL is generally good for “good bugs”, but connection to spillover, and therefore ecosystem service + yield, is tenuous -->
+<!-- - Jess’ paper, Elena Benett’s paper (J of Appl Ecol) -->
+<!-- - Does scale beyond the field matter vs within the field? -->
+<!-- - The variables that moderate spillover and yield needs to be re-investigated -->
+<!-- - Under what circumstance might we expect this? -->
+<!-- - Emphasis: implications of natural areas for yield -->
+<!-- - Variable curves are “consistent with  -->
+
 \newpage
@@
 However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
-Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios.
-Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems.
+Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios.
 
-SNL in and around crops is important for both agricultural production and conservation [@prietoBenitez2011], as it has the potential to create better biotic and abiotic conditions for crop growth [@case2019; @weninger2021; but see @lowe2021].
-They are habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control [@albrecht2020; @gardner2021], but this can vary depending on the system [@gray2014; @quinn2017]
+Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.
+SNL in agroecosystems is important for both agricultural production and conservation [@prietoBenitez2011], and can create better biotic and abiotic conditions for crop growth [@case2019; @weninger2021; but see @lowe2021].
+SNL consitutes habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control [@albrecht2020; @gardner2021], but this can vary depending on the system [@gray2014; @quinn2017].
 They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed [@kort1988; @brandle2004].
-North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL
+North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL.
 
@@
 For example, pollination services from central place foragers (bees) that nest in SNL but forage in crops drops rapidly with distance [@garibaldi2011].
-Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects [@vanVooren2017].
-This suggests a "goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
+Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects [@vanVooren2017; Figure \@ref(fig:hypotheses)].
+This also suggests a "goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
 <!-- In North America there are financial incentives to make fields large, homogeneous, and geometrically simple [@clare2021], but  -->
@@
 Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality [@baker2018; but see also @redhead2020].
-Meta-analyses have been the standard solution to this problem [e.g. @albrecht2020; @lowe2021], but precision yield data holds enormous promise.
+Meta-analyses have been the standard solution to this problem [e.g.
+@albrecht2020; @lowe2021], but precision yield data holds enormous promise because of its common usage and relatively cheap operating costs.
 However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
@@
 We converted raw data to a standard csv format using Ag Leader SMS.
-`r round(100*sum(grepl('(Wheat|Canola|Peas)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_), canola (_Brassica napus_), or peas (_Pisum sativum_), three of the most common crops in rotation in Alberta.
+`r round(100*sum(grepl('(Wheat|Canola|Peas)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_), canola (_Brassica napus_), or peas (_Pisum sativum_), three of the most common crops in rotation in Alberta [@statscan_canola2014].
 The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`), canola (`r sum(datSourceFinal$crop=='Canola')`), or peas (`r sum(datSourceFinal$crop=='Peas')`), for a total of `r nrow(datSourceFinal)` field-years of data.
@@
 Sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time.
-This may be caused by sensors accumulating debris during harvest (especially in canola, pers. comm. Trent Clark), leading to changes in accuracy and bias over time.
+This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time.
 To model this, we fit the following model to each field-year of data:
@@
 
-```{r wheatPlot, echo=FALSE, fig.cap='Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
+```{r wheatPlot, echo=FALSE, fig.cap='Field boundary effect on wheat yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
 Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
@@
 At the field level, both yield average and yield variability (SD of yield) of yield were spatially (average p-values: `r with(peaMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(peaMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`) and temporally dependent (`r with(peaMods,formatC(meanP[Smoother=='s(r)'],3))`, `r with(peaMods,formatC(meanP[Smoother=='s.1(r)'],3))`),
-There was a strong negative effect of ground speed on yield and variability of peas (Figure \@ref(fig:peasPlotSpeed)), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \@ref(fig:wheatPlotSpeed), right panel).
+There was a strong negative effect of ground speed on yield and variability of peas (Figure \@ref(fig:peasPlotSpeed)), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \@ref(fig:peasPlotSpeed), right panel).
 
@@
 Interestingly, the reduction at the edge was slightly higher for Wetlands, but yield increased with distance more quickly, i.e. yields were higher at distance for Wetlands.
-Variability decreased with distance from Grassland and Standard boundaries
+Yield variability decreased with distance from Grassland and Standard boundaries, and was slightly lower overall at Standard boundaries.
 

### 2021-09-10 Manuscript/manuscript.tex
Added notes from meeting

@@
 However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
-Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios.
-Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems.
+Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios.
 
-SNL in and around crops is important for both agricultural production and conservation (Prieto-Benítez \& Méndez 2011), as it has the potential to create better biotic and abiotic conditions for crop growth (Case \emph{et al.} 2019; Weninger \emph{et al.} 2021; but see Lowe \emph{et al.} 2021).
-They are habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this can vary depending on the system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017)
+Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.
+SNL in agroecosystems is important for both agricultural production and conservation (Prieto-Benítez \& Méndez 2011), and can create better biotic and abiotic conditions for crop growth (Case \emph{et al.} 2019; Weninger \emph{et al.} 2021; but see Lowe \emph{et al.} 2021).
+SNL consitutes habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this can vary depending on the system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017).
 They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (Kort 1988; Brandle \emph{et al.} 2004).
-North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL
+North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL.
 
@@
 For example, pollination services from central place foragers (bees) that nest in SNL but forage in crops drops rapidly with distance (Garibaldi \emph{et al.} 2011).
-Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects (Van Vooren \emph{et al.} 2017).
-This suggests a ``goldilocks'' field size, where negative edge effects are canceled out by ecosystem services.
+Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects {[}Van Vooren \emph{et al.} (2017); Figure \ref{fig:hypotheses}{]}.
+This also suggests a ``goldilocks'' field size, where negative edge effects are canceled out by ecosystem services.
 
@@
 Studies of SNL effects on crop yield also suffer from limited scope (e.g.~few crop types) and small sample sizes, limiting inference and reducing generality (Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
-Meta-analyses have been the standard solution to this problem (e.g. Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise.
+Meta-analyses have been the standard solution to this problem (e.g.
+Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its common usage and relatively cheap operating costs.
 However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
@@
 We converted raw data to a standard csv format using Ag Leader SMS.
-85\% of the crop types where either wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}), three of the most common crops in rotation in Alberta.
+85\% of the crop types where either wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}), three of the most common crops in rotation in Alberta (Statistics Canada 2014).
 The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_wheat} \caption{Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
+\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_wheat} \caption{Field boundary effect on wheat yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
 Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
@@
 At the field level, both yield average and yield variability (SD of yield) of yield were spatially (average p-values: 2e-16, 2e-16) and temporally dependent (2e-16, 3.62e-07),
-There was a strong negative effect of ground speed on yield and variability of peas (Figure \ref{fig:peasPlotSpeed}), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \ref{fig:wheatPlotSpeed}, right panel).
+There was a strong negative effect of ground speed on yield and variability of peas (Figure \ref{fig:peasPlotSpeed}), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \ref{fig:peasPlotSpeed}, right panel).
 
@@
 Interestingly, the reduction at the edge was slightly higher for Wetlands, but yield increased with distance more quickly, i.e.~yields were higher at distance for Wetlands.
-Variability decreased with distance from Grassland and Standard boundaries
+Yield variability decreased with distance from Grassland and Standard boundaries, and was slightly lower overall at Standard boundaries.
 
@@
 
+\leavevmode\hypertarget{ref-statscan_canola2014}{}%
+Statistics Canada. (2014). Table 001-0010 - estimated areas, yield, production and average farm price of principal field crops, in metric units, annual.
+
 \leavevmode\hypertarget{ref-sutter2018b}{}%

### 2021-09-13 Manuscript/manuscript.Rmd
Worked on Intro and Discussion

@@
 ---
-title: "Living on the edge: crop boundary effects on canola and wheat yields in Alberta"
+title: "Living on the edge: crop boundary effects on precision canola and wheat yields in Alberta"
 author:
@@
 
-<!-- Things to think about: -->
+<!-- Writing notes: -->
+<!-- - SNL vs shelterbelt/field boundaries: they're related, but not the same thing -->
 <!-- - incorporate precision yield in title -->
@@
 <!-- - setup: SNL is generally good for “good bugs”, but connection to spillover, and therefore ecosystem service + yield, is tenuous -->
-<!-- - Jess’ paper, Elena Benett’s paper (J of Appl Ecol) -->
 <!-- - Does scale beyond the field matter vs within the field? -->
@@
 
+<!-- Things to do on main machine: -->
+<!-- - Update "Field Locations" map to only have canola, wheat, pea fields -->
+<!-- - Update other figures with more "realistic" estimates - what is the average combine ground speed? -->
+
 \newpage
@@
 
-Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth.
-This has allowed increases in human population and increased global stability in production.
+Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth [@ramankutty2018].
+This has allowed increases in human population and increased global stability in production [@foley2011].
 However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
-Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios.
+Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ["multifunctionality" _sensu_ @frei2018].
+Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.
 
-Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.
 SNL in agroecosystems is important for both agricultural production and conservation [@prietoBenitez2011], and can create better biotic and abiotic conditions for crop growth [@case2019; @weninger2021; but see @lowe2021].
-SNL consitutes habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control [@albrecht2020; @gardner2021], but this can vary depending on the system [@gray2014; @quinn2017].
-They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed [@kort1988; @brandle2004].
+SNL consitutes habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control [@albrecht2020; @gardner2021], but this can vary depending on the crop system [@gray2014; @quinn2017] and landscape context [@marja2019; @boetzl2020].
+They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion [@kort1988; @brandle2004; @wratten2012].
 North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL.
+Semi-natural field boundaries (e.g.
+grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are of particular interest, as they are extremely common and require less land than field-sized agricultural set-asides.
 
 <!-- - SNL has conflicting effects on crop yield because of ecosystem service provision as well as negative edge effects -->
-SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate.
-Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds.
-At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges [@kowalchuk1995].
-For example, pollination services from central place foragers (bees) that nest in SNL but forage in crops drops rapidly with distance [@garibaldi2011].
-Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects [@vanVooren2017; Figure \@ref(fig:hypotheses)].
-This also suggests a "goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
+Field boundaries may affect crop yield at intermediate distances, depending on the spatial scale at which ecosystem services operate, as they can induce positive and negative effects on crops.
+Field boundaries can cause low yields at the edge of crops because of patchy seed emergence, poor microclimate, shading, and greater competition with weeds.
+At the same time field boundaries can provide helpful ecosystem services, such as pest suppression or pollination, but these tend to be concentrated around the field edge.
+For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance [@kowalchuk1995; @garibaldi2011].
+Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects [@mitchell2014; @vanVooren2017; Figure \@ref(fig:hypotheses)].
+This suggests a "goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
 <!-- In North America there are financial incentives to make fields large, homogeneous, and geometrically simple [@clare2021], but  -->
@@
 
-Studies of SNL effects on crop yield also suffer from limited scope (e.g. few crop types) and small sample sizes, limiting inference and reducing generality [@baker2018; but see also @redhead2020].
-Meta-analyses have been the standard solution to this problem [e.g.
-@albrecht2020; @lowe2021], but precision yield data holds enormous promise because of its common usage and relatively cheap operating costs.
+Studies of semi-natural field boundaries on crop yield tend to suffer from limited spatial scope and small sample sizes, limiting inference and reducing generality [@baker2018; but see also @redhead2020].
+Meta-analyses are a common solution to this problem [e.g. @albrecht2020; @lowe2021], but precision yield data holds enormous promise because of its relatively common usage and low operating costs. 
 However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
-Ecosystem services can influence both the mean and variability of yield in agroecosystems [@redhead2020]; typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from a grower's perspective.
+Ecosystem services from field boundaries can influence both the mean and variability of yield in agroecosystems [@redhead2020]; typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from a grower's perspective.
 There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
 
-In this paper we ask:
+In this paper use a precision yield dataset from Alberta, Canada to ask the following questions:
   1. How does crop yield change with distance from the edge of field?
@@
   3. Is there an intermediate distance where yield is maximized or variance is minimized?
-To our knowledge, no other study has addressed the effect of field boundaries on productivity using precision yield, making this study highly novel.
+To our knowledge, no other study has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.
 
@@
 We used dry yield (tonnes of seed/hectare minus recorded crop moisture) as our measure of crop yield
-Precision yield data can be highly variable and is prone to extreme outliers (both positive and negative), especially when combine ground speeds are low and at the beginning of rows
+Precision yield data can be highly variable and is prone to extreme outliers (both positive and negative), especially when combine ground speeds are low and at the beginning of rows [@]
 Therefore, we filtered data that lay outside the 95th percentiles of yield within each field, values less than or equal to zero, and values outside of the 95th percentiles of ground speed.
@@
 
+Our work examined only the first and last links in the causal chain _Field boundary_ $\rightarrow$ _Insect Abundance_ $\rightarrow$ _Crop Yield_, but there are two places where this chain could be broken.
+First, field boundaries may not have any influence on insect abundance if separated from the larger landscape context.
+Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are low.
+However, our work does show that precision yield data could easily be joined to data on insect abundance to model this more completely.
+
 Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.

### 2021-09-16 Manuscript/manuscript.Rmd
Worked on Intro/Discussion

@@
 <!-- - Does scale beyond the field matter vs within the field? -->
-<!-- - The variables that moderate spillover and yield needs to be re-investigated -->
+<!-- - The variables that moderate spillover and yield need to be re-investigated -->
 <!-- - Under what circumstance might we expect this? -->
@@
 <!-- - SNL has conflicting effects on crop yield because of ecosystem service provision as well as negative edge effects -->
-Field boundaries may affect crop yield at intermediate distances, depending on the spatial scale at which ecosystem services operate, as they can induce positive and negative effects on crops.
-Field boundaries can cause low yields at the edge of crops because of patchy seed emergence, poor microclimate, shading, and greater competition with weeds.
-At the same time field boundaries can provide helpful ecosystem services, such as pest suppression or pollination, but these tend to be concentrated around the field edge.
+Field boundaries can cause both positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
+Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds.
+Field boundaries can also provide helpful ecosystem services, such as pest suppression or pollination, and these also tend to be occur close to the field edge [@duelli2003]. 
 For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance [@kowalchuk1995; @garibaldi2011].
-Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects [@mitchell2014; @vanVooren2017; Figure \@ref(fig:hypotheses)].
+If negative edge effects of the edge improve with distance and ecosystem services decrease with distance, yield may be maximized at intermediate distances where the ecosystem services cancel out negative edge effects [@mitchell2014; @vanVooren2017; Figure \@ref(fig:hypotheses)a].
 This suggests a "goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
+Alternatively, if ecosystem services are constant with distance (or increase, Figure \@ref(fig:hypotheses)b) then yield will increase with distance from a field boundary and then plateau.
+
 <!-- In North America there are financial incentives to make fields large, homogeneous, and geometrically simple [@clare2021], but  -->
@@
 
+Boundary effects are difficult to observe without large amounts of crop yield data, which is typically expensive and time-consuming to collect.    
 Studies of semi-natural field boundaries on crop yield tend to suffer from limited spatial scope and small sample sizes, limiting inference and reducing generality [@baker2018; but see also @redhead2020].
@@
 However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
-Ecosystem services from field boundaries can influence both the mean and variability of yield in agroecosystems [@redhead2020]; typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from a grower's perspective.
+Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems [@redhead2020]; typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
 There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
@@
 Why did we find few consistent effects of boundary type on crop yield?
-We consider a few potential reasons for this below:
-1. Landscape influence
-2. Boundary variables
-3. Ecosystem disservices
-4. Yearly/location variation
-5. Lack of pest pressure/pollination deficit
+Our work examined only the first and last links in the causal chain _Field boundary_ $\rightarrow$ _Insect Abundance_ $\rightarrow$ _Crop Yield_, meaning that are two mediating relationships where the overall relationship could be broken or obscured.
+First, field boundaries may not have any influence on insect abundance if separated from the larger landscape context, or spillover of beneficial organisms may be low.
+Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are low, which may vary yearly, or by location.
+<!-- However, our work does show that precision yield data could easily be joined to data on insect abundance to model this more completely. -->
+Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) when organisms in crops provide tangible ecosystem services.
+We consider these reasons below:
 
-Our analysis only examined boundary types, without considering the landscape context that fields were embedded in.
-Some studies show the influence of landscape composition at larger spatial scales (e.g.
-Lan's paper) which could be particularly important when beneficial arthropods are highly mobile (e.g.
-bumblebee foraging) or disperse very broadly [@oberg2007].
-There could also be interactions between landscape composition at different scales [@motzke2016; @robinson2021], but this is not well-studied.
+## Boundaries may not provide spillover
+
+SNL generally has a positive effect on spillover, but field boundaries only represent a fraction of the SNL that surrounds a given field.
+Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales (e.g.
+Lan's paper) which could be particularly important when beneficial arthropods are disperse very broadly [@oberg2007].
+There could also be interactions between landscape composition at different scales [@motzke2016; @robinson2021], but this is not broadly studied.
 Landscape complexity [@fahrig2011] and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others).
 This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat [but see @fahrig2019].
-Our definitions of boundary type were also (necessarily) loose, as we used satellite imagery to manually classify them.
-However, it may be that other boundary traits, such as size or plant species composition, are more important for spillover of ecosystems services into fields.
+
+Field boundary traits may play an important role in mediating spillover, but our definitions of boundaries were (necessarily) loose, as we used satellite imagery to manually classify them.
+Therefore, it may be that other aspects of field boundaries, such as size or plant community composition, are more important for spillover of ecosystems services into fields.
 For example, larger [@lecq2018; @stasiov2020] or more functionally diverse [@bonifacio2011; @boughey2011] hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction matters to predatory beetle communities [@sutter2018b].
@@
 Our first scenario (Scenario 1, Figure \@ref(fig:hypotheses)) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
-However, field boundaries may not have a positive effect if a) pests (ecosystem dis-services) also disperse along with beneficials, b) if beneficials do not disperse from the boundary environment, or if c) the effect of beneficial organisms is small.
+However, field boundaries may not have a positive effect if a) pests (ecosystem dis-services) also disperse along with beneficials, b) if beneficials do not disperse from the boundary environment
 Ecosystem disservices such as pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services [@zhang2007].
@@
 Ecosystem services from field boundaries may be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales [@oberg2007; @zamorano2020, but see @robinson2021].
-Finally, pest pressure or pollination deficit may be low enough that ecosystem service provision does not significantly change yield.
+
+## Beneficial organisms may not provide ecosystem services
+
+Pest pressure or pollination deficit may be low enough that ecosystem service provision does not significantly change yield.
 Canola is largely self-pollinating [at least in hybrid canola varieties, @lindstrom2016] and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop [@woodcock2016].
 Pests are also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty.
-
-Our work examined only the first and last links in the causal chain _Field boundary_ $\rightarrow$ _Insect Abundance_ $\rightarrow$ _Crop Yield_, but there are two places where this chain could be broken.
-First, field boundaries may not have any influence on insect abundance if separated from the larger landscape context.
-Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are low.
-However, our work does show that precision yield data could easily be joined to data on insect abundance to model this more completely.
 

### 2021-09-17 Manuscript/manuscript.Rmd
Worked on filter functions

@@
 We used dry yield (tonnes of seed/hectare minus recorded crop moisture) as our measure of crop yield
-Precision yield data can be highly variable and is prone to extreme outliers (both positive and negative), especially when combine ground speeds are low and at the beginning of rows [@]
-Therefore, we filtered data that lay outside the 95th percentiles of yield within each field, values less than or equal to zero, and values outside of the 95th percentiles of ground speed.
+Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows [@griffin2007; @whelan2013].
+We filtered data that lay outside the 95th percentiles of yield within each field, values less than or equal to zero, and values outside of the 95th percentiles of ground speed.
 Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time.

### 2021-09-17 filtering.R

@@ -19,20 +19,19 @@ datSource <- read.csv('./Data/datSource.csv') %>% #Read previous datasource file
 
 nSubSamp <- 50000 #Number of points to start with
 
-# fieldName <- 'Alvin_French C41 2020'
-fieldName <- 'Gibbons Bill Visser 2017'
-# fieldName <- 'Trent_Clark E 06 2018'
+fieldName <- 'Alvin_French C41 2020' #Canola - extremely messy - 1st filter seems better, but hard to tell
+# fieldName <- 'Gibbons Bill Visser 2017' #Peas - 2nd filter seems better
+# fieldName <- 'Trent_Clark E 06 2018' #Canola - chunks missing - 2nd filter seems better
+# fieldName <- 'Dean_Hubbard E_21_11_25 2018' #Wheat
+# fieldName <- 'Dean_Hubbard E_21_11_25 2020' #Wheat
 
 rownum <- which(datSource$filename==fieldName)
 csvPath <- datSource$dataPath[rownum] #Get file paths
 boundaryPath <- datSource$boundaryPath[rownum]
 boundaryPath2 <- datSource$boundaryPath2[rownum]
+(cropType <- datSource$crop[rownum])
 
 dat <- read.csv(csvPath,stringsAsFactors=TRUE,fileEncoding='latin1') 
-
-if(nrow(dat)>nSubSamp){
-  dat <- dat %>% slice(round(seq(1,nrow(dat),length.out=nSubSamp)))
-}
 
 dat <- dat %>% rename_with(.fn = ~gsub('..L.ha.$','_lHa',.x)) %>%
   rename_with(.fn = ~gsub('..tonne.ha.$','_tHa',.x)) %>%
@@ -46,19 +45,18 @@ dat <- dat %>% rename_with(.fn = ~gsub('..L.ha.$','_lHa',.x)) %>%
   rename('ID'='ObjId','DryYield'='YldMassDry_tHa','Lon'='Longitude','Lat'='Latitude','Pass'='PassNum','Speed'='Speed__m') %>%
   st_as_sf(coords=c('Lon','Lat')) %>% #Add spatial feature info
   st_set_crs(4326) %>% st_transform(3401) %>% #Lat-lon -> UTM
-  makePolys(width='SwthWdth_m',dist='Distance_m',angle='TrackAngle') %>%    
+  makePolys(width='SwthWdth_m',dist='Distance_m',angle='TrackAngle') %>%
[…]

### 2021-09-17 helperFunctions.R

@@ -3,6 +3,16 @@
 #Scales x between 0 and 1
 range01 <- function(x) (x-min(x))/(max(x)-min(x))
 
+#Difference in compass bearings (in degrees)
+bearingDiff <- function(x1,x2){
+  x <- x1-x2
+  x <- ifelse(abs(x)>180,x-(360*sign(x)),x) #Angle differences can't be >180
+  return(x)
+}
+# bearingDiff(0,15) #Should be the same as below
+# bearingDiff(355,10)
+# bearingDiff(NA,10)
+
 seqGroup <- function(x,singles=TRUE){ #Turns sequential IDs into numbered groups.
   #Singles = TRUE, sequential values must be exactly 1 greater (gaps matter)
   #Singles = FALSE, sequential values must be greater (gaps don't matter)
@@ -43,7 +53,7 @@ makePolys <- function(dat,width='w',dist='d',angle='a',backwards=FALSE){
     st_polygon(list(r))
   })
   #Combine dat with new polygon geometry, and add original CRS
-  dat2 <- st_sf(st_drop_geometry(dat),st_sfc(polys)) %>% st_set_crs(datCRS)
+  dat2 <- st_sf(st_drop_geometry(dat),geometry=st_sfc(polys)) %>% st_set_crs(datCRS)
   return(dat2)
 }
 

### 2021-09-17 prelim.R

@@ -10,7 +10,7 @@ library(sf)
 library(beepr)
 
 # datLoc <- "C:\\Users\\Samuel\\Documents\\Ag Leader Technology\\SMS\\Export\\Trent Clark\\2019\\SW02.csv" #Multivac
-datLoc <- "/media/rsamuel/Storage/geoData/Rasters/yieldData/csv files/Trent Clark/2019/SW 02.csv" #Galpern machine
+datLoc <- "/media/rsamuel/Storage/geoData/Rasters/yieldData/csv files/Trent_Clark SW_02 2019.csv" #Galpern machine
 
 #Name of field
 fieldName <- unlist(strsplit(datLoc,split='/'))
@@ -20,7 +20,6 @@ fieldName <- gsub('\\.csv','',fieldName)
 source('helperFunctions.R')
 
 dat <- read.csv(datLoc,stringsAsFactors=TRUE,fileEncoding='latin1') %>% 
-  rename_with(.fn = ~gsub('..L.ha.$','_lHa',.x)) %>%
   rename_with(.fn = ~gsub('..tonne.ha.$','_tHa',.x)) %>%
   rename_with(.fn = ~gsub('..m.s.$','_ms',.x)) %>%
   rename_with(.fn = ~gsub('.km.h.$','_kmh',.x)) %>%
@@ -30,21 +29,18 @@ dat <- read.csv(datLoc,stringsAsFactors=TRUE,fileEncoding='latin1') %>%
   rename_with(.fn = ~gsub('.deg.','Angle',.x)) %>%
   rename_with(.fn = ~gsub('\\.','',.x)) %>%
   rename('ID'='ObjId','DryYield'='YldMassDry_tHa','Lon'='Longitude','Lat'='Latitude','Pass'='PassNum','Speed'='Speed__m') %>%
-  # select(Lon:Distance,TrackAngle,Pass.Num,DryYield,Date) %>% 
   st_as_sf(coords=c('Lon','Lat')) %>% #Add spatial feature info
   st_set_crs(4326) %>% st_transform(3401) %>% #Lat-lon -> UTM
-  makePolys(width='SwthWdth_m',dist='Distance_m',angle='TrackAngle') %>%  
-  mutate(pArea=as.numeric(st_area(.))) %>% 
-  mutate(YieldMass=convertYield(DryYield,'tpha','gpm2')*pArea) %>% 
[…]

### 2021-09-29 Manuscript/manuscript.Rmd
Updated figures

@@
 However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
-Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems [@redhead2020]; typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
+Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems [@redhead2020; @hunicken2021]; typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
 There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
@@
 
-Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted to account for overlap with adjacent swaths).
-We extracted the size of each polygon (m$^2$), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
-Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
-Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
-We used dry yield (tonnes of seed/hectare minus recorded crop moisture) as our measure of crop yield
-Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows [@griffin2007; @whelan2013].
-We filtered data that lay outside the 95th percentiles of yield within each field, values less than or equal to zero, and values outside of the 95th percentiles of ground speed.
-Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time.
+Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
+We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m$^2$), dry yield (tonnes), and the spatial location.
+Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
+Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
+Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows [@griffin2007; @whelan2013], or when the combine is accelerating or decelerating [@arslan2002].
+Therefore, we used the following filtering procedure to clean the data prior to model fitting:
+
+1. Filtered spatial outliers and "inliers" using the procedure from [@vega2019] 
+2. Filtered unrealistic yield values (10.75 T/ha for wheat, 8 T/ha for canola and peas), or values less than or equal to zero
+3. Filtered extremely low and high outliers (below 1st percentile, above 99th percentile) for: dry yield, steering angle differences, ground speed
+4. Filtered points with large changes (>20%) in ground speed
+5. Filtered extremely low and high changes in position, indicating that the combine had stopped or moved to another part of the field between points
+
+Due to the large number of data at each field, we sub-sampled filtered data to 50,000 data points per field to reduce computation time.
 
@@
 Yield data from each field contains both random and systematic errors that must be modeled in order to reveal underlying changes in yield.
-We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the data.
+We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation.
 Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices.      <!-- - such as strip farming -->
-Ground speed is extremely important to yield monitor accuracy [@arslan2002], with low ground speed registering higher yields.
+Ground speed is extremely important to yield monitor accuracy, with low ground speed registering higher yields.
 Sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time.
@@
   \sqrt{yield} \sim & Normal (\mu, \sigma)\\
-  \mu = & Intercept + log(\text{Polygon Size}) + f(\text{Edge Distance}, b=12)_i + \\
-   & f(\text{Easting}, \text{Northing}, b=60) + f(\text{Sequence}, b=60) \\
-  log(\sigma) =  & Intercept + log(\text{Polygon Size}) + f(\text{Edge Distance}, b=12)_i + \\
-   & f(\text{Easting}, \text{Northing}, b=60) + f(\text{Sequence}, b=60) \\
+  \mu = & Intercept + f(\text{Edge Distance}, b=5)_i + \\
+   & f(\text{Easting}, \text{Northing}, b=60)\\
+  log(\sigma) =  & Intercept + f(\text{Edge Distance}, b=5)_i + \\
+   & f(\text{Easting}, \text{Northing}, b=60)\\
   \end{split}
@@
 
-Despite the use of spatial and temporal smoothers, the residuals in most of the models displayed a large amount of spatial and temporal autocorrelation.
-_Not really sure how we can get around this.
+The residuals in most of the models displayed both spatial and temporal autocorrelation in the residuals, representing "micro-scale" variation [@cr] that was not captured by the spatial smoothers.
+
+This is extremely common in precision yield data.
+
+We did not include temporal smoothers because spatial and temporal processes are confounded in our model; that is, combining of fields occurs in patterns (e.g.
+east-to-west) so it is impossible to tell whether lower or higher yields in certain areas of the field were caused by deviations in the yield monitor sensors over time.
+Therefore, we assigned all random variation as spatial in nature; customized combine driving patterns may be able to detect temporal variation, but this is beyond the scope of this research.
+
+<!-- _Not really sure how we can get around this.
 In practice, this means that SEs for all coefficients are too small (i.e.
-we gathered 1000 points, but because they're spatially related, it's only equivalent to 100 points)._
-
-<!-- - This is extremely common in precision yield data, but  -->
+we gathered 1000 points, but because they're spatially related, it's only equivalent to 100 points)._ -->
 
@@
 Lan's paper) which could be particularly important when beneficial arthropods are disperse very broadly [@oberg2007].
-There could also be interactions between landscape composition at different scales [@motzke2016; @robinson2021], but this is not broadly studied.
+There could also be interactions between landscape composition at different scales [@motzke2016; @robinson2021; @haslem2021], but this is not broadly studied.
 Landscape complexity [@fahrig2011] and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others).

### 2021-09-29 helperFunctions.R

@@ -506,7 +506,7 @@ runModI <- function(i,dS,nSubSamp=50000,kPar=c(12,60,60,12,60,60),modelCheckDir=
 
 #Function to run the ith model, but with boundary type
 #useClosest = use only closest boundary? Otherwise, uses distance from all boundaries
-runModII <- function(i,dS,nSubSamp=50000,kPar=c(12,60,60,12,60,60),useClosest=TRUE,modelCheckDir='./Figures/ModelCheck2',resultsDir='./Figures/YieldMaps2',filterData=TRUE){ 
+runModII <- function(i,dS,nSubSamp=50000,kPar=c(5,80,5,80),useClosest=TRUE,modelCheckDir='./Figures/ModelCheck2',resultsDir='./Figures/YieldMaps2',filterData=TRUE){ 
   # i <- 295 #Debugging - "Trent_Clark W 34 2019"
   # i <- 5 #Alvin French Zoltan - 1.2 million points
   # dS <- datSource
@@ -608,10 +608,10 @@ runModII <- function(i,dS,nSubSamp=50000,kPar=c(12,60,60,12,60,60),useClosest=TR
   #Make model formula for non-isotropic gam
   if(useClosest){
     f2 <- paste0('~ s(dist,k=',kPar[1],',bs="ts",by=boundaryType) + s(E,N,k=',kPar[2],')')
-    f <- paste0('sqrt(DryYield)~ s(dist,k=',kPar[4],',bs="ts",by=boundaryType) + s(E,N,k=',kPar[5],')')
+    f <- paste0('sqrt(DryYield)~ s(dist,k=',kPar[3],',bs="ts",by=boundaryType) + s(E,N,k=',kPar[4],')')
   } else {
     f2 <- paste0('~ ',paste0('s(dist_',boundaryTypes,',k=',kPar[1],',bs="ts")',collapse=' + '),' + s(E,N,k=',kPar[2],')')
-    f <- paste0('sqrt(DryYield) ~ ',paste0('s(dist_',boundaryTypes,',k=',kPar[4],',bs="ts")',collapse=' + '),' + s(E,N,k=',kPar[5],')')
+    f <- paste0('sqrt(DryYield) ~ ',paste0('s(dist_',boundaryTypes,',k=',kPar[3],',bs="ts")',collapse=' + '),' + s(E,N,k=',kPar[4],')')
   }
   
   # if(useClosest){ #These ones include a time smoother

### 2021-09-29 runModels.R

@@ -300,11 +300,11 @@ ggsave(paste0('./Figures/ModelSummary1a.png'),p,height=6,width=12,dpi=300)
 
 # Run second set of models - boundary type included ---------------------------
 
-# Test with single model
-
-# which(datSource$filename=='Gibbons Lafonda 2016')
+# # Test with single model
+# which(datSource$filename=='Gibbons Lafonda 2018') #18 mins
+# which(datSource$filename=='Alvin_French Al_Jr 2020') #12 mins
 # a <- Sys.time()
-# runModII(111,dS=datSource)
+# runModII(113,dS=datSource,kPar=c(5,60,5,60))
 # Sys.time()-a
 # # beep(1)
 # debugonce(runModII)
@@ -313,7 +313,7 @@ library(parallel)
 detectCores()
 cluster <- makeCluster(15) #10 procs uses about 30% of memory - could probably max it out
 a <- Sys.time()
-runWhich <- c(1:nrow(datSource))[with(datSource,use&!modelComplete2)] #Models to use, 
+runWhich <- c(1:nrow(datSource))[with(datSource,use&!modelComplete2&grepl('(Canola|Wheat|Peas)',crop))] #Models to use, 
 parLapply(cl=cluster,runWhich,runModII,dS=datSource,useClosest=TRUE) #All models
 beep(1)
 stopCluster(cluster)
@@ -420,9 +420,12 @@ ggsave(paste0('./Figures/ModelSummary2a.png'),p,height=6,width=12,dpi=300)
 
 # Function to sample from posterior of each prediction, amalgamate, fit meta-model, return results of meta-model
 # My way of getting around the problem of random effects
+# a = arguments from parLapply, not used
 # useRows = rows of ds to be used, ds = datSource csv, nX = number of samples along range of X
 # rCutoff = restrict point number (some fields have hundreds of thousands of points, but all data uses approximately the same size yield rectangles, so cuts off meta-smoother at "about" the maximum field size - i.e. the most common number of points per field)
-samplePreds <- function(useRows=NULL,ds=datSource,nX=c(30,100,500),rCutoff=50000,margInt=c(FALSE,FALSE,FALSE),Nsamp=1,ncore=NA){ 
+
[…]

### 2021-09-29 yield simulations.R

@@ -54,7 +54,7 @@ curve(leq(x,1,exp(-2))*leq(x,1,-0.1),0,100,ylim=c(0,1)) #Both
 curve(leq2(x,1,exp(-2),1,-0.1),0,100,ylim=c(0,1)) #Both
 
 
-png(paste0('./Figures/ExamplePlots/hypotheses.png'),width=12,height=6,units='in',res=200)
+# png(paste0('./Figures/ExamplePlots/hypotheses.png'),width=12,height=6,units='in',res=200)
 
 par(mfrow=c(1,2))
 #Scenario 1 - beneficial bugs decrease with distance into field
@@ -73,7 +73,35 @@ curve(leq(x,a1,b1),0,100,xlim=range(dist),col='red',add=TRUE,lwd=2)
 curve(leq(x,a2,b2),0,100,xlim=range(dist),col='blue',add=TRUE,lwd=2)
 par(mfrow=c(1,1))
 
-dev.off()
+# dev.off()
+
+#Approximately how many basis functions needed to represent this?
+library(tidyverse)
+library(mgcv)
+
+a1 <- -2; b1 <- 0.2; a2 <- 3; b2 <- -0.03
+scen1 <- data.frame(dist=1:100,yield=leq2(1:100,a1,b1,a2,b2)) %>% mutate(y=yield+rnorm(n(),0,0.1))
+a2 <- 0; b2 <- 0.03
+scen2 <- data.frame(dist=1:100,yield=leq2(1:100,a1,b1,a2,b2)) %>% mutate(y=yield+rnorm(n(),0,0.1))
+a2 <- 1; b2 <- 0
+scen3 <- data.frame(dist=1:100,yield=leq2(1:100,a1,b1,a2,b2)) %>% mutate(y=yield+rnorm(n(),0,0.1))
+
+m1 <- gam(y~s(dist,k=5,bs='cs'),data=scen1) #Looks like about 5 is necessary
+m2 <- gam(y~s(dist,k=5),data=scen2)
+m3 <- gam(y~s(dist,k=5),data=scen3)
+summary(m1)
+
+scen1 %>% mutate(pred=predict(m1)) %>% 
+  ggplot(aes(x=dist,y=y))+geom_point()+
+  geom_line(aes(y=yield))+geom_line(aes(y=pred),col='red')
+
+scen2 %>% mutate(pred=predict(m2)) %>% 
+  ggplot(aes(x=dist,y=y))+geom_point()+
[…]

### 2021-09-30 Manuscript/manuscript.Rmd
Updated manuscript

@@
 However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
-Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ["multifunctionality" _sensu_ @frei2018].
+Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ["multifunctionality" _sensu_ @mitchell2014; @frei2018].
 Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.
@@
 
-1. Filtered spatial outliers and "inliers" using the procedure from [@vega2019] 
-2. Filtered unrealistic yield values (10.75 T/ha for wheat, 8 T/ha for canola and peas), or values less than or equal to zero
-3. Filtered extremely low and high outliers (below 1st percentile, above 99th percentile) for: dry yield, steering angle differences, ground speed
-4. Filtered points with large changes (>20%) in ground speed
-5. Filtered extremely low and high changes in position, indicating that the combine had stopped or moved to another part of the field between points
-
-Due to the large number of data at each field, we sub-sampled filtered data to 50,000 data points per field to reduce computation time.
+1. Removed outliers and "inliers" using the spatial filtering procedure from [@vega2019] 
+2. Removed unrealistic yield values (10.75 T/ha for wheat, 8 T/ha for canola and peas), and values of zero or less
+3. Removed extremely low and high outliers (below 1st percentile, above 99th percentile) for: dry yield, steering angle differences, ground speed
+4. Removed points with large changes (>20%) in ground speed
+5. Removed extremely low and high changes in position, indicating that the combine had stopped or moved to another part of the field between data collection
+6. Sub-sampled filtered data to 50,000 data points per field (reduce computation time)
 
@@
 We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation.
-Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices.      <!-- - such as strip farming -->
-Ground speed is extremely important to yield monitor accuracy, with low ground speed registering higher yields.
-Sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time.
-This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time.
+Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, previous agricultural practices, and other factors not directly related to boundary distance.
+<!-- - such as strip farming -->
+<!-- Ground speed is extremely important to yield monitor accuracy, with low ground speed registering higher yields. -->
+<!-- Sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time. -->
+<!-- This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time. -->
 To model this, we fit the following model to each field-year of data:
@@
   \begin{split}
-  \sqrt{yield} \sim & Normal (\mu, \sigma)\\
-  \mu = & Intercept + f(\text{Edge Distance}, b=5)_i + \\
+  \sqrt{Yield} \sim & Normal (\mu, \sigma)\\
+  \mu = & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
    & f(\text{Easting}, \text{Northing}, b=60)\\
-  log(\sigma) =  & Intercept + f(\text{Edge Distance}, b=5)_i + \\
+  log(\sigma) =  & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
    & f(\text{Easting}, \text{Northing}, b=60)\\
@@
   - where:
-  1. Polygon Size = distance traveled $\times$ width of header bar ($m^2$)
-  2. Edge Distance = distance from field edge type _i_ (m)
+  1. Yield = weight of crop gathered per area of harvest (T/ha)
+  2. Boundary Distance = distance from field boundary type _i_ (m)
   3. Easting, Northing = distance from centre of field (m)
@@
 
-In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in harvested area, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest.
-Spatial or temporal variation can be modelled using a Gaussian Process Model (e.g. Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
+In addition to modeling edge effects (our variable of interest), this model also accounts for within-field spatial variation not related to edges.
+Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g. Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
 Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but using a reduced number of dimensions and computation time.
-The number of basis dimensions was checked with the _gam.check_ function from _mgcv_.
-The relationship between polygon size and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions.
-Because polygon size is not intuitive, we converted it into ground speed by regressing speed on distance traveled per polygon, then using this to back-calculate the ground speed for a given polygon size (given a fixed width of header).
+We used 5 basis dimensions for the distance smoother and 60 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \@ref(fig:hypotheses)).
+<!-- The number of basis dimensions was checked with the _gam.check_ function from _mgcv_. -->
 All models were fit in _R_ using the _mgcv_ library [version `r packageVersion('mgcv')`, @wood2017], and figures were created with _ggplot2_ and _ggpubr_ [versions `r packageVersion('ggplot2')`, @wickham2016; and `r packageVersion('ggpubr')`, @kassambara2020].
 
-The residuals in most of the models displayed both spatial and temporal autocorrelation in the residuals, representing "micro-scale" variation [@cr] that was not captured by the spatial smoothers.
-
-This is extremely common in precision yield data.
-
-We did not include temporal smoothers because spatial and temporal processes are confounded in our model; that is, combining of fields occurs in patterns (e.g.
-east-to-west) so it is impossible to tell whether lower or higher yields in certain areas of the field were caused by deviations in the yield monitor sensors over time.
-Therefore, we assigned all random variation as spatial in nature; customized combine driving patterns may be able to detect temporal variation, but this is beyond the scope of this research.
+The residuals in most of the models displayed both spatial autocorrelation in the residuals, indicating "micro-scale" variation that was not captured by the spatial smoothers [@cressie2011; @wickle2019].
+This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.
+extremely small p-values), and c) results did not change strongly when additional basis functions were added, we decided to leave the models as-is.
+Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data. 
+That is, combining of fields occurs in patterns (e.g.
+east-to-west) so it is impossible to tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
+We considered the latter to be a more realistic assumption, so we assigned all random variation as spatial in nature. 
+Customized combine driving patterns may be able to detect temporal variation, but this is beyond the scope of this research.
 
@@
 
-To consider results from all field-level models together, we fit models independent of each other and fit an "overall" smoother as averages of the field-level smoothers.
+To consider results from all field-level models together, we fit models independently and then fit "overall" smoothers as averages of the field-level smoothers.
 However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models:
@@
 This gives coverage intervals (CIs) for the "average" smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
-Spatial variability was not averaged between fields because differences in field geometry make it impossible to compare fields.
+Spatial variability was not averaged between fields because within-field variability is not directly comparable between fields.
 
@@
 The filtered data from each field still contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
-To illustrate this, we show the results from a single wheat field, where we show the relative impact of ground speed, harvest sequence, boundary distance, and spatial variation on both yield mean and yield SD ("patchiness").
+To illustrate this, we show the results from a single wheat field, where we show the relative impact of boundary distance and spatial variation on both yield mean and yield SD ("patchiness").
 Figure \@ref(fig:rawData) shows the raw data collected from a single field, demonstrating a large amount of variation even among the filtered data.
-Combine ground speed had a large effect on both yield mean and SD, and there was also a strong effect of harvest sequence on yield mean and SD (Figure \@ref(fig:orderSmooths).
 Yield mean and SD tended to decrease with distance from field boundaries, but there were no consistent differences between boundary types (Figure \@ref(fig:distSmooths).
@@
 
-### Canola fields
+Overall,both yield average and yield variability were extremely spatially dependent across all field types (average p-values: `r with(canolaMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(canolaMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`), meaning that within-field variability was the primary driver of yield across canola, wheat, and pea crops.
 
-Overall, there was a strong negative effect of combine ground speed on both the average and variability of canola yield.
-At the field level, both yield average and yield variability (SD of yield) of yield were extremely spatially (average p-values: `r with(canolaMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(canolaMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`) and temporally dependent (`r with(canolaMods,formatC(meanP[Smoother=='s(r)'],3))`, `r with(canolaMods,formatC(meanP[Smoother=='s.1(r)'],3))`), meaning that average and variability changed with location in the field, as well as sequentially during the harvest.
-However, there was no difference in average or variability with harvest sequence when all fields were considered, meaning that combines did not appear to distort yield data from start to finish in any systematic way.
+Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha  (37-44 bushels/acre, Figure \@ref(fig:canolaPlot) top row).
+Shelterbelt and Standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but Shelterbelts had a small intermediate maximum at around 50m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50m and remained roughly constant with distance.
+The effect of other boundary types was less consistent, but average yield tended to increase with distance away from Grassland, Other Crop, and Wetland, and was roughly the same with distance from Bare boundary types.
+Yield variability was essentially constant with distance for Shelterbelt, Standard, and Other Crop field boundaries, but showed some evidence of decreasing with distance from field boundary at Wetlands, Bare, and Grassland
 
-Average canola yield tended to increase with distance from Standard and Other Crop boundaries, leveling off at about 25 m (Figure \@ref(fig:canolaPlot) top row, Other crop and Standard panels); yield was approximately 0.1 T/ha (1.5 bu/ac) lower at the edge for both of these boundary types.
-Interestingly, Shelterbelts were well-represented in the data (Figure \@ref(fig:boundaryTypes)) but the negative edge effect was not strong (Figure \@ref(fig:canolaPlot) top row, Shelterbelt panel), indicating that they may provide better micro-climates for crops located near them.
-Other boundary types were less common in the data and had correspondingly weaker effects on mean yield, but all showed a small negative edge effect.
-Yield variability tended to decrease with distance from the edge of the field, indicating that edge effects contribute to yield "patchiness" as well as averages.
-At Standard, Bare, and Wetland field boundaries, variability decreased until about 100 m, after which variability was constant, while the pattern was less consistent in Grassland, Other Crop, and Shelterbelt.
-
-```{r canolaPlot, echo=FALSE, fig.cap='Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
-Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
-"patchiness").', out.width='100%'}
+```{r canolaPlot, echo=FALSE, fig.cap='Field boundary effect on canola yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and % refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
   include_graphics('../Figures/ModelSummary3a_canola.png',dpi=NA)
@@
 
-### Wheat fields
+<!-- Similar to canola, both yield average and yield variability of wheat was spatially dependent (average p-values: `r with(wheatMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(wheatMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`).
+-->
 
-Similar to canola, there was a strong negative effect of ground speed on both yield and variability of wheat yield (Figure \@ref(fig:wheatPlotSpeed), left panels).
-At the field level, both yield average and yield variability (SD of yield) of yield were spatially (average p-values: `r with(wheatMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(wheatMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`) and temporally dependent (`r with(wheatMods,formatC(meanP[Smoother=='s(r)'],3))`, `r with(wheatMods,formatC(meanP[Smoother=='s.1(r)'],3))`), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \@ref(fig:wheatPlotSpeed), right panel).
+The effect of boundary distance on average wheat yield was much more variable than the patterns seen in canola yield (Figure \@ref(fig:wheatPlot)).
+Average yield roughly increased with distance away from Wetland, Standard, and Grasslands, but both Shelterbelts and Bare ground displayed an intermediate maximum at around 50m (similar to canola), and decreased slightly with distance from Other Crops.
+Yield variability, on the other hand, tended to decrease with distance from Bare, Grassland, and Wetland (before 100 m, at least), increased slightly with distance from Shelterbelt and Standard, and was roughly constant for Other Crops.
 
-Average wheat yield increased slightly with distance from the field boundary, but this was much more variable than the patterns seen in canola yield (Figure \@ref(fig:wheatPlot)).
-There was also a slight increase in variability away from the field boundary, at least for Standard and Other Crop boundaries, but this was inconsistent among other boundary types.
-
-```{r wheatPlot, echo=FALSE, fig.cap='Field boundary effect on wheat yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
-Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
-"patchiness").', out.width='100%'}
+```{r wheatPlot, echo=FALSE, fig.cap='Field boundary effect on wheat yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and % refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
   include_graphics('../Figures/ModelSummary3a_wheat.png',dpi=NA)
@@
 
-### Pea fields
+<!-- Both average and variability of pea yields were spatially dependent (average p-values: `r with(peaMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(peaMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`).
+-->
+Pea crops were less well-represented in our data (N = XXX fields), and the effects of boundaries tended to be less consistant. 
+Average pea yields were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from Shelterbelt, Standard, and Other Crop boundaries, while the effect of Bare, Grassland, and Wetland boundaries was inconsistent.
+Conversely, yield variability decreased with distance from Grassland, Shelterbelt, Standard, and Wetland boundaries, while the pattern with distance from Other Crop and Bare was inconsistent. 
+Unlike canola and wheat crops, there was no obvious evidence of an intermediate increase in yield with distance from the field boundary.
 
-At the field level, both yield average and yield variability (SD of yield) of yield were spatially (average p-values: `r with(peaMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(peaMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`) and temporally dependent (`r with(peaMods,formatC(meanP[Smoother=='s(r)'],3))`, `r with(peaMods,formatC(meanP[Smoother=='s.1(r)'],3))`),
-There was a strong negative effect of ground speed on yield and variability of peas (Figure \@ref(fig:peasPlotSpeed)), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \@ref(fig:peasPlotSpeed), right panel).
-
-Average pea yields increased with distance from Shelterbelt, Standard, and Wetland field boundaries.
-Interestingly, the reduction at the edge was slightly higher for Wetlands, but yield increased with distance more quickly, i.e. yields were higher at distance for Wetlands.
-Yield variability decreased with distance from Grassland and Standard boundaries, and was slightly lower overall at Standard boundaries.
-
-```{r peaPlot, echo=FALSE, fig.cap='Field boundary effect on pea yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
-Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
-"patchiness").', out.width='100%'}
+```{r peaPlot, echo=FALSE, fig.cap='Field boundary effect on pea yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and % refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
   include_graphics('../Figures/ModelSummary3a_peas.png',dpi=NA)
@@
 
-At the level of individual fields, both average and SD of yield varied with distance from the field boundary.
-Both metrics also varied with ground speed and harvest sequence, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
-Yield tended to increase with distance from field boundaries before plateauing, but there was little evidence of intermediate boosts in yield, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \@ref(fig:hypotheses)).
-Some crops displayed a larger decrease in yield next to field boundaries; for example, pea yield directly adjacent to Wetland was lower overall, but increased with distance more than other crop types, indicating that Wetlands may locally decrease pea yields but are beneficial to the crop in general.
-Canola yields next to Other Crops had a reduced edge effect compared to Standard field boundaries, indicating that Other Crops may provide a more consistent microclimate for canola than grassy field boundaries.
-However, the difference between boundary types tended to be both small and variable, especially for wheat.
+At the level of individual fields, both average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
+Yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from Shelterbelts next to canola and wheat.
+This increase tended to be small in magnitude in canola crops, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \@ref(fig:hypotheses)).
+There was a fairly large ($\sim$0.75 T/ha) intermediate increase in average wheat yields away from Shelterbelts, indicating a potential reduction in ecosystem services with distance.
+However, there was a similar (albeit weaker) pattern associated with Bare ground, so this pattern may be caused by something other than the presence of Shelterbelts.
+
 <!-- For example, average canola yield was about 0.5 T/ha lower close to Standard field edges, and yield SD was about 0.3 T/ha SD higher. -->
@@
 
-## Boundaries may not provide spillover
+<!-- ## Boundaries may not provide spillover -->
 
@@
 
-## Beneficial organisms may not provide ecosystem services
+<!-- ## Beneficial organisms may not provide ecosystem services -->
 
@@
 
-We used harvest time as smoothed variable to deal with temporal autocorrelation in the data, but this is a crude way to deal with this.
-This is because in-field combine driving patterns _almost always_ induce correlation between spatial and temporal measurements (e.g.
+<!-- We used harvest time as smoothed variable to deal with temporal autocorrelation in the data, but this is a crude way to deal with this. -->
+<!-- This is because in-field combine driving patterns _almost always_ induce correlation between spatial and temporal measurements (e.g.
 northern section of the field was harvested before the southern section), meaning that temporal and spatial random effects are difficult to separate.
-There are few solutions to this using harvest data as-is, other than to ignore temporal random effects entirely, thereby constraining all random variation to be spatial.
-However, it is possible to remove correlation between spatial and temporal measurements by using specially designed harvest patterns within a field (harvesting alternating strips sequentially across the field, then harvesting the remaining strips on a second pass).
-Future studies could use this to estimate drift in combine sensor measurements over time.
+-->
+<!-- There are few solutions to this using harvest data as-is, other than to ignore temporal random effects entirely, thereby constraining all random variation to be spatial. -->
+<!-- However, it is possible to remove correlation between spatial and temporal measurements by using specially designed harvest patterns within a field (harvesting alternating strips sequentially across the field, then harvesting the remaining strips on a second pass).
+-->
+<!-- Future studies could use this to estimate drift in combine sensor measurements over time. -->
 

### 2021-10-06 Manuscript/abstract.Rmd
Updated figures

@@
-Field boundaries can improve crop yields by creating better abiotic conditions for crop growth (e.g. mosture, temperature), and can act as refuges for beneficial arthropods.
+Field boundaries can improve crop yields by creating better abiotic conditions for crop growth (e.g. moisture, temperature), and can act as refuges for beneficial arthropods.
 This suggests that beneficial crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services (e.g.
@@
 Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is relatively cheap and commonly measured by growers.
-<!-- In this study, we used 252 field-years of yield monitor data from three common crops -- wheat (_Triticum aestivum_), canola (_Brassica napus_), or peas (_Pisum sativum_) -- recorded across Alberta, Canada, and examined how yield varied with distances from common crop boundary types.
--->
-Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease concurrently with distance.
-Differences between boundary types tended to be weak and variable, especially for wheat; there was some evidence of certain field boundaries causing a greater decrease in yield at the edge, but these differences tended to be small in magnitude (<5% of yield).
-Furthermore, there was little evidence of an intermediate increase in yield, suggesting that a) ecosystem services do not strongly depend on distance or b) their effect is relatively small.
+In this study, we used 252 field-years of yield monitor data from three common crops -- wheat (_Triticum aestivum_), canola (_Brassica napus_), or peas (_Pisum sativum_) -- recorded across Alberta, Canada, and examined how yield varied with distances from common crop boundary types.
+Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease with distance.
+There was evidence of an intermediate increase in yield for wheat away from shelterbelts, and a weak increase in canola, but no other boundary type showed evidence of this.
 This study represents one of the first uses of precision yield data to measure ecosystem service provision at large spatial scales.

### 2021-10-06 Manuscript/manuscript.Rmd
Updated figures

@@
 ---
-title: "Living on the edge: crop boundary effects on precision canola and wheat yields in Alberta"
+title: "Living on the edge: crop boundary effects on precision crop yields in Alberta, Canada"
 author:
@@
 
-useThese <- with(datSource,use&grepl('Wheat|Canola|Peas',crop)) #Rows to use
+useThese <- with(datSource,use&grepl('Wheat|Canola|Peas',crop)&modelComplete2) #Rows to use
 
@@
   bind_rows(.id='field') %>% 
-  mutate(pval=as.numeric(pval)) %>% 
-  group_by(Smoother) %>% 
-  summarize(meanP=mean(pval),maxP=max(pval),propSig=sum(pval<0.05)/n(),meanEDF=mean(edf),meanRef=mean(Ref.df))
+  mutate(pval=as.numeric(pval)) 
+# %>% 
+#   group_by(Smoother) %>% 
+#   summarize(meanP=mean(pval),maxP=max(pval),propSig=sum(pval<0.05)/n(),meanEDF=mean(edf),meanRef=mean(Ref.df))
 
@@
 
-rm(use,paths)
+# rm(use,paths)
 
@@
 SNL in agroecosystems is important for both agricultural production and conservation [@prietoBenitez2011], and can create better biotic and abiotic conditions for crop growth [@case2019; @weninger2021; but see @lowe2021].
-SNL consitutes habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control [@albrecht2020; @gardner2021], but this can vary depending on the crop system [@gray2014; @quinn2017] and landscape context [@marja2019; @boetzl2020].
+SNL constitutes habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control [@albrecht2020; @gardner2021], but this can vary depending on the crop system [@gray2014; @quinn2017] and landscape context [@marja2019; @boetzl2020].
 They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion [@kort1988; @brandle2004; @wratten2012].
@@
 If negative edge effects of the edge improve with distance and ecosystem services decrease with distance, yield may be maximized at intermediate distances where the ecosystem services cancel out negative edge effects [@mitchell2014; @vanVooren2017; Figure \@ref(fig:hypotheses)a].
-This suggests a "goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
+This suggests a "Goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
 Alternatively, if ecosystem services are constant with distance (or increase, Figure \@ref(fig:hypotheses)b) then yield will increase with distance from a field boundary and then plateau.
@@
 Farmers were solicited for yield data through local agronomists, and we received `r nrow(datSource)` field-years of data from `r length(unique(datSource$grower))` growers across a total of `r length(unique(datSource$year))` years (`r min(unique(datSource$year))`-`r max(unique(datSource$year))`).
-We converted raw data to a standard csv format using Ag Leader SMS.
+We converted raw data to a standard csv format using Ag Leader SMS®.
 `r round(100*sum(grepl('(Wheat|Canola|Peas)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_), canola (_Brassica napus_), or peas (_Pisum sativum_), three of the most common crops in rotation in Alberta [@statscan_canola2014].
@@
 Individual fields contained between `r min(yearsPerField)` and `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`), containing a total of `r round(sum(datSourceFinal$npoints)/1e6,1)` million data points.
+
+```{r fieldLocataions, echo=FALSE, fig.cap='Locations of fields that data were collected from', out.width='100%'}  
+  include_graphics('../Figures/Field Locations.png',dpi=NA)
+```
 
@@
 To illustrate this, we show the results from a single wheat field, where we show the relative impact of boundary distance and spatial variation on both yield mean and yield SD ("patchiness").
-Figure \@ref(fig:rawData) shows the raw data collected from a single field, demonstrating a large amount of variation even among the filtered data.
-Yield mean and SD tended to decrease with distance from field boundaries, but there were no consistent differences between boundary types (Figure \@ref(fig:distSmooths).
-Finally, the random spatial smoothers in Figure \@ref(fig:spatialSmooths) show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions (Figure \@ref(fig:spatialSmooths).
+Figure \@ref(fig:spatialSmooths)a shows the raw data collected from a single field, demonstrating a large amount of variation even among the filtered data.
+Yield mean and SD tended to decrease with distance from field boundaries, but there were no consistent differences between boundary types (Figure \@ref(fig:distSmooths)).
+Finally, the random spatial smoothers in Figure \@ref(fig:spatialSmooths)b and c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.
 
-```{r rawData, echo=FALSE, fig.cap='Example output from a single field.', out.width='75%'}  
-  include_graphics('../Figures/ExamplePlots/rawData.png',dpi=NA)
+```{r spatialSmooths, echo=FALSE, fig.cap='Raw data and spatial smoothers from a single field.
+Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.', out.width='100%'}
+  include_graphics('../Figures/ExamplePlots/spatialSmooths.png',dpi=NA)
 ```
 
-```{r orderSmooths, echo=FALSE, fig.cap='Ground speed and harvest sequence (random temporal variation) smoothers from a single field.', out.width='100%'}  
-  include_graphics('../Figures/ExamplePlots/orderSmooths.png',dpi=NA)
-```
-
-```{r distSmooths, echo=FALSE, fig.cap='Distance smoothers from a single field.', out.width='100%'}  
+```{r distSmooths, echo=FALSE, fig.cap='Distance smoothers from a single field, showing a negative effect of distance on average yield (first row).
+The variability smoothers show an intermediate reduction in variability at around 90 m.', out.width='100%'}
   include_graphics('../Figures/ExamplePlots/distSmooths.png',dpi=NA)
-```
-
-```{r spatialSmooths, echo=FALSE, fig.cap='Spatial (random spatial variation) smoothers from a single field.', out.width='100%'}  
-  include_graphics('../Figures/ExamplePlots/spatialSmooths.png',dpi=NA)
 ```
@@
 
-Across all fields, Standard (grassy border) field boundaries were the most common boundary type, followed by Shelterbelts and Other Crops (Figure \@ref(fig:boundaryTypes)).
+Across all fields, standard (grassy border) field boundaries were the most common boundary type, followed by shelterbelts and other crops (Figure \@ref(fig:boundaryTypes)).
 <!-- Not sure if we should talk about this. Not very many fields from southern regions, so it's hard to compare. Perhaps mention in Discussion -->
@@
 ```{r boundaryTypes, echo=FALSE, fig.cap='Percentage of field boundaries across sampled fields.', out.width='100%'}
+
 bLengths %>% 
@@
   group_by(location) %>% mutate(nFields=length(unique(filename))) %>% ungroup() %>% 
-  mutate(location=paste0(location,' (N field-years = ',nFields,')')) %>%
+  mutate(location=paste0(location,' (N = ',nFields,')')) %>%
   select(filename,croptype,location,contains('prop_')) %>% 
@@
   mutate(name=factor(name,labels=c('Bare','Grassland','Other\nCrop','Shelterbelt','Standard','Wetland'))) %>% 
+  group_by(croptype) %>% mutate(croptype=paste0(croptype,'\n(N = ',length(unique(filename)),')')) %>% 
   ggplot(aes(x=name,y=value,fill=croptype))+geom_boxplot()+
@@
 
-Overall,both yield average and yield variability were extremely spatially dependent across all field types (average p-values: `r with(canolaMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(canolaMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`), meaning that within-field variability was the primary driver of yield across canola, wheat, and pea crops.
+Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both <0.0001), meaning that within-field variability was the primary driver of yield across canola, wheat, and pea crops (Table \@ref(tab:smoothSummary), row 6).
+Distances from boundary types were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
 
-Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha  (37-44 bushels/acre, Figure \@ref(fig:canolaPlot) top row).
-Shelterbelt and Standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but Shelterbelts had a small intermediate maximum at around 50m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50m and remained roughly constant with distance.
-The effect of other boundary types was less consistent, but average yield tended to increase with distance away from Grassland, Other Crop, and Wetland, and was roughly the same with distance from Bare boundary types.
-Yield variability was essentially constant with distance for Shelterbelt, Standard, and Other Crop field boundaries, but showed some evidence of decreasing with distance from field boundary at Wetlands, Bare, and Grassland
+<!-- `r with(canolaMods,formatC(mean(pval[Smoother=='s(E,N)'])))`, `r with(canolaMods,formatC(mean(pval[Smoother=='s.1(E,N)'],3)))` -->
+
+```{r smoothSummary, echo=FALSE, out.width='100%'}
+
+canolaMods %>% group_by(Smoother) %>% 
+  summarize(N=n(),mean=mean(pval),propSig=sum(pval<(0.05/N))/n()) %>% 
+  ungroup() %>% 
+  mutate(par=ifelse(grepl('\\.',Smoother),'Variability','Mean')) %>% 
+  mutate(type=case_when(
+    grepl('boundaryType',Smoother) ~ gsub('s[\\.1]*\\(dist\\)\\:boundaryType','',Smoother),
+    grepl('E,N',Smoother) ~ 'Spatial Smoother'
+  )) %>% 
+  select(-N,-Smoother) %>% relocate(type,par,mean,propSig) %>% 
+  pivot_wider(id_cols=type,names_from=par,values_from=c(mean,propSig)) %>% 
+  relocate(type,mean_Mean,propSig_Mean,mean_Variability,propSig_Variability) %>% 
+  mutate(type=c('Bare','Grassland','Other Crop','Shelterbelt','Standard','Wetland','Spatial Smoother')) %>% 
+  kable(.,format='latex',caption='Mean p-values and proportion of p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all fields.',col.names = c('Smoother Type','Average p-value','Proportion <0.05','Average p-value','Proportion <0.05'),
+        digits = c(1,3,3,3,3),escape = FALSE) %>% 
+  add_header_above(c('', 'Mean' = 2, 'Variability' = 2))
+
+```
+
+Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha  (Figure \@ref(fig:canolaPlot) top row).
+Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50m and remained roughly constant with distance.
+The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
+Yield variability (Figure \@ref(fig:canolaPlot) bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
+Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from Wetland boundaries, indicating a potential stabilizing effect.
 
@@
 Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and % refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
+N refers to number of fields containing this boundary type, and \\% refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
   include_graphics('../Figures/ModelSummary3a_canola.png',dpi=NA)
@@
 
-The effect of boundary distance on average wheat yield was much more variable than the patterns seen in canola yield (Figure \@ref(fig:wheatPlot)).
-Average yield roughly increased with distance away from Wetland, Standard, and Grasslands, but both Shelterbelts and Bare ground displayed an intermediate maximum at around 50m (similar to canola), and decreased slightly with distance from Other Crops.
-Yield variability, on the other hand, tended to decrease with distance from Bare, Grassland, and Wetland (before 100 m, at least), increased slightly with distance from Shelterbelt and Standard, and was roughly constant for Other Crops.
+The effect of boundary distance on average wheat yield was much more variable than the patterns seen in canola yield.
+Average yield (Figure \@ref(fig:wheatPlot), top row) roughly increased with distance away from wetland, standard, and grasslands, but both shelterbelts and bare ground displayed an intermediate maximum at around 50m (similar to canola), and decreased slightly with distance from Other Crops.
+Yield variability, on the other hand, tended to decrease with distance from bare, grassland, and wetland (before 100 m, at least), increased slightly with distance from shelterbelt and standard boundaries, and was roughly constant with distance from other crops (Figure \@ref(fig:wheatPlot), bottom row).
 
@@
 Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and % refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
+N refers to number of fields containing this boundary type, and \\% refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
   include_graphics('../Figures/ModelSummary3a_wheat.png',dpi=NA)
@@
 -->
-Pea crops were less well-represented in our data (N = XXX fields), and the effects of boundaries tended to be less consistant. 
-Average pea yields were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from Shelterbelt, Standard, and Other Crop boundaries, while the effect of Bare, Grassland, and Wetland boundaries was inconsistent.
-Conversely, yield variability decreased with distance from Grassland, Shelterbelt, Standard, and Wetland boundaries, while the pattern with distance from Other Crop and Bare was inconsistent. 
+
+Pea crops were less well-represented in our data (N = `r sum(with(datSource,use&modelComplete2&crop=='Peas'))` fields, compared to ), and the effects of boundaries tended to be less consistent. 
+Average pea yields (Figure \@ref(fig:peaPlot), top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
+Conversely, yield variability (Figure \@ref(fig:peaPlot), bottom row) decreased with distance from grassland, shelterbelt, standard, and wetland boundaries, while the pattern with distance from other crop and bare was inconsistent.
 Unlike canola and wheat crops, there was no obvious evidence of an intermediate increase in yield with distance from the field boundary.
@@
 Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and % refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
+N refers to number of fields containing this boundary type, and \\% refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
   include_graphics('../Figures/ModelSummary3a_peas.png',dpi=NA)
@@
 At the level of individual fields, both average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
-Yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from Shelterbelts next to canola and wheat.
+Yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat.
 This increase tended to be small in magnitude in canola crops, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \@ref(fig:hypotheses)).
-There was a fairly large ($\sim$0.75 T/ha) intermediate increase in average wheat yields away from Shelterbelts, indicating a potential reduction in ecosystem services with distance.
-However, there was a similar (albeit weaker) pattern associated with Bare ground, so this pattern may be caused by something other than the presence of Shelterbelts.
+There was a fairly large (0.75 T/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
+However, there was a similar (albeit weaker) pattern associated with bare ground, so this pattern may be caused by something other than the presence of shelterbelts.
+This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this across a larger geographic domain in order to inform growers about how best to manage the landscapes surrounding their fields.
 
@@
 
-Why did we find few consistent effects of boundary type on crop yield?
+Why did we find few consistent effects of boundary type on crop yield, aside from shelterbelts?
 Our work examined only the first and last links in the causal chain _Field boundary_ $\rightarrow$ _Insect Abundance_ $\rightarrow$ _Crop Yield_, meaning that are two mediating relationships where the overall relationship could be broken or obscured.
@@
 SNL generally has a positive effect on spillover, but field boundaries only represent a fraction of the SNL that surrounds a given field.
-Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales (e.g.
-Lan's paper) which could be particularly important when beneficial arthropods are disperse very broadly [@oberg2007].
+Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales [@nguyen2021] which could be particularly important when beneficial arthropods are disperse very broadly [@oberg2007].
 There could also be interactions between landscape composition at different scales [@motzke2016; @robinson2021; @haslem2021], but this is not broadly studied.
-Landscape complexity [@fahrig2011] and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others).
+Landscape complexity [@fahrig2011] and connectivity are also associated with higher abundance of beneficial arthropods.
 This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat [but see @fahrig2019].
@@
 Our first scenario (Scenario 1, Figure \@ref(fig:hypotheses)) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
-However, field boundaries may not have a positive effect if a) pests (ecosystem dis-services) also disperse along with beneficials, b) if beneficials do not disperse from the boundary environment
-Ecosystem disservices such as pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services [@zhang2007].
-@gray2014 showed that remnant forest fragments do not provide significant pest control services , but also do not depress yield.
+However, field boundaries may not have a positive effect if a) ecosystem dis-services -- pest insects or weed seeds -- also disperse along with beneficials, or if b) beneficials do not disperse from the boundary environment.
+Pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services [@zhang2007].
+@gray2014 showed that remnant forest fragments do not provide significant pest control services, but also do not depress yield.
 However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data _alone_.
@@
 
+Our work is some of the first to use a large precision yield dataset to answer questions about how landscapes influence yield, but this study has only scratched the surface of some of the richest agroecological datasets ever created.
 Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
-Most fields were located in Central Alberta [Dry Mixedwood region, @ABRegions2006], where boundaries contained a higher proportion of Shelterbelts (Figure \@ref(fig:boundaryTypes)), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
+Most fields were located in Central Alberta [Dry Mixedwood region, @ABRegions2006], where boundaries contained a higher proportion of shelterbelts (Figure \@ref(fig:boundaryTypes)), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
 Between-year effects may be important, especially during dry years [@kort1988], but this was not considered in our analysis.
-Future work could examine whether boundaries have differing effects during hot or cool years.
+Future work could examine whether boundaries have differing effects during hot or cool years, as shading or cooling effects may increase yield during hot years but have negative (or no) effect during cooler years.
 
@@
 
-SVJR and PG conceived of the project.
+SVJR, LHN, and PG conceived of the project.
 SVJR collected data, conducted analysis, and wrote the manuscript.
@@
 
-```{r child = 'supplemental.Rmd'}
-```
+<!-- ```{r child = 'supplemental.Rmd'} -->
+<!-- ``` -->

### 2021-10-06 Manuscript/manuscript.tex
Updated figures

@@
             pdfauthor={},
-            pdftitle={Living on the edge: crop boundary effects on canola and wheat yields in Alberta},
+            pdftitle={Living on the edge: crop boundary effects on precision crop yields in Alberta, Canada},
             colorlinks=false,
@@
 
-  \title{Living on the edge: crop boundary effects on canola and wheat yields in Alberta}
+  \title{Living on the edge: crop boundary effects on precision crop yields in Alberta, Canada}
     \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
@@
   \begin{abstract}
-  Field boundaries can improve crop yields by creating better abiotic conditions for crop growth (e.g.~mosture, temperature), and can act as refuges for beneficial arthropods.
+  Field boundaries can improve crop yields by creating better abiotic conditions for crop growth (e.g.~moisture, temperature), and can act as refuges for beneficial arthropods.
 This suggests that beneficial crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services (e.g.~pollination or pest suppression) from the field boundary.
@@
 Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is relatively cheap and commonly measured by growers.
-Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease concurrently with distance.
-Differences between boundary types tended to be weak and variable, especially for wheat; there was some evidence of certain field boundaries causing a greater decrease in yield at the edge, but these differences tended to be small in magnitude (\textless5\% of yield).
-Furthermore, there was little evidence of an intermediate increase in yield, suggesting that a) ecosystem services do not strongly depend on distance or b) their effect is relatively small.
+In this study, we used 252 field-years of yield monitor data from three common crops -- wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}) -- recorded across Alberta, Canada, and examined how yield varied with distances from common crop boundary types.
+Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease with distance.
+There was evidence of an intermediate increase in yield for wheat away from shelterbelts, and a weak increase in canola, but no other boundary type showed evidence of this.
 This study represents one of the first uses of precision yield data to measure ecosystem service provision at large spatial scales.
@@
 
-Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth.
-This has allowed increases in human population and increased global stability in production.
+Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth (Ramankutty \emph{et al.} 2018).
+This has allowed increases in human population and increased global stability in production (Foley \emph{et al.} 2011).
 However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
-Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios.
+Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ({``multifunctionality''} \emph{sensu} Mitchell \emph{et al.} 2014; Frei \emph{et al.} 2018).
+Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.
 
-Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.
 SNL in agroecosystems is important for both agricultural production and conservation (Prieto-Benítez \& Méndez 2011), and can create better biotic and abiotic conditions for crop growth (Case \emph{et al.} 2019; Weninger \emph{et al.} 2021; but see Lowe \emph{et al.} 2021).
-SNL consitutes habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this can vary depending on the system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017).
-They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed (Kort 1988; Brandle \emph{et al.} 2004).
+SNL constitutes habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this can vary depending on the crop system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017) and landscape context (Marja \emph{et al.} 2019; Boetzl \emph{et al.} 2020).
+They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion (Kort 1988; Brandle \emph{et al.} 2004; Wratten \emph{et al.} 2012).
 North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL.
+Semi-natural field boundaries (e.g.~grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are of particular interest, as they are extremely common and require less land than field-sized agricultural set-asides.
 
-SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate.
-Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds.
-At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges (Kowalchuk \& Jong 1995).
-For example, pollination services from central place foragers (bees) that nest in SNL but forage in crops drops rapidly with distance (Garibaldi \emph{et al.} 2011).
-Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects {[}Van Vooren \emph{et al.} (2017); Figure \ref{fig:hypotheses}{]}.
-This also suggests a ``goldilocks'' field size, where negative edge effects are canceled out by ecosystem services.
+Field boundaries can cause both positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
+Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds.
+Field boundaries can also provide helpful ecosystem services, such as pest suppression or pollination, and these also tend to be occur close to the field edge (Duelli \& Obrist 2003).
+For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Kowalchuk \& Jong 1995; Garibaldi \emph{et al.} 2011).
+If negative edge effects of the edge improve with distance and ecosystem services decrease with distance, yield may be maximized at intermediate distances where the ecosystem services cancel out negative edge effects {[}Mitchell \emph{et al.} (2014); Van Vooren \emph{et al.} (2017); Figure \ref{fig:hypotheses}a{]}.
+This suggests a ``Goldilocks'' field size, where negative edge effects are canceled out by ecosystem services.
+Alternatively, if ecosystem services are constant with distance (or increase, Figure \ref{fig:hypotheses}b) then yield will increase with distance from a field boundary and then plateau.
 
@@
 
-Studies of SNL effects on crop yield also suffer from limited scope (e.g.~few crop types) and small sample sizes, limiting inference and reducing generality (Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
-Meta-analyses have been the standard solution to this problem (e.g.
-Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its common usage and relatively cheap operating costs.
+Boundary effects are difficult to observe without large amounts of crop yield data, which is typically expensive and time-consuming to collect.\\
+Studies of semi-natural field boundaries on crop yield tend to suffer from limited spatial scope and small sample sizes, limiting inference and reducing generality (Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
+Meta-analyses are a common solution to this problem (e.g.
+Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs.
 However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
-Ecosystem services can influence both the mean and variability of yield in agroecosystems (Redhead \emph{et al.} 2020); typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable from a grower's perspective.
+Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems (Redhead \emph{et al.} 2020; Hünicken \emph{et al.} 2021); typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
 There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
 
-In this paper we ask:
+In this paper use a precision yield dataset from Alberta, Canada to ask the following questions:
 1. How does crop yield change with distance from the edge of field?
@@
 3. Is there an intermediate distance where yield is maximized or variance is minimized?
-To our knowledge, no other study has addressed the effect of field boundaries on productivity using precision yield, making this study highly novel.
+To our knowledge, no other study has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.
 
@@
 Farmers were solicited for yield data through local agronomists, and we received 298 field-years of data from 5 growers across a total of 7 years (2014-2020).
-We converted raw data to a standard csv format using Ag Leader SMS.
+We converted raw data to a standard csv format using Ag Leader SMS®.
 85\% of the crop types where either wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}), three of the most common crops in rotation in Alberta (Statistics Canada 2014).
-The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
-Individual fields contained between 1 and 6 years of data (mean: 3.1), containing a total of 14.4 million data points.
+The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (91), canola (118), or peas (38), for a total of 247 field-years of data.
+Individual fields contained between 1 and 6 years of data (mean: 3), containing a total of 13.9 million data points.
 
-Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted to account for overlap with adjacent swaths).
-We extracted the size of each polygon (m\(^2\)), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
-Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
-Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
-We used dry yield (tonnes of seed/hectare minus recorded crop moisture) as our measure of crop yield
-Precision yield data can be highly variable and is prone to extreme outliers (both positive and negative), especially when combine ground speeds are low and at the beginning of rows
-Therefore, we filtered data that lay outside the 95th percentiles of yield within each field, values less than or equal to zero, and values outside of the 95th percentiles of ground speed.
-Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time.
+\begin{figure}
+\includegraphics[width=1\linewidth]{../Figures/Field Locations} \caption{Locations of fields that data were collected from}\label{fig:fieldLocataions}
+\end{figure}
+
+Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
+We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m\(^2\)), dry yield (tonnes), and the spatial location.
+Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
+Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
+Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows (Griffin \emph{et al.} 2007; Whelan \& Taylor 2013), or when the combine is accelerating or decelerating (Arslan \& Colvin 2002).
+Therefore, we used the following filtering procedure to clean the data prior to model fitting:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Removed outliers and ``inliers'' using the spatial filtering procedure from (Vega \emph{et al.} 2019)
+\item
+  Removed unrealistic yield values (10.75 T/ha for wheat, 8 T/ha for canola and peas), and values of zero or less
+\item
+  Removed extremely low and high outliers (below 1st percentile, above 99th percentile) for: dry yield, steering angle differences, ground speed
+\item
+  Removed points with large changes (\textgreater20\%) in ground speed
+\item
+  Removed extremely low and high changes in position, indicating that the combine had stopped or moved to another part of the field between data collection
+\item
+  Sub-sampled filtered data to 50,000 data points per field (reduce computation time)
+\end{enumerate}
 
@@
 Yield data from each field contains both random and systematic errors that must be modeled in order to reveal underlying changes in yield.
-We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the data.
-Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices.
-Ground speed is extremely important to yield monitor accuracy (Arslan \& Colvin 2002), with low ground speed registering higher yields.
-Sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time.
-This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time.
+We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation.
+Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, previous agricultural practices, and other factors not directly related to boundary distance.
 To model this, we fit the following model to each field-year of data:
@@
   \begin{split}
-  \sqrt{yield} \sim & Normal (\mu, \sigma)\\
-  \mu = & Intercept + log(\text{Polygon Size}) + f(\text{Edge Distance}, b=12)_i + \\
-   & f(\text{Easting}, \text{Northing}, b=60) + f(\text{Sequence}, b=60) \\
-  log(\sigma) =  & Intercept + log(\text{Polygon Size}) + f(\text{Edge Distance}, b=12)_i + \\
-   & f(\text{Easting}, \text{Northing}, b=60) + f(\text{Sequence}, b=60) \\
+  \sqrt{Yield} \sim & Normal (\mu, \sigma)\\
+  \mu = & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
+   & f(\text{Easting}, \text{Northing}, b=60)\\
+  log(\sigma) =  & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
+   & f(\text{Easting}, \text{Northing}, b=60)\\
   \end{split}
@@
 \item
-  Polygon Size = distance traveled \(\times\) width of header bar (\(m^2\))
+  Yield = weight of crop gathered per area of harvest (T/ha)
 \item
-  Edge Distance = distance from field edge type \emph{i} (m)
+  Boundary Distance = distance from field boundary type \emph{i} (m)
 \item
@@
 
-In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in harvested area, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest.
-Spatial or temporal variation can be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
+In addition to modeling edge effects (our variable of interest), this model also accounts for within-field spatial variation not related to edges.
+Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
 Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but using a reduced number of dimensions and computation time.
-The number of basis dimensions was checked with the \emph{gam.check} function from \emph{mgcv}.
-The relationship between polygon size and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions.
-Because polygon size is not intuitive, we converted it into ground speed by regressing speed on distance traveled per polygon, then using this to back-calculate the ground speed for a given polygon size (given a fixed width of header).
+We used 5 basis dimensions for the distance smoother and 60 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \ref{fig:hypotheses}).
 All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).
 
-Despite the use of spatial and temporal smoothers, the residuals in most of the models displayed a large amount of spatial and temporal autocorrelation.
-\emph{Not really sure how we can get around this.
-In practice, this means that SEs for all coefficients are too small (i.e.~we gathered 1000 points, but because they're spatially related, it's only equivalent to 100 points).}
+The residuals in most of the models displayed both spatial autocorrelation in the residuals, indicating ``micro-scale'' variation that was not captured by the spatial smoothers (Cressie \& Wikle 2011; Wikle \emph{et al.} 2019).
+This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.~extremely small p-values), and c) results did not change strongly when additional basis functions were added, we decided to leave the models as-is.
+Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
+That is, combining of fields occurs in patterns (e.g.~east-to-west) so it is impossible to tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
+We considered the latter to be a more realistic assumption, so we assigned all random variation as spatial in nature.
+Customized combine driving patterns may be able to detect temporal variation, but this is beyond the scope of this research.
 
-To consider results from all field-level models together, we fit models independent of each other and fit an ``overall'' smoother as averages of the field-level smoothers.
+To consider results from all field-level models together, we fit models independently and then fit ``overall'' smoothers as averages of the field-level smoothers.
 However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models:
@@
 This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
-Spatial variability was not averaged between fields because differences in field geometry make it impossible to compare fields.
+Spatial variability was not averaged between fields because within-field variability is not directly comparable between fields.
 
@@
 The filtered data from each field still contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
-To illustrate this, we show the results from a single wheat field, where we show the relative impact of ground speed, harvest sequence, boundary distance, and spatial variation on both yield mean and yield SD (``patchiness'').
-Figure \ref{fig:rawData} shows the raw data collected from a single field, demonstrating a large amount of variation even among the filtered data.
-Combine ground speed had a large effect on both yield mean and SD, and there was also a strong effect of harvest sequence on yield mean and SD (Figure \ref{fig:orderSmooths}.
-Yield mean and SD tended to decrease with distance from field boundaries, but there were no consistent differences between boundary types (Figure \ref{fig:distSmooths}.
-Finally, the random spatial smoothers in Figure \ref{fig:spatialSmooths} show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions (Figure \ref{fig:spatialSmooths}.
+To illustrate this, we show the results from a single wheat field, where we show the relative impact of boundary distance and spatial variation on both yield mean and yield SD (``patchiness'').
+Figure \ref{fig:spatialSmooths}a shows the raw data collected from a single field, demonstrating a large amount of variation even among the filtered data.
+Yield mean and SD tended to decrease with distance from field boundaries, but there were no consistent differences between boundary types (Figure \ref{fig:distSmooths}).
+Finally, the random spatial smoothers in Figure \ref{fig:spatialSmooths}b and c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.
 
 \begin{figure}
-\includegraphics[width=0.75\linewidth]{../Figures/ExamplePlots/rawData} \caption{Example output from a single field.}\label{fig:rawData}
+\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/spatialSmooths} \caption{Raw data and spatial smoothers from a single field.
+Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.}\label{fig:spatialSmooths}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/orderSmooths} \caption{Ground speed and harvest sequence (random temporal variation) smoothers from a single field.}\label{fig:orderSmooths}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/distSmooths} \caption{Distance smoothers from a single field.}\label{fig:distSmooths}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/spatialSmooths} \caption{Spatial (random spatial variation) smoothers from a single field.}\label{fig:spatialSmooths}
+\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/distSmooths} \caption{Distance smoothers from a single field, showing a negative effect of distance on average yield (first row).
+The variability smoothers show an intermediate reduction in variability at around 90 m.}\label{fig:distSmooths}
 \end{figure}
@@
 
-Across all fields, Standard (grassy border) field boundaries were the most common boundary type, followed by Shelterbelts and Other Crops (Figure \ref{fig:boundaryTypes}).
+Across all fields, standard (grassy border) field boundaries were the most common boundary type, followed by shelterbelts and other crops (Figure \ref{fig:boundaryTypes}).
 
@@
 
-\hypertarget{canola-fields}{%
-\subsubsection{Canola fields}\label{canola-fields}}
+Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variability was the primary driver of yield across canola, wheat, and pea crops (Table \ref{tab:smoothSummary}, row 6).
+Distances from boundary types were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
 
-Overall, there was a strong negative effect of combine ground speed on both the average and variability of canola yield.
-At the field level, both yield average and yield variability (SD of yield) of yield were extremely spatially (average p-values: 2e-16, 1.77e-05) and temporally dependent (2e-16, 0.000243), meaning that average and variability changed with location in the field, as well as sequentially during the harvest.
-However, there was no difference in average or variability with harvest sequence when all fields were considered, meaning that combines did not appear to distort yield data from start to finish in any systematic way.
+\begin{table}
 
-Average canola yield tended to increase with distance from Standard and Other Crop boundaries, leveling off at about 25 m (Figure \ref{fig:canolaPlot} top row, Other crop and Standard panels); yield was approximately 0.1 T/ha (1.5 bu/ac) lower at the edge for both of these boundary types.
-Interestingly, Shelterbelts were well-represented in the data (Figure \ref{fig:boundaryTypes}) but the negative edge effect was not strong (Figure \ref{fig:canolaPlot} top row, Shelterbelt panel), indicating that they may provide better micro-climates for crops located near them.
-Other boundary types were less common in the data and had correspondingly weaker effects on mean yield, but all showed a small negative edge effect.
-Yield variability tended to decrease with distance from the edge of the field, indicating that edge effects contribute to yield ``patchiness'' as well as averages.
-At Standard, Bare, and Wetland field boundaries, variability decreased until about 100 m, after which variability was constant, while the pattern was less consistent in Grassland, Other Crop, and Shelterbelt.
+\caption{\label{tab:smoothSummary}Mean p-values and proportion of p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all fields.}
+\centering
+\begin{tabular}[t]{l|r|r|r|r}
+\hline
+\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Mean} & \multicolumn{2}{c}{Variability} \\
+\cline{2-3} \cline{4-5}
+Smoother Type & Average p-value & Proportion <0.05 & Average p-value & Proportion <0.05\\
+\hline
+Bare & 0.036 & 0.875 & 0.055 & 0.857\\
+\hline
+Grassland & 0.079 & 0.794 & 0.061 & 0.794\\
+\hline
+Other Crop & 0.063 & 0.860 & 0.044 & 0.840\\
+\hline
+Shelterbelt & 0.014 & 0.935 & 0.059 & 0.815\\
+\hline
+Standard & 0.021 & 0.938 & 0.055 & 0.839\\
+\hline
+Wetland & 0.045 & 0.833 & 0.013 & 0.857\\
+\hline
+Spatial Smoother & 0.000 & 1.000 & 0.000 & 1.000\\
+\hline
+\end{tabular}
+\end{table}
+
+Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure \ref{fig:canolaPlot} top row).
+Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50m and remained roughly constant with distance.
+The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
+Yield variability (Figure \ref{fig:canolaPlot} bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
+Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from Wetland boundaries, indicating a potential stabilizing effect.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_canola} \caption{Field boundary effect on canola yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
-Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
-"patchiness").}\label{fig:canolaPlot}
+\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_canola} \caption{Field boundary effect on canola yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:canolaPlot}
 \end{figure}
 
-\hypertarget{wheat-fields}{%
-\subsubsection{Wheat fields}\label{wheat-fields}}
-
-Similar to canola, there was a strong negative effect of ground speed on both yield and variability of wheat yield (Figure \ref{fig:wheatPlotSpeed}, left panels).
-At the field level, both yield average and yield variability (SD of yield) of yield were spatially (average p-values: 2e-16, 2e-16) and temporally dependent (2e-16, 2e-16), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \ref{fig:wheatPlotSpeed}, right panel).
-
-Average wheat yield increased slightly with distance from the field boundary, but this was much more variable than the patterns seen in canola yield (Figure \ref{fig:wheatPlot}).
-There was also a slight increase in variability away from the field boundary, at least for Standard and Other Crop boundaries, but this was inconsistent among other boundary types.
+The effect of boundary distance on average wheat yield was much more variable than the patterns seen in canola yield.
+Average yield (Figure \ref{fig:wheatPlot}, top row) roughly increased with distance away from wetland, standard, and grasslands, but both shelterbelts and bare ground displayed an intermediate maximum at around 50m (similar to canola), and decreased slightly with distance from Other Crops.
+Yield variability, on the other hand, tended to decrease with distance from bare, grassland, and wetland (before 100 m, at least), increased slightly with distance from shelterbelt and standard boundaries, and was roughly constant with distance from other crops (Figure \ref{fig:wheatPlot}, bottom row).
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_wheat} \caption{Field boundary effect on wheat yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
-Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
-"patchiness").}\label{fig:wheatPlot}
+\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_wheat} \caption{Field boundary effect on wheat yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:wheatPlot}
 \end{figure}
 
-\hypertarget{pea-fields}{%
-\subsubsection{Pea fields}\label{pea-fields}}
-
-At the field level, both yield average and yield variability (SD of yield) of yield were spatially (average p-values: 2e-16, 2e-16) and temporally dependent (2e-16, 3.62e-07),
-There was a strong negative effect of ground speed on yield and variability of peas (Figure \ref{fig:peasPlotSpeed}), but harvest sequence did not distort the average or SD of wheat yield in a systematic way (Figure \ref{fig:peasPlotSpeed}, right panel).
-
-Average pea yields increased with distance from Shelterbelt, Standard, and Wetland field boundaries.
-Interestingly, the reduction at the edge was slightly higher for Wetlands, but yield increased with distance more quickly, i.e.~yields were higher at distance for Wetlands.
-Yield variability decreased with distance from Grassland and Standard boundaries, and was slightly lower overall at Standard boundaries.
+Pea crops were less well-represented in our data (N = 38 fields, compared to ), and the effects of boundaries tended to be less consistent.
+Average pea yields (Figure \ref{fig:peaPlot}, top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
+Conversely, yield variability (Figure \ref{fig:peaPlot}, bottom row) decreased with distance from grassland, shelterbelt, standard, and wetland boundaries, while the pattern with distance from other crop and bare was inconsistent.
+Unlike canola and wheat crops, there was no obvious evidence of an intermediate increase in yield with distance from the field boundary.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_peas} \caption{Field boundary effect on pea yield, accounting for the effect of combine speed, spatial variation, and harvest sequence.
-Upper panel represents mean yield, while the lower panel represents yield variation (i.e.
-"patchiness").}\label{fig:peaPlot}
+\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_peas} \caption{Field boundary effect on pea yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:peaPlot}
 \end{figure}
@@
 
-At the level of individual fields, both average and SD of yield varied with distance from the field boundary.
-Both metrics also varied with ground speed and harvest sequence, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
-Yield tended to increase with distance from field boundaries before plateauing, but there was little evidence of intermediate boosts in yield, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \ref{fig:hypotheses}).
-Some crops displayed a larger decrease in yield next to field boundaries; for example, pea yield directly adjacent to Wetland was lower overall, but increased with distance more than other crop types, indicating that Wetlands may locally decrease pea yields but are beneficial to the crop in general.
-Canola yields next to Other Crops had a reduced edge effect compared to Standard field boundaries, indicating that Other Crops may provide a more consistent microclimate for canola than grassy field boundaries.
-However, the difference between boundary types tended to be both small and variable, especially for wheat.
+At the level of individual fields, both average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
+Yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat.
+This increase tended to be small in magnitude in canola crops, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \ref{fig:hypotheses}).
+There was a fairly large (0.75 T/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
+However, there was a similar (albeit weaker) pattern associated with bare ground, so this pattern may be caused by something other than the presence of shelterbelts.
+This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this across a larger geographic domain in order to inform growers about how best to manage the landscapes surrounding their fields.
 
-Why did we find few consistent effects of boundary type on crop yield?
-We consider a few potential reasons for this below:
-1. Landscape influence
-2. Boundary variables
-3. Ecosystem disservices
-4. Yearly/location variation
-5. Lack of pest pressure/pollination deficit
+Why did we find few consistent effects of boundary type on crop yield, aside from shelterbelts?
+Our work examined only the first and last links in the causal chain \emph{Field boundary} \(\rightarrow\) \emph{Insect Abundance} \(\rightarrow\) \emph{Crop Yield}, meaning that are two mediating relationships where the overall relationship could be broken or obscured.
+First, field boundaries may not have any influence on insect abundance if separated from the larger landscape context, or spillover of beneficial organisms may be low.
+Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are low, which may vary yearly, or by location.
+Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) when organisms in crops provide tangible ecosystem services.
+We consider these reasons below:
 
-Our analysis only examined boundary types, without considering the landscape context that fields were embedded in.
-Some studies show the influence of landscape composition at larger spatial scales (e.g.~Lan's paper) which could be particularly important when beneficial arthropods are highly mobile (e.g.~bumblebee foraging) or disperse very broadly (Öberg \emph{et al.} 2007).
-There could also be interactions between landscape composition at different scales (Motzke \emph{et al.} 2016; Robinson \emph{et al.} 2021), but this is not well-studied.
-Landscape complexity (Fahrig \emph{et al.} 2011) and connectivity are also associated with higher abundance of beneficial arthropods (Paul's new paper + others).
+SNL generally has a positive effect on spillover, but field boundaries only represent a fraction of the SNL that surrounds a given field.
+Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales (Nguyen \emph{et al.} 2021) which could be particularly important when beneficial arthropods are disperse very broadly (Öberg \emph{et al.} 2007).
+There could also be interactions between landscape composition at different scales (Motzke \emph{et al.} 2016; Haslem \emph{et al.} 2021; Robinson \emph{et al.} 2021), but this is not broadly studied.
+Landscape complexity (Fahrig \emph{et al.} 2011) and connectivity are also associated with higher abundance of beneficial arthropods.
 This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat (but see Fahrig \emph{et al.} 2019).
-Our definitions of boundary type were also (necessarily) loose, as we used satellite imagery to manually classify them.
-However, it may be that other boundary traits, such as size or plant species composition, are more important for spillover of ecosystems services into fields.
+
+Field boundary traits may play an important role in mediating spillover, but our definitions of boundaries were (necessarily) loose, as we used satellite imagery to manually classify them.
+Therefore, it may be that other aspects of field boundaries, such as size or plant community composition, are more important for spillover of ecosystems services into fields.
 For example, larger (Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020) or more functionally diverse (Bonifacio \emph{et al.} 2011; Boughey \emph{et al.} 2011) hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction matters to predatory beetle communities (Sutter \emph{et al.} 2018).
@@
 Our first scenario (Scenario 1, Figure \ref{fig:hypotheses}) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
-However, field boundaries may not have a positive effect if a) pests (ecosystem dis-services) also disperse along with beneficials, b) if beneficials do not disperse from the boundary environment, or if c) the effect of beneficial organisms is small.
-Ecosystem disservices such as pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang \emph{et al.} 2007).
-Gray \& Lewis (2014) showed that remnant forest fragments do not provide significant pest control services , but also do not depress yield.
+However, field boundaries may not have a positive effect if a) ecosystem dis-services -- pest insects or weed seeds -- also disperse along with beneficials, or if b) beneficials do not disperse from the boundary environment.
+Pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang \emph{et al.} 2007).
+Gray \& Lewis (2014) showed that remnant forest fragments do not provide significant pest control services, but also do not depress yield.
 However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data \emph{alone}.
 Ecosystem services from field boundaries may be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales Robinson \emph{et al.} (2021).
-Finally, pest pressure or pollination deficit may be low enough that ecosystem service provision does not significantly change yield.
+
+Pest pressure or pollination deficit may be low enough that ecosystem service provision does not significantly change yield.
 Canola is largely self-pollinating (at least in hybrid canola varieties, Lindström \emph{et al.} 2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop (Woodcock \emph{et al.} 2016).
@@
 
+Our work is some of the first to use a large precision yield dataset to answer questions about how landscapes influence yield, but this study has only scratched the surface of some of the richest agroecological datasets ever created.
 Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
-Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee 2006), where boundaries contained a higher proportion of Shelterbelts (Figure \ref{fig:boundaryTypes}), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
+Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee 2006), where boundaries contained a higher proportion of shelterbelts (Figure \ref{fig:boundaryTypes}), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
 Between-year effects may be important, especially during dry years (Kort 1988), but this was not considered in our analysis.
-Future work could examine whether boundaries have differing effects during hot or cool years.
-
-We used harvest time as smoothed variable to deal with temporal autocorrelation in the data, but this is a crude way to deal with this.
-This is because in-field combine driving patterns \emph{almost always} induce correlation between spatial and temporal measurements (e.g.~northern section of the field was harvested before the southern section), meaning that temporal and spatial random effects are difficult to separate.
-There are few solutions to this using harvest data as-is, other than to ignore temporal random effects entirely, thereby constraining all random variation to be spatial.
-However, it is possible to remove correlation between spatial and temporal measurements by using specially designed harvest patterns within a field (harvesting alternating strips sequentially across the field, then harvesting the remaining strips on a second pass).
-Future studies could use this to estimate drift in combine sensor measurements over time.
+Future work could examine whether boundaries have differing effects during hot or cool years, as shading or cooling effects may increase yield during hot years but have negative (or no) effect during cooler years.
 
@@
 Baker, T.P., Moroni, M.T., Mendham, D.S., Smith, R. \& Hunt, M.A. (2018). Impacts of windbreak shelter on crop and livestock production. \emph{Crop and Pasture Science}, 69, 785.
+
+\leavevmode\hypertarget{ref-boetzl2020}{}%
+Boetzl, F.A., Schuele, M., Krauss, J.
+\& Steffan-Dewenter, I.
+(2020).
+Pest control potential of adjacent agri-environment schemes varies with crop type and is shaped by landscape context and within-field position.
+\emph{Journal of Applied Ecology}, 57, 1482--1493.
 
@@
 
+\leavevmode\hypertarget{ref-cressie2011}{}%
+Cressie, N. \& Wikle, C.K. (2011). \emph{Statistics for spatio-temporal data}. Wiley.
+
+\leavevmode\hypertarget{ref-duelli2003}{}%
+Duelli, P. \& Obrist, M.K. (2003). Regional biodiversity in an agricultural landscape: The contribution of seminatural habitat islands. \emph{Basic and Applied Ecology}, 4, 129--138.
+
 \leavevmode\hypertarget{ref-fahrig2019}{}%
@@
 \emph{Ecology Letters}, 14, 101--112.
+
+\leavevmode\hypertarget{ref-foley2011}{}%
+Foley, J.A., Ramankutty, N., Brauman, K.A., Cassidy, E.S., Gerber, J.S., Johnston, M., \emph{et al.} (2011). Solutions for a cultivated planet. \emph{Nature}, 478, 337--342.
+
+\leavevmode\hypertarget{ref-frei2018}{}%
+Frei, B., Renard, D., Mitchell, M.G.E., Seufert, V., Chaplin-Kramer, R., Rhemtulla, J.M., \emph{et al.} (2018).
+Bright spots in agricultural landscapes: Identifying areas exceeding expectations for multifunctionality and biodiversity.
+\emph{Journal of Applied Ecology}, 55, 2731--2743.
 
@@
 Gray, C.L. \& Lewis, O.T. (2014). Do riparian forest fragments provide ecosystem services or disservices in surrounding oil palm plantations? \emph{Basic and Applied Ecology}, 15, 693--700.
+
+\leavevmode\hypertarget{ref-griffin2007}{}%
+Griffin, T.W., Brown, J.P. \& Lowenberg-DeBoerc, J. (2007). \emph{Yield monitor data analysis protocol: A primer in the management and analysis of precision agriculture data}.
+
+\leavevmode\hypertarget{ref-haslem2021}{}%
+Haslem, A., Clarke, R.H., Holland, G.J., Radford, J.Q., Stewart, A.
+\& Bennett, A.F.
+(2021).
+Local management or wider context: What determines the value of farm revegetation plantings for birds?
+\emph{Journal of Applied Ecology}.
+
+\leavevmode\hypertarget{ref-hunicken2021}{}%
+Hünicken, P.L., Morales, C.L., Aizen, M.A., Anderson, G.K.S., García, N.
+\& Garibaldi, L.A.
+(2021).
+Insect pollination enhances yield stability in two pollinator-dependent crops.
+\emph{Agriculture, Ecosystems {\&} Environment}, 320, 107573.
 
@@
 
+\leavevmode\hypertarget{ref-marja2019}{}%
+Marja, R., Kleijn, D., Tscharntke, T., Klein, A.-M., Frank, T.
+\& Batáry, P.
+(2019).
+Effectiveness of agri-environmental management on pollinators is moderated more by ecological contrast than by landscape structure or land-use intensity.
+\emph{Ecology Letters}, In press.
+
+\leavevmode\hypertarget{ref-mitchell2014}{}%
+Mitchell, M.G.E., Bennett, E.M. \& Gonzalez, A. (2014). Forest fragments modulate the provision of multiple ecosystem services. \emph{Journal of Applied Ecology}, 51, 909--918.
+
 \leavevmode\hypertarget{ref-motzke2016}{}%
@@
 Natural Regions Committee. (2006). \emph{Natural regions and subregions of {Alberta}}. Government of Alberta, Edmonton.
+
+\leavevmode\hypertarget{ref-nguyen2021}{}%
+Nguyen, L.H., Robinson, S.V.J.
+\& Galpern, P.
+(2021).
+Medium-resolution multispectral satellite imagery in precision agriculture: Mapping precision canola (\emph{brassica napus} {L.}) Yield using {Sentinel-2} time series.
+\emph{Precision Agriculture}, In Review (Preprint: https://www.essoar.org/doi/10.1002/essoar.10507479.2).
 
@@
 Quinn, N.F., Brainard, D.C. \& Szendrei, Z. (2017). Floral strips attract beneficial insects but do not enhance yield in cucumber fields. \emph{Journal of Economic Entomology}, 110, 517--524.
+
+\leavevmode\hypertarget{ref-ramankutty2018}{}%
+Ramankutty, N., Mehrabi, Z., Waha, K., Jarvis, L., Kremen, C., Herrero, M., \emph{et al.} (2018).
+Trends in global agricultural land use: Implications for environmental health and food security.
+\emph{Annual Review of Plant Biology}, 69, 789--815.
 
@@
 
+\leavevmode\hypertarget{ref-vega2019}{}%
+Vega, A., Córdoba, M., Castro-Franco, M. \& Balzarini, M. (2019). Protocol for automating error removal from yield maps. \emph{Precision Agriculture}, 20, 1030--1044.
+
 \leavevmode\hypertarget{ref-weninger2021}{}%
@@
 
+\leavevmode\hypertarget{ref-whelan2013}{}%
+Whelan, B. \& Taylor, J. (2013). \emph{Precision agriculture for grain production systems}. CSIRO Publishing.
+
 \leavevmode\hypertarget{ref-wickham2016}{}%
 Wickham, H. (2016). \emph{{ggplot2}: Elegant graphics for data analysis}. Springer-Verlag New York.
+
+\leavevmode\hypertarget{ref-wickle2019}{}%
+Wikle, C.K., Zammit-Mangion, A. \& Cressie, N. (2019). \emph{Spatio-temporal statistics with r}. Chapman; Hall: CRC Press.
 
@@
 \emph{Agriculture, Ecosystems \& Environment}, 231, 15--23.
+
+\leavevmode\hypertarget{ref-wratten2012}{}%
+Wratten, S.D., Gillespie, M., Decourtye, A., Mader, E.
+\& Desneux, N.
+(2012).
+Pollinator habitat enhancement: Benefits to other ecosystem services.
+\emph{Agriculture, Ecosystems {\&} Environment}, 159, 112--122.
 
@@
 
-\newpage
-
-\hypertarget{appendix-a-supplementary-material}{%
-\section*{Appendix A: Supplementary Material}\label{appendix-a-supplementary-material}}
-\addcontentsline{toc}{section}{Appendix A: Supplementary Material}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/groundSpeed_canola} \caption{Effect of ground speed and order (sequence of harvest) on canola yield.}\label{fig:canolaPlotSpeed}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/groundSpeed_wheat} \caption{Effect of ground speed and order (sequence of harvest) on wheat yield.}\label{fig:wheatPlotSpeed}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/groundSpeed_peas} \caption{Effect of ground speed and order (sequence of harvest) on pea yield.}\label{fig:peasPlotSpeed}
-\end{figure}
-
 \end{document}

### 2021-10-06 helperFunctions.R

@@ -1,5 +1,23 @@
 # Helper functions
 
+backTrans <- function(x){ #Back-transform units - data was sqrt transformed, so logSD is actually (log(sqrt(sd)))
+  pow <- function(x,p) x^p
+  expow <- function(x,p) log(exp(x)^p)
+  # x$pArea$predMean <- pow(x$pArea$predMean,2)
+  # x$pArea$predLogSD <- expow(x$pArea$predLogSD,2)
+  
+  x$coverDist <- lapply(x$coverDist,function(y){
+    y$predMean <- pow(y$predMean,2)
+    y$predLogSD <- expow(y$predLogSD,2)
+    return(y)
+  })
+  
+  # x$r$predMean <- pow(x$r$predMean,2)
+  # x$r$predLogSD <- expow(x$r$predLogSD,2)
+  
+  return(x)
+}
+
 #Scales x between 0 and 1
 range01 <- function(x) (x-min(x))/(max(x)-min(x))
 
@@ -1170,4 +1188,6 @@ lag2 <- function(x,n){
     } else {
       lead(x,abs(n)) #Negative lag
     } 
-  } 
+} 
+
+

### 2021-10-06 runModels.R

@@ -559,7 +559,7 @@ samplePreds <- function(a=NA,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutoff=
 
 # Add to current samples - canola
 isCanola <- which(with(datSource,use & crop=='Canola' & modelComplete2)) #Canola crops only
-Nsamp <- 200 #Number of samples
+Nsamp <- 500 #Number of samples
 library(parallel)
 cluster <- makeCluster(15) #Memory usage is OK, so could probably max it out
 clusterExport(cluster,c('datSource'))
@@ -571,7 +571,7 @@ save(samp,file='./Data/postSamples_canola.Rdata')
 
 # Add to current samples - wheat
 isWheat <- which(datSource$crop=='Wheat') #Wheat
-Nsamp <- 200 #Number of samples
+Nsamp <- 500 #Number of samples
 library(parallel)
 cluster <- makeCluster(15)
 clusterExport(cluster,c('datSource'))
@@ -583,7 +583,7 @@ save(samp,file='./Data/postSamples_wheat.Rdata')
 
 # Add to current samples - peas
 isPeas <- which(datSource$crop=='Peas') #Peas
-Nsamp <- 200 #Number of samples
+Nsamp <- 500 #Number of samples
 library(parallel)
 cluster <- makeCluster(15)
 clusterExport(cluster,c('datSource'))
@@ -594,23 +594,6 @@ stopCluster(cluster)
 save(samp,file='./Data/postSamples_peas.Rdata')
 
 #Get samples from storage
-backTrans <- function(x){ #Back-transform units - data was sqrt transformed, so logSD is actually (log(sqrt(sd)))
-  pow <- function(x,p) x^p
-  expow <- function(x,p) log(exp(x)^p)
-  # x$pArea$predMean <- pow(x$pArea$predMean,2)
-  # x$pArea$predLogSD <- expow(x$pArea$predLogSD,2)
-  
-  x$coverDist <- lapply(x$coverDist,function(y){
-    y$predMean <- pow(y$predMean,2)
-    y$predLogSD <- expow(y$predLogSD,2)
-    return(y)
-  })
-  
-  # x$r$predMean <- pow(x$r$predMean,2)
-  # x$r$predLogSD <- expow(x$r$predLogSD,2)
-  
-  return(x)
-}
[…]

### 2021-10-12 yield simulations.R
Updated hypothesis figure

@@ -56,24 +56,36 @@ curve(leq2(x,1,exp(-2),1,-0.1),0,100,ylim=c(0,1)) #Both
 
 # png(paste0('./Figures/ExamplePlots/hypotheses.png'),width=12,height=6,units='in',res=200)
 
-par(mfrow=c(1,2))
+png("C:\\Users\\Samuel\\Desktop\\hypothesisPlot.png",width=12,height=6,units='in',res=200) #Path on Multivac
+
+savepar <- par(no.readonly = TRUE)
+par(mfrow=c(1,2),mai=c(1,0.5,0.8,0.4))
 #Scenario 1 - beneficial bugs decrease with distance into field
 a1 <- -2; b1 <- 0.2; a2 <- 3; b2 <- -0.03
-curve(leq2(x,a1,b1,a2,b2),0,100,ylim=c(0,1),
-      lwd=4,ylab='Yield',xlab='Distance from Edge',main='Scenario 1')#,main='Scenario 1: bugs - with dist')
+
+plot(0,0,type='n',xlim=c(0,100),ylim=c(0,1),axes=FALSE,
+     ylab=NA,xlab='Distance from Edge',main='Scenario 1',xaxt='n')
+axis(2,at = c(0,1),labels = c('Low','High')); axis(1,at=seq(0,100,by=20))
+
+curve(leq2(x,a1,b1,a2,b2),0,100,lwd=4,add=TRUE)#,main='Scenario 1: bugs - with dist')
 curve(leq(x,a1,b1),0,100,xlim=range(dist),col='red',add=TRUE,lwd=2)
 curve(leq(x,a2,b2),0,100,xlim=range(dist),col='blue',add=TRUE,lwd=2)
-text(5,0.85,'Ecosystem\nservices',col='blue'); text(80,0.9,'Environmental\nEffect',col='red'); text(80,0.6,'Yield')
+text(10,0.83,'Ecosystem\nservice\neffect',col='blue'); text(70,0.95,'Environmental effect',col='red'); text(80,0.75,'Crop Yield')
 
 #Scenario 2 - beneficial bugs increase with distance into field (or pest insects at edge)
 a2 <- 0; b2 <- 0.03
+
+plot(0,0,type='n',xlim=c(0,100),ylim=c(0,1),axes=FALSE,
+     ylab=NA,xlab='Distance from Edge',main='Scenario 2',xaxt='n')
[…]

### 2021-10-13 Manuscript/abstract.Rmd
Updated runModels

@@
 Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease with distance.
-There was evidence of an intermediate increase in yield for wheat away from shelterbelts, and a weak increase in canola, but no other boundary type showed evidence of this.
+There was evidence of an intermediate increase in yield for wheat away from shelterbelts, and a weak increase in canola, but this was not seen for other crop types or boundary types.
 This study represents one of the first uses of precision yield data to measure ecosystem service provision at large spatial scales.

### 2021-10-13 runModels.R

@@ -557,41 +557,41 @@ samplePreds <- function(a=NA,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutoff=
 
 # Get crop-specific smoother info from second set of models 
 
-# Add to current samples - canola
-isCanola <- which(with(datSource,use & crop=='Canola' & modelComplete2)) #Canola crops only
-Nsamp <- 500 #Number of samples
-library(parallel)
-cluster <- makeCluster(15) #Memory usage is OK, so could probably max it out
-clusterExport(cluster,c('datSource'))
-a <- Sys.time()
-samp <- parLapply(cl=cluster,1:Nsamp,fun=samplePreds,useRows=isCanola,margInt=c(FALSE,FALSE,FALSE),kPar=10)
-Sys.time()-a
-stopCluster(cluster)
-save(samp,file='./Data/postSamples_canola.Rdata')
-
-# Add to current samples - wheat
-isWheat <- which(datSource$crop=='Wheat') #Wheat
-Nsamp <- 500 #Number of samples
-library(parallel)
-cluster <- makeCluster(15)
-clusterExport(cluster,c('datSource'))
-a <- Sys.time()
-samp <- parLapply(cl=cluster,1:Nsamp,fun=samplePreds,useRows=isWheat,margInt=c(FALSE,FALSE,FALSE),kPar=10)
-Sys.time()-a
-stopCluster(cluster)
-save(samp,file='./Data/postSamples_wheat.Rdata')
-
-# Add to current samples - peas
-isPeas <- which(datSource$crop=='Peas') #Peas
-Nsamp <- 500 #Number of samples
-library(parallel)
-cluster <- makeCluster(15)
-clusterExport(cluster,c('datSource'))
-a <- Sys.time()
-samp <- parLapply(cl=cluster,1:Nsamp,fun=samplePreds,useRows=isPeas,margInt=c(FALSE,FALSE,FALSE),kPar=10)
-Sys.time()-a
-stopCluster(cluster)
-save(samp,file='./Data/postSamples_peas.Rdata')
+# # Add to current samples - canola
+# isCanola <- which(with(datSource,use & crop=='Canola' & modelComplete2)) #Canola crops only
+# Nsamp <- 500 #Number of samples
+# library(parallel)
[…]

### 2021-10-18 runModels.R
Updated figures

@@ -729,7 +729,8 @@ for(i in 1:length(croptype)){
 # Example figures (Trent Clark Johnson 2014) ---------------------------------------------------------
 
 use <- which(datSource$filename == 'Trent_Clark W 34 2014')
-#Could also use 'Trent_Clark JOHNSON 2014'
+use <- which(datSource$filename == 'Trent_Clark JOHNSON 2014') #Could also use this
+
 load(datSource$modelPath2[use]) #Load model
 
 #Spatial smoothers
@@ -764,13 +765,15 @@ hexGrid <- hexGrid %>%  #Center coordinates
 
 theme_set(theme_bw())
 
-palette <- 'RdYlGn'
+# palette <- 'RdYlBu'
+# palette <- 'Spectral'
+palette <- 'YlGnBu'
 
 #Raw data + 
 (p1 <-ggplot()+
     geom_sf(data=dat,aes(geometry=geometry,col=DryYield),size=0.5,show.legend = 'point')+
     geom_sf(data=fieldBoundaryType,aes(geometry=geometry),show.legend= 'line')+
-    scale_colour_distiller(type='div',palette = palette, direction = 1)+
+    scale_colour_distiller(type='div',palette = palette, direction = -1)+
     labs(col='Yield (T/ha)',title='Raw Yield Data')+
     theme(legend.position='bottom'))
 
@@ -796,7 +799,7 @@ useSmooths <- which(grepl('(E,N)',sapply(modList$smooths,function(x) x$label),fi
     ggplot()+geom_sf(col=NA,aes(fill=pred))+
     geom_sf(data=fieldBoundaryType,aes(geometry=geometry),show.legend= 'line')+
     labs(fill='Yield (T/ha)',title='Yield Average Smoother')+
-    scale_fill_distiller(type='div',palette = palette, direction = 1) +
+    scale_fill_distiller(type='div',palette = palette, direction = -1) +
     theme(legend.position='bottom')
   )
 
@@ -805,8 +808,9 @@ useSmooths <- which(grepl('(E,N)',sapply(modList$smooths,function(x) x$label),fi
   mutate(pred=log(exp(pred)^2)) %>% #Back-transform
   ggplot()+geom_sf(col=NA,aes(fill=pred))+
[…]

### 2021-10-18 yield simulations.R

@@ -54,9 +54,9 @@ curve(leq(x,1,exp(-2))*leq(x,1,-0.1),0,100,ylim=c(0,1)) #Both
 curve(leq2(x,1,exp(-2),1,-0.1),0,100,ylim=c(0,1)) #Both
 
 
-# png(paste0('./Figures/ExamplePlots/hypotheses.png'),width=12,height=6,units='in',res=200)
+png(paste0('./Figures/ExamplePlots/hypotheses.png'),width=12,height=6,units='in',res=200)
 
-png("C:\\Users\\Samuel\\Desktop\\hypothesisPlot.png",width=12,height=6,units='in',res=200) #Path on Multivac
+# png("C:\\Users\\Samuel\\Desktop\\hypothesisPlot.png",width=12,height=6,units='in',res=200) #Path on Multivac
 
 savepar <- par(no.readonly = TRUE)
 par(mfrow=c(1,2),mai=c(1,0.5,0.8,0.4))

### 2021-10-18 Manuscript/manuscript.Rmd
Worked on paper draft

@@
 
-Potential journals: Agriculture Ecosystems and Environment, Basic and Applied Ecology, Ecological Economics, Journal of Applied Ecology (probably won't like it)
+<!-- Potential journals: Agriculture Ecosystems and Environment, Basic and Applied Ecology, Ecological Economics, Journal of Applied Ecology (probably won't like it) -->
 
-Potential reviewers: Teja Tscharntke, Ignasi Bartomeus, Ben Woodcock, John Redhead, some Canadian/Albertan reviewers who know the crop systems?
+<!-- Potential reviewers: Teja Tscharntke, Ignasi Bartomeus, Ben Woodcock, John Redhead, some Canadian/Albertan reviewers who know the crop systems? -->
 
@@
 SNL in agroecosystems is important for both agricultural production and conservation [@prietoBenitez2011], and can create better biotic and abiotic conditions for crop growth [@case2019; @weninger2021; but see @lowe2021].
-SNL constitutes habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control [@albrecht2020; @gardner2021], but this can vary depending on the crop system [@gray2014; @quinn2017] and landscape context [@marja2019; @boetzl2020].
-They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion [@kort1988; @brandle2004; @wratten2012].
-North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL.
-Semi-natural field boundaries (e.g.
-grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are of particular interest, as they are extremely common and require less land than field-sized agricultural set-asides.
+SNL acts as habitat for mobile organisms, and can act as a source of ecosystem services, such as pollination or pest control [@albrecht2020; @gardner2021], but this depends on both the crop system [@gray2014; @quinn2017] and landscape context [@marja2019; @boetzl2020].
+SNL can also create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion [@kort1988; @brandle2004; @wratten2012].
+<!-- North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL. -->
+Field boundaries (e.g.
+grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are SNL of particular interest, as they are extremely common and require less land than field-sized agricultural set-asides.
 
 <!-- - SNL has conflicting effects on crop yield because of ecosystem service provision as well as negative edge effects -->
-Field boundaries can cause both positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
-Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds.
-Field boundaries can also provide helpful ecosystem services, such as pest suppression or pollination, and these also tend to be occur close to the field edge [@duelli2003]. 
+Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
+Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure \@ref(fig:hypotheses), Scenario 2).
+Alternatively, yield may be higher near field boundaries (Scenario 3) if they provide helpful ecosystem services, as these also tend to be occur close to the field edge [@duelli2003]. 
 For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance [@kowalchuk1995; @garibaldi2011].
-If negative edge effects of the edge improve with distance and ecosystem services decrease with distance, yield may be maximized at intermediate distances where the ecosystem services cancel out negative edge effects [@mitchell2014; @vanVooren2017; Figure \@ref(fig:hypotheses)a].
-This suggests a "Goldilocks" field size, where negative edge effects are canceled out by ecosystem services.
-Alternatively, if ecosystem services are constant with distance (or increase, Figure \@ref(fig:hypotheses)b) then yield will increase with distance from a field boundary and then plateau.
+Depending on the rate at which positive and negative ecosystems services decline with distance, yet another scenario is possible.
+Yield may also be maximized at intermediate distances (Scenario 4), suggesting a "Goldilocks" field size where negative edge effects are canceled out by ecosystem services.
+However, this scenario is rarely considered within studies of SNL or boundary effects [but see @mitchell2014; @vanVooren2017].
 
@@
 ```{r hypotheses, echo=FALSE, fig.cap='Potential yield patterns, depending on ecosystem service effects with distance.
-In both figures, negative environmental effects (edge effects) are shown in red, and decrease with distance from the edge of the crop, creating an increase in yield with distance.
-The left figure shows a decline in ecosystem services with distance, leading to an intermediate peak in yield, while the right figure shows an increase in services with distance, leading to a peak at the centre of the field (note: Scenario 2 also occurs if ecosystem services do not change with distance).', out.width='100%'}
+Scenario 1: no effect of boundary.
+Scenario 2: negative effect on yield.
+Scenario 3: positive effect on yield.
+Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.', out.width='100%'}
   include_graphics('../Figures/ExamplePlots/hypotheses.png',dpi=NA)
@@
 
-Boundary effects are difficult to observe without large amounts of crop yield data, which is typically expensive and time-consuming to collect.    
-Studies of semi-natural field boundaries on crop yield tend to suffer from limited spatial scope and small sample sizes, limiting inference and reducing generality [@baker2018; but see also @redhead2020].
+Boundary effects are difficult to observe without large amounts of crop data, which is typically expensive and time-consuming to collect.    
+Because of this, studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality [@baker2018; but see also @redhead2020].
 Meta-analyses are a common solution to this problem [e.g. @albrecht2020; @lowe2021], but precision yield data holds enormous promise because of its relatively common usage and low operating costs. 
-However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
+However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
 Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems [@redhead2020; @hunicken2021]; typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
@@
 
-Field boundaries were automatically digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC. 
-Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g. flooding during some years).
-Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
-This makes accurate and consistent classification of field boundaries difficult.
-We defined the following general categories for field boundaries:
+Field boundaries were digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data [@AAFC_data]. 
+Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions.
+For example, ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
+This makes accurate and consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:
     1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide) 
@@
 Yield data from each field contains both random and systematic errors that must be modeled in order to reveal underlying changes in yield.
-We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation.
 Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, previous agricultural practices, and other factors not directly related to boundary distance.
-<!-- - such as strip farming -->
+To account for this, we used additive models of the effect of boundary distance on crop yield while accounting for within-field spatial variation.
+
 <!-- Ground speed is extremely important to yield monitor accuracy, with low ground speed registering higher yields. -->
@@
 <!-- This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time. -->
-To model this, we fit the following model to each field-year of data:
+We fit the following model to each field-year of data:
 
@@
 Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g. Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
-Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but using a reduced number of dimensions and computation time.
+Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
 We used 5 basis dimensions for the distance smoother and 60 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \@ref(fig:hypotheses)).
@@
 This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.
-extremely small p-values), and c) results did not change strongly when additional basis functions were added, we decided to leave the models as-is.
+extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 60), we decided to leave the models as-is.
 Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
@@
 
-To consider results from all field-level models together, we fit models independently and then fit "overall" smoothers as averages of the field-level smoothers.
-However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models:
-  1. Extract single posterior sample (_rnorm_ in R) of smoother parameters from each field-level model using coefficient estimates and standard error
-  2. Use posterior sample to create a new smoother from each field
-  3. Fit new model of new smoothers from all fields, and save this "meta-smoother"
+To consider results from all field-level models together, we fit models of each field-year independently and then fit overall smoothers as averages of the field-level smoothers.
+This does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping:
+  1. Extract single posterior sample (_rnorm_ in R) of smoother parameters from each field-level model using coefficient estimates and standard errors
+  2. Use posterior sample to simulate a new smoother from each field
+  3. Fit model of simulated smoothers from all fields, and save this "meta-smoother"
   4. Repeat 1000 times, and calculate coverage intervals (5-95% percentiles) on saved meta-smoothers
@@
 
-The filtered data from each field still contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
+The filtered data from each field contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
 To illustrate this, we show the results from a single wheat field, where we show the relative impact of boundary distance and spatial variation on both yield mean and yield SD ("patchiness").
-Figure \@ref(fig:spatialSmooths)a shows the raw data collected from a single field, demonstrating a large amount of variation even among the filtered data.
-Yield mean and SD tended to decrease with distance from field boundaries, but there were no consistent differences between boundary types (Figure \@ref(fig:distSmooths)).
+Figure \@ref(fig:spatialSmooths)a shows the filtered data collected from a single field, demonstrating a large amount of variation even after filtering.
+Yield mean tended to increase with distance from field boundaries while yield SD tended to decrease, but there were no consistent differences between boundary types (Figure \@ref(fig:distSmooths)).
 Finally, the random spatial smoothers in Figure \@ref(fig:spatialSmooths)b and c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.
@@
 
-```{r distSmooths, echo=FALSE, fig.cap='Distance smoothers from a single field, showing a negative effect of distance on average yield (first row).
-The variability smoothers show an intermediate reduction in variability at around 90 m.', out.width='100%'}
+```{r distSmooths, echo=FALSE, fig.cap='Distance smoothers from a single field, showing a positive saturating effect of distance on average yield (first row), while the variability smoothers show a general decrease in yield SD (second row).', out.width='100%'}
   include_graphics('../Figures/ExamplePlots/distSmooths.png',dpi=NA)
@@
 
-Across all fields, standard (grassy border) field boundaries were the most common boundary type, followed by shelterbelts and other crops (Figure \@ref(fig:boundaryTypes)).
+Across all fields, standard field boundaries (grassy borders) were the most common boundary type, followed by shelterbelts and other crops (Figure \@ref(fig:boundaryTypes)).
 <!-- Not sure if we should talk about this. Not very many fields from southern regions, so it's hard to compare. Perhaps mention in Discussion -->
@@
 
-Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both <0.0001), meaning that within-field variability was the primary driver of yield across canola, wheat, and pea crops (Table \@ref(tab:smoothSummary), row 6).
+Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both <0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table \@ref(tab:smoothSummary), row 6).
 Distances from boundary types were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
@@
 canolaMods %>% group_by(Smoother) %>% 
-  summarize(N=n(),mean=mean(pval),propSig=sum(pval<(0.05/N))/n()) %>% 
+  summarize(N=n(),meanChi=mean(Chisq),meanP=mean(pval),propSig=sum(pval<(0.05/N))/n()) %>% 
   ungroup() %>% 
@@
   )) %>% 
-  select(-N,-Smoother) %>% relocate(type,par,mean,propSig) %>% 
-  pivot_wider(id_cols=type,names_from=par,values_from=c(mean,propSig)) %>% 
-  relocate(type,mean_Mean,propSig_Mean,mean_Variability,propSig_Variability) %>% 
+  select(-N,-Smoother) %>% relocate(type,par,meanChi,meanP,propSig) %>% 
+  pivot_wider(id_cols=type,names_from=par,values_from=c(meanChi,meanP,propSig)) %>% 
+  relocate(type,meanChi_Mean,meanP_Mean,propSig_Mean,meanChi_Variability,meanP_Variability,propSig_Variability) %>% 
   mutate(type=c('Bare','Grassland','Other Crop','Shelterbelt','Standard','Wetland','Spatial Smoother')) %>% 
-  kable(.,format='latex',caption='Mean p-values and proportion of p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all fields.',col.names = c('Smoother Type','Average p-value','Proportion <0.05','Average p-value','Proportion <0.05'),
-        digits = c(1,3,3,3,3),escape = FALSE) %>% 
-  add_header_above(c('', 'Mean' = 2, 'Variability' = 2))
+  kable(.,format='latex',caption='Mean $\\chi^2$, p-values, and proportion of smoother p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all models.',col.names = c('Smoother Type','$\\chi^2$-value','p-value','Prop.
+<0.05','$\\chi^2$-value','p-value','Prop.
+<0.05'),
+        digits = c(1,1,3,3,1,3,3),escape = FALSE) %>% 
+  add_header_above(c('', 'Mean Yield' = 3, 'Yield Variability' = 3))
 

### 2021-10-18 Manuscript/manuscript.tex
Worked on paper draft

@@
  \end{frontmatter}
-
-Potential journals: Agriculture Ecosystems and Environment, Basic and Applied Ecology, Ecological Economics, Journal of Applied Ecology (probably won't like it)
-
-Potential reviewers: Teja Tscharntke, Ignasi Bartomeus, Ben Woodcock, John Redhead, some Canadian/Albertan reviewers who know the crop systems?
 
@@
 SNL in agroecosystems is important for both agricultural production and conservation (Prieto-Benítez \& Méndez 2011), and can create better biotic and abiotic conditions for crop growth (Case \emph{et al.} 2019; Weninger \emph{et al.} 2021; but see Lowe \emph{et al.} 2021).
-SNL constitutes habitat for mobile organisms, and can act as sources of ecosystem services such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this can vary depending on the crop system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017) and landscape context (Marja \emph{et al.} 2019; Boetzl \emph{et al.} 2020).
-They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion (Kort 1988; Brandle \emph{et al.} 2004; Wratten \emph{et al.} 2012).
-North American agroecosystems have larger fields, different crop varieties, and agronomic practices, all of which could negate or confound potential effects of SNL.
-Semi-natural field boundaries (e.g.~grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are of particular interest, as they are extremely common and require less land than field-sized agricultural set-asides.
+SNL acts as habitat for mobile organisms, and can act as a source of ecosystem services, such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this depends on both the crop system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017) and landscape context (Marja \emph{et al.} 2019; Boetzl \emph{et al.} 2020).
+SNL can also create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion (Kort 1988; Brandle \emph{et al.} 2004; Wratten \emph{et al.} 2012).
+Field boundaries (e.g.~grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are SNL of particular interest, as they are extremely common and require less land than field-sized agricultural set-asides.
 
-Field boundaries can cause both positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
-Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds.
-Field boundaries can also provide helpful ecosystem services, such as pest suppression or pollination, and these also tend to be occur close to the field edge (Duelli \& Obrist 2003).
+Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
+Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure \ref{fig:hypotheses}, Scenario 2).
+Alternatively, yield may be higher near field boundaries (Scenario 3) if they provide helpful ecosystem services, as these also tend to be occur close to the field edge (Duelli \& Obrist 2003).
 For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Kowalchuk \& Jong 1995; Garibaldi \emph{et al.} 2011).
-If negative edge effects of the edge improve with distance and ecosystem services decrease with distance, yield may be maximized at intermediate distances where the ecosystem services cancel out negative edge effects {[}Mitchell \emph{et al.} (2014); Van Vooren \emph{et al.} (2017); Figure \ref{fig:hypotheses}a{]}.
-This suggests a ``Goldilocks'' field size, where negative edge effects are canceled out by ecosystem services.
-Alternatively, if ecosystem services are constant with distance (or increase, Figure \ref{fig:hypotheses}b) then yield will increase with distance from a field boundary and then plateau.
+Depending on the rate at which positive and negative ecosystems services decline with distance, yet another scenario is possible.
+Yield may also be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services.
+However, this scenario is rarely considered within studies of SNL or boundary effects (but see Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).
 
@@
 \includegraphics[width=1\linewidth]{../Figures/ExamplePlots/hypotheses} \caption{Potential yield patterns, depending on ecosystem service effects with distance.
-In both figures, negative environmental effects (edge effects) are shown in red, and decrease with distance from the edge of the crop, creating an increase in yield with distance.
-The left figure shows a decline in ecosystem services with distance, leading to an intermediate peak in yield, while the right figure shows an increase in services with distance, leading to a peak at the centre of the field (note: Scenario 2 also occurs if ecosystem services do not change with distance).}\label{fig:hypotheses}
+Scenario 1: no effect of boundary.
+Scenario 2: negative effect on yield.
+Scenario 3: positive effect on yield.
+Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.}\label{fig:hypotheses}
 \end{figure}
 
-Boundary effects are difficult to observe without large amounts of crop yield data, which is typically expensive and time-consuming to collect.\\
-Studies of semi-natural field boundaries on crop yield tend to suffer from limited spatial scope and small sample sizes, limiting inference and reducing generality (Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
+Boundary effects are difficult to observe without large amounts of crop data, which is typically expensive and time-consuming to collect.\\
+Because of this, studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality (Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
 Meta-analyses are a common solution to this problem (e.g.
 Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs.
-However, its use is limited for several reasons, namely: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistics.
+However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
 Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems (Redhead \emph{et al.} 2020; Hünicken \emph{et al.} 2021); typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
@@
 85\% of the crop types where either wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}), three of the most common crops in rotation in Alberta (Statistics Canada 2014).
-The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (91), canola (118), or peas (38), for a total of 247 field-years of data.
-Individual fields contained between 1 and 6 years of data (mean: 3), containing a total of 13.9 million data points.
+The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
+Individual fields contained between 1 and 6 years of data (mean: 3.1), containing a total of 14.4 million data points.
 
@@
 
-Field boundaries were automatically digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC.
-Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g.~flooding during some years).
-Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
-This makes accurate and consistent classification of field boundaries difficult.
-We defined the following general categories for field boundaries:
+Field boundaries were digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data (Agriculture and Agri-Food Canada 2018).
+Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions.
+For example, ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
+This makes accurate and consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:
 1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
@@
 Yield data from each field contains both random and systematic errors that must be modeled in order to reveal underlying changes in yield.
-We fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation.
 Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, previous agricultural practices, and other factors not directly related to boundary distance.
-To model this, we fit the following model to each field-year of data:
+To account for this, we used additive models of the effect of boundary distance on crop yield while accounting for within-field spatial variation.
+
+We fit the following model to each field-year of data:
 
@@
 Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
-Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but using a reduced number of dimensions and computation time.
+Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
 We used 5 basis dimensions for the distance smoother and 60 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \ref{fig:hypotheses}).
@@
 The residuals in most of the models displayed both spatial autocorrelation in the residuals, indicating ``micro-scale'' variation that was not captured by the spatial smoothers (Cressie \& Wikle 2011; Wikle \emph{et al.} 2019).
-This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.~extremely small p-values), and c) results did not change strongly when additional basis functions were added, we decided to leave the models as-is.
+This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.~extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 60), we decided to leave the models as-is.
 Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
@@
 
-To consider results from all field-level models together, we fit models independently and then fit ``overall'' smoothers as averages of the field-level smoothers.
-However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models:
-1. Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard error
-2. Use posterior sample to create a new smoother from each field
-3. Fit new model of new smoothers from all fields, and save this ``meta-smoother''
+To consider results from all field-level models together, we fit models of each field-year independently and then fit overall smoothers as averages of the field-level smoothers.
+This does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping:
+1. Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard errors
+2. Use posterior sample to simulate a new smoother from each field
+3. Fit model of simulated smoothers from all fields, and save this ``meta-smoother''
 4. Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers
@@
 
-The filtered data from each field still contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
+The filtered data from each field contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
 To illustrate this, we show the results from a single wheat field, where we show the relative impact of boundary distance and spatial variation on both yield mean and yield SD (``patchiness'').
-Figure \ref{fig:spatialSmooths}a shows the raw data collected from a single field, demonstrating a large amount of variation even among the filtered data.
-Yield mean and SD tended to decrease with distance from field boundaries, but there were no consistent differences between boundary types (Figure \ref{fig:distSmooths}).
+Figure \ref{fig:spatialSmooths}a shows the filtered data collected from a single field, demonstrating a large amount of variation even after filtering.
+Yield mean tended to increase with distance from field boundaries while yield SD tended to decrease, but there were no consistent differences between boundary types (Figure \ref{fig:distSmooths}).
 Finally, the random spatial smoothers in Figure \ref{fig:spatialSmooths}b and c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/distSmooths} \caption{Distance smoothers from a single field, showing a negative effect of distance on average yield (first row).
-The variability smoothers show an intermediate reduction in variability at around 90 m.}\label{fig:distSmooths}
+\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/distSmooths} \caption{Distance smoothers from a single field, showing a positive saturating effect of distance on average yield (first row), while the variability smoothers show a general decrease in yield SD (second row).}\label{fig:distSmooths}
 \end{figure}
@@
 
-Across all fields, standard (grassy border) field boundaries were the most common boundary type, followed by shelterbelts and other crops (Figure \ref{fig:boundaryTypes}).
+Across all fields, standard field boundaries (grassy borders) were the most common boundary type, followed by shelterbelts and other crops (Figure \ref{fig:boundaryTypes}).
 
@@
 
-Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variability was the primary driver of yield across canola, wheat, and pea crops (Table \ref{tab:smoothSummary}, row 6).
+Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table \ref{tab:smoothSummary}, row 6).
 Distances from boundary types were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
@@
 
-\caption{\label{tab:smoothSummary}Mean p-values and proportion of p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all fields.}
+\caption{\label{tab:smoothSummary}Mean $\chi^2$, p-values, and proportion of smoother p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all models.}
 \centering
-\begin{tabular}[t]{l|r|r|r|r}
+\begin{tabular}[t]{l|r|r|r|r|r|r}
 \hline
-\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{Mean} & \multicolumn{2}{c}{Variability} \\
-\cline{2-3} \cline{4-5}
-Smoother Type & Average p-value & Proportion <0.05 & Average p-value & Proportion <0.05\\
+\multicolumn{1}{c|}{} & \multicolumn{3}{c|}{Mean Yield} & \multicolumn{3}{c}{Yield Variability} \\
+\cline{2-4} \cline{5-7}
+Smoother Type & $\chi^2$-value & p-value & Prop. <0.05 & $\chi^2$-value & p-value & Prop. <0.05\\
 \hline
-Bare & 0.036 & 0.875 & 0.055 & 0.857\\
+Bare & 154.7 & 0.036 & 0.877 & 121.4 & 0.054 & 0.860\\
 \hline
-Grassland & 0.079 & 0.794 & 0.061 & 0.794\\
+Grassland & 303.1 & 0.079 & 0.794 & 93.2 & 0.061 & 0.794\\
 \hline
-Other Crop & 0.063 & 0.860 & 0.044 & 0.840\\
+Other Crop & 174.9 & 0.063 & 0.860 & 104.5 & 0.044 & 0.840\\
 \hline
-Shelterbelt & 0.014 & 0.935 & 0.059 & 0.815\\
+Shelterbelt & 328.1 & 0.014 & 0.936 & 137.6 & 0.058 & 0.817\\
 \hline
-Standard & 0.021 & 0.938 & 0.055 & 0.839\\
+Standard & 281.9 & 0.021 & 0.938 & 159.6 & 0.057 & 0.832\\
 \hline
-Wetland & 0.045 & 0.833 & 0.013 & 0.857\\
+Wetland & 148.9 & 0.045 & 0.833 & 106.7 & 0.013 & 0.857\\
 \hline
-Spatial Smoother & 0.000 & 1.000 & 0.000 & 1.000\\
+Spatial Smoother & 17357.5 & 0.000 & 1.000 & 4739.1 & 0.000 & 1.000\\
 \hline
@@
 
-Pea crops were less well-represented in our data (N = 38 fields, compared to ), and the effects of boundaries tended to be less consistent.
+Pea crops were less well-represented in our data (N = 39 fields, compared to ), and the effects of boundaries tended to be less consistent.
 Average pea yields (Figure \ref{fig:peaPlot}, top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-AAFC_data}{}%
+Agriculture and Agri-Food Canada. (2018). \emph{{Annual Crop Inventory}}. Available at: \url{https://open.canada.ca/data/en/dataset/ba2645d5-4458-414d-b196-6303ac06c1c9}. Last accessed.
+
 \leavevmode\hypertarget{ref-albrecht2020}{}%

### 2021-10-18 yield simulations.R

@@ -54,35 +54,67 @@ curve(leq(x,1,exp(-2))*leq(x,1,-0.1),0,100,ylim=c(0,1)) #Both
 curve(leq2(x,1,exp(-2),1,-0.1),0,100,ylim=c(0,1)) #Both
 
 
-png(paste0('./Figures/ExamplePlots/hypotheses.png'),width=12,height=6,units='in',res=200)
+png(paste0('./Figures/ExamplePlots/hypotheses.png'),width=10,height=10,units='in',res=200)
 
 # png("C:\\Users\\Samuel\\Desktop\\hypothesisPlot.png",width=12,height=6,units='in',res=200) #Path on Multivac
 
 savepar <- par(no.readonly = TRUE)
-par(mfrow=c(1,2),mai=c(1,0.5,0.8,0.4))
-#Scenario 1 - beneficial bugs decrease with distance into field
-a1 <- -2; b1 <- 0.2; a2 <- 3; b2 <- -0.03
+par(mfrow=c(2,2),mai=c(1,0.5,0.8,0.4))
 
+
+#Scenario: no effect environmental effect
+a1 <- 0; b1 <- 0
 plot(0,0,type='n',xlim=c(0,100),ylim=c(0,1),axes=FALSE,
      ylab=NA,xlab='Distance from Edge',main='Scenario 1',xaxt='n')
 axis(2,at = c(0,1),labels = c('Low','High')); axis(1,at=seq(0,100,by=20))
+curve(leq(x,a1,b1)+0.05,0,100,xlim=range(dist),col='red',add=TRUE,lwd=2)
+curve(leq(x,a1,b1)-0.05,0,100,xlim=range(dist),col='black',add=TRUE,lwd=2)
+text(70,0.6,'Edge effect',col='red'); text(70,0.4,'Crop Yield')
 
-curve(leq2(x,a1,b1,a2,b2),0,100,lwd=4,add=TRUE)#,main='Scenario 1: bugs - with dist')
-curve(leq(x,a1,b1),0,100,xlim=range(dist),col='red',add=TRUE,lwd=2)
-curve(leq(x,a2,b2),0,100,xlim=range(dist),col='blue',add=TRUE,lwd=2)
-text(10,0.83,'Ecosystem\nservice\neffect',col='blue'); text(70,0.95,'Environmental effect',col='red'); text(80,0.75,'Crop Yield')
-
[…]

### 2021-10-20 gam_TMB/examplePlot.R
Added plots of basis functions

@@ -1,0 +1,33 @@
+#Example plot of basis functions
+
+library(mgcv)
+library(tidyverse)
+
+N <- 100
+dat <- data.frame(x=runif(N,-10,10)) %>% 
+  mutate(y=1-0.5*x+0.2*x^2+rnorm(N)) %>% 
+  arrange(x)
+
+plot(y~x,data=dat)
+
+m1 <- gam(y~s(x,k=6,bs = 'cr'),data=dat)
+
+par(mfrow=c(2,1),mar=c(4,4.1, 1, 1))
+plot(dat$x,model.matrix(m1)[,2],col='red',type='l',ylim=c(-0.6,1),xlab='Distance',ylab='Z')
+lines(dat$x,model.matrix(m1)[,3],col='blue')
+lines(dat$x,model.matrix(m1)[,4],col='green')
+lines(dat$x,model.matrix(m1)[,5],col='magenta')
+lines(dat$x,model.matrix(m1)[,6],col='grey40')
+
+plot(dat$x,model.matrix(m1)[,c(2:6)]%*%coef(m1)[2:6],type='l',lwd=2,ylim=c(-10,20),
+     xlab='Distance',ylab='Zu')
+lines(dat$x,model.matrix(m1)[,2]*coef(m1)[2],col='red')
+lines(dat$x,model.matrix(m1)[,3]*coef(m1)[3],col='blue')
+lines(dat$x,model.matrix(m1)[,4]*coef(m1)[4],col='green')
+lines(dat$x,model.matrix(m1)[,5]*coef(m1)[5],col='magenta')
+lines(dat$x,model.matrix(m1)[,6]*coef(m1)[6],col='grey40')
+
+
+m1$smooth[[1]]
+
+Z <- m1$smooth[[1]]$UZ

### 2021-10-20 helperFunctions.R

@@ -584,14 +584,18 @@ runModII <- function(i,dS,nSubSamp=50000,kPar=c(5,80,5,80),useClosest=TRUE,model
   NrawDat <- nrow(dat)
   
   #Filter data
-  dat <- dat %>% mutate(vegaFilt = vegaFilter(.,DryYield,nDist = 30)) #Vega et al 2019 spatial "inlier" filter - takes about a minute
+  dat <- dat %>% 
+    #Vega et al 2019 spatial "inlier" filter - takes about a minute
+    mutate(vegaFilt = vegaFilter(.,DryYield,nDist = 30)) 
   
   dat <- dat %>% 
-    mutate(noBS = DryYield<ifelse(cropType=='Wheat',10.75,8)) %>% #JP recommends maximum filters of 160 bu (10.75 T/ha) for wheat, 120 bu (8 T/ha) for peas & canola
+    #BS filter - JP recommends maximum filters of 160 bu (10.75 T/ha) for wheat, 120 bu (8 T/ha) for peas & canola
+    mutate(noBS = DryYield<ifelse(cropType=='Wheat',10.75,8)) %>% 
     mutate(Qfilt = QuantileFilter(DryYield,q=0.98)) %>% #Trim dry yield outliers
     mutate(bFilt = bearingFilter(TrackAngle,q=0.98)) %>% #Trim extreme bearing changes (turning)
     mutate(speedFilt = QuantileFilter(Speed,q=0.98)) %>% #Trim absolute speed outliers
-    mutate(dSpeedFilt = dSpeedFilter(Speed,l=c(-2,-1,1,2),perc = 0.2)) %>% #Trim speed differences (>20% change 2 steps forward and backward, suggested by Lyle et al 2014)
+    #Trim speed differences (>20% change 2 steps forward and backward, suggested by Lyle et al 2014)
+    mutate(dSpeedFilt = dSpeedFilter(Speed,l=c(-2,-1,1,2),perc = 0.2)) %>% 
     mutate(posFilt = posFilter(.,q=0.98)) %>% #Trim points that are far away from eachother
     #Combine filter criteria
     mutate(allFilt = noBS & vegaFilt & Qfilt & bFilt & speedFilt & dSpeedFilt & posFilt) %>%  
@@ -1077,7 +1081,6 @@ vegaFilter <- function(data,ycol,pvalCutoff=0.05,nDist=40){
   # nDist <- 30
   # yield <- data$DryYield
   
-  # if(is.null(ycol)) stop('Specify yield column')
[…]

### 2021-10-20 yield simulations.R

@@ -119,6 +119,17 @@ par(savepar)
 
 dev.off()
 
+#Version of figure for Lan:
+plot(0,0,type='n',xlim=c(0,100),ylim=c(0,1),axes=FALSE,
+     ylab=NA,xlab='Distance from Edge',xaxt='n')
+axis(2,at = c(0,1),labels = c('Low','High')); axis(1,at=c(0,100),labels=c('Near','Far'))
+
+curve(leq2(x,a1,b1,a2,b2)+0.1,0,100,lwd=2,add=TRUE)#,main='Scenario 3: bugs - with dist')
+curve(leq(x,a1,b1)-0.1,0,100,xlim=range(dist),col='red',add=TRUE,lwd=2)
+curve(leq(x,a2,b2)+0.2,0,100,xlim=range(dist),col='blue',add=TRUE,lwd=2)
+text(70,0.65,'Ecosystem\nservice effect',col='blue'); text(70,0.95,'Edge effect',col='red'); text(70,0.3,'Crop Yield')
+
+
 #Approximately how many basis functions needed to represent this?
 library(tidyverse)
 library(mgcv)

### 2021-10-22 Manuscript/manuscript.Rmd
Added appendix with filter

@@
 Scenario 1: no effect of boundary.
-Scenario 2: negative effect on yield.
-Scenario 3: positive effect on yield.
+Scenario 2: negative edge effect on yield.
+Scenario 3: positive ecosystem service effect on yield.
 Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.', out.width='100%'}
@@
 Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows [@griffin2007; @whelan2013], or when the combine is accelerating or decelerating [@arslan2002].
-Therefore, we used the following filtering procedure to clean the data prior to model fitting:
+Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix A for R code):
 
@@
 
-Pea crops were less well-represented in our data (N = `r sum(with(datSource,use&modelComplete2&crop=='Peas'))` fields, compared to ), and the effects of boundaries tended to be less consistent. 
+Pea crops were less well-represented in our data (N = `r sum(with(datSource,use&modelComplete2&crop=='Peas'))` fields, compared to ), and therefore the effects of boundaries tended to be less consistent.
 Average pea yields (Figure \@ref(fig:peaPlot), top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
@@
 
-At the level of individual fields, both average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
-Yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat.
-This increase tended to be small in magnitude in canola crops, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \@ref(fig:hypotheses)).
+Within individual fields, both the average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
+Mean yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat, while peas showed no evidence of this.
+However, this intermediate increase was small in canola crops, meaning that ecosystem services likely did not vary strongly with distance from field boundaries (Scenario 2, Figure \@ref(fig:hypotheses)).
 There was a fairly large (0.75 T/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
@@
 
+<!-- Need a paragraph on wheat + shelterbelt effect -->
+
 <!-- For example, average canola yield was about 0.5 T/ha lower close to Standard field edges, and yield SD was about 0.3 T/ha SD higher. -->
-
 Why did we find few consistent effects of boundary type on crop yield, aside from shelterbelts?

### 2021-10-22 Manuscript/supplemental.Rmd
Added appendix with filter

--- 2021-09-01 Manuscript/supplemental.Rmd
@@
 
-# Appendix A: Supplementary Material {-}
+# Appendix A: R Code for Filtering Procedure {-}
 
-```{r , echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on canola yield.', out.width='100%'}  
-  include_graphics('../Figures/groundSpeed_canola.png',dpi=NA)
+```{r filterCode, eval=FALSE, include=TRUE}
+# Helper functions -----------------
+
+#"Inlier" spatial filtering procedure from Vega et al 2019, rewritten to work with sf + dplyr. Returns boolean
+vegaFilter <- function(data,ycol,pvalCutoff=0.05,nDist=40){
+  require(sf)
+  require(sp)
+  require(spdep)
+  require(tidyverse)
+
+  if(!any(class(data) %in% 'sf')) stop('Dataframe must be sf object')
+
+  coords <- st_coordinates(data) #Get coordinates in matrix form
+
+  #Get neighbourhood weights from 0 to ndist meters
+  nWeights <- dnearneigh(coords,0,nDist)
+
+  if(any(sapply(nWeights,length)==1)) warning('Some points had no neighbours and were removed')
+
+  #Get neighbourhood indices for each point (which other points are in this point's neighbourhood?)
+  nIndices <- nb2listw(nWeights, style = "W",zero.policy = TRUE)
+
+  yield <- pull(data,{{ycol}}) #Get yield data column
+
+  #Local Moran's I for each neighbourhood
+  LM <- localmoran(yield,nIndices,p.adjust.method="bonferroni",alternative ="less")
+
+  results <- data.frame(LM) %>% 
+    rename('pval'=contains('Pr.z.')) %>% 
+    mutate(keepThese= Ii > 0 | pval > pvalCutoff) %>% #Filter negative Ii and pvals < 0.05
+    mutate(keepThese=ifelse(is.na(keepThese),FALSE,keepThese)) #NAs (points had no neighbours)
+
+  ret <- pull(results)
+  return(ret)
+}
+
+#Function to filter anything above a certain z-score. Returns boolean
+ZscoreFilter <- function(x,z=3){
+  xmean <- mean(x,na.rm=TRUE)
+  xsd <- sd(x,na.rm=TRUE)
+  zscore <- abs(((x-xmean)/xsd))
+  zscore<z
+}
+
+#Function to filter anything above certain quantiles. Returns boolean
+QuantileFilter <- function(x,quant=0.99){ 
+  l <- c((1-quant)/2,1-(1-quant)/2) #Symmetric quantiles
+  x>quantile(x,l[1]) & x<quantile(x,l[2]) 
+}
+
+#Function to filter large changes in bearing. Returns boolean unless returnDiffs==TRUE
+bearingFilter <- function(bearing,q=NULL,z=NULL,returnDiffs=FALSE){
+  if(!xor(is.null(q),is.null(z))&!returnDiffs) stop('Input quantiles or Z-score')
+  #Difference in compass bearings (in degrees)
+  bearingDiff <- function(x1,x2){
+    x <- x1-x2
+    x <- ifelse(abs(x)>180,x-(360*sign(x)),x) #Angle differences can't be >180
+    return(x)
+  }
+
+  #Looks 1 point ahead and behind
+  bd <- cbind(bearingDiff(lag(bearing),bearing),
+              bearingDiff(lead(bearing),bearing))
+  bd <- apply(bd,1,function(x) max(abs(x),na.rm=TRUE)*sign(x[which.max(abs(x))])) #Maximum bearing difference ahead and behind
+
+  if(returnDiffs) return(bd) #Return bearing differences only, without filtering
+
+  if(!is.null(q)){
+    ret <- QuantileFilter(bd,q=q)
+  } else {
+    ret <- ZscoreFilter(bd,z=z)
+  }
+  return(ret)
+}
+
+#Positional difference filter - filters out very distant and very close points. Returns boolean unless returnDiffs==TRUE
+posFilter <- function(data,q=NULL,returnDiffs=FALSE){ 
+  require(sf)
+  if(is.null(q)&!returnDiffs) stop('Input upper quantile')
+
+  if(units(st_distance(data[1,],data[2,]))$numerator!='m') warning('Position differences not in meters')
+  coords <- st_coordinates(data) #Get coordinates
+  pdiff <- sapply(1:(nrow(coords)-1),function(i) as.numeric(dist(coords[i:(i+1),]))) #Distances between points
+  pdiff <- cbind(c(pdiff,NA),c(NA,pdiff)) #Forward and backward lags
+  pdiff <- apply(pdiff,1,max,na.rm=TRUE) #Maximum distance ahead and behind
+  if(returnDiffs){
+    return(pdiff)
+  } else {
+    return(QuantileFilter(pdiff,q)) #Note: uses 2-sided quantiles
+  }
+}
+
+#Filter for (forward and backward) lagged speed differences. Returns boolean
+dSpeedFilter <- function(speed,l=c(-1,1),perc=0.2){
+
+  llist <- sapply(l,function(x) (lag2(speed,x)-speed)/lag2(speed,x)) #Matrix of % diffs
+
+  #Are any lagged speed values > percent change threshold?
+  ret <- !apply(llist,1,function(y) any(abs(y)[!is.na(y)]>perc))
+
+  return(ret)
+}
+
+#Overloaded lag function that takes negative values
+lag2 <- function(x,n){
+  require(dplyr)
+  if(n==0) {
+    return(x) #No lag
+  } else if(n>0){
+    lag(x,n) #Positive lag
+  } else {
+    lead(x,abs(n)) #Negative lag
+  } 
+}
+
+#Filtering pipeline  --------------------------
+
+library(tidyverse)
+library(sf)
+
+#Note: dat == sf dataframe of yield data with columns: DryYield (T/ha), TrackAngle (compass bearing), Speed (km/hr), cropType (Wheat, Canola, Peas)
+
+dat <- dat %>% 
+  #Vega et al 2019 spatial "inlier" filter - takes a few minutes...
+  mutate(vegaFilt = vegaFilter(.,DryYield,nDist = 30))
+
+dat <- dat %>% 
+  #"BS" filter - removes values >10.75 T/ha for wheat, >8 T/ha for peas & canola
+  mutate(noBS = DryYield<ifelse(cropType=='Wheat',10.75,8)) %>% 
+  mutate(Qfilt = QuantileFilter(DryYield,q=0.98)) %>% #Trim dry yield outliers
+  mutate(bFilt = bearingFilter(TrackAngle,q=0.98)) %>% #Trim extreme bearing changes (turning)
+  mutate(speedFilt = QuantileFilter(Speed,q=0.98)) %>% #Trim absolute speed outliers
+  #Trim speed differences (>20% change 2 steps forward and backward, suggested by Lyle et al 2014)
+  mutate(dSpeedFilt = dSpeedFilter(Speed,l=c(-2,-1,1,2),perc = 0.2)) %>% 
+  mutate(posFilt = posFilter(.,q=0.98)) %>% #Trim points that are far away from eachother
 ```
-
-```{r , echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on wheat yield.', out.width='100%'}  
-  include_graphics('../Figures/groundSpeed_wheat.png',dpi=NA)
-```

### 2021-10-25 Manuscript/manuscript.Rmd
Worked on Discussion

@@
 Boundary effects are difficult to observe without large amounts of crop data, which is typically expensive and time-consuming to collect.    
-Because of this, studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality [@baker2018; but see also @redhead2020].
-Meta-analyses are a common solution to this problem [e.g. @albrecht2020; @lowe2021], but precision yield data holds enormous promise because of its relatively common usage and low operating costs. 
+Because of this, studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality [@brandle1984; @baker2018; but see also @redhead2020].
+Meta-analyses are a common solution to this problem [e.g.
+@albrecht2020; @lowe2021], but precision yield data holds enormous promise because of its relatively common usage and low operating costs [@osorio2018].
 However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
@@
   3. Is there an intermediate distance where yield is maximized or variance is minimized?
-To our knowledge, no other study has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.
+To our knowledge, only [@osorio2018] has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.
 
@@
 
-```{r boundaryTypes, echo=FALSE, fig.cap='Percentage of field boundaries across sampled fields.', out.width='100%'}
+```{r boundaryTypes, echo=FALSE, fig.cap='Boxplot of field boundary compositions across sampled fields.
+The x-axis indicates the boundary type, while the y-axis indicates the percentage of all field boundaries covered by that boundary type.', out.width='100%'}
 
@@
 Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both <0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table \@ref(tab:smoothSummary), row 6).
-Distances from boundary types were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
+Divvstances from boundary types were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
 
@@
 
+While we found an intermediate positive effect of shelterbelts on wheat, other studies have shown variable impacts of shelterbelts, ranging from no effect [@akbar1990; @crawford1998] to a positive effect [@staple1955; @kowalchuk1995; @bao2012].
+This is complicated by the fact that their impact also depends on year-to-year water stress, shelterbelt orientation, as well as the physical structure of the shelterbelt [@pelton1967; @kort1988; @singh1999].
+However, @kowalchuk1995 found that positive effects of shelterbelts on wheat yield occur at around 20 m, while our work shows a peak at around 50 m, and were tied to increased snow deposition and reduced water stress.
+This also may be related to soil deposition from shelterbelts [@deJong1995]
+Strangely, we also found this intermediate yield maximum with distance away from bare (flooded or fallow) field boundaries, indicating that the patterns we found may be at least somewhat related to water stress.
+Shelterbelts can harbour predators or parasitoids of wheat crops [@dong2015; @ward2021]
+
+Our work also did not examine orientation-specific effects driven by prevailing wind (e.g. east-west distance rather than absolute distance)
+
 <!-- For example, average canola yield was about 0.5 T/ha lower close to Standard field edges, and yield SD was about 0.3 T/ha SD higher. -->
-Why did we find few consistent effects of boundary type on crop yield, aside from shelterbelts?
+Wheat and shelterbelts aside, why did we find few consistent effects of boundary type on crop yield?
 Our work examined only the first and last links in the causal chain _Field boundary_ $\rightarrow$ _Insect Abundance_ $\rightarrow$ _Crop Yield_, meaning that are two mediating relationships where the overall relationship could be broken or obscured.

### 2021-10-27 Manuscript/abstract.Rmd
Finished Discussion and submitted to preprint server

@@
-Field boundaries can improve crop yields by creating better abiotic conditions for crop growth (e.g. moisture, temperature), and can act as refuges for beneficial arthropods.
-This suggests that beneficial crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services (e.g.
-pollination or pest suppression) from the field boundary.
-However, there is little large-scale evidence showing this, largely because yield is costly and time-consuming to measure.
-Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is relatively cheap and commonly measured by growers.
+Field boundaries can provide ecosystem services to crops by creating better abiotic conditions for crop growth, and can also act as habitat for beneficial arthropods.
+This suggests that crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services from the field boundary.
+However, there is little large-scale evidence showing this, largely because plot-scale crop yields are costly and time-consuming to measure.
+Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is widespread and data is frequently recorded by growers.
 In this study, we used 252 field-years of yield monitor data from three common crops -- wheat (_Triticum aestivum_), canola (_Brassica napus_), or peas (_Pisum sativum_) -- recorded across Alberta, Canada, and examined how yield varied with distances from common crop boundary types.

### 2021-10-27 Manuscript/manuscript.Rmd
Finished Discussion and submitted to preprint server

--- 2021-10-22 Manuscript/manuscript.Rmd
@@
 ---
-title: "Living on the edge: crop boundary effects on precision crop yields in Alberta, Canada"
+title: "Livin' on the edge: Precision yield data shows evidence of ecosystem services from field boundaries"
 author:
@@
   - landscape;
+  - shelterbelt;
 
@@
     fig_caption: true
+    highlight: haddock
 csl: ecology-letters.csl    
@@
   \usepackage{setspace}
-  \linenumbers
+  %\linenumbers
+
 ---
@@
 
-<!-- Things to do on main machine: -->
-<!-- - Update "Field Locations" map to only have canola, wheat, pea fields -->
-<!-- - Update other figures with more "realistic" estimates - what is the average combine ground speed? -->
+<!-- Potential titles (best to worst): -->
+<!-- "Livin' on the edge: Precision yield data shows evidence of ecosystem services from field boundaries" -->
+<!-- "Field boundaries and ecosystem service spillover: evidence from precision yield data" -->
+<!-- "Living on the edge: field boundary effects on precision yield" -->
+<!-- "Ecosystem services for crop yields: do field boundaries matter?" -->
+<!-- "Crop boundary effects on precision yield data" -->
 
 \newpage
-\doublespacing
+<!-- \doublespacing -->
+\onehalfspacing
 
@@
 Field boundaries (e.g.
-grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are SNL of particular interest, as they are extremely common and require less land than field-sized agricultural set-asides.
+grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are SNL of particular interest, as they are extremely common and require less land than larger agricultural set-asides.
 
@@
 Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure \@ref(fig:hypotheses), Scenario 2).
-Alternatively, yield may be higher near field boundaries (Scenario 3) if they provide helpful ecosystem services, as these also tend to be occur close to the field edge [@duelli2003]. 
-For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance [@kowalchuk1995; @garibaldi2011].
-Depending on the rate at which positive and negative ecosystems services decline with distance, yet another scenario is possible.
-Yield may also be maximized at intermediate distances (Scenario 4), suggesting a "Goldilocks" field size where negative edge effects are canceled out by ecosystem services.
-However, this scenario is rarely considered within studies of SNL or boundary effects [but see @mitchell2014; @vanVooren2017].
+Alternatively, yield may be higher near field boundaries (Scenario 3) if boundaries provide helpful ecosystem services [@duelli2003]. 
+For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance [@garibaldi2011] and wind speed is reduced directly downwind of treelines [@deJong1995], so positive effects will be concentrated close to the boundary.
+Depending on the rate at which positive and negative boundary effects decline with distance, another scenario is also possible.
+Yield may be maximized at intermediate distances (Scenario 4), suggesting a "Goldilocks" field size where negative edge effects are canceled out by ecosystem services [@kowalchuk1995; @helmers2005; @mitchell2014; @vanVooren2017].
 
@@
 
-Boundary effects are difficult to observe without large amounts of crop data, which is typically expensive and time-consuming to collect.    
-Because of this, studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality [@baker2018; but see also @redhead2020].
-Meta-analyses are a common solution to this problem [e.g. @albrecht2020; @lowe2021], but precision yield data holds enormous promise because of its relatively common usage and low operating costs. 
+Unfortunately, data to inform these decisions are often minimal, as observing boundary effects requires large amounts of expensive and time-consuming crop yield data.
+This means that studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality [@brandle1984; @baker2018; but see also @redhead2020].
+Meta-analyses are a common solution to this problem [e.g.
+@albrecht2020; @lowe2021], but precision yield data holds enormous promise because of its relatively common usage and low operating costs [@steel2017; @osorio2018].
 However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
@@
 
-In this paper use a precision yield dataset from Alberta, Canada to ask the following questions:
+In this paper we use a precision yield dataset from Alberta, Canada to ask the following questions:
   1. How does crop yield change with distance from the edge of field?
@@
 This makes accurate and consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:
+
 1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide) 
@@
   \mu = & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
-   & f(\text{Easting}, \text{Northing}, b=60)\\
+   & f(\text{Easting}, \text{Northing}, b=80)\\
   log(\sigma) =  & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
-   & f(\text{Easting}, \text{Northing}, b=60)\\
+   & f(\text{Easting}, \text{Northing}, b=80)\\
   \end{split}
@@
   3. Easting, Northing = distance from centre of field (m)
-  4. Sequence = order of harvest within field (1--N points)
-  5. f(x,b) = penalized thin-plate regression spline, where _x_ is the predictor and _b_ is the number of basis dimensions
+  4. f(x,b) = penalized thin-plate regression spline, where _x_ is the predictor and _b_ is the number of basis dimensions
 
@@
 Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
-We used 5 basis dimensions for the distance smoother and 60 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \@ref(fig:hypotheses)).
+We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \@ref(fig:hypotheses)).
 <!-- The number of basis dimensions was checked with the _gam.check_ function from _mgcv_. -->
@@
 This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.
-extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 60), we decided to leave the models as-is.
+extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 80), we decided to leave the models as-is.
 Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
 That is, combining of fields occurs in patterns (e.g.
-east-to-west) so it is impossible to tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
+east-to-west) so we cannot tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
 We considered the latter to be a more realistic assumption, so we assigned all random variation as spatial in nature. 
-Customized combine driving patterns may be able to detect temporal variation, but this is beyond the scope of this research.
+Non-standard combine driving patterns may be able to gather data in a way that can detect sensor drift, but this is beyond the scope of this research.
 
@@
 This does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping:
-  1. Extract single posterior sample (_rnorm_ in R) of smoother parameters from each field-level model using coefficient estimates and standard errors
-  2. Use posterior sample to simulate a new smoother from each field
-  3. Fit model of simulated smoothers from all fields, and save this "meta-smoother"
-  4. Repeat 1000 times, and calculate coverage intervals (5-95% percentiles) on saved meta-smoothers
+
+  1. Extract single posterior sample (_rnorm_ in R) of smoother parameters from each field-level model using coefficient estimates and standard errors.
+  2. Use posterior sample to simulate a new smoother from each field.
+  3. Fit new model of simulated smoothers from all fields (with a field-year random intercept), and save this "meta-smoother".
+  4. Repeat 1000 times, and calculate coverage intervals (5-95% percentiles) on saved meta-smoothers.
 This gives coverage intervals (CIs) for the "average" smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
@@
 
-```{r spatialSmooths, echo=FALSE, fig.cap='Raw data and spatial smoothers from a single field.
-Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.', out.width='100%'}
+```{r spatialSmooths, echo=FALSE, fig.cap="Raw data and spatial smoothers from a single field.
+Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.
+Field dimensions are approximately 800 x 800 m, but coordinates are hidden to protect data privacy.", out.width='100%'}
   include_graphics('../Figures/ExamplePlots/spatialSmooths.png',dpi=NA)
@@
 Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both <0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table \@ref(tab:smoothSummary), row 6).
-Distances from boundary types were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
+Distances from boundaries were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
 
@@
 Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha  (Figure \@ref(fig:canolaPlot) top row).
-Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50m and remained roughly constant with distance.
+Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50 m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50 m and remained roughly constant with distance.
 The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
 Yield variability (Figure \@ref(fig:canolaPlot) bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
-Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from Wetland boundaries, indicating a potential stabilizing effect.
+Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from wetland boundaries, indicating a potential stabilizing effect, but this was offset by the large increase in yield variability at the edge of wetlands.
 
@@
 
-Pea crops were less well-represented in our data (N = `r sum(with(datSource,use&modelComplete2&crop=='Peas'))` fields, compared to ), and therefore the effects of boundaries tended to be less consistent.
+Pea crops were less well-represented in our data (N = `r sum(with(datSource,use&modelComplete2&crop=='Peas'))` fields), and therefore the effects of boundaries tended to be less consistent. 
 Average pea yields (Figure \@ref(fig:peaPlot), top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
@@
 However, there was a similar (albeit weaker) pattern associated with bare ground, so this pattern may be caused by something other than the presence of shelterbelts.
-This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this across a larger geographic domain in order to inform growers about how best to manage the landscapes surrounding their fields.
+This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this within relevant geographic sub-regions in order to locally inform growers about how best to manage the landscapes surrounding their fields.
 
-<!-- Need a paragraph on wheat + shelterbelt effect -->
+Our work showed an intermediate increase in wheat yield with distance from shelterbelts and bare ground, but what is the cause of these patterns?
+We hypothesize that the potential yield-increasing mechanisms may be a) microclimate effects, b) soil effects, or c) beneficial animal spillover. 
+Shelterbelts can act to capture snow and moderate downwind microclimate, reducing water stress on crop plants and increasing yield [@staple1955; @kowalchuk1995; @bao2012]. 
+However, @kowalchuk1995 found that positive microclimate effects of shelterbelts on wheat yield occur at around 20 m (3-4 times the shelterbelt height), while our work shows a peak at around 50 m.
+This additionally complicated by the fact that water stress varies from year-to-year, and that the impact of the shelterbelt also depends on the orientation to prevailing winds and the physical structure of the shelterbelt [@pelton1967; @kort1988; @singh1999].
+Unexpectedly, we also found this intermediate yield maximum with distance away from bare field boundaries, indicating that the patterns we found may be at least somewhat related to water stress.
+While we consider microclimate to be the most likely cause of the intermediate increase, it may also be related to soil effects from shelterbelts, where erosion is reduced and leaf litter from the shelterbelt can cause an increase in soil deposition [@deJong1995].
+Finally, shelterbelts may also act as sources of predators or parasitoids of wheat crop pests [@dong2015; @ward2021] but their effect depends on pest pressures, which are typically suppressed by pesticide application.
+We recommend that follow-up studies a) use precision yield data to confirm that this pattern exists across broader geographical areas and b) examine the influence of the three potential mechanisms considered above (as well as their interaction).
+<!-- While we found an intermediate positive effect of shelterbelts on wheat, other studies have shown variable effects, ranging from no effect [@akbar1990; @crawford1998]  -->
+<!-- Our work also did not examine orientation-specific effects driven by prevailing wind (e.g. east-west distance rather than absolute distance) -->
 
 <!-- For example, average canola yield was about 0.5 T/ha lower close to Standard field edges, and yield SD was about 0.3 T/ha SD higher. -->
-Why did we find few consistent effects of boundary type on crop yield, aside from shelterbelts?
-Our work examined only the first and last links in the causal chain _Field boundary_ $\rightarrow$ _Insect Abundance_ $\rightarrow$ _Crop Yield_, meaning that are two mediating relationships where the overall relationship could be broken or obscured.
-First, field boundaries may not have any influence on insect abundance if separated from the larger landscape context, or spillover of beneficial organisms may be low.
-Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are low, which may vary yearly, or by location.
+Wheat and shelterbelts aside, why did we find few consistent effects of boundary type on crop yield?
+Our work examined only the first and last links in the causal chain _Field Boundary_ $\rightarrow$ _Ecosystem Service_ $\rightarrow$ _Crop Yield_, meaning that are two mediating relationships where the overall relationship could be broken or obscured.
+First, field boundaries may not have any influence on beneficial insect abundance if separated from the larger landscape context, or spillover of beneficial organisms may be low.
+Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are low, which may vary among year or location.
 <!-- However, our work does show that precision yield data could easily be joined to data on insect abundance to model this more completely. -->
-Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) when organisms in crops provide tangible ecosystem services.
+Abiotic ecosystem services, such as reduced wind speed or temperature, also depend on the structure of the boundary and whether wind or temperature are limiting crop production.
+Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) under what conditions field boundaries provide tangible ecosystem services.
 We consider these reasons below:
@@
 
-SNL generally has a positive effect on spillover, but field boundaries only represent a fraction of the SNL that surrounds a given field.
+SNL generally has a positive effect on beneficial insect abundance, but field boundaries only represent a fraction of the SNL that surrounds a given field.
 Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales [@nguyen2021] which could be particularly important when beneficial arthropods are disperse very broadly [@oberg2007].
@@
 
-Field boundary traits may play an important role in mediating spillover, but our definitions of boundaries were (necessarily) loose, as we used satellite imagery to manually classify them.
-Therefore, it may be that other aspects of field boundaries, such as size or plant community composition, are more important for spillover of ecosystems services into fields.
-For example, larger [@lecq2018; @stasiov2020] or more functionally diverse [@bonifacio2011; @boughey2011] hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction matters to predatory beetle communities [@sutter2018b].
+Field boundary characteristics, such as height or species composition, may play an important role in mediating spillover, but our definitions of boundaries were necessarily loose, as we used satellite imagery to manually classify them.
+Therefore, it may be that other aspects of field boundaries are more important for spillover of ecosystems services into fields.
+For example, larger or more functionally diverse [@bonifacio2011; @boughey2011; @lecq2018; @stasiov2020] hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction influences predatory beetle communities [@sutter2018b].
 These fine-scale boundary characteristics were not measured in our study, but could be used as variables in more biologically-meaningful classifications of field boundaries.
+Finally, the relevant spatial scale of boundary effects are often related to the height or windward orientation of the shelterbelt [@kort1988; @kowalchuk1995], and precision yield data has the potential to show this in far greater detail than previous studies.
 
@@
 Pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services [@zhang2007].
-@gray2014 showed that remnant forest fragments do not provide significant pest control services, but also do not depress yield.
+<!-- @gray2014 showed that remnant forest fragments do not provide significant pest control services, but also do not depress yield. -->
 However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data _alone_.
-Ecosystem services from field boundaries may be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales [@oberg2007; @zamorano2020, but see @robinson2021].
+Ecosystem services from field boundaries may also be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales [@oberg2007; @zamorano2020, but see @robinson2021].
+Lastly, pest pressure or pollination requirements may be low enough that ecosystem service provision does not significantly change yield.
+Canola is largely self-pollinating [at least in hybrid canola varieties, @lindstrom2016] and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop [@woodcock2016].
+Crop pests in most western Canadian fields are suppressed by insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty, meaning spillover of beneficial predators and parasitoids may only matter during outbreak years.
+<!-- no effect [@akbar1990; @crawford1998] -->
 
-<!-- ## Beneficial organisms may not provide ecosystem services -->
-
-Pest pressure or pollination deficit may be low enough that ecosystem service provision does not significantly change yield.
-Canola is largely self-pollinating [at least in hybrid canola varieties, @lindstrom2016] and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop [@woodcock2016].
-Pests are also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty.
-
-Our work is some of the first to use a large precision yield dataset to answer questions about how landscapes influence yield, but this study has only scratched the surface of some of the richest agroecological datasets ever created.
+Our work is some of the first to use a large precision yield dataset to answer questions about how landscapes influence yield.
 Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
 Most fields were located in Central Alberta [Dry Mixedwood region, @ABRegions2006], where boundaries contained a higher proportion of shelterbelts (Figure \@ref(fig:boundaryTypes)), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
-Between-year effects may be important, especially during dry years [@kort1988], but this was not considered in our analysis.
+Between-year effects may be important, especially during hot dry years [@kort1988], but this was not considered in our analysis.
 Future work could examine whether boundaries have differing effects during hot or cool years, as shading or cooling effects may increase yield during hot years but have negative (or no) effect during cooler years.
+Finally, about 50\% of growers in western Canada have collected precision agriculture data, and about 85\% have yield monitoring capability [@steel2017], which means that our work has only scratched the surface of what could be (collectively) one of the richest agroecological datasets ever created.
+We urge agronomists, growers, and ecologists to adopt more wide-spread use of precision yield data, as it represents a vast source of incredibly detailed information which can be used to answer a variety of important questions.
 
@@
 
-<!-- ```{r child = 'supplemental.Rmd'} -->
-<!-- ``` -->
+\singlespacing
+\nolinenumbers
+
+```{r child = 'supplemental.Rmd'}
+```

### 2021-10-27 Manuscript/manuscript.tex
Finished Discussion and submitted to preprint server

@@
 \bibliographystyle{elsarticle-harv}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\newenvironment{Shaded}{}{}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
+\newcommand{\AttributeTok}[1]{#1}
+\newcommand{\BaseNTok}[1]{#1}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
+\newcommand{\ConstantTok}[1]{#1}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.00,1.00}{#1}}
+\newcommand{\DataTypeTok}[1]{#1}
+\newcommand{\DecValTok}[1]{#1}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{#1}
+\newcommand{\FunctionTok}[1]{#1}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.00,1.00}{#1}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{#1}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{1.00,0.25,0.00}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{1.00,0.25,0.00}{#1}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
+\newcommand{\VariableTok}[1]{#1}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
 \usepackage{longtable}
@@
             pdfauthor={},
-            pdftitle={Living on the edge: crop boundary effects on precision crop yields in Alberta, Canada},
+            pdftitle={Livin' on the edge: Precision yield data shows evidence of ecosystem services from field boundaries},
             colorlinks=false,
@@
 % Pandoc header
-\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} \usepackage{setspace} \linenumbers
+\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} \usepackage{setspace} %\linenumbers
 \usepackage{booktabs}
@@
 
-  \title{Living on the edge: crop boundary effects on precision crop yields in Alberta, Canada}
+  \title{Livin' on the edge: Precision yield data shows evidence of ecosystem services from field boundaries}
     \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
@@
   \begin{abstract}
-  Field boundaries can improve crop yields by creating better abiotic conditions for crop growth (e.g.~moisture, temperature), and can act as refuges for beneficial arthropods.
-This suggests that beneficial crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services (e.g.~pollination or pest suppression) from the field boundary.
-However, there is little large-scale evidence showing this, largely because yield is costly and time-consuming to measure.
-Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is relatively cheap and commonly measured by growers.
+  Field boundaries can provide ecosystem services to crops by creating better abiotic conditions for crop growth, and can also act as habitat for beneficial arthropods.
+This suggests that crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services from the field boundary.
+However, there is little large-scale evidence showing this, largely because plot-scale crop yields are costly and time-consuming to measure.
+Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is widespread and data is frequently recorded by growers.
 In this study, we used 252 field-years of yield monitor data from three common crops -- wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}) -- recorded across Alberta, Canada, and examined how yield varied with distances from common crop boundary types.
@@
   \end{abstract}
-   \begin{keyword} precision yield; spillover; landscape;\end{keyword}
+   \begin{keyword} precision yield; spillover; landscape; shelterbelt;\end{keyword}
  \end{frontmatter}
@@
 \newpage
-\doublespacing
+
+\onehalfspacing
 
@@
 SNL can also create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion (Kort 1988; Brandle \emph{et al.} 2004; Wratten \emph{et al.} 2012).
-Field boundaries (e.g.~grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are SNL of particular interest, as they are extremely common and require less land than field-sized agricultural set-asides.
+Field boundaries (e.g.~grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are SNL of particular interest, as they are extremely common and require less land than larger agricultural set-asides.
 
@@
 Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure \ref{fig:hypotheses}, Scenario 2).
-Alternatively, yield may be higher near field boundaries (Scenario 3) if they provide helpful ecosystem services, as these also tend to be occur close to the field edge (Duelli \& Obrist 2003).
-For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Kowalchuk \& Jong 1995; Garibaldi \emph{et al.} 2011).
-Depending on the rate at which positive and negative ecosystems services decline with distance, yet another scenario is possible.
-Yield may also be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services.
-However, this scenario is rarely considered within studies of SNL or boundary effects (but see Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).
+Alternatively, yield may be higher near field boundaries (Scenario 3) if boundaries provide helpful ecosystem services (Duelli \& Obrist 2003).
+For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Garibaldi \emph{et al.} 2011) and wind speed is reduced directly downwind of treelines (de Jong \& Kowalchuk 1995), so positive effects will be concentrated close to the boundary.
+Depending on the rate at which positive and negative boundary effects decline with distance, another scenario is also possible.
+Yield may be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services (Kowalchuk \& de Jong 1995; Helmers \& Brandle 2005; Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).
 
@@
 Scenario 1: no effect of boundary.
-Scenario 2: negative effect on yield.
-Scenario 3: positive effect on yield.
+Scenario 2: negative edge effect on yield.
+Scenario 3: positive ecosystem service effect on yield.
 Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.}\label{fig:hypotheses}
@@
 
-Boundary effects are difficult to observe without large amounts of crop data, which is typically expensive and time-consuming to collect.\\
-Because of this, studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality (Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
+Unfortunately, data to inform these decisions are often minimal, as observing boundary effects requires large amounts of expensive and time-consuming crop yield data.
+This means that studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality (Brandle \emph{et al.} 1984; Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
 Meta-analyses are a common solution to this problem (e.g.
-Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs.
+Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs (Steel 2017; Osorio \emph{et al.} 2018).
 However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
@@
 
-In this paper use a precision yield dataset from Alberta, Canada to ask the following questions:
+In this paper we use a precision yield dataset from Alberta, Canada to ask the following questions:
 1. How does crop yield change with distance from the edge of field?
@@
 3. Is there an intermediate distance where yield is maximized or variance is minimized?
-To our knowledge, no other study has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.
+To our knowledge, only (Osorio \emph{et al.} 2018) has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.
 
@@
 Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows (Griffin \emph{et al.} 2007; Whelan \& Taylor 2013), or when the combine is accelerating or decelerating (Arslan \& Colvin 2002).
-Therefore, we used the following filtering procedure to clean the data prior to model fitting:
+Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix A for R code):
 
@@
 This makes accurate and consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:
-1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
-2. Wetland: permanent wetland; borders are largely unchanged from year-to-year
-3. Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
-4. Other crop: annual crop or pasture with little or no visible boundary between planted areas
-5. Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
-6. Grassland: permanent seminatural grassland or pasture (not in rotation)
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
+\item
+  Wetland: permanent wetland; borders are largely unchanged from year-to-year
+\item
+  Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
+\item
+  Other crop: annual crop or pasture with little or no visible boundary between planted areas
+\item
+  Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
+\item
+  Grassland: permanent seminatural grassland or pasture (not in rotation)
+\end{enumerate}
 
@@
 \item
-  Sequence = order of harvest within field (1--N points)
-\item
   f(x,b) = penalized thin-plate regression spline, where \emph{x} is the predictor and \emph{b} is the number of basis dimensions
@@
 Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
-We used 5 basis dimensions for the distance smoother and 60 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \ref{fig:hypotheses}).
+We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \ref{fig:hypotheses}).
 All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).
@@
 The residuals in most of the models displayed both spatial autocorrelation in the residuals, indicating ``micro-scale'' variation that was not captured by the spatial smoothers (Cressie \& Wikle 2011; Wikle \emph{et al.} 2019).
-This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.~extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 60), we decided to leave the models as-is.
+This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.~extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 80), we decided to leave the models as-is.
 Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
-That is, combining of fields occurs in patterns (e.g.~east-to-west) so it is impossible to tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
+That is, combining of fields occurs in patterns (e.g.~east-to-west) so we cannot tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
 We considered the latter to be a more realistic assumption, so we assigned all random variation as spatial in nature.
-Customized combine driving patterns may be able to detect temporal variation, but this is beyond the scope of this research.
+Non-standard combine driving patterns may be able to gather data in a way that can detect sensor drift, but this is beyond the scope of this research.
 
@@
 This does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping:
-1. Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard errors
-2. Use posterior sample to simulate a new smoother from each field
-3. Fit model of simulated smoothers from all fields, and save this ``meta-smoother''
-4. Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard errors.
+\item
+  Use posterior sample to simulate a new smoother from each field.
+\item
+  Fit new model of simulated smoothers from all fields (with a field-year random intercept), and save this ``meta-smoother.''
+\item
+  Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers.
   This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
   Spatial variability was not averaged between fields because within-field variability is not directly comparable between fields.
+\end{enumerate}
 
@@
 \includegraphics[width=1\linewidth]{../Figures/ExamplePlots/spatialSmooths} \caption{Raw data and spatial smoothers from a single field.
-Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.}\label{fig:spatialSmooths}
+Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.
+Field dimensions are approximately 800 x 800 m, but coordinates are hidden to protect data privacy.}\label{fig:spatialSmooths}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/boundaryTypes-1} \caption{Percentage of field boundaries across sampled fields.}\label{fig:boundaryTypes}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/boundaryTypes-1} \caption{Boxplot of field boundary compositions across sampled fields.
+The x-axis indicates the boundary type, while the y-axis indicates the percentage of all field boundaries covered by that boundary type.}\label{fig:boundaryTypes}
 \end{figure}
@@
 Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table \ref{tab:smoothSummary}, row 6).
-Distances from boundary types were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
+Distances from boundaries were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
 
@@
 Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure \ref{fig:canolaPlot} top row).
-Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50m and remained roughly constant with distance.
+Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50 m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50 m and remained roughly constant with distance.
 The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
 Yield variability (Figure \ref{fig:canolaPlot} bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
-Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from Wetland boundaries, indicating a potential stabilizing effect.
+Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from wetland boundaries, indicating a potential stabilizing effect, but this was offset by the large increase in yield variability at the edge of wetlands.
 
@@
 
-Pea crops were less well-represented in our data (N = 39 fields, compared to ), and the effects of boundaries tended to be less consistent.
+Pea crops were less well-represented in our data (N = 39 fields), and therefore the effects of boundaries tended to be less consistent.
 Average pea yields (Figure \ref{fig:peaPlot}, top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
@@
 
-At the level of individual fields, both average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
-Yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat.
-This increase tended to be small in magnitude in canola crops, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \ref{fig:hypotheses}).
+Within individual fields, both the average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
+Mean yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat, while peas showed no evidence of this.
+However, this intermediate increase was small in canola crops, meaning that ecosystem services likely did not vary strongly with distance from field boundaries (Scenario 2, Figure \ref{fig:hypotheses}).
 There was a fairly large (0.75 T/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
 However, there was a similar (albeit weaker) pattern associated with bare ground, so this pattern may be caused by something other than the presence of shelterbelts.
-This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this across a larger geographic domain in order to inform growers about how best to manage the landscapes surrounding their fields.
+This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this within relevant geographic sub-regions in order to locally inform growers about how best to manage the landscapes surrounding their fields.
 
-Why did we find few consistent effects of boundary type on crop yield, aside from shelterbelts?
-Our work examined only the first and last links in the causal chain \emph{Field boundary} \(\rightarrow\) \emph{Insect Abundance} \(\rightarrow\) \emph{Crop Yield}, meaning that are two mediating relationships where the overall relationship could be broken or obscured.
-First, field boundaries may not have any influence on insect abundance if separated from the larger landscape context, or spillover of beneficial organisms may be low.
-Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are low, which may vary yearly, or by location.
-Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) when organisms in crops provide tangible ecosystem services.
+Our work showed an intermediate increase in wheat yield with distance from shelterbelts and bare ground, but what is the cause of these patterns?
+We hypothesize that the potential yield-increasing mechanisms may be a) microclimate effects, b) soil effects, or c) beneficial animal spillover.
+Shelterbelts can act to capture snow and moderate downwind microclimate, reducing water stress on crop plants and increasing yield (Staple \& Lehane 1955; Kowalchuk \& de Jong 1995; Bao \emph{et al.} 2012).
+However, Kowalchuk \& de Jong (1995) found that positive microclimate effects of shelterbelts on wheat yield occur at around 20 m (3-4 times the shelterbelt height), while our work shows a peak at around 50 m.
+This additionally complicated by the fact that water stress varies from year-to-year, and that the impact of the shelterbelt also depends on the orientation to prevailing winds and the physical structure of the shelterbelt (Pelton 1967; Kort 1988; Singh \emph{et al.} 1999).
+Unexpectedly, we also found this intermediate yield maximum with distance away from bare field boundaries, indicating that the patterns we found may be at least somewhat related to water stress.
+While we consider microclimate to be the most likely cause of the intermediate increase, it may also be related to soil effects from shelterbelts, where erosion is reduced and leaf litter from the shelterbelt can cause an increase in soil deposition (de Jong \& Kowalchuk 1995).
+Finally, shelterbelts may also act as sources of predators or parasitoids of wheat crop pests (Dong \emph{et al.} 2015; Ward \emph{et al.} 2021) but their effect depends on pest pressures, which are typically suppressed by pesticide application.
+We recommend that follow-up studies a) use precision yield data to confirm that this pattern exists across broader geographical areas and b) examine the influence of the three potential mechanisms considered above (as well as their interaction).
+
+Wheat and shelterbelts aside, why did we find few consistent effects of boundary type on crop yield?
+Our work examined only the first and last links in the causal chain \emph{Field Boundary} \(\rightarrow\) \emph{Ecosystem Service} \(\rightarrow\) \emph{Crop Yield}, meaning that are two mediating relationships where the overall relationship could be broken or obscured.
+First, field boundaries may not have any influence on beneficial insect abundance if separated from the larger landscape context, or spillover of beneficial organisms may be low.
+Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are low, which may vary among year or location.
+Abiotic ecosystem services, such as reduced wind speed or temperature, also depend on the structure of the boundary and whether wind or temperature are limiting crop production.
+Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) under what conditions field boundaries provide tangible ecosystem services.
 We consider these reasons below:
 
-SNL generally has a positive effect on spillover, but field boundaries only represent a fraction of the SNL that surrounds a given field.
+SNL generally has a positive effect on beneficial insect abundance, but field boundaries only represent a fraction of the SNL that surrounds a given field.
 Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales (Nguyen \emph{et al.} 2021) which could be particularly important when beneficial arthropods are disperse very broadly (Öberg \emph{et al.} 2007).
@@
 
-Field boundary traits may play an important role in mediating spillover, but our definitions of boundaries were (necessarily) loose, as we used satellite imagery to manually classify them.
-Therefore, it may be that other aspects of field boundaries, such as size or plant community composition, are more important for spillover of ecosystems services into fields.
-For example, larger (Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020) or more functionally diverse (Bonifacio \emph{et al.} 2011; Boughey \emph{et al.} 2011) hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction matters to predatory beetle communities (Sutter \emph{et al.} 2018).
+Field boundary characteristics, such as height or species composition, may play an important role in mediating spillover, but our definitions of boundaries were necessarily loose, as we used satellite imagery to manually classify them.
+Therefore, it may be that other aspects of field boundaries are more important for spillover of ecosystems services into fields.
+For example, larger or more functionally diverse (Bonifacio \emph{et al.} 2011; Boughey \emph{et al.} 2011; Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020) hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction influences predatory beetle communities (Sutter \emph{et al.} 2018).
 These fine-scale boundary characteristics were not measured in our study, but could be used as variables in more biologically-meaningful classifications of field boundaries.
+Finally, the relevant spatial scale of boundary effects are often related to the height or windward orientation of the shelterbelt (Kort 1988; Kowalchuk \& de Jong 1995), and precision yield data has the potential to show this in far greater detail than previous studies.
 
@@
 Pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang \emph{et al.} 2007).
-Gray \& Lewis (2014) showed that remnant forest fragments do not provide significant pest control services, but also do not depress yield.
 However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data \emph{alone}.
-Ecosystem services from field boundaries may be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales Robinson \emph{et al.} (2021).
+Ecosystem services from field boundaries may also be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales Robinson \emph{et al.} (2021).
+Lastly, pest pressure or pollination requirements may be low enough that ecosystem service provision does not significantly change yield.
+Canola is largely self-pollinating (at least in hybrid canola varieties, Lindström \emph{et al.} 2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop (Woodcock \emph{et al.} 2016).
+Crop pests in most western Canadian fields are suppressed by insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty, meaning spillover of beneficial predators and parasitoids may only matter during outbreak years.
 
-Pest pressure or pollination deficit may be low enough that ecosystem service provision does not significantly change yield.
-Canola is largely self-pollinating (at least in hybrid canola varieties, Lindström \emph{et al.} 2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop (Woodcock \emph{et al.} 2016).
-Pests are also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty.
-
-Our work is some of the first to use a large precision yield dataset to answer questions about how landscapes influence yield, but this study has only scratched the surface of some of the richest agroecological datasets ever created.
+Our work is some of the first to use a large precision yield dataset to answer questions about how landscapes influence yield.
 Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
 Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee 2006), where boundaries contained a higher proportion of shelterbelts (Figure \ref{fig:boundaryTypes}), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
-Between-year effects may be important, especially during dry years (Kort 1988), but this was not considered in our analysis.
+Between-year effects may be important, especially during hot dry years (Kort 1988), but this was not considered in our analysis.
 Future work could examine whether boundaries have differing effects during hot or cool years, as shading or cooling effects may increase yield during hot years but have negative (or no) effect during cooler years.
+Finally, about 50\% of growers in western Canada have collected precision agriculture data, and about 85\% have yield monitoring capability (Steel 2017), which means that our work has only scratched the surface of what could be (collectively) one of the richest agroecological datasets ever created.
+We urge agronomists, growers, and ecologists to adopt more wide-spread use of precision yield data, as it represents a vast source of incredibly detailed information which can be used to answer a variety of important questions.
 
@@
 
+\leavevmode\hypertarget{ref-bao2012}{}%
+Bao, Y., Li, H.
+\& Zhao, H.
+(2012).
+Effect of shelterbelts on winter wheat yields in sanded farmland of north-western shandong province, china.
+\emph{Journal of Food Agriculture \& Environment}, 10, 1399--1403.
+
 \leavevmode\hypertarget{ref-boetzl2020}{}%
@@
 
+\leavevmode\hypertarget{ref-brandle1984}{}%
+Brandle, J.R., Johnson, B.B. \& Dearmont, D.D. (1984). Windbreak economics: The case of winter wheat production in eastern {Nebraska}. \emph{Journal of Soil and Water Conservation}, 39, 339--343.
+
 \leavevmode\hypertarget{ref-case2019}{}%
@@
 Cressie, N. \& Wikle, C.K. (2011). \emph{Statistics for spatio-temporal data}. Wiley.
+
+\leavevmode\hypertarget{ref-deJong1995}{}%
+de Jong, E. \& Kowalchuk, T.E. (1995). The effect of shelterbelts on erosion and soil properties. \emph{Soil Science}, 159, 337--345.
+
+\leavevmode\hypertarget{ref-dong2015}{}%
+Dong, Z., Ouyang, F., Lu, F. \& Ge, F. (2015). Shelterbelts in agricultural landscapes enhance ladybeetle abundance in spillover from cropland to adjacent habitats. \emph{BioControl}, 60, 351--361.
 
@@
 
+\leavevmode\hypertarget{ref-helmers2005}{}%
+Helmers, G. \& Brandle, J.R. (2005). Optimum windbreak spacing in great plains agriculture. \emph{Great Plains Research}, 15, 179--198.
+
 \leavevmode\hypertarget{ref-hunicken2021}{}%
@@
 \leavevmode\hypertarget{ref-kowalchuk1995}{}%
-Kowalchuk, T.E. \& Jong, E. de. (1995). Shelterbelts and their effect on crop yield. \emph{Canadian Journal of Soil Science}, 75, 543--550.
+Kowalchuk, T.E. \& de Jong, E. (1995). Shelterbelts and their effect on crop yield. \emph{Canadian Journal of Soil Science}, 75, 543--550.
 
@@
 
+\leavevmode\hypertarget{ref-osorio2018}{}%
+Osorio, R.J., Barden, C.J. \& Ciampitti, I.A. (2018). {GIS} approach to estimate windbreak crop yield effects in {Kansas}{{}}{Nebraska}. \emph{Agroforestry Systems}, 93, 1567--1576.
+
 \leavevmode\hypertarget{ref-oberg2007}{}%
@@
 \emph{Agriculture, Ecosystems {\&} Environment}, 122, 211--219.
+
+\leavevmode\hypertarget{ref-pelton1967}{}%
+Pelton, W.L. (1967). The effect of a windbreak on wind travel, evaporation and wheat yield. \emph{Canadian Journal of Plant Science}, 47, 209--214.
 
@@
 
+\leavevmode\hypertarget{ref-singh1999}{}%
+Singh, H.P., Kohli, R.K.
+\& Batish, D.R.
+(1999).
+Impact of \emph{populus deltoides} and \emph{dalbergia sissoo} shelterbelts on wheat {}a comparative study.
+\emph{International Tree Crops Journal}, 10, 51--60.
+
+\leavevmode\hypertarget{ref-staple1955}{}%
+Staple, W.J. \& Lehane, J.J. (1955). The influence of field shelterbelts on wind velocity, evaporation, soil moisture and crop yield. \emph{Canadian Journal of Agricultural Science}, 35, 440--453.
+
 \leavevmode\hypertarget{ref-stasiov2020}{}%
@@
 Statistics Canada. (2014). Table 001-0010 - estimated areas, yield, production and average farm price of principal field crops, in metric units, annual.
+
+\leavevmode\hypertarget{ref-steel2017}{}%
+Steel, D. (2017). \emph{Analysis of precision agriculture adoption \& barriers in western {Canada}}. Agriculture; Agri-Food Canada.
 
@@
 Vega, A., Córdoba, M., Castro-Franco, M. \& Balzarini, M. (2019). Protocol for automating error removal from yield maps. \emph{Precision Agriculture}, 20, 1030--1044.
+
+\leavevmode\hypertarget{ref-ward2021}{}%
+Ward, S.E., Umina, P.A., Macfadyen, S. \& Hoffmann, A.A. (2021). Hymenopteran parasitoids of aphid pests within {Australian} grain production landscapes. \emph{Insects}, 12, 44.
 
@@
 
+\singlespacing
+\nolinenumbers
+
+\newpage
+
+\hypertarget{appendix-a-r-code-for-data-filtering}{%
+\section*{Appendix A: R Code for Data Filtering}\label{appendix-a-r-code-for-data-filtering}}
+\addcontentsline{toc}{section}{Appendix A: R Code for Data Filtering}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Helper functions {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+
+\CommentTok{\#"Inlier" spatial filtering procedure from Vega et al 2019, rewritten to work}
+\CommentTok{\#with sf + dplyr. Returns boolean}
+\NormalTok{vegaFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data,ycol,}\AttributeTok{pvalCutoff=}\FloatTok{0.05}\NormalTok{,}\AttributeTok{nDist=}\DecValTok{40}\NormalTok{)\{}
+  \FunctionTok{require}\NormalTok{(sf)}
+  \FunctionTok{require}\NormalTok{(sp)}
+  \FunctionTok{require}\NormalTok{(spdep)}
+  \FunctionTok{require}\NormalTok{(tidyverse)}
+
+  \ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{any}\NormalTok{(}\FunctionTok{class}\NormalTok{(data) }\SpecialCharTok{\%in\%} \StringTok{\textquotesingle{}sf\textquotesingle{}}\NormalTok{)) }\FunctionTok{stop}\NormalTok{(}\StringTok{\textquotesingle{}Dataframe must be sf object\textquotesingle{}}\NormalTok{)}
+
+\NormalTok{  coords }\OtherTok{\textless{}{-}} \FunctionTok{st\_coordinates}\NormalTok{(data) }\CommentTok{\#Get coordinates in matrix form}
+
+  \CommentTok{\#Get neighbourhood weights from 0 to ndist meters}
+\NormalTok{  nWeights }\OtherTok{\textless{}{-}} \FunctionTok{dnearneigh}\NormalTok{(coords,}\DecValTok{0}\NormalTok{,nDist) }
+
+  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{any}\NormalTok{(}\FunctionTok{sapply}\NormalTok{(nWeights,length)}\SpecialCharTok{==}\DecValTok{1}\NormalTok{))\{}
+    \FunctionTok{warning}\NormalTok{(}\StringTok{\textquotesingle{}Some points had no neighbours and were removed\textquotesingle{}}\NormalTok{)}
+\NormalTok{  \} }
+
+  \CommentTok{\#Get neighbourhood indices for each point (which other points are in this}
+  \CommentTok{\#point\textquotesingle{}s neighbourhood?)}
+\NormalTok{  nIndices }\OtherTok{\textless{}{-}} \FunctionTok{nb2listw}\NormalTok{(nWeights, }\AttributeTok{style =} \StringTok{"W"}\NormalTok{,}\AttributeTok{zero.policy =} \ConstantTok{TRUE}\NormalTok{) }
+
+\NormalTok{  yield }\OtherTok{\textless{}{-}} \FunctionTok{pull}\NormalTok{(data,\{\{ycol\}\}) }\CommentTok{\#Get yield data column}
+
+  \CommentTok{\#Local Moran\textquotesingle{}s I for each neighbourhood}
+\NormalTok{  LM }\OtherTok{\textless{}{-}} \FunctionTok{localmoran}\NormalTok{(yield,nIndices,}\AttributeTok{p.adjust.method=}\StringTok{"bonferroni"}\NormalTok{,}
+                   \AttributeTok{alternative =}\StringTok{"less"}\NormalTok{)}
+
+\NormalTok{  results }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(LM) }\SpecialCharTok{\%\textgreater{}\%} 
+    \FunctionTok{rename}\NormalTok{(}\StringTok{\textquotesingle{}pval\textquotesingle{}}\OtherTok{=}\FunctionTok{contains}\NormalTok{(}\StringTok{\textquotesingle{}Pr.z.\textquotesingle{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+    \CommentTok{\#Filter negative Ii and pvals \textless{} 0.05}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{keepThese=}\NormalTok{ Ii }\SpecialCharTok{\textgreater{}} \DecValTok{0} \SpecialCharTok{|}\NormalTok{ pval }\SpecialCharTok{\textgreater{}}\NormalTok{ pvalCutoff) }\SpecialCharTok{\%\textgreater{}\%}
+    \CommentTok{\#NAs (points had no neighbours)}
+    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{keepThese=}\FunctionTok{ifelse}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(keepThese),}\ConstantTok{FALSE}\NormalTok{,keepThese)) }
+
+\NormalTok{  ret }\OtherTok{\textless{}{-}} \FunctionTok{pull}\NormalTok{(results)}
+  \FunctionTok{return}\NormalTok{(ret)}
+\NormalTok{\}}
+
+\CommentTok{\#Function to filter anything above a certain z{-}score. Returns boolean}
+\NormalTok{ZscoreFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x,}\AttributeTok{z=}\DecValTok{3}\NormalTok{)\{}
+\NormalTok{  xmean }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(x,}\AttributeTok{na.rm=}\ConstantTok{TRUE}\NormalTok{)}
+\NormalTok{  xsd }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(x,}\AttributeTok{na.rm=}\ConstantTok{TRUE}\NormalTok{)}
+\NormalTok{  zscore }\OtherTok{\textless{}{-}} \FunctionTok{abs}\NormalTok{(((x}\SpecialCharTok{{-}}\NormalTok{xmean)}\SpecialCharTok{/}\NormalTok{xsd))}
+\NormalTok{  zscore}\SpecialCharTok{\textless{}}\NormalTok{z}
+\NormalTok{\} }
+
+\CommentTok{\#Function to filter anything above certain quantiles. Returns boolean}
+\NormalTok{QuantileFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x,}\AttributeTok{quant=}\FloatTok{0.99}\NormalTok{)\{ }
+\NormalTok{  l }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{((}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{quant)}\SpecialCharTok{/}\DecValTok{2}\NormalTok{,}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{(}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{quant)}\SpecialCharTok{/}\DecValTok{2}\NormalTok{) }\CommentTok{\#Symmetric quantiles}
+\NormalTok{  x}\SpecialCharTok{\textgreater{}}\FunctionTok{quantile}\NormalTok{(x,l[}\DecValTok{1}\NormalTok{]) }\SpecialCharTok{\&}\NormalTok{ x}\SpecialCharTok{\textless{}}\FunctionTok{quantile}\NormalTok{(x,l[}\DecValTok{2}\NormalTok{]) }
+\NormalTok{\}}
+
+\CommentTok{\#Function to filter large changes in bearing. Returns boolean unless}
+\CommentTok{\#returnDiffs==TRUE}
+\NormalTok{bearingFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(bearing,}\AttributeTok{q=}\ConstantTok{NULL}\NormalTok{,}\AttributeTok{z=}\ConstantTok{NULL}\NormalTok{,}\AttributeTok{returnDiffs=}\ConstantTok{FALSE}\NormalTok{)\{}
+  \ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{xor}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(q),}\FunctionTok{is.null}\NormalTok{(z))}\SpecialCharTok{\&!}\NormalTok{returnDiffs) }\FunctionTok{stop}\NormalTok{(}\StringTok{\textquotesingle{}Input quantiles or Z{-}score\textquotesingle{}}\NormalTok{)}
+  \CommentTok{\#Difference in compass bearings (in degrees)}
+\NormalTok{  bearingDiff }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x1,x2)\{}
+\NormalTok{    x }\OtherTok{\textless{}{-}}\NormalTok{ x1}\SpecialCharTok{{-}}\NormalTok{x2}
+\NormalTok{    x }\OtherTok{\textless{}{-}} \FunctionTok{ifelse}\NormalTok{(}\FunctionTok{abs}\NormalTok{(x)}\SpecialCharTok{\textgreater{}}\DecValTok{180}\NormalTok{,x}\SpecialCharTok{{-}}\NormalTok{(}\DecValTok{360}\SpecialCharTok{*}\FunctionTok{sign}\NormalTok{(x)),x) }\CommentTok{\#Angle differences can\textquotesingle{}t be \textgreater{}180}
+    \FunctionTok{return}\NormalTok{(x)}
+\NormalTok{  \}}
+
+  \CommentTok{\#Looks 1 point ahead and behind}
+\NormalTok{  bd }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{bearingDiff}\NormalTok{(}\FunctionTok{lag}\NormalTok{(bearing),bearing),}
+              \FunctionTok{bearingDiff}\NormalTok{(}\FunctionTok{lead}\NormalTok{(bearing),bearing))}
+  \CommentTok{\#Maximum bearing difference ahead and behind}
+\NormalTok{  bd }\OtherTok{\textless{}{-}} \FunctionTok{apply}\NormalTok{(bd,}\DecValTok{1}\NormalTok{,}\ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{max}\NormalTok{(}\FunctionTok{abs}\NormalTok{(x),}\AttributeTok{na.rm=}\ConstantTok{TRUE}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{sign}\NormalTok{(x[}\FunctionTok{which.max}\NormalTok{(}\FunctionTok{abs}\NormalTok{(x))])) }
+
+  \ControlFlowTok{if}\NormalTok{(returnDiffs) }\FunctionTok{return}\NormalTok{(bd) }\CommentTok{\#Return bearing differences only, without filtering}
+
+  \ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.null}\NormalTok{(q))\{}
+\NormalTok{    ret }\OtherTok{\textless{}{-}} \FunctionTok{QuantileFilter}\NormalTok{(bd,}\AttributeTok{q=}\NormalTok{q)}
+\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
+\NormalTok{    ret }\OtherTok{\textless{}{-}} \FunctionTok{ZscoreFilter}\NormalTok{(bd,}\AttributeTok{z=}\NormalTok{z)}
+\NormalTok{  \}}
+  \FunctionTok{return}\NormalTok{(ret)}
+\NormalTok{\} }
+
+\CommentTok{\#Positional difference filter {-} filters out very distant and very close points.}
+\CommentTok{\#Returns boolean unless returnDiffs==TRUE}
+\NormalTok{posFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data,}\AttributeTok{q=}\ConstantTok{NULL}\NormalTok{,}\AttributeTok{returnDiffs=}\ConstantTok{FALSE}\NormalTok{)\{ }
+  \FunctionTok{require}\NormalTok{(sf)}
+  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(q)}\SpecialCharTok{\&!}\NormalTok{returnDiffs) }\FunctionTok{stop}\NormalTok{(}\StringTok{\textquotesingle{}Input upper quantile\textquotesingle{}}\NormalTok{)}
+
+  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{units}\NormalTok{(}\FunctionTok{st\_distance}\NormalTok{(data[}\DecValTok{1}\NormalTok{,],data[}\DecValTok{2}\NormalTok{,]))}\SpecialCharTok{$}\NormalTok{numerator}\SpecialCharTok{!=}\StringTok{\textquotesingle{}m\textquotesingle{}}\NormalTok{)\{}
+    \FunctionTok{warning}\NormalTok{(}\StringTok{\textquotesingle{}Position differences not in meters\textquotesingle{}}\NormalTok{)}
+\NormalTok{  \} }
+\NormalTok{  coords }\OtherTok{\textless{}{-}} \FunctionTok{st\_coordinates}\NormalTok{(data) }\CommentTok{\#Get coordinates}
+  \CommentTok{\#Distances between points}
+\NormalTok{  pdiff }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(coords)}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{),}\ControlFlowTok{function}\NormalTok{(i) }\FunctionTok{as.numeric}\NormalTok{(}\FunctionTok{dist}\NormalTok{(coords[i}\SpecialCharTok{:}\NormalTok{(i}\SpecialCharTok{+}\DecValTok{1}\NormalTok{),]))) }
+\NormalTok{  pdiff }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(pdiff,}\ConstantTok{NA}\NormalTok{),}\FunctionTok{c}\NormalTok{(}\ConstantTok{NA}\NormalTok{,pdiff)) }\CommentTok{\#Forward and backward lags}
+\NormalTok{  pdiff }\OtherTok{\textless{}{-}} \FunctionTok{apply}\NormalTok{(pdiff,}\DecValTok{1}\NormalTok{,max,}\AttributeTok{na.rm=}\ConstantTok{TRUE}\NormalTok{) }\CommentTok{\#Maximum distance ahead and behind}
+  \ControlFlowTok{if}\NormalTok{(returnDiffs)\{}
+    \FunctionTok{return}\NormalTok{(pdiff)}
+\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
+    \FunctionTok{return}\NormalTok{(}\FunctionTok{QuantileFilter}\NormalTok{(pdiff,q)) }\CommentTok{\#Note: uses 2{-}sided quantiles}
+\NormalTok{  \}}
+\NormalTok{\}}
+
+\CommentTok{\#Filter for (forward and backward) lagged speed differences. Returns boolean}
+\NormalTok{dSpeedFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(speed,}\AttributeTok{l=}\FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{),}\AttributeTok{perc=}\FloatTok{0.2}\NormalTok{)\{ }
+
+  \CommentTok{\#Matrix of \% diffs}
+\NormalTok{  llist }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(l,}\ControlFlowTok{function}\NormalTok{(x) (}\FunctionTok{lag2}\NormalTok{(speed,x)}\SpecialCharTok{{-}}\NormalTok{speed)}\SpecialCharTok{/}\FunctionTok{lag2}\NormalTok{(speed,x)) }
+
+  \CommentTok{\#Are any lagged speed values \textgreater{} percent change threshold?}
+\NormalTok{  ret }\OtherTok{\textless{}{-}} \SpecialCharTok{!}\FunctionTok{apply}\NormalTok{(llist,}\DecValTok{1}\NormalTok{,}\ControlFlowTok{function}\NormalTok{(y) }\FunctionTok{any}\NormalTok{(}\FunctionTok{abs}\NormalTok{(y)[}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(y)]}\SpecialCharTok{\textgreater{}}\NormalTok{perc))}
+
+  \FunctionTok{return}\NormalTok{(ret)}
+\NormalTok{\}}
+
+\CommentTok{\#Overloaded lag function that takes negative values}
+\NormalTok{lag2 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x,n)\{}
+  \FunctionTok{require}\NormalTok{(dplyr)}
+  \ControlFlowTok{if}\NormalTok{(n}\SpecialCharTok{==}\DecValTok{0}\NormalTok{) \{}
+    \FunctionTok{return}\NormalTok{(x) }\CommentTok{\#No lag}
+\NormalTok{  \} }\ControlFlowTok{else} \ControlFlowTok{if}\NormalTok{(n}\SpecialCharTok{\textgreater{}}\DecValTok{0}\NormalTok{)\{}
+    \FunctionTok{lag}\NormalTok{(x,n) }\CommentTok{\#Positive lag}
+\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
+    \FunctionTok{lead}\NormalTok{(x,}\FunctionTok{abs}\NormalTok{(n)) }\CommentTok{\#Negative lag}
+\NormalTok{  \} }
+\NormalTok{\} }
+
+\CommentTok{\#Filtering pipeline  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+
+\FunctionTok{library}\NormalTok{(tidyverse)}
+\FunctionTok{library}\NormalTok{(sf)}
+
+\CommentTok{\#Note: dat == sf dataframe of yield data with columns: DryYield (T/ha),}
+\CommentTok{\#TrackAngle (compass bearing), Speed (km/hr), cropType (Wheat, Canola, Peas)}
+
+\NormalTok{dat }\OtherTok{\textless{}{-}}\NormalTok{ dat }\SpecialCharTok{\%\textgreater{}\%} 
+  \CommentTok{\#Vega et al 2019 spatial "inlier" filter {-} takes a few minutes...}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{vegaFilt =} \FunctionTok{vegaFilter}\NormalTok{(.,DryYield,}\AttributeTok{nDist =} \DecValTok{30}\NormalTok{)) }
+
+\NormalTok{dat }\OtherTok{\textless{}{-}}\NormalTok{ dat }\SpecialCharTok{\%\textgreater{}\%} 
+  \CommentTok{\#"BS" filter {-} removes values \textgreater{}10.75 T/ha for wheat, \textgreater{}8 T/ha for peas \& canola}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{noBS =}\NormalTok{ DryYield}\SpecialCharTok{\textless{}}\FunctionTok{ifelse}\NormalTok{(cropType}\SpecialCharTok{==}\StringTok{\textquotesingle{}Wheat\textquotesingle{}}\NormalTok{,}\FloatTok{10.75}\NormalTok{,}\DecValTok{8}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\#Trim dry yield outliers}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Qfilt =} \FunctionTok{QuantileFilter}\NormalTok{(DryYield,}\AttributeTok{q=}\FloatTok{0.98}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \CommentTok{\#Trim extreme bearing changes (turning)}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{bFilt =} \FunctionTok{bearingFilter}\NormalTok{(TrackAngle,}\AttributeTok{q=}\FloatTok{0.98}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \CommentTok{\#Trim absolute speed outliers}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{speedFilt =} \FunctionTok{QuantileFilter}\NormalTok{(Speed,}\AttributeTok{q=}\FloatTok{0.98}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \CommentTok{\#Trim speed differences (\textgreater{}20\% change 2 steps forward and backward, suggested}
+  \CommentTok{\#by Lyle et al 2014)}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dSpeedFilt =} \FunctionTok{dSpeedFilter}\NormalTok{(Speed,}\AttributeTok{l=}\FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{,}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{),}\AttributeTok{perc =} \FloatTok{0.2}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+  \CommentTok{\#Trim points that are far away from eachother}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{posFilt =} \FunctionTok{posFilter}\NormalTok{(.,}\AttributeTok{q=}\FloatTok{0.98}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
+  \CommentTok{\#Combine filter criteria}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{allFilt =}\NormalTok{ noBS }\SpecialCharTok{\&}\NormalTok{ vegaFilt }\SpecialCharTok{\&}\NormalTok{ Qfilt }\SpecialCharTok{\&}\NormalTok{ bFilt }\SpecialCharTok{\&}\NormalTok{ speedFilt }\SpecialCharTok{\&}
+\NormalTok{           dSpeedFilt }\SpecialCharTok{\&}\NormalTok{ posFilt) }\SpecialCharTok{\%\textgreater{}\%}  
+  \CommentTok{\#Remove filtered values}
+  \FunctionTok{filter}\NormalTok{(allFilt)}
+\end{Highlighting}
+\end{Shaded}
+
+\hypertarget{appendix-b-r-code-for-modelling}{%
+\section*{Appendix B: R Code for Modelling}\label{appendix-b-r-code-for-modelling}}
+\addcontentsline{toc}{section}{Appendix B: R Code for Modelling}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Fit models {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
+
+\FunctionTok{library}\NormalTok{(mgcv)}
+
+\CommentTok{\#Note: dat == dataframe of yield data with columns: DryYield (T/ha), dist}
+\CommentTok{\#(distance to nearest field boundary, m), boundaryType (type of boundary,}
+\CommentTok{\#factor), E,N (easting and northing distance from centre of field, m)}
+
+\CommentTok{\#Uses thin{-}plate splines with extra shrinkage (bs = \textquotesingle{}ts\textquotesingle{}), with 5 basis}
+\CommentTok{\#functions for distance smoothers and 80 basis functions for spatial smoothers}
+
+\CommentTok{\#Variance smoother}
+\NormalTok{f2 }\OtherTok{\textless{}{-}} \StringTok{\textquotesingle{}\textasciitilde{} s(dist,k=5,bs="ts",by=boundaryType) + s(E,N,k=80)\textquotesingle{}} 
+\CommentTok{\#Means smoother}
+\NormalTok{f }\OtherTok{\textless{}{-}} \StringTok{\textquotesingle{}sqrt(DryYield)\textasciitilde{} s(dist,k=5,bs="ts",by=boundaryType) + s(E,N,k=80)\textquotesingle{}}\ErrorTok{)} 
+\CommentTok{\#Combine into a list}
+\NormalTok{flist }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(}\FunctionTok{as.formula}\NormalTok{(f),}\FunctionTok{as.formula}\NormalTok{(f2)) }
+
+\CommentTok{\#Fit model {-} takes about 45 mins for a 50000 point dataset using an AMD Ryzen 7}
+\CommentTok{\#3800X 8{-}core processor}
+\NormalTok{mod }\OtherTok{\textless{}{-}} \FunctionTok{gam}\NormalTok{(flist,}\AttributeTok{data=}\NormalTok{dat,}\AttributeTok{family=}\FunctionTok{gaulss}\NormalTok{()) }
+\end{Highlighting}
+\end{Shaded}
+
 \end{document}

### 2021-10-27 Manuscript/supplemental.Rmd
Finished Discussion and submitted to preprint server

--- 2021-09-08 Manuscript/supplemental.Rmd
@@
 
-# Appendix A: Supplementary Material {-}
+# Appendix A: R Code for Data Filtering {-}
 
-```{r canolaPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on canola yield.', out.width='100%'}  
-  include_graphics('../Figures/groundSpeed_canola.png',dpi=NA)
+```{r filterCode, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}
+# Helper functions -----------------
+
+#"Inlier" spatial filtering procedure from Vega et al 2019, rewritten to work
+#with sf + dplyr. Returns boolean
+vegaFilter <- function(data,ycol,pvalCutoff=0.05,nDist=40){
+  require(sf)
+  require(sp)
+  require(spdep)
+  require(tidyverse)
+
+  if(!any(class(data) %in% 'sf')) stop('Dataframe must be sf object')
+
+  coords <- st_coordinates(data) #Get coordinates in matrix form
+
+  #Get neighbourhood weights from 0 to ndist meters
+  nWeights <- dnearneigh(coords,0,nDist)
+
+  if(any(sapply(nWeights,length)==1)){
+    warning('Some points had no neighbours and were removed')
+  }
+
+  #Get neighbourhood indices for each point (which other points are in this
+  #point's neighbourhood?)
+  nIndices <- nb2listw(nWeights, style = "W",zero.policy = TRUE)
+
+  yield <- pull(data,{{ycol}}) #Get yield data column
+
+  #Local Moran's I for each neighbourhood
+  LM <- localmoran(yield,nIndices,p.adjust.method="bonferroni",
+                   alternative ="less")
+
+  results <- data.frame(LM) %>% 
+    rename('pval'=contains('Pr.z.')) %>% 
+    #Filter negative Ii and pvals < 0.05
+    mutate(keepThese= Ii > 0 | pval > pvalCutoff) %>% 
+    #NAs (points had no neighbours)
+    mutate(keepThese=ifelse(is.na(keepThese),FALSE,keepThese))
+
+  ret <- pull(results)
+  return(ret)
+}
+
+#Function to filter anything above a certain z-score. Returns boolean
+ZscoreFilter <- function(x,z=3){
+  xmean <- mean(x,na.rm=TRUE)
+  xsd <- sd(x,na.rm=TRUE)
+  zscore <- abs(((x-xmean)/xsd))
+  zscore<z
+}
+
+#Function to filter anything above certain quantiles. Returns boolean
+QuantileFilter <- function(x,quant=0.99){ 
+  l <- c((1-quant)/2,1-(1-quant)/2) #Symmetric quantiles
+  x>quantile(x,l[1]) & x<quantile(x,l[2]) 
+}
+
+#Function to filter large changes in bearing. Returns boolean unless
+#returnDiffs==TRUE
+bearingFilter <- function(bearing,q=NULL,z=NULL,returnDiffs=FALSE){
+  if(!xor(is.null(q),is.null(z))&!returnDiffs) stop('Input quantiles or Z-score')
+  #Difference in compass bearings (in degrees)
+  bearingDiff <- function(x1,x2){
+    x <- x1-x2
+    x <- ifelse(abs(x)>180,x-(360*sign(x)),x) #Angle differences can't be >180
+    return(x)
+  }
+
+  #Looks 1 point ahead and behind
+  bd <- cbind(bearingDiff(lag(bearing),bearing),
+              bearingDiff(lead(bearing),bearing))
+  #Maximum bearing difference ahead and behind
+  bd <- apply(bd,1,function(x) max(abs(x),na.rm=TRUE)*sign(x[which.max(abs(x))]))
+
+  if(returnDiffs) return(bd) #Return bearing differences only, without filtering
+
+  if(!is.null(q)){
+    ret <- QuantileFilter(bd,q=q)
+  } else {
+    ret <- ZscoreFilter(bd,z=z)
+  }
+  return(ret)
+}
+
+#Positional difference filter - filters out very distant and very close points.
+#Returns boolean unless returnDiffs==TRUE
+posFilter <- function(data,q=NULL,returnDiffs=FALSE){ 
+  require(sf)
+  if(is.null(q)&!returnDiffs) stop('Input upper quantile')
+
+  if(units(st_distance(data[1,],data[2,]))$numerator!='m'){
+    warning('Position differences not in meters')
+  } 
+  coords <- st_coordinates(data) #Get coordinates
+  #Distances between points
+  pdiff <- sapply(1:(nrow(coords)-1),function(i) as.numeric(dist(coords[i:(i+1),]))) 
+  pdiff <- cbind(c(pdiff,NA),c(NA,pdiff)) #Forward and backward lags
+  pdiff <- apply(pdiff,1,max,na.rm=TRUE) #Maximum distance ahead and behind
+  if(returnDiffs){
+    return(pdiff)
+  } else {
+    return(QuantileFilter(pdiff,q)) #Note: uses 2-sided quantiles
+  }
+}
+
+#Filter for (forward and backward) lagged speed differences. Returns boolean
+dSpeedFilter <- function(speed,l=c(-1,1),perc=0.2){
+
+  #Matrix of % diffs
+  llist <- sapply(l,function(x) (lag2(speed,x)-speed)/lag2(speed,x))
+
+  #Are any lagged speed values > percent change threshold?
+  ret <- !apply(llist,1,function(y) any(abs(y)[!is.na(y)]>perc))
+
+  return(ret)
+}
+
+#Overloaded lag function that takes negative values
+lag2 <- function(x,n){
+  require(dplyr)
+  if(n==0) {
+    return(x) #No lag
+  } else if(n>0){
+    lag(x,n) #Positive lag
+  } else {
+    lead(x,abs(n)) #Negative lag
+  } 
+}
+
+#Filtering pipeline  --------------------------
+
+library(tidyverse)
+library(sf)
+
+#Note: dat == sf dataframe of yield data with columns: DryYield (T/ha),
+#TrackAngle (compass bearing), Speed (km/hr), cropType (Wheat, Canola, Peas)
+
+dat <- dat %>% 
+  #Vega et al 2019 spatial "inlier" filter - takes a few minutes...
+  mutate(vegaFilt = vegaFilter(.,DryYield,nDist = 30))
+
+dat <- dat %>% 
+  #"BS" filter - removes values >10.75 T/ha for wheat, >8 T/ha for peas & canola
+  mutate(noBS = DryYield<ifelse(cropType=='Wheat',10.75,8)) %>% 
+  #Trim dry yield outliers
+  mutate(Qfilt = QuantileFilter(DryYield,q=0.98)) %>% 
+  #Trim extreme bearing changes (turning)
+  mutate(bFilt = bearingFilter(TrackAngle,q=0.98)) %>% 
+  #Trim absolute speed outliers
+  mutate(speedFilt = QuantileFilter(Speed,q=0.98)) %>% 
+  #Trim speed differences (>20% change 2 steps forward and backward, suggested
+  #by Lyle et al 2014)
+  mutate(dSpeedFilt = dSpeedFilter(Speed,l=c(-2,-1,1,2),perc = 0.2)) %>% 
+  #Trim points that are far away from eachother
+  mutate(posFilt = posFilter(.,q=0.98)) %>% 
+  #Combine filter criteria
+  mutate(allFilt = noBS & vegaFilt & Qfilt & bFilt & speedFilt & 
+           dSpeedFilt & posFilt) %>%  
+  #Remove filtered values
+  filter(allFilt)
+
 ```
 
-```{r wheatPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on wheat yield.', out.width='100%'}  
-  include_graphics('../Figures/groundSpeed_wheat.png',dpi=NA)
+# Appendix B: R Code for Modelling {-}
+
+```{r modelCode, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}
+
+# Fit models ----------------------------------------------------------------
+
+library(mgcv)
+
+#Note: dat == dataframe of yield data with columns: DryYield (T/ha), dist
+#(distance to nearest field boundary, m), boundaryType (type of boundary,
+#factor), E,N (easting and northing distance from centre of field, m)
+
+#Uses thin-plate splines with extra shrinkage (bs = 'ts'), with 5 basis
+#functions for distance smoothers and 80 basis functions for spatial smoothers
+#(see Wood 2017 for more details)
+
+#Variance smoothers
+f2 <- '~ s(dist,k=5,bs="ts",by=boundaryType) + s(E,N,k=80)' 
+#Mean smoothers
+f <- 'sqrt(DryYield)~ s(dist,k=5,bs="ts",by=boundaryType) + s(E,N,k=80)') 
+#Combine into a list
+flist <- list(as.formula(f),as.formula(f2))
+
+#Fit model - takes about 45 mins for a 50000 point dataset using an AMD Ryzen 7
+#3800X 8-core processor
+mod <- gam(flist,data=dat,family=gaulss())
+
 ```
-
-```{r peasPlotSpeed, echo=FALSE, fig.cap='Effect of ground speed and order (sequence of harvest) on pea yield.', out.width='100%'}  
-  include_graphics('../Figures/groundSpeed_peas.png',dpi=NA)
-```

### 2021-10-27 runModels.R

@@ -732,7 +732,7 @@ for(i in 1:length(croptype)){
 use <- which(datSource$filename == 'Trent_Clark JOHNSON 2014') #Could also use this
 
 load(datSource$modelPath2[use]) #Load model
-load("C:\\Users\\Samuel\\Documents\\Projects\\UofC postdoc\\yield-analysis-2021\\Figures\\ExamplePlots\\Trent_Clark JOHNSON 2014 modList.Rdata") #Path on Multivac
+# load("C:\\Users\\Samuel\\Documents\\Projects\\UofC postdoc\\yield-analysis-2021\\Figures\\ExamplePlots\\Trent_Clark JOHNSON 2014 modList.Rdata") #Path on Multivac
 
 #Spatial smoothers
 
@@ -776,7 +776,8 @@ palette <- 'YlGnBu'
     geom_sf(data=fieldBoundaryType,aes(geometry=geometry),show.legend= 'line')+
     scale_colour_distiller(type='div',palette = palette, direction = -1)+
     labs(col='Yield (T/ha)',title='Raw Yield Data')+
-    theme(legend.position='bottom'))
+    theme(legend.position='bottom',axis.text = element_blank(),axis.ticks = element_blank())
+  )
 
 # temp <- lapply(levels(fieldBoundaryType$type),function(x) filter(fieldBoundaryType,type==x)) %>% 
 #   set_names(nm=levels(fieldBoundaryType$type))
@@ -801,7 +802,7 @@ useSmooths <- which(grepl('(E,N)',sapply(modList$smooths,function(x) x$label),fi
     geom_sf(data=fieldBoundaryType,aes(geometry=geometry),show.legend= 'line')+
     labs(fill='Yield (T/ha)',title='Yield Average Smoother')+
     scale_fill_distiller(type='div',palette = palette, direction = -1) +
-    theme(legend.position='bottom')
+    theme(legend.position='bottom',axis.text = element_blank(),axis.ticks = element_blank())
   )
 
 (p3 <- hexGrid %>% 
@@ -811,7 +812,8 @@ useSmooths <- which(grepl('(E,N)',sapply(modList$smooths,function(x) x$label),fi
     geom_sf(data=fieldBoundaryType,aes(geometry=geometry),show.legend= 'line')+
     labs(fill='log Yield SD',title='Yield Variability Smoother')+
     scale_fill_distiller(type='div',palette = palette, direction = -1) +
[…]

### 2021-11-04 Manuscript/manuscript.Rmd
Updated manuscript for submission

@@
   \usepackage{setspace}
-  %\linenumbers
+  \linenumbers
 
@@
 
+fieldAreas <- sapply(paste0('.',datSource$boundaryPath),function(x){
+  require(sf)
+  read_sf(x) %>% st_area() %>% sum()/10000 #Field areas in hectares
+})
+
 ```
@@
 \newpage
-<!-- \doublespacing -->
-\onehalfspacing
+\doublespacing
+<!-- \onehalfspacing -->
 
@@
 However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
-Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ["multifunctionality" _sensu_ @mitchell2014; @frei2018].
+Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ["multifunctionality" _sensu_ @mitchell2014; @frei2018; but see also @karp2018].
 Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.
@@
 The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`), canola (`r sum(datSourceFinal$crop=='Canola')`), or peas (`r sum(datSourceFinal$crop=='Peas')`), for a total of `r nrow(datSourceFinal)` field-years of data.
-Individual fields contained between `r min(yearsPerField)` and `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`), containing a total of `r round(sum(datSourceFinal$npoints)/1e6,1)` million data points.
+Individual fields ranged from `r round(min(fieldAreas),1)` to `r round(max(fieldAreas),1)` ha in area (mean: `r round(mean(fieldAreas),1)` ha), and contained between `r min(yearsPerField)` and `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`), for a total of `r round(sum(datSourceFinal$npoints)/1e6,1)` million data points.
 
@@
 Abiotic ecosystem services, such as reduced wind speed or temperature, also depend on the structure of the boundary and whether wind or temperature are limiting crop production.
-Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) under what conditions field boundaries provide tangible ecosystem services.
+Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) under what conditions field boundaries provide tangible ecosystem services [@karp2018].
 We consider these reasons below:
@@
 
-# Author Contributions
+<!-- # Author Contributions -->
 
-SVJR, LHN, and PG conceived of the project.
-SVJR collected data, conducted analysis, and wrote the manuscript.
+<!-- SVJR, LHN, and PG conceived of the project. -->
+<!-- SVJR collected data, conducted analysis, and wrote the manuscript. -->
 
@@
 We also thank Autumn Barnes and Keith Gabert of the Canola Council of Canada, and Laurel Thompson and JP Pettyjohn of Lakeland College for connecting us with farmers and agronomists.
+
+# Funding Sources
+
 Funding for this research was provided by Ducks Unlimited Canada's Institute for Wetland and Waterfowl Research, the Alberta Canola Producers Commission, Manitoba Canola Growers Association, SaskCanola, the Alberta Biodiversity Monitoring Institute, and the Alberta Conservation Association.

### 2021-11-04 Manuscript/manuscript.tex
Updated manuscript for submission

@@
 % Pandoc header
-\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} \usepackage{setspace} %\linenumbers
+\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} \usepackage{setspace} \linenumbers
 \usepackage{booktabs}
@@
 \newpage
-
-\onehalfspacing
+\doublespacing
 
@@
 However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
-Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ({``multifunctionality''} \emph{sensu} Mitchell \emph{et al.} 2014; Frei \emph{et al.} 2018).
+Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ({``multifunctionality''} \emph{sensu} Mitchell \emph{et al.} 2014; Frei \emph{et al.} 2018; but see also Karp \emph{et al.} 2018).
 Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.
@@
 The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
-Individual fields contained between 1 and 6 years of data (mean: 3.1), containing a total of 14.4 million data points.
+Individual fields ranged from 5.9 to 369 ha in area (mean: 72.2 ha), and contained between 1 and 6 years of data (mean: 3.1), for a total of 14.4 million data points.
 
@@
 Abiotic ecosystem services, such as reduced wind speed or temperature, also depend on the structure of the boundary and whether wind or temperature are limiting crop production.
-Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) under what conditions field boundaries provide tangible ecosystem services.
+Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) under what conditions field boundaries provide tangible ecosystem services (Karp \emph{et al.} 2018).
 We consider these reasons below:
@@
 
-\hypertarget{author-contributions}{%
-\section{Author Contributions}\label{author-contributions}}
-
-SVJR, LHN, and PG conceived of the project.
-SVJR collected data, conducted analysis, and wrote the manuscript.
-
 \hypertarget{acknowledgements}{%
@@
 We also thank Autumn Barnes and Keith Gabert of the Canola Council of Canada, and Laurel Thompson and JP Pettyjohn of Lakeland College for connecting us with farmers and agronomists.
+
+\hypertarget{funding-sources}{%
+\section{Funding Sources}\label{funding-sources}}
+
 Funding for this research was provided by Ducks Unlimited Canada's Institute for Wetland and Waterfowl Research, the Alberta Canola Producers Commission, Manitoba Canola Growers Association, SaskCanola, the Alberta Biodiversity Monitoring Institute, and the Alberta Conservation Association.
@@
 \leavevmode\hypertarget{ref-AAFC_data}{}%
-Agriculture and Agri-Food Canada. (2018). \emph{{Annual Crop Inventory}}. Available at: \url{https://open.canada.ca/data/en/dataset/ba2645d5-4458-414d-b196-6303ac06c1c9}. Last accessed.
+Agriculture and Agri-Food Canada.
+(2018).
+\emph{{Annual Crop Inventory}}.
+Available at: \url{https://open.canada.ca/data/en/dataset/ba2645d5-4458-414d-b196-6303ac06c1c9}.
+Last accessed 6 February 2019.
 
@@
 The effectiveness of flower strips and hedgerows on pest control, pollination services and crop yield: A quantitative synthesis.
-\emph{Ecology Letters}.
+\emph{Ecology Letters}, 23, 1488--1498.
 
@@
 (2012).
-Effect of shelterbelts on winter wheat yields in sanded farmland of north-western shandong province, china.
+Effect of shelterbelts on winter wheat yields in sanded farmland of north-western {Shandong} province, {China}.
 \emph{Journal of Food Agriculture \& Environment}, 10, 1399--1403.
@@
 Field boundary features can stabilise bee populations and the pollination of mass-flowering crops in rotational systems.
-\emph{Journal of Applied Ecology}, In press.
+\emph{Journal of Applied Ecology}, 58, 2287--2304.
 
@@
 \leavevmode\hypertarget{ref-griffin2007}{}%
-Griffin, T.W., Brown, J.P. \& Lowenberg-DeBoerc, J. (2007). \emph{Yield monitor data analysis protocol: A primer in the management and analysis of precision agriculture data}.
+Griffin, T.W., Brown, J.P. \& Lowenberg-DeBoerc, J. (2007). \emph{Yield monitor data analysis protocol: A primer in the management and analysis of precision agriculture data}. SSRN.
 
@@
 
+\leavevmode\hypertarget{ref-karp2018}{}%
+Karp, D.S., Chaplin-Kramer, R., Meehan, T.D., Martin, E.A., DeClerck, F., Grab, H., \emph{et al.} (2018).
+Crop pests and predators exhibit inconsistent responses to surrounding landscape composition.
+\emph{Proceedings of the National Academy of Sciences}, 115, E7863--E7870.
+
 \leavevmode\hypertarget{ref-kassambara2020}{}%
@@
 (2021).
-Impacts of field-edge flower plantings on pollinator conservation and ecosystem service delivery {{}} a meta-analysis.
+Impacts of field-edge flower plantings on pollinator conservation and ecosystem service delivery -- a meta-analysis.
 \emph{Agriculture, Ecosystems {\&} Environment}, 310, 107290.
@@
 Effectiveness of agri-environmental management on pollinators is moderated more by ecological contrast than by landscape structure or land-use intensity.
-\emph{Ecology Letters}, In press.
+\emph{Ecology Letters}, 22, 1493--1500.
 
@@
 (2021).
-Medium-resolution multispectral satellite imagery in precision agriculture: Mapping precision canola (\emph{brassica napus} {L.}) Yield using {Sentinel-2} time series.
+Medium-resolution multispectral satellite imagery in precision agriculture: Mapping precision canola ({\emph{Brassica napus}} {L.}) Yield using {Sentinel-2} time series.
 \emph{Precision Agriculture}, In Review (Preprint: https://www.essoar.org/doi/10.1002/essoar.10507479.2).
@@
 (1999).
-Impact of \emph{populus deltoides} and \emph{dalbergia sissoo} shelterbelts on wheat {}a comparative study.
+Impact of {\emph{Populus deltoides}} and {\emph{Dalbergia sissoo}} shelterbelts on wheat {}a comparative study.
 \emph{International Tree Crops Journal}, 10, 51--60.
@@
 Ecosystem services of tree windbreaks in rural landscapes -- a systematic review.
-\emph{Environmental Research Letters}, In press.
+\emph{Environmental Research Letters}, 16, 103002.
 
@@
 \CommentTok{\#functions for distance smoothers and 80 basis functions for spatial smoothers}
-
-\CommentTok{\#Variance smoother}
+\CommentTok{\#(see Wood 2017 for more details)}
+
+\CommentTok{\#Variance smoothers}
 \NormalTok{f2 }\OtherTok{\textless{}{-}} \StringTok{\textquotesingle{}\textasciitilde{} s(dist,k=5,bs="ts",by=boundaryType) + s(E,N,k=80)\textquotesingle{}} 
-\CommentTok{\#Means smoother}
+\CommentTok{\#Mean smoothers}
 \NormalTok{f }\OtherTok{\textless{}{-}} \StringTok{\textquotesingle{}sqrt(DryYield)\textasciitilde{} s(dist,k=5,bs="ts",by=boundaryType) + s(E,N,k=80)\textquotesingle{}}\ErrorTok{)} 

### 2022-03-04 Manuscript/manuscript.Rmd
Updated manuscript

@@
 
-```{r hypotheses, echo=FALSE, fig.cap='Potential yield patterns, depending on ecosystem service effects with distance.
-Scenario 1: no effect of boundary.
-Scenario 2: negative edge effect on yield.
-Scenario 3: positive ecosystem service effect on yield.
-Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.', out.width='100%'}
-  include_graphics('../Figures/ExamplePlots/hypotheses.png',dpi=NA)
-```
-
 Unfortunately, data to inform these decisions are often minimal, as observing boundary effects requires large amounts of expensive and time-consuming crop yield data.
@@
 
-Precision yield data were collected directly from farmers across Alberta.
+Precision yield data were collected directly from farmers across Alberta (Figure \@ref(fig:fieldLocations)).
 Farmers were solicited for yield data through local agronomists, and we received `r nrow(datSource)` field-years of data from `r length(unique(datSource$grower))` growers across a total of `r length(unique(datSource$year))` years (`r min(unique(datSource$year))`-`r max(unique(datSource$year))`).
@@
 
-```{r fieldLocataions, echo=FALSE, fig.cap='Locations of fields that data were collected from', out.width='100%'}  
-  include_graphics('../Figures/Field Locations.png',dpi=NA)
-```
-
 Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed $\times$ interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
@@
 Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows [@griffin2007; @whelan2013], or when the combine is accelerating or decelerating [@arslan2002].
-Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix A for R code):
+Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix B for R code):
 
@@
 
-```{r spatialSmooths, echo=FALSE, fig.cap="Raw data and spatial smoothers from a single field.
-Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.
-Field dimensions are approximately 800 x 800 m, but coordinates are hidden to protect data privacy.", out.width='100%'}
-  include_graphics('../Figures/ExamplePlots/spatialSmooths.png',dpi=NA)
-```
-
-```{r distSmooths, echo=FALSE, fig.cap='Distance smoothers from a single field, showing a positive saturating effect of distance on average yield (first row), while the variability smoothers show a general decrease in yield SD (second row).', out.width='100%'}
-  include_graphics('../Figures/ExamplePlots/distSmooths.png',dpi=NA)
-```
-
 ## Overall results
@@
 <!-- Not sure if we should talk about this. Not very many fields from southern regions, so it's hard to compare. Perhaps mention in Discussion -->
-
-```{r boundaryTypes, echo=FALSE, fig.cap='Boxplot of field boundary compositions across sampled fields.
-The x-axis indicates the boundary type, while the y-axis indicates the percentage of all field boundaries covered by that boundary type.', out.width='100%'}
-
-bLengths %>% 
-  mutate(croptype=datSourceFinal$crop) %>% 
-  mutate(location=ifelse(grepl('(Alvin|Dean)',filename),'Southern Alberta','Central Alberta')) %>%
-  group_by(location) %>% mutate(nFields=length(unique(filename))) %>% ungroup() %>% 
-  mutate(location=paste0(location,' (N = ',nFields,')')) %>%
-  select(filename,croptype,location,contains('prop_')) %>% 
-  pivot_longer(contains('prop_')) %>% 
-  mutate(name=gsub('prop_','',name),value=value*100) %>% na.omit() %>%
-  mutate(name=factor(name,labels=c('Bare','Grassland','Other\nCrop','Shelterbelt','Standard','Wetland'))) %>% 
-  group_by(croptype) %>% mutate(croptype=paste0(croptype,'\n(N = ',length(unique(filename)),')')) %>% 
-  ggplot(aes(x=name,y=value,fill=croptype))+geom_boxplot()+
-  facet_wrap(~location,ncol=1) +
-  labs(y='Percentage of field boundary',fill='Crop')+
-  scale_fill_manual(values=c('grey50','grey80','white'))+
-  theme(axis.title.x = element_blank(),legend.position = 'bottom')
-
-# bLengthSummary %>% #Table version
-#   kable(.,format='latex',caption='Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.',col.names = c('Type','Mean','SD','Number of Fields'),digits = c(1,4,4,1),escape = FALSE)
-
-```
 
@@
 
-```{r smoothSummary, echo=FALSE, out.width='100%'}
-
-canolaMods %>% group_by(Smoother) %>% 
-  summarize(N=n(),meanChi=mean(Chisq),meanP=mean(pval),propSig=sum(pval<(0.05/N))/n()) %>% 
-  ungroup() %>% 
-  mutate(par=ifelse(grepl('\\.',Smoother),'Variability','Mean')) %>% 
-  mutate(type=case_when(
-    grepl('boundaryType',Smoother) ~ gsub('s[\\.1]*\\(dist\\)\\:boundaryType','',Smoother),
-    grepl('E,N',Smoother) ~ 'Spatial Smoother'
-  )) %>% 
-  select(-N,-Smoother) %>% relocate(type,par,meanChi,meanP,propSig) %>% 
-  pivot_wider(id_cols=type,names_from=par,values_from=c(meanChi,meanP,propSig)) %>% 
-  relocate(type,meanChi_Mean,meanP_Mean,propSig_Mean,meanChi_Variability,meanP_Variability,propSig_Variability) %>% 
-  mutate(type=c('Bare','Grassland','Other Crop','Shelterbelt','Standard','Wetland','Spatial Smoother')) %>% 
-  kable(.,format='latex',caption='Mean $\\chi^2$, p-values, and proportion of smoother p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all models.',col.names = c('Smoother Type','$\\chi^2$-value','p-value','Prop.
-<0.05','$\\chi^2$-value','p-value','Prop.
-<0.05'),
-        digits = c(1,1,3,3,1,3,3),escape = FALSE) %>% 
-  add_header_above(c('', 'Mean Yield' = 3, 'Yield Variability' = 3))
-
-```
-
 Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha  (Figure \@ref(fig:canolaPlot) top row).
@@
 Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from wetland boundaries, indicating a potential stabilizing effect, but this was offset by the large increase in yield variability at the edge of wetlands.
-
-```{r canolaPlot, echo=FALSE, fig.cap='Field boundary effect on canola yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \\% refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
-  include_graphics('../Figures/ModelSummary3a_canola.png',dpi=NA)
-```
 
@@
 
-```{r wheatPlot, echo=FALSE, fig.cap='Field boundary effect on wheat yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \\% refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
-  include_graphics('../Figures/ModelSummary3a_wheat.png',dpi=NA)
-```
-
 <!-- Both average and variability of pea yields were spatially dependent (average p-values: `r with(peaMods,formatC(meanP[Smoother=='s(E,N)']))`, `r with(peaMods,formatC(meanP[Smoother=='s.1(E,N)'],3))`).
@@
 Unlike canola and wheat crops, there was no obvious evidence of an intermediate increase in yield with distance from the field boundary.
-
-```{r peaPlot, echo=FALSE, fig.cap='Field boundary effect on pea yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \\% refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
-  include_graphics('../Figures/ModelSummary3a_peas.png',dpi=NA)
-```
 
@@
 
+\newpage
+
+# Figures {-}
+
+```{r hypotheses, echo=FALSE, fig.cap='Potential yield patterns, depending on ecosystem service effects with distance.
+Scenario 1: no effect of boundary.
+Scenario 2: negative edge effect on yield.
+Scenario 3: positive ecosystem service effect on yield.
+Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.', out.width='100%'}
+include_graphics('../Figures/ExamplePlots/hypotheses.png',dpi=NA)
+```
+
+```{r spatialSmooths, echo=FALSE, fig.cap="Raw data and spatial smoothers from a single field.
+Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.
+Field dimensions are approximately 800 x 800 m, but coordinates are hidden to protect data privacy.", out.width='100%'}
+include_graphics('../Figures/ExamplePlots/spatialSmooths.png',dpi=NA)
+```
+
+```{r distSmooths, echo=FALSE, fig.cap='Distance smoothers from a single field, showing a positive saturating effect of distance on average yield (first row), while the variability smoothers show a general decrease in yield SD (second row).', out.width='100%'}
+include_graphics('../Figures/ExamplePlots/distSmooths.png',dpi=NA)
+```
+
+```{r canolaPlot, echo=FALSE, fig.cap='Field boundary effect on canola yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and \\% refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
+include_graphics('../Figures/ModelSummary3a_canola.png',dpi=NA)
+```
+
+```{r wheatPlot, echo=FALSE, fig.cap='Field boundary effect on wheat yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and \\% refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
+include_graphics('../Figures/ModelSummary3a_wheat.png',dpi=NA)
+```
+
+```{r peaPlot, echo=FALSE, fig.cap='Field boundary effect on pea yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and \\% refers to the average percentage of field boundary accounted for by this boundary type.', out.width='100%'}
+include_graphics('../Figures/ModelSummary3a_peas.png',dpi=NA)
+```
+
+\newpage
+
+# Tables {-}
+
+```{r smoothSummary, echo=FALSE, out.width='100%'}
+
+canolaMods %>% group_by(Smoother) %>% 
+  summarize(N=n(),meanChi=mean(Chisq),meanP=mean(pval),propSig=sum(pval<(0.05/N))/n()) %>% 
+  ungroup() %>% 
+  mutate(par=ifelse(grepl('\\.',Smoother),'Variability','Mean')) %>% 
+  mutate(type=case_when(
+    grepl('boundaryType',Smoother) ~ gsub('s[\\.1]*\\(dist\\)\\:boundaryType','',Smoother),
+    grepl('E,N',Smoother) ~ 'Spatial Smoother'
+  )) %>% 
+  select(-N,-Smoother) %>% relocate(type,par,meanChi,meanP,propSig) %>% 
+  pivot_wider(id_cols=type,names_from=par,values_from=c(meanChi,meanP,propSig)) %>% 
+  relocate(type,meanChi_Mean,meanP_Mean,propSig_Mean,meanChi_Variability,meanP_Variability,propSig_Variability) %>% 
+  mutate(type=c('Bare','Grassland','Other Crop','Shelterbelt','Standard','Wetland','Spatial Smoother')) %>% 
+  kable(.,format='latex',caption='Mean $\\chi^2$, p-values, and proportion of smoother p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all models.',col.names = c('Smoother Type','$\\chi^2$-value','p-value','Prop.
+<0.05','$\\chi^2$-value','p-value','Prop.
+<0.05'),
+        digits = c(1,1,3,3,1,3,3),escape = FALSE) %>% 
+  add_header_above(c('', 'Mean Yield' = 3, 'Yield Variability' = 3))
+
+```
+
 # Acknowledgements

### 2022-03-04 Manuscript/manuscript_figures.tex
Updated manuscript

--- 2021-10-18 Manuscript/manuscript.tex
@@
 \usepackage{graphicx}
-\usepackage{booktabs} % book-quality tables
 %%%%%%%%%%%%%%%% end my additions to header
@@
 \bibliographystyle{elsarticle-harv}
-\usepackage{longtable}
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\newenvironment{Shaded}{}{}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
+\newcommand{\AttributeTok}[1]{#1}
+\newcommand{\BaseNTok}[1]{#1}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
+\newcommand{\ConstantTok}[1]{#1}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.00,1.00}{#1}}
+\newcommand{\DataTypeTok}[1]{#1}
+\newcommand{\DecValTok}[1]{#1}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{#1}
+\newcommand{\FunctionTok}[1]{#1}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.00,1.00}{#1}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{#1}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{1.00,0.25,0.00}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{1.00,0.25,0.00}{#1}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
+\newcommand{\VariableTok}[1]{#1}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.00,0.50,0.50}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \ifxetex
@@
 % Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
 \newlength{\csllabelwidth}
 \setlength{\csllabelwidth}{3em}
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
 % for Pandoc 2.8 to 2.10.1
@@
 % For Pandoc 2.11+
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
  {% don't indent paragraphs
@@
  {}
-\usepackage{calc} % for calculating minipage widths
+\usepackage{calc}
 \newcommand{\CSLBlock}[1]{#1\hfill\break}
 \newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 % Pandoc header
-\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} \usepackage{setspace} \linenumbers
+\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
 \usepackage{booktabs}
@@
 \begin{document}
-\begin{frontmatter}
 
-  \title{Living on the edge: crop boundary effects on precision crop yields in Alberta, Canada}
-    \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
-   \ead{samuel.robinson@ucalgary.ca} 
-    \author[University of Calgary]{Lan H. Nguyen}
-   \ead{hoanglan.nguyen@ucalgary.ca} 
-    \author[University of Calgary]{Paul Galpern}
-   \ead{paul.galpern@ucalgary.ca} 
-      \address[University of Calgary]{2500 University Drive NW, Calgary, AB}
-      \cortext[1]{Corresponding Author}
+\hypertarget{figures}{%
+\section*{Figures and Tables}\label{figures}}
 
-  \begin{abstract}
-  Field boundaries can improve crop yields by creating better abiotic conditions for crop growth (e.g.~moisture, temperature), and can act as refuges for beneficial arthropods.
-This suggests that beneficial crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services (e.g.~pollination or pest suppression) from the field boundary.
-However, there is little large-scale evidence showing this, largely because yield is costly and time-consuming to measure.
-Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is relatively cheap and commonly measured by growers.
-In this study, we used 252 field-years of yield monitor data from three common crops -- wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}) -- recorded across Alberta, Canada, and examined how yield varied with distances from common crop boundary types.
-Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease with distance.
-There was evidence of an intermediate increase in yield for wheat away from shelterbelts, and a weak increase in canola, but this was not seen for other crop types or boundary types.
-This study represents one of the first uses of precision yield data to measure ecosystem service provision at large spatial scales.
-  \end{abstract}
-   \begin{keyword} precision yield; spillover; landscape;\end{keyword}
- \end{frontmatter}
+\begin{figure}
+\includegraphics[width=1\linewidth]{hypotheses} \caption{Potential yield patterns, depending on ecosystem service effects with distance.
+Scenario 1: no effect of boundary.
+Scenario 2: negative edge effect on yield.
+Scenario 3: positive ecosystem service effect on yield.
+Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.}
+\end{figure}
 
 \newpage
-\doublespacing
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth (Ramankutty \emph{et al.} 2018).
-This has allowed increases in human population and increased global stability in production (Foley \emph{et al.} 2011).
-However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
-Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ({``multifunctionality''} \emph{sensu} Mitchell \emph{et al.} 2014; Frei \emph{et al.} 2018).
-Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.
-
-SNL in agroecosystems is important for both agricultural production and conservation (Prieto-Benítez \& Méndez 2011), and can create better biotic and abiotic conditions for crop growth (Case \emph{et al.} 2019; Weninger \emph{et al.} 2021; but see Lowe \emph{et al.} 2021).
-SNL acts as habitat for mobile organisms, and can act as a source of ecosystem services, such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this depends on both the crop system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017) and landscape context (Marja \emph{et al.} 2019; Boetzl \emph{et al.} 2020).
-SNL can also create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion (Kort 1988; Brandle \emph{et al.} 2004; Wratten \emph{et al.} 2012).
-Field boundaries (e.g.~grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are SNL of particular interest, as they are extremely common and require less land than field-sized agricultural set-asides.
-
-Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
-Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure \ref{fig:hypotheses}, Scenario 2).
-Alternatively, yield may be higher near field boundaries (Scenario 3) if they provide helpful ecosystem services, as these also tend to be occur close to the field edge (Duelli \& Obrist 2003).
-For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Kowalchuk \& Jong 1995; Garibaldi \emph{et al.} 2011).
-Depending on the rate at which positive and negative ecosystems services decline with distance, yet another scenario is possible.
-Yield may also be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services.
-However, this scenario is rarely considered within studies of SNL or boundary effects (but see Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/hypotheses} \caption{Potential yield patterns, depending on ecosystem service effects with distance.
-Scenario 1: no effect of boundary.
-Scenario 2: negative effect on yield.
-Scenario 3: positive effect on yield.
-Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.}\label{fig:hypotheses}
+\includegraphics[width=1\linewidth]{spatialSmooths} \caption{Raw data and spatial smoothers from a single field.
+Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.
+Field dimensions are approximately 800 x 800 m, but coordinates are hidden to protect data privacy.}
 \end{figure}
 
-Boundary effects are difficult to observe without large amounts of crop data, which is typically expensive and time-consuming to collect.\\
-Because of this, studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality (Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
-Meta-analyses are a common solution to this problem (e.g.
-Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs.
-However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
-Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems (Redhead \emph{et al.} 2020; Hünicken \emph{et al.} 2021); typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
-There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
-
-In this paper use a precision yield dataset from Alberta, Canada to ask the following questions:
-1. How does crop yield change with distance from the edge of field?
-2. Does this depend on type of field edge?
-3. Is there an intermediate distance where yield is maximized or variance is minimized?
-To our knowledge, no other study has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{data-collection}{%
-\subsection{Data collection}\label{data-collection}}
-
-Precision yield data were collected directly from farmers across Alberta.
-Farmers were solicited for yield data through local agronomists, and we received 298 field-years of data from 5 growers across a total of 7 years (2014-2020).
-We converted raw data to a standard csv format using Ag Leader SMS®.
-85\% of the crop types where either wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}), three of the most common crops in rotation in Alberta (Statistics Canada 2014).
-The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
-Individual fields contained between 1 and 6 years of data (mean: 3.1), containing a total of 14.4 million data points.
+\newpage
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/Field Locations} \caption{Locations of fields that data were collected from}\label{fig:fieldLocataions}
+\includegraphics[width=1\linewidth]{distSmooths} \caption{Distance smoothers from a single field, showing a positive saturating effect of distance on average yield (first row), while the variability smoothers show a general decrease in yield SD (second row).}
 \end{figure}
 
-Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
-We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m\(^2\)), dry yield (tonnes), and the spatial location.
-Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
-Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
-Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows (Griffin \emph{et al.} 2007; Whelan \& Taylor 2013), or when the combine is accelerating or decelerating (Arslan \& Colvin 2002).
-Therefore, we used the following filtering procedure to clean the data prior to model fitting:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Removed outliers and ``inliers'' using the spatial filtering procedure from (Vega \emph{et al.} 2019)
-\item
-  Removed unrealistic yield values (10.75 T/ha for wheat, 8 T/ha for canola and peas), and values of zero or less
-\item
-  Removed extremely low and high outliers (below 1st percentile, above 99th percentile) for: dry yield, steering angle differences, ground speed
-\item
-  Removed points with large changes (\textgreater20\%) in ground speed
-\item
-  Removed extremely low and high changes in position, indicating that the combine had stopped or moved to another part of the field between data collection
-\item
-  Sub-sampled filtered data to 50,000 data points per field (reduce computation time)
-\end{enumerate}
-
-Field boundaries were digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data (Agriculture and Agri-Food Canada 2018).
-Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions.
-For example, ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
-This makes accurate and consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:
-1. Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
-2. Wetland: permanent wetland; borders are largely unchanged from year-to-year
-3. Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
-4. Other crop: annual crop or pasture with little or no visible boundary between planted areas
-5. Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
-6. Grassland: permanent seminatural grassland or pasture (not in rotation)
-
-\hypertarget{analysis}{%
-\subsection{Analysis}\label{analysis}}
-
-Yield data from each field contains both random and systematic errors that must be modeled in order to reveal underlying changes in yield.
-Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, previous agricultural practices, and other factors not directly related to boundary distance.
-To account for this, we used additive models of the effect of boundary distance on crop yield while accounting for within-field spatial variation.
-
-We fit the following model to each field-year of data:
-
-\begin{equation}
-  \begin{split}
-  \sqrt{Yield} \sim & Normal (\mu, \sigma)\\
-  \mu = & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
-   & f(\text{Easting}, \text{Northing}, b=60)\\
-  log(\sigma) =  & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
-   & f(\text{Easting}, \text{Northing}, b=60)\\
-  \end{split}
-  \end{equation}
-
-\begin{itemize}
-\tightlist
-\item
-  where:
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Yield = weight of crop gathered per area of harvest (T/ha)
-\item
-  Boundary Distance = distance from field boundary type \emph{i} (m)
-\item
-  Easting, Northing = distance from centre of field (m)
-\item
-  Sequence = order of harvest within field (1--N points)
-\item
-  f(x,b) = penalized thin-plate regression spline, where \emph{x} is the predictor and \emph{b} is the number of basis dimensions
-\end{enumerate}
-
-In addition to modeling edge effects (our variable of interest), this model also accounts for within-field spatial variation not related to edges.
-Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
-Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
-We used 5 basis dimensions for the distance smoother and 60 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \ref{fig:hypotheses}).
-All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).
-
-The residuals in most of the models displayed both spatial autocorrelation in the residuals, indicating ``micro-scale'' variation that was not captured by the spatial smoothers (Cressie \& Wikle 2011; Wikle \emph{et al.} 2019).
-This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.~extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 60), we decided to leave the models as-is.
-Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
-That is, combining of fields occurs in patterns (e.g.~east-to-west) so it is impossible to tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
-We considered the latter to be a more realistic assumption, so we assigned all random variation as spatial in nature.
-Customized combine driving patterns may be able to detect temporal variation, but this is beyond the scope of this research.
-
-To consider results from all field-level models together, we fit models of each field-year independently and then fit overall smoothers as averages of the field-level smoothers.
-This does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping:
-1. Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard errors
-2. Use posterior sample to simulate a new smoother from each field
-3. Fit model of simulated smoothers from all fields, and save this ``meta-smoother''
-4. Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers
-This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
-Spatial variability was not averaged between fields because within-field variability is not directly comparable between fields.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{field-level-example}{%
-\subsection{Field-level example}\label{field-level-example}}
-
-The filtered data from each field contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
-To illustrate this, we show the results from a single wheat field, where we show the relative impact of boundary distance and spatial variation on both yield mean and yield SD (``patchiness'').
-Figure \ref{fig:spatialSmooths}a shows the filtered data collected from a single field, demonstrating a large amount of variation even after filtering.
-Yield mean tended to increase with distance from field boundaries while yield SD tended to decrease, but there were no consistent differences between boundary types (Figure \ref{fig:distSmooths}).
-Finally, the random spatial smoothers in Figure \ref{fig:spatialSmooths}b and c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.
+\newpage
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/spatialSmooths} \caption{Raw data and spatial smoothers from a single field.
-Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.}\label{fig:spatialSmooths}
+\includegraphics[width=1\linewidth]{ModelSummary3a_canola} \caption{Field boundary effect on canola yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}
 \end{figure}
 
+\newpage
+
 \begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/distSmooths} \caption{Distance smoothers from a single field, showing a positive saturating effect of distance on average yield (first row), while the variability smoothers show a general decrease in yield SD (second row).}\label{fig:distSmooths}
+\includegraphics[width=1\linewidth]{ModelSummary3a_wheat} \caption{Field boundary effect on wheat yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}
 \end{figure}
 
-\hypertarget{overall-results}{%
-\subsection{Overall results}\label{overall-results}}
-
-Across all fields, standard field boundaries (grassy borders) were the most common boundary type, followed by shelterbelts and other crops (Figure \ref{fig:boundaryTypes}).
+\newpage
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/boundaryTypes-1} \caption{Percentage of field boundaries across sampled fields.}\label{fig:boundaryTypes}
+\includegraphics[width=1\linewidth]{ModelSummary3a_peas} \caption{Field boundary effect on pea yield, accounting for the effect of spatial variation.
+Upper panel represents mean yield, while the lower panel represents yield variation.
+N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}
 \end{figure}
 
-Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table \ref{tab:smoothSummary}, row 6).
-Distances from boundary types were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
+\newpage
 
@@
 
-\caption{\label{tab:smoothSummary}Mean $\chi^2$, p-values, and proportion of smoother p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all models.}
+\caption{Mean $\chi^2$, p-values, and proportion of smoother p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all models.}
 \centering
-\begin{tabular}[t]{l|r|r|r|r|r|r}
+\begin{tabular}[H]{l|r|r|r|r|r|r}
 \hline
@@
 
-Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure \ref{fig:canolaPlot} top row).
-Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50m and remained roughly constant with distance.
-The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
-Yield variability (Figure \ref{fig:canolaPlot} bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
-Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from Wetland boundaries, indicating a potential stabilizing effect.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_canola} \caption{Field boundary effect on canola yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:canolaPlot}
-\end{figure}
-
-The effect of boundary distance on average wheat yield was much more variable than the patterns seen in canola yield.
-Average yield (Figure \ref{fig:wheatPlot}, top row) roughly increased with distance away from wetland, standard, and grasslands, but both shelterbelts and bare ground displayed an intermediate maximum at around 50m (similar to canola), and decreased slightly with distance from Other Crops.
-Yield variability, on the other hand, tended to decrease with distance from bare, grassland, and wetland (before 100 m, at least), increased slightly with distance from shelterbelt and standard boundaries, and was roughly constant with distance from other crops (Figure \ref{fig:wheatPlot}, bottom row).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_wheat} \caption{Field boundary effect on wheat yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:wheatPlot}
-\end{figure}
-
-Pea crops were less well-represented in our data (N = 39 fields, compared to ), and the effects of boundaries tended to be less consistent.
-Average pea yields (Figure \ref{fig:peaPlot}, top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
-Conversely, yield variability (Figure \ref{fig:peaPlot}, bottom row) decreased with distance from grassland, shelterbelt, standard, and wetland boundaries, while the pattern with distance from other crop and bare was inconsistent.
-Unlike canola and wheat crops, there was no obvious evidence of an intermediate increase in yield with distance from the field boundary.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_peas} \caption{Field boundary effect on pea yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:peaPlot}
-\end{figure}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-At the level of individual fields, both average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
-Yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat.
-This increase tended to be small in magnitude in canola crops, meaning that ecosystem services likely increased or did not vary with distance from field boundaries (Scenario 2, Figure \ref{fig:hypotheses}).
-There was a fairly large (0.75 T/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
-However, there was a similar (albeit weaker) pattern associated with bare ground, so this pattern may be caused by something other than the presence of shelterbelts.
-This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this across a larger geographic domain in order to inform growers about how best to manage the landscapes surrounding their fields.
-
-Why did we find few consistent effects of boundary type on crop yield, aside from shelterbelts?
-Our work examined only the first and last links in the causal chain \emph{Field boundary} \(\rightarrow\) \emph{Insect Abundance} \(\rightarrow\) \emph{Crop Yield}, meaning that are two mediating relationships where the overall relationship could be broken or obscured.
-First, field boundaries may not have any influence on insect abundance if separated from the larger landscape context, or spillover of beneficial organisms may be low.
-Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are low, which may vary yearly, or by location.
-Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) when organisms in crops provide tangible ecosystem services.
-We consider these reasons below:
-
-SNL generally has a positive effect on spillover, but field boundaries only represent a fraction of the SNL that surrounds a given field.
-Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales (Nguyen \emph{et al.} 2021) which could be particularly important when beneficial arthropods are disperse very broadly (Öberg \emph{et al.} 2007).
-There could also be interactions between landscape composition at different scales (Motzke \emph{et al.} 2016; Haslem \emph{et al.} 2021; Robinson \emph{et al.} 2021), but this is not broadly studied.
-Landscape complexity (Fahrig \emph{et al.} 2011) and connectivity are also associated with higher abundance of beneficial arthropods.
-This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat (but see Fahrig \emph{et al.} 2019).
-
-Field boundary traits may play an important role in mediating spillover, but our definitions of boundaries were (necessarily) loose, as we used satellite imagery to manually classify them.
-Therefore, it may be that other aspects of field boundaries, such as size or plant community composition, are more important for spillover of ecosystems services into fields.
-For example, larger (Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020) or more functionally diverse (Bonifacio \emph{et al.} 2011; Boughey \emph{et al.} 2011) hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction matters to predatory beetle communities (Sutter \emph{et al.} 2018).
-These fine-scale boundary characteristics were not measured in our study, but could be used as variables in more biologically-meaningful classifications of field boundaries.
-
-Our first scenario (Scenario 1, Figure \ref{fig:hypotheses}) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
-However, field boundaries may not have a positive effect if a) ecosystem dis-services -- pest insects or weed seeds -- also disperse along with beneficials, or if b) beneficials do not disperse from the boundary environment.
-Pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang \emph{et al.} 2007).
-Gray \& Lewis (2014) showed that remnant forest fragments do not provide significant pest control services, but also do not depress yield.
-However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data \emph{alone}.
-Ecosystem services from field boundaries may be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales Robinson \emph{et al.} (2021).
-
-Pest pressure or pollination deficit may be low enough that ecosystem service provision does not significantly change yield.
-Canola is largely self-pollinating (at least in hybrid canola varieties, Lindström \emph{et al.} 2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop (Woodcock \emph{et al.} 2016).
-Pests are also limited by growers' application of insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty.
-
-Our work is some of the first to use a large precision yield dataset to answer questions about how landscapes influence yield, but this study has only scratched the surface of some of the richest agroecological datasets ever created.
-Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
-Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee 2006), where boundaries contained a higher proportion of shelterbelts (Figure \ref{fig:boundaryTypes}), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
-Between-year effects may be important, especially during dry years (Kort 1988), but this was not considered in our analysis.
-Future work could examine whether boundaries have differing effects during hot or cool years, as shading or cooling effects may increase yield during hot years but have negative (or no) effect during cooler years.
-
-\hypertarget{author-contributions}{%
-\section{Author Contributions}\label{author-contributions}}
-
-SVJR, LHN, and PG conceived of the project.
-SVJR collected data, conducted analysis, and wrote the manuscript.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-We thank Trent Clark, Dean Hubbard, Alvin French, and Kristina Polziehn for providing yield data for us, and providing insight on their yield data.
-We also thank Autumn Barnes and Keith Gabert of the Canola Council of Canada, and Laurel Thompson and JP Pettyjohn of Lakeland College for connecting us with farmers and agronomists.
-Funding for this research was provided by Ducks Unlimited Canada's Institute for Wetland and Waterfowl Research, the Alberta Canola Producers Commission, Manitoba Canola Growers Association, SaskCanola, the Alberta Biodiversity Monitoring Institute, and the Alberta Conservation Association.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-AAFC_data}{}%
-Agriculture and Agri-Food Canada. (2018). \emph{{Annual Crop Inventory}}. Available at: \url{https://open.canada.ca/data/en/dataset/ba2645d5-4458-414d-b196-6303ac06c1c9}. Last accessed.
-
-\leavevmode\hypertarget{ref-albrecht2020}{}%
-Albrecht, M., Kleijn, D., Williams, N.M., Tschumi, M., Blaauw, B.R., Bommarco, R., \emph{et al.} (2020).
-The effectiveness of flower strips and hedgerows on pest control, pollination services and crop yield: A quantitative synthesis.
-\emph{Ecology Letters}.
-
-\leavevmode\hypertarget{ref-arslan2002}{}%
-Arslan, S. \& Colvin, T.S. (2002). An evaluation of the response of yield monitors and combines to varying yields. \emph{Precision Agriculture}, 3, 107--122.
-
-\leavevmode\hypertarget{ref-baker2018}{}%
-Baker, T.P., Moroni, M.T., Mendham, D.S., Smith, R. \& Hunt, M.A. (2018). Impacts of windbreak shelter on crop and livestock production. \emph{Crop and Pasture Science}, 69, 785.
-
-\leavevmode\hypertarget{ref-boetzl2020}{}%
-Boetzl, F.A., Schuele, M., Krauss, J.
-\& Steffan-Dewenter, I.
-(2020).
-Pest control potential of adjacent agri-environment schemes varies with crop type and is shaped by landscape context and within-field position.
-\emph{Journal of Applied Ecology}, 57, 1482--1493.
-
-\leavevmode\hypertarget{ref-bonifacio2011}{}%
-Bonifacio, R.S., Kinross, C.M., Gurr, G.M.
-\& Nicol, H.
-(2011).
-The effect of woody plant diversity and other stand and landscape factors on the diversity and abundance of birds using farm shelterbelts.
-\emph{Pacific Conservation Biology}, 17, 22.
-
-\leavevmode\hypertarget{ref-boughey2011}{}%
-Boughey, K.L., Lake, I.R., Haysom, K.A.
-\& Dolman, P.M.
-(2011).
-Improving the biodiversity benefits of hedgerows: How physical characteristics and the proximity of foraging habitat affect the use of linear features by bats.
-\emph{Biological Conservation}, 144, 1790--1798.
-
-\leavevmode\hypertarget{ref-brandle2004}{}%
-Brandle, J.R., Hodges, L.
-\& Zhou, X.H.
-(2004).
-Windbreaks in {North American} agricultural systems.
-In: \emph{New vistas in agroforestry: A compendium for 1st world congress of agroforestry, 2004} (eds.
-Nair, P.K.R., Rao, M.R.
-\& Buck, L.E.).
-Springer Netherlands, Dordrecht, pp.
-65--78.
-
-\leavevmode\hypertarget{ref-case2019}{}%
-Case, B.S., Pannell, J.L., Stanley, M.C., Norton, D.A., Brugman, A., Funaki, M., \emph{et al.} (2019).
-Non-production vegetation has a positive effect on ecological processes in agroecosystems.
-\emph{bioRxiv}, 624635.
-
-\leavevmode\hypertarget{ref-cressie2011}{}%
-Cressie, N. \& Wikle, C.K. (2011). \emph{Statistics for spatio-temporal data}. Wiley.
-
-\leavevmode\hypertarget{ref-duelli2003}{}%
-Duelli, P. \& Obrist, M.K. (2003). Regional biodiversity in an agricultural landscape: The contribution of seminatural habitat islands. \emph{Basic and Applied Ecology}, 4, 129--138.
-
-\leavevmode\hypertarget{ref-fahrig2019}{}%
-Fahrig, L., Arroyo-Rodríguez, V., Bennett, J.R., Boucher-Lalonde, V., Cazetta, E., Currie, D.J., \emph{et al.} (2019).
-Is habitat fragmentation bad for biodiversity?
-\emph{Biological Conservation}, 230, 179--186.
-
-\leavevmode\hypertarget{ref-fahrig2011}{}%
-Fahrig, L., Baudry, J., Brotons, L., Burel, F.G., Crist, T.O., Fuller, R.J., \emph{et al.} (2011).
-Functional landscape heterogeneity and animal biodiversity in agricultural landscapes.
-\emph{Ecology Letters}, 14, 101--112.
-
-\leavevmode\hypertarget{ref-foley2011}{}%
-Foley, J.A., Ramankutty, N., Brauman, K.A., Cassidy, E.S., Gerber, J.S., Johnston, M., \emph{et al.} (2011). Solutions for a cultivated planet. \emph{Nature}, 478, 337--342.
-
-\leavevmode\hypertarget{ref-frei2018}{}%
-Frei, B., Renard, D., Mitchell, M.G.E., Seufert, V., Chaplin-Kramer, R., Rhemtulla, J.M., \emph{et al.} (2018).
-Bright spots in agricultural landscapes: Identifying areas exceeding expectations for multifunctionality and biodiversity.
-\emph{Journal of Applied Ecology}, 55, 2731--2743.
-
-\leavevmode\hypertarget{ref-gardner2021}{}%
-Gardner, E., Breeze, T.D., Clough, Y., Smith, H.G., Baldock, K.C.R., Campbell, A., \emph{et al.} (2021).
-Field boundary features can stabilise bee populations and the pollination of mass-flowering crops in rotational systems.
-\emph{Journal of Applied Ecology}, In press.
-
-\leavevmode\hypertarget{ref-garibaldi2011}{}%
-Garibaldi, L.A., Steffan-Dewenter, I., Kremen, C., Morales, J.M., Bommarco, R., Cunningham, S.A., \emph{et al.} (2011).
-Stability of pollination services decreases with isolation from natural areas despite honey bee visits.
-\emph{Ecology Letters}, 14, 1062--1072.
-
-\leavevmode\hypertarget{ref-gray2014}{}%
-Gray, C.L. \& Lewis, O.T. (2014). Do riparian forest fragments provide ecosystem services or disservices in surrounding oil palm plantations? \emph{Basic and Applied Ecology}, 15, 693--700.
-
-\leavevmode\hypertarget{ref-griffin2007}{}%
-Griffin, T.W., Brown, J.P. \& Lowenberg-DeBoerc, J. (2007). \emph{Yield monitor data analysis protocol: A primer in the management and analysis of precision agriculture data}.
-
-\leavevmode\hypertarget{ref-haslem2021}{}%
-Haslem, A., Clarke, R.H., Holland, G.J., Radford, J.Q., Stewart, A.
-\& Bennett, A.F.
-(2021).
-Local management or wider context: What determines the value of farm revegetation plantings for birds?
-\emph{Journal of Applied Ecology}.
-
-\leavevmode\hypertarget{ref-hunicken2021}{}%
-Hünicken, P.L., Morales, C.L., Aizen, M.A., Anderson, G.K.S., García, N.
-\& Garibaldi, L.A.
-(2021).
-Insect pollination enhances yield stability in two pollinator-dependent crops.
-\emph{Agriculture, Ecosystems {\&} Environment}, 320, 107573.
-
-\leavevmode\hypertarget{ref-kassambara2020}{}%
-Kassambara, A. (2020). \emph{{ggpubr}: 'ggplot2' based publication ready plots}.
-
-\leavevmode\hypertarget{ref-kort1988}{}%
-Kort, J. (1988). 9. Benefits of windbreaks to field and forage crops. \emph{Agriculture, Ecosystems {\&} Environment}, 22-23, 165--190.
-
-\leavevmode\hypertarget{ref-kowalchuk1995}{}%
-Kowalchuk, T.E. \& Jong, E. de. (1995). Shelterbelts and their effect on crop yield. \emph{Canadian Journal of Soil Science}, 75, 543--550.
-
-\leavevmode\hypertarget{ref-lecq2018}{}%
-Lecq, S., Loisel, A., Mullin, S.J.
-\& Bonnet, X.
-(2018).
-Manipulating hedgerow quality: Embankment size influences animal biodiversity in a peri-urban context.
-\emph{Urban Forestry {\&} Urban Greening}, 35, 1--7.
-
-\leavevmode\hypertarget{ref-lindstrom2016}{}%
-Lindström, S.A.M., Herbertsson, L., Rundlöf, M., Smith, H.G.
-\& Bommarco, R.
-(2016).
-Large-scale pollination experiment demonstrates the importance of insect pollination in winter oilseed rape.
-\emph{Oecologia}, 180, 759--769.
-
-\leavevmode\hypertarget{ref-lowe2021}{}%
-Lowe, E.B., Groves, R.
-\& Gratton, C.
-(2021).
-Impacts of field-edge flower plantings on pollinator conservation and ecosystem service delivery {{}} a meta-analysis.
-\emph{Agriculture, Ecosystems {\&} Environment}, 310, 107290.
-
-\leavevmode\hypertarget{ref-marja2019}{}%
-Marja, R., Kleijn, D., Tscharntke, T., Klein, A.-M., Frank, T.
-\& Batáry, P.
-(2019).
-Effectiveness of agri-environmental management on pollinators is moderated more by ecological contrast than by landscape structure or land-use intensity.
-\emph{Ecology Letters}, In press.
-
-\leavevmode\hypertarget{ref-mitchell2014}{}%
-Mitchell, M.G.E., Bennett, E.M. \& Gonzalez, A. (2014). Forest fragments modulate the provision of multiple ecosystem services. \emph{Journal of Applied Ecology}, 51, 909--918.
-
-\leavevmode\hypertarget{ref-motzke2016}{}%
-Motzke, I., Klein, A.-M., Saleh, S., Wanger, T.C.
-\& Tscharntke, T.
-(2016).
-Habitat management on multiple spatial scales can enhance bee pollination and crop yield in tropical homegardens.
-\emph{Agriculture, Ecosystems \& Environment}, 223, 144--151.
-
-\leavevmode\hypertarget{ref-ABRegions2006}{}%
-Natural Regions Committee. (2006). \emph{Natural regions and subregions of {Alberta}}. Government of Alberta, Edmonton.
-
-\leavevmode\hypertarget{ref-nguyen2021}{}%
-Nguyen, L.H., Robinson, S.V.J.
-\& Galpern, P.
-(2021).
-Medium-resolution multispectral satellite imagery in precision agriculture: Mapping precision canola (\emph{brassica napus} {L.}) Yield using {Sentinel-2} time series.
-\emph{Precision Agriculture}, In Review (Preprint: https://www.essoar.org/doi/10.1002/essoar.10507479.2).
-
-\leavevmode\hypertarget{ref-oberg2007}{}%
-Öberg, S., Ekbom, B.
-\& Bommarco, R.
-(2007).
-Influence of habitat type and surrounding landscape on spider diversity in {Swedish} agroecosystems.
-\emph{Agriculture, Ecosystems {\&} Environment}, 122, 211--219.
-
-\leavevmode\hypertarget{ref-prietoBenitez2011}{}%
-Prieto-Benítez, S. \& Méndez, M. (2011). Effects of land management on the abundance and richness of spiders ({Araneae}): A meta-analysis. \emph{Biological Conservation}, 144, 683--691.
-
-\leavevmode\hypertarget{ref-quinn2017}{}%
-Quinn, N.F., Brainard, D.C. \& Szendrei, Z. (2017). Floral strips attract beneficial insects but do not enhance yield in cucumber fields. \emph{Journal of Economic Entomology}, 110, 517--524.
-
-\leavevmode\hypertarget{ref-ramankutty2018}{}%
-Ramankutty, N., Mehrabi, Z., Waha, K., Jarvis, L., Kremen, C., Herrero, M., \emph{et al.} (2018).
-Trends in global agricultural land use: Implications for environmental health and food security.
-\emph{Annual Review of Plant Biology}, 69, 789--815.
-
-\leavevmode\hypertarget{ref-redhead2020}{}%
-Redhead, J.W., Oliver, T.H., Woodcock, B.A.
-\& Pywell, R.F.
-(2020).
-The influence of landscape composition and configuration on crop yield resilience.
-\emph{Journal of Applied Ecology}, 57, 2180--2190.
-
-\leavevmode\hypertarget{ref-robinson2021}{}%
-Robinson, S.V.J., Edwards, D., Vickruck, J.L., Best, L.R.
-\& Galpern, P.
-(2021).
-Non-crop sources of beneficial arthropods vary within-season across a prairie agroecosystem.
-\emph{Agriculture, Ecosystems {\&} Environment}, 320, 107581.
-
-\leavevmode\hypertarget{ref-stasiov2020}{}%
-Stašiov, S., Diviaková, A., Svitok, M., Novikmec, M.
-\& Dovciak, M.
-(2020).
-Hedgerows support rich communities of harvestmen ({Opiliones}) in upland agricultural landscape.
-\emph{Basic and Applied Ecology}, 47, 73--82.
-
-\leavevmode\hypertarget{ref-statscan_canola2014}{}%
-Statistics Canada. (2014). Table 001-0010 - estimated areas, yield, production and average farm price of principal field crops, in metric units, annual.
-
-\leavevmode\hypertarget{ref-sutter2018b}{}%
-Sutter, L., Amato, M., Jeanneret, P.
-\& Albrecht, M.
-(2018).
-Overwintering of pollen beetles and their predators in oilseed rape and semi-natural habitats.
-\emph{Agriculture, Ecosystems {\&} Environment}, 265, 275--281.
-
-\leavevmode\hypertarget{ref-vanVooren2017}{}%
-Van Vooren, L., Bert, R., Steven, B., Pieter, D.F., Victoria, N., Paul, P., \emph{et al.} (2017).
-Ecosystem service delivery of agri-environment measures: A synthesis for hedgerows and grass strips on arable land.
-\emph{Agriculture, Ecosystems {\&} Environment}, 244, 32--51.
-
-\leavevmode\hypertarget{ref-vega2019}{}%
-Vega, A., Córdoba, M., Castro-Franco, M. \& Balzarini, M. (2019). Protocol for automating error removal from yield maps. \emph{Precision Agriculture}, 20, 1030--1044.
-
-\leavevmode\hypertarget{ref-weninger2021}{}%
-Weninger, T., Scheper, S., Lackóová, L., Kitzler, B., Gartner, K., King, N.W., \emph{et al.} (2021).
-Ecosystem services of tree windbreaks in rural landscapes -- a systematic review.
-\emph{Environmental Research Letters}, In press.
-
-\leavevmode\hypertarget{ref-whelan2013}{}%
-Whelan, B. \& Taylor, J. (2013). \emph{Precision agriculture for grain production systems}. CSIRO Publishing.
-
-\leavevmode\hypertarget{ref-wickham2016}{}%
-Wickham, H. (2016). \emph{{ggplot2}: Elegant graphics for data analysis}. Springer-Verlag New York.
-
-\leavevmode\hypertarget{ref-wickle2019}{}%
-Wikle, C.K., Zammit-Mangion, A. \& Cressie, N. (2019). \emph{Spatio-temporal statistics with r}. Chapman; Hall: CRC Press.
-
-\leavevmode\hypertarget{ref-wood2017}{}%
-Wood, S.N. (2017). \emph{Generalized additive models: An introduction with {R}}. CRC press.
-
-\leavevmode\hypertarget{ref-woodcock2016}{}%
-Woodcock, B.A., Bullock, J.M., McCracken, M., Chapman, R.E., Ball, S.L., Edwards, M.E., \emph{et al.} (2016).
-Spill-over of pest control and pollination services into arable crops.
-\emph{Agriculture, Ecosystems \& Environment}, 231, 15--23.
-
-\leavevmode\hypertarget{ref-wratten2012}{}%
-Wratten, S.D., Gillespie, M., Decourtye, A., Mader, E.
-\& Desneux, N.
-(2012).
-Pollinator habitat enhancement: Benefits to other ecosystem services.
-\emph{Agriculture, Ecosystems {\&} Environment}, 159, 112--122.
-
-\leavevmode\hypertarget{ref-zamorano2020}{}%
-Zamorano, J., Bartomeus, I., Grez, A.A.
-\& Garibaldi, L.A.
-(2020).
-Field margin floral enhancements increase pollinator diversity at the field edge but show no consistent spillover into the crop field: A meta-analysis.
-\emph{Insect Conservation and Diversity}, 13, 519--531.
-
-\leavevmode\hypertarget{ref-zhang2007}{}%
-Zhang, W., Ricketts, T.H., Kremen, C., Carney, K. \& Swinton, S.M. (2007). Ecosystem services and dis-services to agriculture. \emph{Ecological Economics}, 64, 253--260.
-
-\end{CSLReferences}
-
 \end{document}

### 2022-03-04 Manuscript/manuscript_supplemental2.tex
Updated manuscript

--- 2021-10-27 Manuscript/manuscript.tex
@@
 \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
-\usepackage{longtable}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \ifxetex
@@
 % Pandoc header
-\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}} \usepackage{setspace} %\linenumbers
+\makeatletter \def\ps@pprintTitle{  \let\@oddhead\@empty  \let\@evenhead\@empty  \def\@oddfoot{\centerline{\thepage}} \let\@evenfoot\@oddfoot} \makeatother \usepackage{float} \floatplacement{figure}{H} \newcommand{\beginsupplement}{\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
 \usepackage{booktabs}
@@
 \begin{document}
-\begin{frontmatter}
 
-  \title{Livin' on the edge: Precision yield data shows evidence of ecosystem services from field boundaries}
-    \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
-   \ead{samuel.robinson@ucalgary.ca} 
-    \author[University of Calgary]{Lan H. Nguyen}
-   \ead{hoanglan.nguyen@ucalgary.ca} 
-    \author[University of Calgary]{Paul Galpern}
-   \ead{paul.galpern@ucalgary.ca} 
-      \address[University of Calgary]{2500 University Drive NW, Calgary, AB}
-      \cortext[1]{Corresponding Author}
+\hypertarget{appendix-a-supplemental-information}{%
+\section*{Appendix A: Supplemental Information}\label{appendix-a-supplemental-information}}
+\addcontentsline{toc}{section}{Appendix A: Supplemental Information}
 
-  \begin{abstract}
-  Field boundaries can provide ecosystem services to crops by creating better abiotic conditions for crop growth, and can also act as habitat for beneficial arthropods.
-This suggests that crop boundaries may create an intermediate hump-shaped increase in crop yield, where negative edge effects are cancelled out by increased ecosystem services from the field boundary.
-However, there is little large-scale evidence showing this, largely because plot-scale crop yields are costly and time-consuming to measure.
-Precision yield data from combine yield monitors has huge potential in this respect, as the equipment is widespread and data is frequently recorded by growers.
-In this study, we used 252 field-years of yield monitor data from three common crops -- wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}) -- recorded across Alberta, Canada, and examined how yield varied with distances from common crop boundary types.
-Average crop yield tended to increase with distance from crop boundaries before plateauing at about 50 m, and yield variation (SD) tended to decrease with distance.
-There was evidence of an intermediate increase in yield for wheat away from shelterbelts, and a weak increase in canola, but this was not seen for other crop types or boundary types.
-This study represents one of the first uses of precision yield data to measure ecosystem service provision at large spatial scales.
-  \end{abstract}
-   \begin{keyword} precision yield; spillover; landscape; shelterbelt;\end{keyword}
- \end{frontmatter}
+\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{Field Locations} \caption{Locations of fields that data were collected from.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{boundaryTypes-1} \caption{Boxplot of field boundary compositions across sampled fields.
+The x-axis indicates the boundary type, while the y-axis indicates the percentage of all field boundaries covered by that boundary type.}
+\end{figure}
 
@@
 
-\onehalfspacing
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth (Ramankutty \emph{et al.} 2018).
-This has allowed increases in human population and increased global stability in production (Foley \emph{et al.} 2011).
-However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms.
-Maintaining both biodiversity and production in agroecosystems represents a goal of conservationists and agronomists, and holds the potential for win-win scenarios ({``multifunctionality''} \emph{sensu} Mitchell \emph{et al.} 2014; Frei \emph{et al.} 2018).
-Key to this goal is the preservation of semi-natural land (SNL) in and around crops, which represents the interface between crops and non-crops within agroecosystems.
-
-SNL in agroecosystems is important for both agricultural production and conservation (Prieto-Benítez \& Méndez 2011), and can create better biotic and abiotic conditions for crop growth (Case \emph{et al.} 2019; Weninger \emph{et al.} 2021; but see Lowe \emph{et al.} 2021).
-SNL acts as habitat for mobile organisms, and can act as a source of ecosystem services, such as pollination or pest control (Albrecht \emph{et al.} 2020; Gardner \emph{et al.} 2021), but this depends on both the crop system (Gray \& Lewis 2014; Quinn \emph{et al.} 2017) and landscape context (Marja \emph{et al.} 2019; Boetzl \emph{et al.} 2020).
-SNL can also create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed and soil erosion (Kort 1988; Brandle \emph{et al.} 2004; Wratten \emph{et al.} 2012).
-Field boundaries (e.g.~grass and flower strips, tree or shrub shelterbelts, road right-of-ways) are SNL of particular interest, as they are extremely common and require less land than larger agricultural set-asides.
-
-Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
-Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure \ref{fig:hypotheses}, Scenario 2).
-Alternatively, yield may be higher near field boundaries (Scenario 3) if boundaries provide helpful ecosystem services (Duelli \& Obrist 2003).
-For example, crop pollination services from bees that nest in field boundaries drops rapidly with distance (Garibaldi \emph{et al.} 2011) and wind speed is reduced directly downwind of treelines (de Jong \& Kowalchuk 1995), so positive effects will be concentrated close to the boundary.
-Depending on the rate at which positive and negative boundary effects decline with distance, another scenario is also possible.
-Yield may be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services (Kowalchuk \& de Jong 1995; Helmers \& Brandle 2005; Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/hypotheses} \caption{Potential yield patterns, depending on ecosystem service effects with distance.
-Scenario 1: no effect of boundary.
-Scenario 2: negative edge effect on yield.
-Scenario 3: positive ecosystem service effect on yield.
-Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.}\label{fig:hypotheses}
-\end{figure}
-
-Unfortunately, data to inform these decisions are often minimal, as observing boundary effects requires large amounts of expensive and time-consuming crop yield data.
-This means that studies of SNL effects on crop yield tend to suffer from limited spatial scope and small sample sizes, reducing statistical power and generality (Brandle \emph{et al.} 1984; Baker \emph{et al.} 2018; but see also Redhead \emph{et al.} 2020).
-Meta-analyses are a common solution to this problem (e.g.
-Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs (Steel 2017; Osorio \emph{et al.} 2018).
-However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
-Ecosystem services from field boundaries may influence both the mean as well as the variability of yield in agroecosystems (Redhead \emph{et al.} 2020; Hünicken \emph{et al.} 2021); typically only means are considered, but lower variability in yield can also be valuable from a grower's perspective.
-There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield.
-
-In this paper we use a precision yield dataset from Alberta, Canada to ask the following questions:
-1. How does crop yield change with distance from the edge of field?
-2. Does this depend on type of field edge?
-3. Is there an intermediate distance where yield is maximized or variance is minimized?
-To our knowledge, only (Osorio \emph{et al.} 2018) has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{data-collection}{%
-\subsection{Data collection}\label{data-collection}}
-
-Precision yield data were collected directly from farmers across Alberta.
-Farmers were solicited for yield data through local agronomists, and we received 298 field-years of data from 5 growers across a total of 7 years (2014-2020).
-We converted raw data to a standard csv format using Ag Leader SMS®.
-85\% of the crop types where either wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}), three of the most common crops in rotation in Alberta (Statistics Canada 2014).
-The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
-Individual fields contained between 1 and 6 years of data (mean: 3.1), containing a total of 14.4 million data points.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/Field Locations} \caption{Locations of fields that data were collected from}\label{fig:fieldLocataions}
-\end{figure}
-
-Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
-We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m\(^2\)), dry yield (tonnes), and the spatial location.
-Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
-Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
-Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows (Griffin \emph{et al.} 2007; Whelan \& Taylor 2013), or when the combine is accelerating or decelerating (Arslan \& Colvin 2002).
-Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix A for R code):
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Removed outliers and ``inliers'' using the spatial filtering procedure from (Vega \emph{et al.} 2019)
-\item
-  Removed unrealistic yield values (10.75 T/ha for wheat, 8 T/ha for canola and peas), and values of zero or less
-\item
-  Removed extremely low and high outliers (below 1st percentile, above 99th percentile) for: dry yield, steering angle differences, ground speed
-\item
-  Removed points with large changes (\textgreater20\%) in ground speed
-\item
-  Removed extremely low and high changes in position, indicating that the combine had stopped or moved to another part of the field between data collection
-\item
-  Sub-sampled filtered data to 50,000 data points per field (reduce computation time)
-\end{enumerate}
-
-Field boundaries were digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data (Agriculture and Agri-Food Canada 2018).
-Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions.
-For example, ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
-This makes accurate and consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
-\item
-  Wetland: permanent wetland; borders are largely unchanged from year-to-year
-\item
-  Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
-\item
-  Other crop: annual crop or pasture with little or no visible boundary between planted areas
-\item
-  Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
-\item
-  Grassland: permanent seminatural grassland or pasture (not in rotation)
-\end{enumerate}
-
-\hypertarget{analysis}{%
-\subsection{Analysis}\label{analysis}}
-
-Yield data from each field contains both random and systematic errors that must be modeled in order to reveal underlying changes in yield.
-Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, previous agricultural practices, and other factors not directly related to boundary distance.
-To account for this, we used additive models of the effect of boundary distance on crop yield while accounting for within-field spatial variation.
-
-We fit the following model to each field-year of data:
-
-\begin{equation}
-  \begin{split}
-  \sqrt{Yield} \sim & Normal (\mu, \sigma)\\
-  \mu = & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
-   & f(\text{Easting}, \text{Northing}, b=80)\\
-  log(\sigma) =  & Intercept + f(\text{Boundary Distance}, b=5)_i + \\
-   & f(\text{Easting}, \text{Northing}, b=80)\\
-  \end{split}
-  \end{equation}
-
-\begin{itemize}
-\tightlist
-\item
-  where:
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Yield = weight of crop gathered per area of harvest (T/ha)
-\item
-  Boundary Distance = distance from field boundary type \emph{i} (m)
-\item
-  Easting, Northing = distance from centre of field (m)
-\item
-  f(x,b) = penalized thin-plate regression spline, where \emph{x} is the predictor and \emph{b} is the number of basis dimensions
-\end{enumerate}
-
-In addition to modeling edge effects (our variable of interest), this model also accounts for within-field spatial variation not related to edges.
-Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
-Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
-We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \ref{fig:hypotheses}).
-All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).
-
-The residuals in most of the models displayed both spatial autocorrelation in the residuals, indicating ``micro-scale'' variation that was not captured by the spatial smoothers (Cressie \& Wikle 2011; Wikle \emph{et al.} 2019).
-This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.~extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 80), we decided to leave the models as-is.
-Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
-That is, combining of fields occurs in patterns (e.g.~east-to-west) so we cannot tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
-We considered the latter to be a more realistic assumption, so we assigned all random variation as spatial in nature.
-Non-standard combine driving patterns may be able to gather data in a way that can detect sensor drift, but this is beyond the scope of this research.
-
-To consider results from all field-level models together, we fit models of each field-year independently and then fit overall smoothers as averages of the field-level smoothers.
-This does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard errors.
-\item
-  Use posterior sample to simulate a new smoother from each field.
-\item
-  Fit new model of simulated smoothers from all fields (with a field-year random intercept), and save this ``meta-smoother.''
-\item
-  Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers.
-  This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
-  Spatial variability was not averaged between fields because within-field variability is not directly comparable between fields.
-\end{enumerate}
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{field-level-example}{%
-\subsection{Field-level example}\label{field-level-example}}
-
-The filtered data from each field contained a large amount of variation, but the non-stationary additive models were able to reveal underlying yield patterns.
-To illustrate this, we show the results from a single wheat field, where we show the relative impact of boundary distance and spatial variation on both yield mean and yield SD (``patchiness'').
-Figure \ref{fig:spatialSmooths}a shows the filtered data collected from a single field, demonstrating a large amount of variation even after filtering.
-Yield mean tended to increase with distance from field boundaries while yield SD tended to decrease, but there were no consistent differences between boundary types (Figure \ref{fig:distSmooths}).
-Finally, the random spatial smoothers in Figure \ref{fig:spatialSmooths}b and c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/spatialSmooths} \caption{Raw data and spatial smoothers from a single field.
-Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.
-Field dimensions are approximately 800 x 800 m, but coordinates are hidden to protect data privacy.}\label{fig:spatialSmooths}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/distSmooths} \caption{Distance smoothers from a single field, showing a positive saturating effect of distance on average yield (first row), while the variability smoothers show a general decrease in yield SD (second row).}\label{fig:distSmooths}
-\end{figure}
-
-\hypertarget{overall-results}{%
-\subsection{Overall results}\label{overall-results}}
-
-Across all fields, standard field boundaries (grassy borders) were the most common boundary type, followed by shelterbelts and other crops (Figure \ref{fig:boundaryTypes}).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/boundaryTypes-1} \caption{Boxplot of field boundary compositions across sampled fields.
-The x-axis indicates the boundary type, while the y-axis indicates the percentage of all field boundaries covered by that boundary type.}\label{fig:boundaryTypes}
-\end{figure}
-
-Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table \ref{tab:smoothSummary}, row 6).
-Distances from boundaries were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
-
-\begin{table}
-
-\caption{\label{tab:smoothSummary}Mean $\chi^2$, p-values, and proportion of smoother p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all models.}
-\centering
-\begin{tabular}[t]{l|r|r|r|r|r|r}
-\hline
-\multicolumn{1}{c|}{} & \multicolumn{3}{c|}{Mean Yield} & \multicolumn{3}{c}{Yield Variability} \\
-\cline{2-4} \cline{5-7}
-Smoother Type & $\chi^2$-value & p-value & Prop. <0.05 & $\chi^2$-value & p-value & Prop. <0.05\\
-\hline
-Bare & 154.7 & 0.036 & 0.877 & 121.4 & 0.054 & 0.860\\
-\hline
-Grassland & 303.1 & 0.079 & 0.794 & 93.2 & 0.061 & 0.794\\
-\hline
-Other Crop & 174.9 & 0.063 & 0.860 & 104.5 & 0.044 & 0.840\\
-\hline
-Shelterbelt & 328.1 & 0.014 & 0.936 & 137.6 & 0.058 & 0.817\\
-\hline
-Standard & 281.9 & 0.021 & 0.938 & 159.6 & 0.057 & 0.832\\
-\hline
-Wetland & 148.9 & 0.045 & 0.833 & 106.7 & 0.013 & 0.857\\
-\hline
-Spatial Smoother & 17357.5 & 0.000 & 1.000 & 4739.1 & 0.000 & 1.000\\
-\hline
-\end{tabular}
-\end{table}
-
-Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure \ref{fig:canolaPlot} top row).
-Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50 m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50 m and remained roughly constant with distance.
-The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
-Yield variability (Figure \ref{fig:canolaPlot} bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
-Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from wetland boundaries, indicating a potential stabilizing effect, but this was offset by the large increase in yield variability at the edge of wetlands.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_canola} \caption{Field boundary effect on canola yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:canolaPlot}
-\end{figure}
-
-The effect of boundary distance on average wheat yield was much more variable than the patterns seen in canola yield.
-Average yield (Figure \ref{fig:wheatPlot}, top row) roughly increased with distance away from wetland, standard, and grasslands, but both shelterbelts and bare ground displayed an intermediate maximum at around 50m (similar to canola), and decreased slightly with distance from Other Crops.
-Yield variability, on the other hand, tended to decrease with distance from bare, grassland, and wetland (before 100 m, at least), increased slightly with distance from shelterbelt and standard boundaries, and was roughly constant with distance from other crops (Figure \ref{fig:wheatPlot}, bottom row).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_wheat} \caption{Field boundary effect on wheat yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:wheatPlot}
-\end{figure}
-
-Pea crops were less well-represented in our data (N = 39 fields), and therefore the effects of boundaries tended to be less consistent.
-Average pea yields (Figure \ref{fig:peaPlot}, top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
-Conversely, yield variability (Figure \ref{fig:peaPlot}, bottom row) decreased with distance from grassland, shelterbelt, standard, and wetland boundaries, while the pattern with distance from other crop and bare was inconsistent.
-Unlike canola and wheat crops, there was no obvious evidence of an intermediate increase in yield with distance from the field boundary.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_peas} \caption{Field boundary effect on pea yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:peaPlot}
-\end{figure}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-Within individual fields, both the average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
-Mean yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat, while peas showed no evidence of this.
-However, this intermediate increase was small in canola crops, meaning that ecosystem services likely did not vary strongly with distance from field boundaries (Scenario 2, Figure \ref{fig:hypotheses}).
-There was a fairly large (0.75 T/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
-However, there was a similar (albeit weaker) pattern associated with bare ground, so this pattern may be caused by something other than the presence of shelterbelts.
-This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this within relevant geographic sub-regions in order to locally inform growers about how best to manage the landscapes surrounding their fields.
-
-Our work showed an intermediate increase in wheat yield with distance from shelterbelts and bare ground, but what is the cause of these patterns?
-We hypothesize that the potential yield-increasing mechanisms may be a) microclimate effects, b) soil effects, or c) beneficial animal spillover.
-Shelterbelts can act to capture snow and moderate downwind microclimate, reducing water stress on crop plants and increasing yield (Staple \& Lehane 1955; Kowalchuk \& de Jong 1995; Bao \emph{et al.} 2012).
-However, Kowalchuk \& de Jong (1995) found that positive microclimate effects of shelterbelts on wheat yield occur at around 20 m (3-4 times the shelterbelt height), while our work shows a peak at around 50 m.
-This additionally complicated by the fact that water stress varies from year-to-year, and that the impact of the shelterbelt also depends on the orientation to prevailing winds and the physical structure of the shelterbelt (Pelton 1967; Kort 1988; Singh \emph{et al.} 1999).
-Unexpectedly, we also found this intermediate yield maximum with distance away from bare field boundaries, indicating that the patterns we found may be at least somewhat related to water stress.
-While we consider microclimate to be the most likely cause of the intermediate increase, it may also be related to soil effects from shelterbelts, where erosion is reduced and leaf litter from the shelterbelt can cause an increase in soil deposition (de Jong \& Kowalchuk 1995).
-Finally, shelterbelts may also act as sources of predators or parasitoids of wheat crop pests (Dong \emph{et al.} 2015; Ward \emph{et al.} 2021) but their effect depends on pest pressures, which are typically suppressed by pesticide application.
-We recommend that follow-up studies a) use precision yield data to confirm that this pattern exists across broader geographical areas and b) examine the influence of the three potential mechanisms considered above (as well as their interaction).
-
-Wheat and shelterbelts aside, why did we find few consistent effects of boundary type on crop yield?
-Our work examined only the first and last links in the causal chain \emph{Field Boundary} \(\rightarrow\) \emph{Ecosystem Service} \(\rightarrow\) \emph{Crop Yield}, meaning that are two mediating relationships where the overall relationship could be broken or obscured.
-First, field boundaries may not have any influence on beneficial insect abundance if separated from the larger landscape context, or spillover of beneficial organisms may be low.
-Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are low, which may vary among year or location.
-Abiotic ecosystem services, such as reduced wind speed or temperature, also depend on the structure of the boundary and whether wind or temperature are limiting crop production.
-Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) under what conditions field boundaries provide tangible ecosystem services.
-We consider these reasons below:
-
-SNL generally has a positive effect on beneficial insect abundance, but field boundaries only represent a fraction of the SNL that surrounds a given field.
-Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales (Nguyen \emph{et al.} 2021) which could be particularly important when beneficial arthropods are disperse very broadly (Öberg \emph{et al.} 2007).
-There could also be interactions between landscape composition at different scales (Motzke \emph{et al.} 2016; Haslem \emph{et al.} 2021; Robinson \emph{et al.} 2021), but this is not broadly studied.
-Landscape complexity (Fahrig \emph{et al.} 2011) and connectivity are also associated with higher abundance of beneficial arthropods.
-This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat (but see Fahrig \emph{et al.} 2019).
-
-Field boundary characteristics, such as height or species composition, may play an important role in mediating spillover, but our definitions of boundaries were necessarily loose, as we used satellite imagery to manually classify them.
-Therefore, it may be that other aspects of field boundaries are more important for spillover of ecosystems services into fields.
-For example, larger or more functionally diverse (Bonifacio \emph{et al.} 2011; Boughey \emph{et al.} 2011; Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020) hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction influences predatory beetle communities (Sutter \emph{et al.} 2018).
-These fine-scale boundary characteristics were not measured in our study, but could be used as variables in more biologically-meaningful classifications of field boundaries.
-Finally, the relevant spatial scale of boundary effects are often related to the height or windward orientation of the shelterbelt (Kort 1988; Kowalchuk \& de Jong 1995), and precision yield data has the potential to show this in far greater detail than previous studies.
-
-Our first scenario (Scenario 1, Figure \ref{fig:hypotheses}) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
-However, field boundaries may not have a positive effect if a) ecosystem dis-services -- pest insects or weed seeds -- also disperse along with beneficials, or if b) beneficials do not disperse from the boundary environment.
-Pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang \emph{et al.} 2007).
-However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data \emph{alone}.
-Ecosystem services from field boundaries may also be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales Robinson \emph{et al.} (2021).
-Lastly, pest pressure or pollination requirements may be low enough that ecosystem service provision does not significantly change yield.
-Canola is largely self-pollinating (at least in hybrid canola varieties, Lindström \emph{et al.} 2016) and wheat is wind-pollinated, meaning that slight increases in pollination services may not change yields significantly in either crop (Woodcock \emph{et al.} 2016).
-Crop pests in most western Canadian fields are suppressed by insecticides (seed-treated or on the crop directly), and insect pest outbreaks tend to be temporally and spatially spotty, meaning spillover of beneficial predators and parasitoids may only matter during outbreak years.
-
-Our work is some of the first to use a large precision yield dataset to answer questions about how landscapes influence yield.
-Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
-Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee 2006), where boundaries contained a higher proportion of shelterbelts (Figure \ref{fig:boundaryTypes}), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
-Between-year effects may be important, especially during hot dry years (Kort 1988), but this was not considered in our analysis.
-Future work could examine whether boundaries have differing effects during hot or cool years, as shading or cooling effects may increase yield during hot years but have negative (or no) effect during cooler years.
-Finally, about 50\% of growers in western Canada have collected precision agriculture data, and about 85\% have yield monitoring capability (Steel 2017), which means that our work has only scratched the surface of what could be (collectively) one of the richest agroecological datasets ever created.
-We urge agronomists, growers, and ecologists to adopt more wide-spread use of precision yield data, as it represents a vast source of incredibly detailed information which can be used to answer a variety of important questions.
-
-\hypertarget{author-contributions}{%
-\section{Author Contributions}\label{author-contributions}}
-
-SVJR, LHN, and PG conceived of the project.
-SVJR collected data, conducted analysis, and wrote the manuscript.
-
-\hypertarget{acknowledgements}{%
-\section{Acknowledgements}\label{acknowledgements}}
-
-We thank Trent Clark, Dean Hubbard, Alvin French, and Kristina Polziehn for providing yield data for us, and providing insight on their yield data.
-We also thank Autumn Barnes and Keith Gabert of the Canola Council of Canada, and Laurel Thompson and JP Pettyjohn of Lakeland College for connecting us with farmers and agronomists.
-Funding for this research was provided by Ducks Unlimited Canada's Institute for Wetland and Waterfowl Research, the Alberta Canola Producers Commission, Manitoba Canola Growers Association, SaskCanola, the Alberta Biodiversity Monitoring Institute, and the Alberta Conservation Association.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-AAFC_data}{}%
-Agriculture and Agri-Food Canada. (2018). \emph{{Annual Crop Inventory}}. Available at: \url{https://open.canada.ca/data/en/dataset/ba2645d5-4458-414d-b196-6303ac06c1c9}. Last accessed.
-
-\leavevmode\hypertarget{ref-albrecht2020}{}%
-Albrecht, M., Kleijn, D., Williams, N.M., Tschumi, M., Blaauw, B.R., Bommarco, R., \emph{et al.} (2020).
-The effectiveness of flower strips and hedgerows on pest control, pollination services and crop yield: A quantitative synthesis.
-\emph{Ecology Letters}.
-
-\leavevmode\hypertarget{ref-arslan2002}{}%
-Arslan, S. \& Colvin, T.S. (2002). An evaluation of the response of yield monitors and combines to varying yields. \emph{Precision Agriculture}, 3, 107--122.
-
-\leavevmode\hypertarget{ref-baker2018}{}%
-Baker, T.P., Moroni, M.T., Mendham, D.S., Smith, R. \& Hunt, M.A. (2018). Impacts of windbreak shelter on crop and livestock production. \emph{Crop and Pasture Science}, 69, 785.
-
-\leavevmode\hypertarget{ref-bao2012}{}%
-Bao, Y., Li, H.
-\& Zhao, H.
-(2012).
-Effect of shelterbelts on winter wheat yields in sanded farmland of north-western shandong province, china.
-\emph{Journal of Food Agriculture \& Environment}, 10, 1399--1403.
-
-\leavevmode\hypertarget{ref-boetzl2020}{}%
-Boetzl, F.A., Schuele, M., Krauss, J.
-\& Steffan-Dewenter, I.
-(2020).
-Pest control potential of adjacent agri-environment schemes varies with crop type and is shaped by landscape context and within-field position.
-\emph{Journal of Applied Ecology}, 57, 1482--1493.
-
-\leavevmode\hypertarget{ref-bonifacio2011}{}%
-Bonifacio, R.S., Kinross, C.M., Gurr, G.M.
-\& Nicol, H.
-(2011).
-The effect of woody plant diversity and other stand and landscape factors on the diversity and abundance of birds using farm shelterbelts.
-\emph{Pacific Conservation Biology}, 17, 22.
-
-\leavevmode\hypertarget{ref-boughey2011}{}%
-Boughey, K.L., Lake, I.R., Haysom, K.A.
-\& Dolman, P.M.
-(2011).
-Improving the biodiversity benefits of hedgerows: How physical characteristics and the proximity of foraging habitat affect the use of linear features by bats.
-\emph{Biological Conservation}, 144, 1790--1798.
-
-\leavevmode\hypertarget{ref-brandle2004}{}%
-Brandle, J.R., Hodges, L.
-\& Zhou, X.H.
-(2004).
-Windbreaks in {North American} agricultural systems.
-In: \emph{New vistas in agroforestry: A compendium for 1st world congress of agroforestry, 2004} (eds.
-Nair, P.K.R., Rao, M.R.
-\& Buck, L.E.).
-Springer Netherlands, Dordrecht, pp.
-65--78.
-
-\leavevmode\hypertarget{ref-brandle1984}{}%
-Brandle, J.R., Johnson, B.B. \& Dearmont, D.D. (1984). Windbreak economics: The case of winter wheat production in eastern {Nebraska}. \emph{Journal of Soil and Water Conservation}, 39, 339--343.
-
-\leavevmode\hypertarget{ref-case2019}{}%
-Case, B.S., Pannell, J.L., Stanley, M.C., Norton, D.A., Brugman, A., Funaki, M., \emph{et al.} (2019).
-Non-production vegetation has a positive effect on ecological processes in agroecosystems.
-\emph{bioRxiv}, 624635.
-
-\leavevmode\hypertarget{ref-cressie2011}{}%
-Cressie, N. \& Wikle, C.K. (2011). \emph{Statistics for spatio-temporal data}. Wiley.
-
-\leavevmode\hypertarget{ref-deJong1995}{}%
-de Jong, E. \& Kowalchuk, T.E. (1995). The effect of shelterbelts on erosion and soil properties. \emph{Soil Science}, 159, 337--345.
-
-\leavevmode\hypertarget{ref-dong2015}{}%
-Dong, Z., Ouyang, F., Lu, F. \& Ge, F. (2015). Shelterbelts in agricultural landscapes enhance ladybeetle abundance in spillover from cropland to adjacent habitats. \emph{BioControl}, 60, 351--361.
-
-\leavevmode\hypertarget{ref-duelli2003}{}%
-Duelli, P. \& Obrist, M.K. (2003). Regional biodiversity in an agricultural landscape: The contribution of seminatural habitat islands. \emph{Basic and Applied Ecology}, 4, 129--138.
-
-\leavevmode\hypertarget{ref-fahrig2019}{}%
-Fahrig, L., Arroyo-Rodríguez, V., Bennett, J.R., Boucher-Lalonde, V., Cazetta, E., Currie, D.J., \emph{et al.} (2019).
-Is habitat fragmentation bad for biodiversity?
-\emph{Biological Conservation}, 230, 179--186.
-
-\leavevmode\hypertarget{ref-fahrig2011}{}%
-Fahrig, L., Baudry, J., Brotons, L., Burel, F.G., Crist, T.O., Fuller, R.J., \emph{et al.} (2011).
-Functional landscape heterogeneity and animal biodiversity in agricultural landscapes.
-\emph{Ecology Letters}, 14, 101--112.
-
-\leavevmode\hypertarget{ref-foley2011}{}%
-Foley, J.A., Ramankutty, N., Brauman, K.A., Cassidy, E.S., Gerber, J.S., Johnston, M., \emph{et al.} (2011). Solutions for a cultivated planet. \emph{Nature}, 478, 337--342.
-
-\leavevmode\hypertarget{ref-frei2018}{}%
-Frei, B., Renard, D., Mitchell, M.G.E., Seufert, V., Chaplin-Kramer, R., Rhemtulla, J.M., \emph{et al.} (2018).
-Bright spots in agricultural landscapes: Identifying areas exceeding expectations for multifunctionality and biodiversity.
-\emph{Journal of Applied Ecology}, 55, 2731--2743.
-
-\leavevmode\hypertarget{ref-gardner2021}{}%
-Gardner, E., Breeze, T.D., Clough, Y., Smith, H.G., Baldock, K.C.R., Campbell, A., \emph{et al.} (2021).
-Field boundary features can stabilise bee populations and the pollination of mass-flowering crops in rotational systems.
-\emph{Journal of Applied Ecology}, In press.
-
-\leavevmode\hypertarget{ref-garibaldi2011}{}%
-Garibaldi, L.A., Steffan-Dewenter, I., Kremen, C., Morales, J.M., Bommarco, R., Cunningham, S.A., \emph{et al.} (2011).
-Stability of pollination services decreases with isolation from natural areas despite honey bee visits.
-\emph{Ecology Letters}, 14, 1062--1072.
-
-\leavevmode\hypertarget{ref-gray2014}{}%
-Gray, C.L. \& Lewis, O.T. (2014). Do riparian forest fragments provide ecosystem services or disservices in surrounding oil palm plantations? \emph{Basic and Applied Ecology}, 15, 693--700.
-
-\leavevmode\hypertarget{ref-griffin2007}{}%
-Griffin, T.W., Brown, J.P. \& Lowenberg-DeBoerc, J. (2007). \emph{Yield monitor data analysis protocol: A primer in the management and analysis of precision agriculture data}.
-
-\leavevmode\hypertarget{ref-haslem2021}{}%
-Haslem, A., Clarke, R.H., Holland, G.J., Radford, J.Q., Stewart, A.
-\& Bennett, A.F.
-(2021).
-Local management or wider context: What determines the value of farm revegetation plantings for birds?
-\emph{Journal of Applied Ecology}.
-
-\leavevmode\hypertarget{ref-helmers2005}{}%
-Helmers, G. \& Brandle, J.R. (2005). Optimum windbreak spacing in great plains agriculture. \emph{Great Plains Research}, 15, 179--198.
-
-\leavevmode\hypertarget{ref-hunicken2021}{}%
-Hünicken, P.L., Morales, C.L., Aizen, M.A., Anderson, G.K.S., García, N.
-\& Garibaldi, L.A.
-(2021).
-Insect pollination enhances yield stability in two pollinator-dependent crops.
-\emph{Agriculture, Ecosystems {\&} Environment}, 320, 107573.
-
-\leavevmode\hypertarget{ref-kassambara2020}{}%
-Kassambara, A. (2020). \emph{{ggpubr}: 'ggplot2' based publication ready plots}.
-
-\leavevmode\hypertarget{ref-kort1988}{}%
-Kort, J. (1988). 9. Benefits of windbreaks to field and forage crops. \emph{Agriculture, Ecosystems {\&} Environment}, 22-23, 165--190.
-
-\leavevmode\hypertarget{ref-kowalchuk1995}{}%
-Kowalchuk, T.E. \& de Jong, E. (1995). Shelterbelts and their effect on crop yield. \emph{Canadian Journal of Soil Science}, 75, 543--550.
-
-\leavevmode\hypertarget{ref-lecq2018}{}%
-Lecq, S., Loisel, A., Mullin, S.J.
-\& Bonnet, X.
-(2018).
-Manipulating hedgerow quality: Embankment size influences animal biodiversity in a peri-urban context.
-\emph{Urban Forestry {\&} Urban Greening}, 35, 1--7.
-
-\leavevmode\hypertarget{ref-lindstrom2016}{}%
-Lindström, S.A.M., Herbertsson, L., Rundlöf, M., Smith, H.G.
-\& Bommarco, R.
-(2016).
-Large-scale pollination experiment demonstrates the importance of insect pollination in winter oilseed rape.
-\emph{Oecologia}, 180, 759--769.
-
-\leavevmode\hypertarget{ref-lowe2021}{}%
-Lowe, E.B., Groves, R.
-\& Gratton, C.
-(2021).
-Impacts of field-edge flower plantings on pollinator conservation and ecosystem service delivery {{}} a meta-analysis.
-\emph{Agriculture, Ecosystems {\&} Environment}, 310, 107290.
-
-\leavevmode\hypertarget{ref-marja2019}{}%
-Marja, R., Kleijn, D., Tscharntke, T., Klein, A.-M., Frank, T.
-\& Batáry, P.
-(2019).
-Effectiveness of agri-environmental management on pollinators is moderated more by ecological contrast than by landscape structure or land-use intensity.
-\emph{Ecology Letters}, In press.
-
-\leavevmode\hypertarget{ref-mitchell2014}{}%
-Mitchell, M.G.E., Bennett, E.M. \& Gonzalez, A. (2014). Forest fragments modulate the provision of multiple ecosystem services. \emph{Journal of Applied Ecology}, 51, 909--918.
-
-\leavevmode\hypertarget{ref-motzke2016}{}%
-Motzke, I., Klein, A.-M., Saleh, S., Wanger, T.C.
-\& Tscharntke, T.
-(2016).
-Habitat management on multiple spatial scales can enhance bee pollination and crop yield in tropical homegardens.
-\emph{Agriculture, Ecosystems \& Environment}, 223, 144--151.
-
-\leavevmode\hypertarget{ref-ABRegions2006}{}%
-Natural Regions Committee. (2006). \emph{Natural regions and subregions of {Alberta}}. Government of Alberta, Edmonton.
-
-\leavevmode\hypertarget{ref-nguyen2021}{}%
-Nguyen, L.H., Robinson, S.V.J.
-\& Galpern, P.
-(2021).
-Medium-resolution multispectral satellite imagery in precision agriculture: Mapping precision canola (\emph{brassica napus} {L.}) Yield using {Sentinel-2} time series.
-\emph{Precision Agriculture}, In Review (Preprint: https://www.essoar.org/doi/10.1002/essoar.10507479.2).
-
-\leavevmode\hypertarget{ref-osorio2018}{}%
-Osorio, R.J., Barden, C.J. \& Ciampitti, I.A. (2018). {GIS} approach to estimate windbreak crop yield effects in {Kansas}{{}}{Nebraska}. \emph{Agroforestry Systems}, 93, 1567--1576.
-
-\leavevmode\hypertarget{ref-oberg2007}{}%
-Öberg, S., Ekbom, B.
-\& Bommarco, R.
-(2007).
-Influence of habitat type and surrounding landscape on spider diversity in {Swedish} agroecosystems.
-\emph{Agriculture, Ecosystems {\&} Environment}, 122, 211--219.
-
-\leavevmode\hypertarget{ref-pelton1967}{}%
-Pelton, W.L. (1967). The effect of a windbreak on wind travel, evaporation and wheat yield. \emph{Canadian Journal of Plant Science}, 47, 209--214.
-
-\leavevmode\hypertarget{ref-prietoBenitez2011}{}%
-Prieto-Benítez, S. \& Méndez, M. (2011). Effects of land management on the abundance and richness of spiders ({Araneae}): A meta-analysis. \emph{Biological Conservation}, 144, 683--691.
-
-\leavevmode\hypertarget{ref-quinn2017}{}%
-Quinn, N.F., Brainard, D.C. \& Szendrei, Z. (2017). Floral strips attract beneficial insects but do not enhance yield in cucumber fields. \emph{Journal of Economic Entomology}, 110, 517--524.
-
-\leavevmode\hypertarget{ref-ramankutty2018}{}%
-Ramankutty, N., Mehrabi, Z., Waha, K., Jarvis, L., Kremen, C., Herrero, M., \emph{et al.} (2018).
-Trends in global agricultural land use: Implications for environmental health and food security.
-\emph{Annual Review of Plant Biology}, 69, 789--815.
-
-\leavevmode\hypertarget{ref-redhead2020}{}%
-Redhead, J.W., Oliver, T.H., Woodcock, B.A.
-\& Pywell, R.F.
-(2020).
-The influence of landscape composition and configuration on crop yield resilience.
-\emph{Journal of Applied Ecology}, 57, 2180--2190.
-
-\leavevmode\hypertarget{ref-robinson2021}{}%
-Robinson, S.V.J., Edwards, D., Vickruck, J.L., Best, L.R.
-\& Galpern, P.
-(2021).
-Non-crop sources of beneficial arthropods vary within-season across a prairie agroecosystem.
-\emph{Agriculture, Ecosystems {\&} Environment}, 320, 107581.
-
-\leavevmode\hypertarget{ref-singh1999}{}%
-Singh, H.P., Kohli, R.K.
-\& Batish, D.R.
-(1999).
-Impact of \emph{populus deltoides} and \emph{dalbergia sissoo} shelterbelts on wheat {}a comparative study.
-\emph{International Tree Crops Journal}, 10, 51--60.
-
-\leavevmode\hypertarget{ref-staple1955}{}%
-Staple, W.J. \& Lehane, J.J. (1955). The influence of field shelterbelts on wind velocity, evaporation, soil moisture and crop yield. \emph{Canadian Journal of Agricultural Science}, 35, 440--453.
-
-\leavevmode\hypertarget{ref-stasiov2020}{}%
-Stašiov, S., Diviaková, A., Svitok, M., Novikmec, M.
-\& Dovciak, M.
-(2020).
-Hedgerows support rich communities of harvestmen ({Opiliones}) in upland agricultural landscape.
-\emph{Basic and Applied Ecology}, 47, 73--82.
-
-\leavevmode\hypertarget{ref-statscan_canola2014}{}%
-Statistics Canada. (2014). Table 001-0010 - estimated areas, yield, production and average farm price of principal field crops, in metric units, annual.
-
-\leavevmode\hypertarget{ref-steel2017}{}%
-Steel, D. (2017). \emph{Analysis of precision agriculture adoption \& barriers in western {Canada}}. Agriculture; Agri-Food Canada.
-
-\leavevmode\hypertarget{ref-sutter2018b}{}%
-Sutter, L., Amato, M., Jeanneret, P.
-\& Albrecht, M.
-(2018).
-Overwintering of pollen beetles and their predators in oilseed rape and semi-natural habitats.
-\emph{Agriculture, Ecosystems {\&} Environment}, 265, 275--281.
-
-\leavevmode\hypertarget{ref-vanVooren2017}{}%
-Van Vooren, L., Bert, R., Steven, B., Pieter, D.F., Victoria, N., Paul, P., \emph{et al.} (2017).
-Ecosystem service delivery of agri-environment measures: A synthesis for hedgerows and grass strips on arable land.
-\emph{Agriculture, Ecosystems {\&} Environment}, 244, 32--51.
-
-\leavevmode\hypertarget{ref-vega2019}{}%
-Vega, A., Córdoba, M., Castro-Franco, M. \& Balzarini, M. (2019). Protocol for automating error removal from yield maps. \emph{Precision Agriculture}, 20, 1030--1044.
-
-\leavevmode\hypertarget{ref-ward2021}{}%
-Ward, S.E., Umina, P.A., Macfadyen, S. \& Hoffmann, A.A. (2021). Hymenopteran parasitoids of aphid pests within {Australian} grain production landscapes. \emph{Insects}, 12, 44.
-
-\leavevmode\hypertarget{ref-weninger2021}{}%
-Weninger, T., Scheper, S., Lackóová, L., Kitzler, B., Gartner, K., King, N.W., \emph{et al.} (2021).
-Ecosystem services of tree windbreaks in rural landscapes -- a systematic review.
-\emph{Environmental Research Letters}, In press.
-
-\leavevmode\hypertarget{ref-whelan2013}{}%
-Whelan, B. \& Taylor, J. (2013). \emph{Precision agriculture for grain production systems}. CSIRO Publishing.
-
-\leavevmode\hypertarget{ref-wickham2016}{}%
-Wickham, H. (2016). \emph{{ggplot2}: Elegant graphics for data analysis}. Springer-Verlag New York.
-
-\leavevmode\hypertarget{ref-wickle2019}{}%
-Wikle, C.K., Zammit-Mangion, A. \& Cressie, N. (2019). \emph{Spatio-temporal statistics with r}. Chapman; Hall: CRC Press.
-
-\leavevmode\hypertarget{ref-wood2017}{}%
-Wood, S.N. (2017). \emph{Generalized additive models: An introduction with {R}}. CRC press.
-
-\leavevmode\hypertarget{ref-woodcock2016}{}%
-Woodcock, B.A., Bullock, J.M., McCracken, M., Chapman, R.E., Ball, S.L., Edwards, M.E., \emph{et al.} (2016).
-Spill-over of pest control and pollination services into arable crops.
-\emph{Agriculture, Ecosystems \& Environment}, 231, 15--23.
-
-\leavevmode\hypertarget{ref-wratten2012}{}%
-Wratten, S.D., Gillespie, M., Decourtye, A., Mader, E.
-\& Desneux, N.
-(2012).
-Pollinator habitat enhancement: Benefits to other ecosystem services.
-\emph{Agriculture, Ecosystems {\&} Environment}, 159, 112--122.
-
-\leavevmode\hypertarget{ref-zamorano2020}{}%
-Zamorano, J., Bartomeus, I., Grez, A.A.
-\& Garibaldi, L.A.
-(2020).
-Field margin floral enhancements increase pollinator diversity at the field edge but show no consistent spillover into the crop field: A meta-analysis.
-\emph{Insect Conservation and Diversity}, 13, 519--531.
-
-\leavevmode\hypertarget{ref-zhang2007}{}%
-Zhang, W., Ricketts, T.H., Kremen, C., Carney, K. \& Swinton, S.M. (2007). Ecosystem services and dis-services to agriculture. \emph{Ecological Economics}, 64, 253--260.
-
-\end{CSLReferences}
-
-\singlespacing
-\nolinenumbers
-
-\newpage
-
-\hypertarget{appendix-a-r-code-for-data-filtering}{%
-\section*{Appendix A: R Code for Data Filtering}\label{appendix-a-r-code-for-data-filtering}}
-\addcontentsline{toc}{section}{Appendix A: R Code for Data Filtering}
+\hypertarget{appendix-b-r-code-for-data-filtering}{%
+\section*{Appendix B: R Code for Data Filtering}\label{appendix-b-r-code-for-data-filtering}}
+\addcontentsline{toc}{section}{Appendix B: R Code for Data Filtering}
 
@@
 
-\hypertarget{appendix-b-r-code-for-modelling}{%
-\section*{Appendix B: R Code for Modelling}\label{appendix-b-r-code-for-modelling}}
-\addcontentsline{toc}{section}{Appendix B: R Code for Modelling}
+\newpage
+
+\hypertarget{appendix-c-r-code-for-modelling}{%
+\section*{Appendix C: R Code for Modelling}\label{appendix-c-r-code-for-modelling}}
+\addcontentsline{toc}{section}{Appendix C: R Code for Modelling}
 
@@
 \CommentTok{\#functions for distance smoothers and 80 basis functions for spatial smoothers}
+\CommentTok{\#(see Wood 2017 for more details)}
 
-\CommentTok{\#Variance smoother}
+\CommentTok{\#Variance smoothers}
 \NormalTok{f2 }\OtherTok{\textless{}{-}} \StringTok{\textquotesingle{}\textasciitilde{} s(dist,k=5,bs="ts",by=boundaryType) + s(E,N,k=80)\textquotesingle{}} 
-\CommentTok{\#Means smoother}
+\CommentTok{\#Mean smoothers}
 \NormalTok{f }\OtherTok{\textless{}{-}} \StringTok{\textquotesingle{}sqrt(DryYield)\textasciitilde{} s(dist,k=5,bs="ts",by=boundaryType) + s(E,N,k=80)\textquotesingle{}}\ErrorTok{)} 

### 2022-03-04 Manuscript/manuscript_text.tex
Updated manuscript

--- 2021-11-04 Manuscript/manuscript.tex
@@
 Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
-Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure \ref{fig:hypotheses}, Scenario 2).
+Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure 1, Scenario 2).
 Alternatively, yield may be higher near field boundaries (Scenario 3) if boundaries provide helpful ecosystem services (Duelli \& Obrist 2003).
@@
 Yield may be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services (Kowalchuk \& de Jong 1995; Helmers \& Brandle 2005; Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/hypotheses} \caption{Potential yield patterns, depending on ecosystem service effects with distance.
-Scenario 1: no effect of boundary.
-Scenario 2: negative edge effect on yield.
-Scenario 3: positive ecosystem service effect on yield.
-Scenario 4: edge effects are shown in red, ecosystem service effect shown blue, leading to an intermediate peak in yield.}\label{fig:hypotheses}
-\end{figure}
 
@@
 
-Precision yield data were collected directly from farmers across Alberta.
+Precision yield data were collected directly from farmers across Alberta (Figure S1).
 Farmers were solicited for yield data through local agronomists, and we received 298 field-years of data from 5 growers across a total of 7 years (2014-2020).
@@
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/Field Locations} \caption{Locations of fields that data were collected from}\label{fig:fieldLocataions}
-\end{figure}
-
 Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
@@
 Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows (Griffin \emph{et al.} 2007; Whelan \& Taylor 2013), or when the combine is accelerating or decelerating (Arslan \& Colvin 2002).
-Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix A for R code):
+Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix B for R code):
 
@@
 Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
-We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure \ref{fig:hypotheses}).
-All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).
+We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure 1).
+All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.38, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.5, Wickham 2016; and 0.4.0, Kassambara 2020).
 
@@
 To illustrate this, we show the results from a single wheat field, where we show the relative impact of boundary distance and spatial variation on both yield mean and yield SD (``patchiness'').
-Figure \ref{fig:spatialSmooths}a shows the filtered data collected from a single field, demonstrating a large amount of variation even after filtering.
-Yield mean tended to increase with distance from field boundaries while yield SD tended to decrease, but there were no consistent differences between boundary types (Figure \ref{fig:distSmooths}).
-Finally, the random spatial smoothers in Figure \ref{fig:spatialSmooths}b and c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/spatialSmooths} \caption{Raw data and spatial smoothers from a single field.
-Yield averages (means) and variability (log SD) were modeled separately, and both show large spatial dependence within the field.
-Field dimensions are approximately 800 x 800 m, but coordinates are hidden to protect data privacy.}\label{fig:spatialSmooths}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ExamplePlots/distSmooths} \caption{Distance smoothers from a single field, showing a positive saturating effect of distance on average yield (first row), while the variability smoothers show a general decrease in yield SD (second row).}\label{fig:distSmooths}
-\end{figure}
+Figure 2a shows the filtered data collected from a single field, demonstrating a large amount of variation even after filtering.
+Yield mean tended to increase with distance from field boundaries while yield SD tended to decrease, but there were no consistent differences between boundary types (Figure 3).
+Finally, the random spatial smoothers in Figure 2b and c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.
 
@@
 
-Across all fields, standard field boundaries (grassy borders) were the most common boundary type, followed by shelterbelts and other crops (Figure \ref{fig:boundaryTypes}).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/boundaryTypes-1} \caption{Boxplot of field boundary compositions across sampled fields.
-The x-axis indicates the boundary type, while the y-axis indicates the percentage of all field boundaries covered by that boundary type.}\label{fig:boundaryTypes}
-\end{figure}
+Across all fields, standard field boundaries (grassy borders) were the most common boundary type, followed by shelterbelts and other crops (Figure S2).
 
@@
 
-\begin{table}
-
-\caption{\label{tab:smoothSummary}Mean $\chi^2$, p-values, and proportion of smoother p-values less than 0.05 (after Bonneferoni correction) for mean and variability smoothers across all models.}
-\centering
-\begin{tabular}[t]{l|r|r|r|r|r|r}
-\hline
-\multicolumn{1}{c|}{} & \multicolumn{3}{c|}{Mean Yield} & \multicolumn{3}{c}{Yield Variability} \\
-\cline{2-4} \cline{5-7}
-Smoother Type & $\chi^2$-value & p-value & Prop. <0.05 & $\chi^2$-value & p-value & Prop. <0.05\\
-\hline
-Bare & 154.7 & 0.036 & 0.877 & 121.4 & 0.054 & 0.860\\
-\hline
-Grassland & 303.1 & 0.079 & 0.794 & 93.2 & 0.061 & 0.794\\
-\hline
-Other Crop & 174.9 & 0.063 & 0.860 & 104.5 & 0.044 & 0.840\\
-\hline
-Shelterbelt & 328.1 & 0.014 & 0.936 & 137.6 & 0.058 & 0.817\\
-\hline
-Standard & 281.9 & 0.021 & 0.938 & 159.6 & 0.057 & 0.832\\
-\hline
-Wetland & 148.9 & 0.045 & 0.833 & 106.7 & 0.013 & 0.857\\
-\hline
-Spatial Smoother & 17357.5 & 0.000 & 1.000 & 4739.1 & 0.000 & 1.000\\
-\hline
-\end{tabular}
-\end{table}
-
-Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure \ref{fig:canolaPlot} top row).
+Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure 4 top row).
 Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50 m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50 m and remained roughly constant with distance.
 The effect of other boundary types was less consistent, but average yield tended to increase with distance away from grassland, other crop, and wetland, and was roughly the same with distance from bare boundary types.
-Yield variability (Figure \ref{fig:canolaPlot} bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
+Yield variability (Figure 4 bottom row) was essentially constant with distance for shelterbelt, standard, and other crop field boundaries, but showed some evidence of decreasing with distance from field boundary at wetlands, bare, and grassland.
 Interestingly, there was an intermediate reduction in canola yield variability at about 75 m away from wetland boundaries, indicating a potential stabilizing effect, but this was offset by the large increase in yield variability at the edge of wetlands.
 
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_canola} \caption{Field boundary effect on canola yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:canolaPlot}
-\end{figure}
-
 The effect of boundary distance on average wheat yield was much more variable than the patterns seen in canola yield.
-Average yield (Figure \ref{fig:wheatPlot}, top row) roughly increased with distance away from wetland, standard, and grasslands, but both shelterbelts and bare ground displayed an intermediate maximum at around 50m (similar to canola), and decreased slightly with distance from Other Crops.
-Yield variability, on the other hand, tended to decrease with distance from bare, grassland, and wetland (before 100 m, at least), increased slightly with distance from shelterbelt and standard boundaries, and was roughly constant with distance from other crops (Figure \ref{fig:wheatPlot}, bottom row).
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_wheat} \caption{Field boundary effect on wheat yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:wheatPlot}
-\end{figure}
+Average yield (Figure 5, top row) roughly increased with distance away from wetland, standard, and grasslands, but both shelterbelts and bare ground displayed an intermediate maximum at around 50m (similar to canola), and decreased slightly with distance from Other Crops.
+Yield variability, on the other hand, tended to decrease with distance from bare, grassland, and wetland (before 100 m, at least), increased slightly with distance from shelterbelt and standard boundaries, and was roughly constant with distance from other crops (Figure 5, bottom row).
 
 Pea crops were less well-represented in our data (N = 39 fields), and therefore the effects of boundaries tended to be less consistent.
-Average pea yields (Figure \ref{fig:peaPlot}, top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
-Conversely, yield variability (Figure \ref{fig:peaPlot}, bottom row) decreased with distance from grassland, shelterbelt, standard, and wetland boundaries, while the pattern with distance from other crop and bare was inconsistent.
+Average pea yields (Figure 6, top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
+Conversely, yield variability (Figure 6, bottom row) decreased with distance from grassland, shelterbelt, standard, and wetland boundaries, while the pattern with distance from other crop and bare was inconsistent.
 Unlike canola and wheat crops, there was no obvious evidence of an intermediate increase in yield with distance from the field boundary.
-
-\begin{figure}
-\includegraphics[width=1\linewidth]{../Figures/ModelSummary3a_peas} \caption{Field boundary effect on pea yield, accounting for the effect of spatial variation.
-Upper panel represents mean yield, while the lower panel represents yield variation.
-N refers to number of fields containing this boundary type, and \% refers to the average percentage of field boundary accounted for by this boundary type.}\label{fig:peaPlot}
-\end{figure}
 
@@
 Mean yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat, while peas showed no evidence of this.
-However, this intermediate increase was small in canola crops, meaning that ecosystem services likely did not vary strongly with distance from field boundaries (Scenario 2, Figure \ref{fig:hypotheses}).
+However, this intermediate increase was small in canola crops, meaning that ecosystem services likely did not vary strongly with distance from field boundaries (Scenario 2, Figure 1).
 There was a fairly large (0.75 T/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
@@
 
-Our first scenario (Scenario 1, Figure \ref{fig:hypotheses}) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
+Our first scenario (Scenario 1, Figure 1) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
 However, field boundaries may not have a positive effect if a) ecosystem dis-services -- pest insects or weed seeds -- also disperse along with beneficials, or if b) beneficials do not disperse from the boundary environment.
@@
 Since our sampling represents a limited amount of spatial and temporal coverage, there could be additional confounding variables that obscure signals of yield.
-Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee 2006), where boundaries contained a higher proportion of shelterbelts (Figure \ref{fig:boundaryTypes}), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
+Most fields were located in Central Alberta (Dry Mixedwood region, Natural Regions Committee 2006), where boundaries contained a higher proportion of shelterbelts (Figure S2), but differences from Southern AB (Foothills Fescue and Mixedgrass region) may be confounding the signal.
 Between-year effects may be important, especially during hot dry years (Kort 1988), but this was not considered in our analysis.
@@
 
-\singlespacing
-\nolinenumbers
-
-\newpage
-
-\hypertarget{appendix-a-r-code-for-data-filtering}{%
-\section*{Appendix A: R Code for Data Filtering}\label{appendix-a-r-code-for-data-filtering}}
-\addcontentsline{toc}{section}{Appendix A: R Code for Data Filtering}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Helper functions {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-
-\CommentTok{\#"Inlier" spatial filtering procedure from Vega et al 2019, rewritten to work}
-\CommentTok{\#with sf + dplyr. Returns boolean}
-\NormalTok{vegaFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data,ycol,}\AttributeTok{pvalCutoff=}\FloatTok{0.05}\NormalTok{,}\AttributeTok{nDist=}\DecValTok{40}\NormalTok{)\{}
-  \FunctionTok{require}\NormalTok{(sf)}
-  \FunctionTok{require}\NormalTok{(sp)}
-  \FunctionTok{require}\NormalTok{(spdep)}
-  \FunctionTok{require}\NormalTok{(tidyverse)}
-
-  \ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{any}\NormalTok{(}\FunctionTok{class}\NormalTok{(data) }\SpecialCharTok{\%in\%} \StringTok{\textquotesingle{}sf\textquotesingle{}}\NormalTok{)) }\FunctionTok{stop}\NormalTok{(}\StringTok{\textquotesingle{}Dataframe must be sf object\textquotesingle{}}\NormalTok{)}
-
-\NormalTok{  coords }\OtherTok{\textless{}{-}} \FunctionTok{st\_coordinates}\NormalTok{(data) }\CommentTok{\#Get coordinates in matrix form}
-
-  \CommentTok{\#Get neighbourhood weights from 0 to ndist meters}
-\NormalTok{  nWeights }\OtherTok{\textless{}{-}} \FunctionTok{dnearneigh}\NormalTok{(coords,}\DecValTok{0}\NormalTok{,nDist) }
-
-  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{any}\NormalTok{(}\FunctionTok{sapply}\NormalTok{(nWeights,length)}\SpecialCharTok{==}\DecValTok{1}\NormalTok{))\{}
-    \FunctionTok{warning}\NormalTok{(}\StringTok{\textquotesingle{}Some points had no neighbours and were removed\textquotesingle{}}\NormalTok{)}
-\NormalTok{  \} }
-
-  \CommentTok{\#Get neighbourhood indices for each point (which other points are in this}
-  \CommentTok{\#point\textquotesingle{}s neighbourhood?)}
-\NormalTok{  nIndices }\OtherTok{\textless{}{-}} \FunctionTok{nb2listw}\NormalTok{(nWeights, }\AttributeTok{style =} \StringTok{"W"}\NormalTok{,}\AttributeTok{zero.policy =} \ConstantTok{TRUE}\NormalTok{) }
-
-\NormalTok{  yield }\OtherTok{\textless{}{-}} \FunctionTok{pull}\NormalTok{(data,\{\{ycol\}\}) }\CommentTok{\#Get yield data column}
-
-  \CommentTok{\#Local Moran\textquotesingle{}s I for each neighbourhood}
-\NormalTok{  LM }\OtherTok{\textless{}{-}} \FunctionTok{localmoran}\NormalTok{(yield,nIndices,}\AttributeTok{p.adjust.method=}\StringTok{"bonferroni"}\NormalTok{,}
-                   \AttributeTok{alternative =}\StringTok{"less"}\NormalTok{)}
-
-\NormalTok{  results }\OtherTok{\textless{}{-}} \FunctionTok{data.frame}\NormalTok{(LM) }\SpecialCharTok{\%\textgreater{}\%} 
-    \FunctionTok{rename}\NormalTok{(}\StringTok{\textquotesingle{}pval\textquotesingle{}}\OtherTok{=}\FunctionTok{contains}\NormalTok{(}\StringTok{\textquotesingle{}Pr.z.\textquotesingle{}}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-    \CommentTok{\#Filter negative Ii and pvals \textless{} 0.05}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{keepThese=}\NormalTok{ Ii }\SpecialCharTok{\textgreater{}} \DecValTok{0} \SpecialCharTok{|}\NormalTok{ pval }\SpecialCharTok{\textgreater{}}\NormalTok{ pvalCutoff) }\SpecialCharTok{\%\textgreater{}\%}
-    \CommentTok{\#NAs (points had no neighbours)}
-    \FunctionTok{mutate}\NormalTok{(}\AttributeTok{keepThese=}\FunctionTok{ifelse}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(keepThese),}\ConstantTok{FALSE}\NormalTok{,keepThese)) }
-
-\NormalTok{  ret }\OtherTok{\textless{}{-}} \FunctionTok{pull}\NormalTok{(results)}
-  \FunctionTok{return}\NormalTok{(ret)}
-\NormalTok{\}}
-
-\CommentTok{\#Function to filter anything above a certain z{-}score. Returns boolean}
-\NormalTok{ZscoreFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x,}\AttributeTok{z=}\DecValTok{3}\NormalTok{)\{}
-\NormalTok{  xmean }\OtherTok{\textless{}{-}} \FunctionTok{mean}\NormalTok{(x,}\AttributeTok{na.rm=}\ConstantTok{TRUE}\NormalTok{)}
-\NormalTok{  xsd }\OtherTok{\textless{}{-}} \FunctionTok{sd}\NormalTok{(x,}\AttributeTok{na.rm=}\ConstantTok{TRUE}\NormalTok{)}
-\NormalTok{  zscore }\OtherTok{\textless{}{-}} \FunctionTok{abs}\NormalTok{(((x}\SpecialCharTok{{-}}\NormalTok{xmean)}\SpecialCharTok{/}\NormalTok{xsd))}
-\NormalTok{  zscore}\SpecialCharTok{\textless{}}\NormalTok{z}
-\NormalTok{\} }
-
-\CommentTok{\#Function to filter anything above certain quantiles. Returns boolean}
-\NormalTok{QuantileFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x,}\AttributeTok{quant=}\FloatTok{0.99}\NormalTok{)\{ }
-\NormalTok{  l }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{((}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{quant)}\SpecialCharTok{/}\DecValTok{2}\NormalTok{,}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{(}\DecValTok{1}\SpecialCharTok{{-}}\NormalTok{quant)}\SpecialCharTok{/}\DecValTok{2}\NormalTok{) }\CommentTok{\#Symmetric quantiles}
-\NormalTok{  x}\SpecialCharTok{\textgreater{}}\FunctionTok{quantile}\NormalTok{(x,l[}\DecValTok{1}\NormalTok{]) }\SpecialCharTok{\&}\NormalTok{ x}\SpecialCharTok{\textless{}}\FunctionTok{quantile}\NormalTok{(x,l[}\DecValTok{2}\NormalTok{]) }
-\NormalTok{\}}
-
-\CommentTok{\#Function to filter large changes in bearing. Returns boolean unless}
-\CommentTok{\#returnDiffs==TRUE}
-\NormalTok{bearingFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(bearing,}\AttributeTok{q=}\ConstantTok{NULL}\NormalTok{,}\AttributeTok{z=}\ConstantTok{NULL}\NormalTok{,}\AttributeTok{returnDiffs=}\ConstantTok{FALSE}\NormalTok{)\{}
-  \ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{xor}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(q),}\FunctionTok{is.null}\NormalTok{(z))}\SpecialCharTok{\&!}\NormalTok{returnDiffs) }\FunctionTok{stop}\NormalTok{(}\StringTok{\textquotesingle{}Input quantiles or Z{-}score\textquotesingle{}}\NormalTok{)}
-  \CommentTok{\#Difference in compass bearings (in degrees)}
-\NormalTok{  bearingDiff }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x1,x2)\{}
-\NormalTok{    x }\OtherTok{\textless{}{-}}\NormalTok{ x1}\SpecialCharTok{{-}}\NormalTok{x2}
-\NormalTok{    x }\OtherTok{\textless{}{-}} \FunctionTok{ifelse}\NormalTok{(}\FunctionTok{abs}\NormalTok{(x)}\SpecialCharTok{\textgreater{}}\DecValTok{180}\NormalTok{,x}\SpecialCharTok{{-}}\NormalTok{(}\DecValTok{360}\SpecialCharTok{*}\FunctionTok{sign}\NormalTok{(x)),x) }\CommentTok{\#Angle differences can\textquotesingle{}t be \textgreater{}180}
-    \FunctionTok{return}\NormalTok{(x)}
-\NormalTok{  \}}
-
-  \CommentTok{\#Looks 1 point ahead and behind}
-\NormalTok{  bd }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{bearingDiff}\NormalTok{(}\FunctionTok{lag}\NormalTok{(bearing),bearing),}
-              \FunctionTok{bearingDiff}\NormalTok{(}\FunctionTok{lead}\NormalTok{(bearing),bearing))}
-  \CommentTok{\#Maximum bearing difference ahead and behind}
-\NormalTok{  bd }\OtherTok{\textless{}{-}} \FunctionTok{apply}\NormalTok{(bd,}\DecValTok{1}\NormalTok{,}\ControlFlowTok{function}\NormalTok{(x) }\FunctionTok{max}\NormalTok{(}\FunctionTok{abs}\NormalTok{(x),}\AttributeTok{na.rm=}\ConstantTok{TRUE}\NormalTok{)}\SpecialCharTok{*}\FunctionTok{sign}\NormalTok{(x[}\FunctionTok{which.max}\NormalTok{(}\FunctionTok{abs}\NormalTok{(x))])) }
-
-  \ControlFlowTok{if}\NormalTok{(returnDiffs) }\FunctionTok{return}\NormalTok{(bd) }\CommentTok{\#Return bearing differences only, without filtering}
-
-  \ControlFlowTok{if}\NormalTok{(}\SpecialCharTok{!}\FunctionTok{is.null}\NormalTok{(q))\{}
-\NormalTok{    ret }\OtherTok{\textless{}{-}} \FunctionTok{QuantileFilter}\NormalTok{(bd,}\AttributeTok{q=}\NormalTok{q)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-\NormalTok{    ret }\OtherTok{\textless{}{-}} \FunctionTok{ZscoreFilter}\NormalTok{(bd,}\AttributeTok{z=}\NormalTok{z)}
-\NormalTok{  \}}
-  \FunctionTok{return}\NormalTok{(ret)}
-\NormalTok{\} }
-
-\CommentTok{\#Positional difference filter {-} filters out very distant and very close points.}
-\CommentTok{\#Returns boolean unless returnDiffs==TRUE}
-\NormalTok{posFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(data,}\AttributeTok{q=}\ConstantTok{NULL}\NormalTok{,}\AttributeTok{returnDiffs=}\ConstantTok{FALSE}\NormalTok{)\{ }
-  \FunctionTok{require}\NormalTok{(sf)}
-  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(q)}\SpecialCharTok{\&!}\NormalTok{returnDiffs) }\FunctionTok{stop}\NormalTok{(}\StringTok{\textquotesingle{}Input upper quantile\textquotesingle{}}\NormalTok{)}
-
-  \ControlFlowTok{if}\NormalTok{(}\FunctionTok{units}\NormalTok{(}\FunctionTok{st\_distance}\NormalTok{(data[}\DecValTok{1}\NormalTok{,],data[}\DecValTok{2}\NormalTok{,]))}\SpecialCharTok{$}\NormalTok{numerator}\SpecialCharTok{!=}\StringTok{\textquotesingle{}m\textquotesingle{}}\NormalTok{)\{}
-    \FunctionTok{warning}\NormalTok{(}\StringTok{\textquotesingle{}Position differences not in meters\textquotesingle{}}\NormalTok{)}
-\NormalTok{  \} }
-\NormalTok{  coords }\OtherTok{\textless{}{-}} \FunctionTok{st\_coordinates}\NormalTok{(data) }\CommentTok{\#Get coordinates}
-  \CommentTok{\#Distances between points}
-\NormalTok{  pdiff }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(coords)}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{),}\ControlFlowTok{function}\NormalTok{(i) }\FunctionTok{as.numeric}\NormalTok{(}\FunctionTok{dist}\NormalTok{(coords[i}\SpecialCharTok{:}\NormalTok{(i}\SpecialCharTok{+}\DecValTok{1}\NormalTok{),]))) }
-\NormalTok{  pdiff }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(pdiff,}\ConstantTok{NA}\NormalTok{),}\FunctionTok{c}\NormalTok{(}\ConstantTok{NA}\NormalTok{,pdiff)) }\CommentTok{\#Forward and backward lags}
-\NormalTok{  pdiff }\OtherTok{\textless{}{-}} \FunctionTok{apply}\NormalTok{(pdiff,}\DecValTok{1}\NormalTok{,max,}\AttributeTok{na.rm=}\ConstantTok{TRUE}\NormalTok{) }\CommentTok{\#Maximum distance ahead and behind}
-  \ControlFlowTok{if}\NormalTok{(returnDiffs)\{}
-    \FunctionTok{return}\NormalTok{(pdiff)}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \FunctionTok{return}\NormalTok{(}\FunctionTok{QuantileFilter}\NormalTok{(pdiff,q)) }\CommentTok{\#Note: uses 2{-}sided quantiles}
-\NormalTok{  \}}
-\NormalTok{\}}
-
-\CommentTok{\#Filter for (forward and backward) lagged speed differences. Returns boolean}
-\NormalTok{dSpeedFilter }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(speed,}\AttributeTok{l=}\FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{),}\AttributeTok{perc=}\FloatTok{0.2}\NormalTok{)\{ }
-
-  \CommentTok{\#Matrix of \% diffs}
-\NormalTok{  llist }\OtherTok{\textless{}{-}} \FunctionTok{sapply}\NormalTok{(l,}\ControlFlowTok{function}\NormalTok{(x) (}\FunctionTok{lag2}\NormalTok{(speed,x)}\SpecialCharTok{{-}}\NormalTok{speed)}\SpecialCharTok{/}\FunctionTok{lag2}\NormalTok{(speed,x)) }
-
-  \CommentTok{\#Are any lagged speed values \textgreater{} percent change threshold?}
-\NormalTok{  ret }\OtherTok{\textless{}{-}} \SpecialCharTok{!}\FunctionTok{apply}\NormalTok{(llist,}\DecValTok{1}\NormalTok{,}\ControlFlowTok{function}\NormalTok{(y) }\FunctionTok{any}\NormalTok{(}\FunctionTok{abs}\NormalTok{(y)[}\SpecialCharTok{!}\FunctionTok{is.na}\NormalTok{(y)]}\SpecialCharTok{\textgreater{}}\NormalTok{perc))}
-
-  \FunctionTok{return}\NormalTok{(ret)}
-\NormalTok{\}}
-
-\CommentTok{\#Overloaded lag function that takes negative values}
-\NormalTok{lag2 }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x,n)\{}
-  \FunctionTok{require}\NormalTok{(dplyr)}
-  \ControlFlowTok{if}\NormalTok{(n}\SpecialCharTok{==}\DecValTok{0}\NormalTok{) \{}
-    \FunctionTok{return}\NormalTok{(x) }\CommentTok{\#No lag}
-\NormalTok{  \} }\ControlFlowTok{else} \ControlFlowTok{if}\NormalTok{(n}\SpecialCharTok{\textgreater{}}\DecValTok{0}\NormalTok{)\{}
-    \FunctionTok{lag}\NormalTok{(x,n) }\CommentTok{\#Positive lag}
-\NormalTok{  \} }\ControlFlowTok{else}\NormalTok{ \{}
-    \FunctionTok{lead}\NormalTok{(x,}\FunctionTok{abs}\NormalTok{(n)) }\CommentTok{\#Negative lag}
-\NormalTok{  \} }
-\NormalTok{\} }
-
-\CommentTok{\#Filtering pipeline  {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-
-\FunctionTok{library}\NormalTok{(tidyverse)}
-\FunctionTok{library}\NormalTok{(sf)}
-
-\CommentTok{\#Note: dat == sf dataframe of yield data with columns: DryYield (T/ha),}
-\CommentTok{\#TrackAngle (compass bearing), Speed (km/hr), cropType (Wheat, Canola, Peas)}
-
-\NormalTok{dat }\OtherTok{\textless{}{-}}\NormalTok{ dat }\SpecialCharTok{\%\textgreater{}\%} 
-  \CommentTok{\#Vega et al 2019 spatial "inlier" filter {-} takes a few minutes...}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{vegaFilt =} \FunctionTok{vegaFilter}\NormalTok{(.,DryYield,}\AttributeTok{nDist =} \DecValTok{30}\NormalTok{)) }
-
-\NormalTok{dat }\OtherTok{\textless{}{-}}\NormalTok{ dat }\SpecialCharTok{\%\textgreater{}\%} 
-  \CommentTok{\#"BS" filter {-} removes values \textgreater{}10.75 T/ha for wheat, \textgreater{}8 T/ha for peas \& canola}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{noBS =}\NormalTok{ DryYield}\SpecialCharTok{\textless{}}\FunctionTok{ifelse}\NormalTok{(cropType}\SpecialCharTok{==}\StringTok{\textquotesingle{}Wheat\textquotesingle{}}\NormalTok{,}\FloatTok{10.75}\NormalTok{,}\DecValTok{8}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \CommentTok{\#Trim dry yield outliers}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{Qfilt =} \FunctionTok{QuantileFilter}\NormalTok{(DryYield,}\AttributeTok{q=}\FloatTok{0.98}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
-  \CommentTok{\#Trim extreme bearing changes (turning)}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{bFilt =} \FunctionTok{bearingFilter}\NormalTok{(TrackAngle,}\AttributeTok{q=}\FloatTok{0.98}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
-  \CommentTok{\#Trim absolute speed outliers}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{speedFilt =} \FunctionTok{QuantileFilter}\NormalTok{(Speed,}\AttributeTok{q=}\FloatTok{0.98}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
-  \CommentTok{\#Trim speed differences (\textgreater{}20\% change 2 steps forward and backward, suggested}
-  \CommentTok{\#by Lyle et al 2014)}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{dSpeedFilt =} \FunctionTok{dSpeedFilter}\NormalTok{(Speed,}\AttributeTok{l=}\FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{2}\NormalTok{,}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{),}\AttributeTok{perc =} \FloatTok{0.2}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-  \CommentTok{\#Trim points that are far away from eachother}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{posFilt =} \FunctionTok{posFilter}\NormalTok{(.,}\AttributeTok{q=}\FloatTok{0.98}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%} 
-  \CommentTok{\#Combine filter criteria}
-  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{allFilt =}\NormalTok{ noBS }\SpecialCharTok{\&}\NormalTok{ vegaFilt }\SpecialCharTok{\&}\NormalTok{ Qfilt }\SpecialCharTok{\&}\NormalTok{ bFilt }\SpecialCharTok{\&}\NormalTok{ speedFilt }\SpecialCharTok{\&}
-\NormalTok{           dSpeedFilt }\SpecialCharTok{\&}\NormalTok{ posFilt) }\SpecialCharTok{\%\textgreater{}\%}  
-  \CommentTok{\#Remove filtered values}
-  \FunctionTok{filter}\NormalTok{(allFilt)}
-\end{Highlighting}
-\end{Shaded}
-
-\hypertarget{appendix-b-r-code-for-modelling}{%
-\section*{Appendix B: R Code for Modelling}\label{appendix-b-r-code-for-modelling}}
-\addcontentsline{toc}{section}{Appendix B: R Code for Modelling}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# Fit models {-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}{-}}
-
-\FunctionTok{library}\NormalTok{(mgcv)}
-
-\CommentTok{\#Note: dat == dataframe of yield data with columns: DryYield (T/ha), dist}
-\CommentTok{\#(distance to nearest field boundary, m), boundaryType (type of boundary,}
-\CommentTok{\#factor), E,N (easting and northing distance from centre of field, m)}
-
-\CommentTok{\#Uses thin{-}plate splines with extra shrinkage (bs = \textquotesingle{}ts\textquotesingle{}), with 5 basis}
-\CommentTok{\#functions for distance smoothers and 80 basis functions for spatial smoothers}
-\CommentTok{\#(see Wood 2017 for more details)}
-
-\CommentTok{\#Variance smoothers}
-\NormalTok{f2 }\OtherTok{\textless{}{-}} \StringTok{\textquotesingle{}\textasciitilde{} s(dist,k=5,bs="ts",by=boundaryType) + s(E,N,k=80)\textquotesingle{}} 
-\CommentTok{\#Mean smoothers}
-\NormalTok{f }\OtherTok{\textless{}{-}} \StringTok{\textquotesingle{}sqrt(DryYield)\textasciitilde{} s(dist,k=5,bs="ts",by=boundaryType) + s(E,N,k=80)\textquotesingle{}}\ErrorTok{)} 
-\CommentTok{\#Combine into a list}
-\NormalTok{flist }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(}\FunctionTok{as.formula}\NormalTok{(f),}\FunctionTok{as.formula}\NormalTok{(f2)) }
-
-\CommentTok{\#Fit model {-} takes about 45 mins for a 50000 point dataset using an AMD Ryzen 7}
-\CommentTok{\#3800X 8{-}core processor}
-\NormalTok{mod }\OtherTok{\textless{}{-}} \FunctionTok{gam}\NormalTok{(flist,}\AttributeTok{data=}\NormalTok{dat,}\AttributeTok{family=}\FunctionTok{gaulss}\NormalTok{()) }
-\end{Highlighting}
-\end{Shaded}
-
 \end{document}

### 2022-03-04 Manuscript/manuscript_text2.tex
Updated manuscript

--- 2022-03-04 Manuscript/manuscript_text.tex
@@
 
-Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth (Ramankutty \emph{et al.} 2018).
+Intensive agricultural production has increased over the last 100 years -- both in terms of area and intensity, such as enhanced tillage or the addition of synthetic fertilizers and pesticides -- and agricultural land now makes up over a third of ice-free land on Earth (Ramankutty \emph{et al.} 2018).
 This has allowed increases in human population and increased global stability in production (Foley \emph{et al.} 2011).
@@
 Semi-natural field boundaries may cause positive or negative effects on crops, depending on the spatial scale at which ecosystem services operate.
-Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, or greater competition with weeds (Figure 1, Scenario 2).
+Low yields near may occur near the edge of the field because of patchy seed emergence, poor microclimate, shading from trees, soil compaction, or patchy herbicide application leading to higher weed competition (Figure 1, Scenario 2).
 Alternatively, yield may be higher near field boundaries (Scenario 3) if boundaries provide helpful ecosystem services (Duelli \& Obrist 2003).
@@
 Depending on the rate at which positive and negative boundary effects decline with distance, another scenario is also possible.
-Yield may be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services (Kowalchuk \& de Jong 1995; Helmers \& Brandle 2005; Mitchell \emph{et al.} 2014; Van Vooren \emph{et al.} 2017).
+Yield may be maximized at intermediate distances (Scenario 4), suggesting a ``Goldilocks'' field size where negative edge effects are canceled out by ecosystem services (Kowalchuk \& de Jong 1995; Helmers \& Brandle 2005; Mitchell \emph{et al.} 2014; Pywell \emph{et al.} 2015; Van Vooren \emph{et al.} 2017).
 
@@
 Meta-analyses are a common solution to this problem (e.g.
-Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs (Steel 2017; Osorio \emph{et al.} 2018).
+Albrecht \emph{et al.} 2020; Lowe \emph{et al.} 2021), but precision yield data holds enormous promise because of its relatively common usage and low operating costs (Steel 2017; Pywell \emph{et al.} 2015; Osorio \emph{et al.} 2018).
 However, its use is limited for several reasons, namely 1) a lack of standardized formats between equipment types, 2) sensor calibration required for field-level accuracy, and 3) low familiarity with spatial statistical methods.
@@
 
-In this paper we use a precision yield dataset from Alberta, Canada to ask the following questions:
+Within the Canadian prairie provinces, Alberta has roughly 58300 km$^2$ of cultivated agriculture (excluding pasture), of which the majority is spring wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), barley (\emph{Hordeum vulgare}), and peas (\emph{Pisum sativum}) (36.6\%, 30.8\%, 17.1\%, and 7.0\% in 2020, respectively; Agriculture and Agri-Food Canada 2022).
+Roughly 85\% of growers in western Canada report having precision yield monitoring capability (Steel 2017), but these data are seldom used to answer large-scale agroecological questions.
+In this paper we use a precision yield dataset from Alberta to ask the following questions:
 1. How does crop yield change with distance from the edge of field?
@@
 3. Is there an intermediate distance where yield is maximized or variance is minimized?
-To our knowledge, only (Osorio \emph{et al.} 2018) has studied the effects of field boundaries using precision yield data, making this study one of the first of its kind.
+To our knowledge, only Pywell \emph{et al.} (2015) and Osorio \emph{et al.} (2018) have studied the effects of field boundaries using precision yield data, making this study one of the first of its kind, and even more so within a Canadian context.
 
@@
 
-Precision yield data were collected directly from farmers across Alberta (Figure S1).
-Farmers were solicited for yield data through local agronomists, and we received 298 field-years of data from 5 growers across a total of 7 years (2014-2020).
+Precision yield data were collected directly from farmers across Alberta.
+Farmers were solicited for yield data through local agronomists, and we received data from 5 growers for 87 individual fields across 7 years (2014-2020), for a total of 298 field-years of data (Figure S1).
 We converted raw data to a standard csv format using Ag Leader SMS®.
-85\% of the crop types where either wheat (\emph{Triticum aestivum}), canola (\emph{Brassica napus}), or peas (\emph{Pisum sativum}), three of the most common crops in rotation in Alberta (Statistics Canada 2014).
+85\% of the crop types where either wheat, canola, or peas, three of the most common crops in rotation in Alberta (Agriculture and Agri-Food Canada 2022).
+Crop rotations in Alberta usually consist of a combination of wheat - canola - pulses - barley, with other crops being included dependent on the regional growing conditions (e.g.
+sunflowers, flax, corn).
 The remaining crop types had low replication, so we constrained our analysis to only field-years containing wheat (94), canola (119), or peas (39), for a total of 252 field-years of data.
@@
 Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
-Seeding and application rates were constant across fields, so we did not consider inputs in our analysis.
-Precision yield data can be highly variable and is prone to extreme outliers, especially when combine ground speeds are low and at the beginning of rows (Griffin \emph{et al.} 2007; Whelan \& Taylor 2013), or when the combine is accelerating or decelerating (Arslan \& Colvin 2002).
+Fields were all conventional (standard herbicide and fungicide application), and seeding and application rates were constant within fields, so we did not consider these inputs in our analysis.
+Precision yield data can be highly variable and is prone to extreme outliers, especially when ground speeds are low, during turning on headlands, or when the combine is changing speed (Arslan \& Colvin 2002; Griffin \emph{et al.} 2007; Whelan \& Taylor 2013).
+Recording data between different fields can also cause extreme ``jumps" in location, and we verified that all the data points were located within the given field boundaries.
 Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix B for R code):
@@
 \tightlist
-\item
-  Removed outliers and ``inliers'' using the spatial filtering procedure from (Vega \emph{et al.} 2019)
-\item
-  Removed unrealistic yield values (10.75 T/ha for wheat, 8 T/ha for canola and peas), and values of zero or less
-\item
-  Removed extremely low and high outliers (below 1st percentile, above 99th percentile) for: dry yield, steering angle differences, ground speed
-\item
-  Removed points with large changes (\textgreater20\%) in ground speed
-\item
-  Removed extremely low and high changes in position, indicating that the combine had stopped or moved to another part of the field between data collection
-\item
-  Sub-sampled filtered data to 50,000 data points per field (reduce computation time)
+\item Removed outliers and ``inliers'' using the spatial filtering procedure from Vega \emph{et al.} (2019)
+\item Removed unrealistic high yield values (10.75 t/ha for wheat, 8 t/ha for canola and peas), and values $\leq$ zero 
+\item Removed extremely low and high outliers (below 1st percentile, above 99th percentile) for: dry yield, steering angle differences, ground speed
+\item Removed points with large changes (\textgreater20\%) in ground speed
+\item Removed extremely low and high changes in position, indicating that the combine had stopped or moved to another part of the field between data collection
+\item Sub-sampled filtered data to 50,000 data points per field (reduce computation time)
 \end{enumerate}
 
-Field boundaries were digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data (Agriculture and Agri-Food Canada 2018).
-Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions.
+Field boundaries were digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data.
+Edge location and type were usually clearly visible from the Google Earth imagery (CNES/Airbus imagery, 1.5 m resolution), but if they were ambiguous we used classified Agriculture and Agri-Food Canada (2022, 30 m resolution) data to confirm the land cover types surrounding each field.
+The location of a given crop edge is flexible and can change yearly, depending on planting and emergence conditions.
 For example, ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years, and grass boundaries can change if fields are used for as haying or pasture during crop rotation.
-This makes accurate and consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:
+This makes consistent classification of field boundaries difficult, but we used the following general categories for field boundaries:
 
@@
 \tightlist
-\item
-  Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
-\item
-  Wetland: permanent wetland; borders are largely unchanged from year-to-year
-\item
-  Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
-\item
-  Other crop: annual crop or pasture with little or no visible boundary between planted areas
-\item
-  Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
-\item
-  Grassland: permanent seminatural grassland or pasture (not in rotation)
+\item Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
+\item Wetland: permanent wetland; borders are largely unchanged from year-to-year
+\item Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
+\item Other crop: annual crop or pasture with little or no visible boundary between planted areas
+\item Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
+\item Grassland: permanent seminatural grassland or pasture (not in rotation)
 \end{enumerate}
@@
 \tightlist
-\item
-  Yield = weight of crop gathered per area of harvest (T/ha)
-\item
-  Boundary Distance = distance from field boundary type \emph{i} (m)
-\item
-  Easting, Northing = distance from centre of field (m)
-\item
-  f(x,b) = penalized thin-plate regression spline, where \emph{x} is the predictor and \emph{b} is the number of basis dimensions
+\item Yield = weight of crop gathered per area of harvest (t/ha)
+\item Boundary Distance = distance from field boundary type \emph{i} (m)
+\item Easting, Northing = distance from centre of field (m)
+\item f(x,b) = penalized thin-plate regression spline, where \emph{x} is the predictor and \emph{b} is the number of basis dimensions
 \end{enumerate}
@@
 In addition to modeling edge effects (our variable of interest), this model also accounts for within-field spatial variation not related to edges.
-Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g.~Kriging, SPDE Approximations) but this was computationally unfeasible with 50,000 data points per field.
+Spatiotemporal variation can also be modelled using a Gaussian Process Model (e.g. Kriging, SPDE approximations) but this was computationally unfeasible with 50,000 data points per field.
 Penalized splines offer a compromise, as they account for nonlinear relationships in the same way as Gaussian processes using a reduced number of dimensions.
-We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that boundary effects would not be as complex as within-field variation (as in Figure 1).
+We used 5 basis dimensions for the distance smoother and 80 for the spatial smoothers, as we hypothesized that the functional form of boundary effects would not be as complex as within-field variation (as in Figure 1).
 All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.38, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.5, Wickham 2016; and 0.4.0, Kassambara 2020).
@@
 The residuals in most of the models displayed both spatial autocorrelation in the residuals, indicating ``micro-scale'' variation that was not captured by the spatial smoothers (Cressie \& Wikle 2011; Wikle \emph{et al.} 2019).
-This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.~extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 80), we decided to leave the models as-is.
+This is extremely common in precision yield data, and can bias standard errors of regression terms, but because a) there were a large number of data points per field, b) standard errors were relatively small (i.e.
+extremely small p-values), and c) results did not change with the addition of more basis functions (120 rather than 80), we decided to leave the models as-is.
 Temporal autocorrelation in residuals was also present, but we did not include temporal smoothers because spatial and temporal processes are confounded in precision yield data.
-That is, combining of fields occurs in patterns (e.g.~east-to-west) so we cannot tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
+That is, combining of fields occurs in patterns (e.g.
+east-to-west) so we cannot tell whether changes in yield in certain areas of the field were caused by deviations in the yield monitor sensors over time, or whether they represent changes in growing conditions.
 We considered the latter to be a more realistic assumption, so we assigned all random variation as spatial in nature.
@@
 \tightlist
-\item
-  Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard errors.
-\item
-  Use posterior sample to simulate a new smoother from each field.
-\item
-  Fit new model of simulated smoothers from all fields (with a field-year random intercept), and save this ``meta-smoother.''
-\item
-  Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers.
+\item Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard errors.
+\item Use posterior sample to simulate a new smoother from each field.
+\item Fit new model of simulated smoothers from all fields (with a field-year random intercept), and save this ``meta-smoother.''
+\item Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers.
+\end{enumerate}
+
 This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability, and is conceptually similar to a meta-analysis.
 Spatial variability was not averaged between fields because within-field variability is not directly comparable between fields.
-\end{enumerate}
 
@@
 
-Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table \ref{tab:smoothSummary}, row 6).
+Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table 1, row 6).
 Distances from boundaries were also important, but a lower proportion of the fields had strong effects (rows 1-5, cols 3 + 5).
 
-Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 T/ha at the field edge and leveling off at about 3 T/ha (Figure 4 top row).
+Average canola yield tended to increase with distance from all crop boundaries, starting at 2.5 t/ha at the field edge and leveling off at about 3 t/ha (Figure 4 top row).
 Shelterbelt and standard boundaries were well-represented in the data, and average yield quickly increased with distance away from the field boundary, but shelterbelts had a small intermediate maximum at around 50 m before a weak decrease with distance, while average yield away from Standard field boundaries plateaued at 50 m and remained roughly constant with distance.
@@
 Pea crops were less well-represented in our data (N = 39 fields), and therefore the effects of boundaries tended to be less consistent.
-Average pea yields (Figure 6, top row) were roughly 4.25 T/ha at the field boundary and increased to just below 5 T/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
+Average pea yields (Figure 6, top row) were roughly 4.25 t/ha at the field boundary and increased to just below 5 t/ha away from shelterbelt, standard, and other crop boundaries, while the effect of bare, grassland, and wetland boundaries was inconsistent.
 Conversely, yield variability (Figure 6, bottom row) decreased with distance from grassland, shelterbelt, standard, and wetland boundaries, while the pattern with distance from other crop and bare was inconsistent.
@@
 However, this intermediate increase was small in canola crops, meaning that ecosystem services likely did not vary strongly with distance from field boundaries (Scenario 2, Figure 1).
-There was a fairly large (0.75 T/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
+There was a fairly large (0.75 t/ha) intermediate increase in average wheat yields away from shelterbelts, indicating a potential reduction in ecosystem services with distance.
 However, there was a similar (albeit weaker) pattern associated with bare ground, so this pattern may be caused by something other than the presence of shelterbelts.
-This indicates that shelterbelts may have a positive effect on wheat yields, and future work should expand this within relevant geographic sub-regions in order to locally inform growers about how best to manage the landscapes surrounding their fields.
+This indicates that shelterbelts may have a positive effect on wheat yields (and possibly other crops not considered in this study), and future work should expand this within relevant geographic sub-regions in order to locally inform growers about how best to manage the landscapes surrounding their fields.
 
@@
 
-Field boundary characteristics, such as height or species composition, may play an important role in mediating spillover, but our definitions of boundaries were necessarily loose, as we used satellite imagery to manually classify them.
+Field boundary characteristics, such as vegetation height or species composition, may play an important role in mediating spillover, but our definitions of boundaries were necessarily loose, as we used satellite imagery to manually classify them.
 Therefore, it may be that other aspects of field boundaries are more important for spillover of ecosystems services into fields.
-For example, larger or more functionally diverse (Bonifacio \emph{et al.} 2011; Boughey \emph{et al.} 2011; Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020) hedgerows tend to support more biodiversity than smaller hedgerows, and some studies have shown that leaf litter and soil compaction influences predatory beetle communities (Sutter \emph{et al.} 2018).
+For example, larger or more functionally diverse hedgerows tend to support more biodiversity than smaller hedgerows (Bonifacio \emph{et al.} 2011; Boughey \emph{et al.} 2011; Lecq \emph{et al.} 2018; Stašiov \emph{et al.} 2020), and some studies have shown that leaf litter and soil compaction influences predatory beetle communities (Sutter \emph{et al.} 2018).
 These fine-scale boundary characteristics were not measured in our study, but could be used as variables in more biologically-meaningful classifications of field boundaries.
@@
 We also thank Autumn Barnes and Keith Gabert of the Canola Council of Canada, and Laurel Thompson and JP Pettyjohn of Lakeland College for connecting us with farmers and agronomists.
+Finally, we thank Ben Woodcock and an anonymous second reviewer for their comments on the original manuscript.
 
@@
 \leavevmode\hypertarget{ref-AAFC_data}{}%
-Agriculture and Agri-Food Canada.
-(2018).
-\emph{{Annual Crop Inventory}}.
-Available at: \url{https://open.canada.ca/data/en/dataset/ba2645d5-4458-414d-b196-6303ac06c1c9}.
-Last accessed 6 February 2019.
+Agriculture and Agri-Food Canada. (2022). \emph{{Annual Crop Inventory}}. Available at: \url{https://open.canada.ca/data/en/dataset/ba2645d5-4458-414d-b196-6303ac06c1c9}. Accessed 28 February 2022.
 
@@
 
+\leavevmode\hypertarget{ref-pywell2015}{}%
+Pywell, R.F., Heard, M.S., Woodcock, B.A., Hinsley, S., Ridding, L., Nowakowski, M.
+\& Bullock, J.M.
+(2015).
+Wildlife-friendly farming increases crop yield: evidence for ecological intensification.
+\emph{Proceedings of the Royal Society B: Biological Sciences}, 282, 20151740.
+
 \leavevmode\hypertarget{ref-quinn2017}{}%
@@
 \emph{Basic and Applied Ecology}, 47, 73--82.
-
-\leavevmode\hypertarget{ref-statscan_canola2014}{}%
-Statistics Canada. (2014). Table 001-0010 - estimated areas, yield, production and average farm price of principal field crops, in metric units, annual.
 

### 2022-03-04 Manuscript/supplemental.Rmd
Updated manuscript

@@
 
-# Appendix A: R Code for Data Filtering {-}
+# Appendix A: Supplemental Information {-}
+
+\beginsupplement
+
+```{r fieldLocations, echo=FALSE, fig.cap='Locations of fields that data were collected from', out.width='100%'}  
+  include_graphics('../Figures/Field Locations.png',dpi=NA)
+```
+
+```{r boundaryTypes, echo=FALSE, fig.cap='Boxplot of field boundary compositions across sampled fields.
+The x-axis indicates the boundary type, while the y-axis indicates the percentage of all field boundaries covered by that boundary type.', out.width='100%'}
+
+bLengths %>% 
+  mutate(croptype=datSourceFinal$crop) %>% 
+  mutate(location=ifelse(grepl('(Alvin|Dean)',filename),'Southern Alberta','Central Alberta')) %>%
+  group_by(location) %>% mutate(nFields=length(unique(filename))) %>% ungroup() %>% 
+  mutate(location=paste0(location,' (N = ',nFields,')')) %>%
+  select(filename,croptype,location,contains('prop_')) %>% 
+  pivot_longer(contains('prop_')) %>% 
+  mutate(name=gsub('prop_','',name),value=value*100) %>% na.omit() %>%
+  mutate(name=factor(name,labels=c('Bare','Grassland','Other\nCrop','Shelterbelt','Standard','Wetland'))) %>% 
+  group_by(croptype) %>% mutate(croptype=paste0(croptype,'\n(N = ',length(unique(filename)),')')) %>% 
+  ggplot(aes(x=name,y=value,fill=croptype))+geom_boxplot()+
+  facet_wrap(~location,ncol=1) +
+  labs(y='Percentage of field boundary',fill='Crop')+
+  scale_fill_manual(values=c('grey50','grey80','white'))+
+  theme(axis.title.x = element_blank(),legend.position = 'bottom')
+
+# bLengthSummary %>% #Table version
+#   kable(.,format='latex',caption='Mean and SD of proportion of boundary types across all fields, along with the number of fields bordered by at least some of the boundary type.',col.names = c('Type','Mean','SD','Number of Fields'),digits = c(1,4,4,1),escape = FALSE)
+
+```
+
+\newpage
+
+# Appendix B: R Code for Data Filtering {-}
 
@@
 
-# Appendix B: R Code for Modelling {-}
+\newpage
+
+# Appendix C: R Code for Modelling {-}
 

### 2022-03-04 helperFunctions.R

@@ -1193,4 +1193,137 @@ lag2 <- function(x,n){
     } 
 } 
 
+# Function to sample from posterior of each prediction, amalgamate, fit meta-model, return results of meta-model
+# My way of getting around the problem of random effects
+# a = arguments from parLapply, not used
+# useRows = rows of ds to be used, ds = datSource csv, nX = number of samples along range of X
+# rCutoff = restrict point number (some fields have hundreds of thousands of points, but all data uses approximately the same size yield rectangles, so cuts off meta-smoother at "about" the maximum field size - i.e. the most common number of points per field)
 
+#Note: this is more memory-intensive, but quicker. Looks like it maxes out around 12 GB
+samplePreds <- function(a=NA,useRows=NULL,ds=datSource,nX=c(30,100,500),rCutoff=50000,margInt=c(FALSE,FALSE,FALSE),Nsamp=1,ncore=NA,silent=TRUE,samp=TRUE,kPar=5){ 
+  # #Debugging
+  # useRows <- which(datSource$crop=='Canola')
+  # ds <- datSource
+  # nX=c(30,100,500)
+  # rCutoff=50000
+  # margInt=c(FALSE,FALSE,FALSE)
+  # Nsamp <- 10
+  # ncore <- 5
+  
+  if(is.null(useRows)){
+    warning('Rows not specified. Using entire dataset.')
+    useRows <- 1:nrow(datSource)
+  } 
+  require(tidyverse); require(mgcv); require(lme4)
+  source('helperFunctions.R')
+  getFiles <- ds %>% slice(useRows) %>% filter(use,modelComplete2) #Completed models
+  
+  if(!silent) print('Sampling...')
+  #Sample from posterior, get new lines for each field
+  if(is.na(ncore)){
+    allSmooths <- lapply(getFiles$modelPath2,getPreds,l=nX,margInt=margInt,samp=samp,reps=Nsamp) %>% set_names(getFiles$filename) #Serial version
+    if(Nsamp==1) allSmooths <- lapply(allSmooths,function(x){ret <- list(x)})
+  } else { #Parallel version
+    require(parallel)
+    cluster <- makeCluster(ncore) #Memory usage is OK, so could probably max it out
[…]

### 2022-03-04 runModels.R

@@ -320,12 +320,10 @@ stopCluster(cluster)
 Sys.time()-a #Takes ~ 10 hrs for 50 models at 15 procs
 #Takes 2.364893 days to do all fields at 15 procs
 
-
-
 # Get smoother info from second set of models ---------------------------
 
 #Estimate at single field
-debugonce(getPreds)
+# debugonce(getPreds)
 est <- getPreds(paste0('./Figures/ModelCheck2/',datSource$filename[13],' modList.Rdata'),
                 margInt=c(FALSE,TRUE,TRUE),
                 samp=FALSE)$distDat %>% bind_rows(.id='dist_type')
@@ -360,7 +358,7 @@ getFiles <- datSource %>% filter(use,modelComplete2) #Completed models
 allSmooths <- lapply(getFiles$modelPath2,getPreds,margInt=c(FALSE,TRUE,TRUE)) %>% #Takes about 20 seconds
   set_names(getFiles$filename)
 
-#Get df of predictions for each variable
+#Get dataframe of predictions for each variable
 allEff <- lapply(names(allSmooths[[1]]),function(y){
   if(class(allSmooths[[1]][[y]])=='data.frame'){
     lapply(allSmooths,function(x) x[[y]]) %>% 
@@ -418,143 +416,6 @@ ggsave(paste0('./Figures/ModelSummary2a.png'),p,height=6,width=12,dpi=300)
 
 #Draw posterior samples from each field and fit curve at each field
 
-# Function to sample from posterior of each prediction, amalgamate, fit meta-model, return results of meta-model
-# My way of getting around the problem of random effects
-# a = arguments from parLapply, not used
-# useRows = rows of ds to be used, ds = datSource csv, nX = number of samples along range of X
-# rCutoff = restrict point number (some fields have hundreds of thousands of points, but all data uses approximately the same size yield rectangles, so cuts off meta-smoother at "about" the maximum field size - i.e. the most common number of points per field)
-
-#Note: this is more memory-intensive, but quicker. Looks like it maxes out around 12 GB
[…]

### 2022-03-11 Manuscript/manuscript_text2.tex
Added simulation to show whole-field effect

@@
 
-Precision yield data were collected directly from farmers across Alberta.
-Farmers were solicited for yield data through local agronomists, and we received data from 5 growers for 87 individual fields across 7 years (2014-2020), for a total of 298 field-years of data (Figure S1).
-We converted raw data to a standard csv format using Ag Leader SMS®.
+Precision yield data were collected directly from farmers across Alberta (Figure S1).
+Farmers were solicited for yield data through local agronomists, and we received data from 4 individual growers and 1 agronomist with 87 individual fields across 7 years (2014-2020), for a total of 298 field-years of data.
+%31 growers from single agronomist (Kristina)
+Raw data were converted to csv format using Ag Leader SMS®.
 85\% of the crop types where either wheat, canola, or peas, three of the most common crops in rotation in Alberta (Agriculture and Agri-Food Canada 2022).
-Crop rotations in Alberta usually consist of a combination of wheat - canola - pulses - barley, with other crops being included dependent on the regional growing conditions (e.g.
+Crop rotations in Alberta usually consist of a combination of wheat \(\rightarrow\) canola \(\rightarrow\) pulses \(\rightarrow\) barley, with other crops being included dependent on the regional growing conditions (e.g.
 sunflowers, flax, corn).
@@
 Yield data was collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m, adjusted by the combine to account for overlap with adjacent swaths).
-We used dry yield (tonnes of seed per hectare minus - crop moisture) as our measure of yield, using the size of each rectangle (m\(^2\)), dry yield (tonnes), and the spatial location.
+We used dry yield (tonnes of seed per hectare minus crop moisture) as our measure of yield, using the area of each rectangle (m\(^2\)), dry yield (tonnes), and the spatial location.
 Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data.
@@
 Precision yield data can be highly variable and is prone to extreme outliers, especially when ground speeds are low, during turning on headlands, or when the combine is changing speed (Arslan \& Colvin 2002; Griffin \emph{et al.} 2007; Whelan \& Taylor 2013).
-Recording data between different fields can also cause extreme ``jumps" in location, and we verified that all the data points were located within the given field boundaries.
+Recording data between different fields can also cause extreme ``jumps'' in location, and we verified that all the data points were located within the given field boundaries.
 Therefore, we used the following filtering procedure to clean the data prior to model fitting (see Appendix B for R code):
@@
 \tightlist
-\item Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
+\item Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10 m wide)
 \item Wetland: permanent wetland; borders are largely unchanged from year-to-year
@@
 \item Fit new model of simulated smoothers from all fields (with a field-year random intercept), and save this ``meta-smoother.''
-\item Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers.
+\item Repeat 500 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers.
 \end{enumerate}
@@
 Yield mean tended to increase with distance from field boundaries while yield SD tended to decrease, but there were no consistent differences between boundary types (Figure 3).
-Finally, the random spatial smoothers in Figure 2b and c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.
+Finally, the random spatial smoothers in Figures 2b and 2c show that yield mean and SD systematically vary within fields, likely due to moisture, elevation, or soil conditions.
 
@@
 Across all fields, standard field boundaries (grassy borders) were the most common boundary type, followed by shelterbelts and other crops (Figure S2).
-
 Overall, both yield average and yield variability were extremely spatially dependent across all fields (average p-values both \textless0.0001), meaning that within-field variation was the primary driver of yield across canola, wheat, and pea crops (Table 1, row 6).
@@
 The effect of boundary distance on average wheat yield was much more variable than the patterns seen in canola yield.
-Average yield (Figure 5, top row) roughly increased with distance away from wetland, standard, and grasslands, but both shelterbelts and bare ground displayed an intermediate maximum at around 50m (similar to canola), and decreased slightly with distance from Other Crops.
+Average yield (Figure 5, top row) roughly increased with distance away from wetland, standard, and grasslands, but both shelterbelts and bare ground displayed an intermediate maximum at around 50 m (similar to canola), and decreased slightly with distance from Other Crops.
 Yield variability, on the other hand, tended to decrease with distance from bare, grassland, and wetland (before 100 m, at least), increased slightly with distance from shelterbelt and standard boundaries, and was roughly constant with distance from other crops (Figure 5, bottom row).
@@
 Within individual fields, both the average and variability of yield changed with distance from the field boundary, and there were strong spatial relationships within each field, likely related to underlying soil or moisture conditions.
-Mean yield tended to increase with distance from field boundaries before plateauing, and there was some evidence of intermediate boosts in yield from shelterbelts next to canola and wheat, while peas showed no evidence of this.
+Mean yield tended to increase with distance from field boundaries before plateauing, which was not surprising; however, we also found evidence of intermediate boosts in yield from shelterbelts next to canola and wheat.
 However, this intermediate increase was small in canola crops, meaning that ecosystem services likely did not vary strongly with distance from field boundaries (Scenario 2, Figure 1).
@@
 This indicates that shelterbelts may have a positive effect on wheat yields (and possibly other crops not considered in this study), and future work should expand this within relevant geographic sub-regions in order to locally inform growers about how best to manage the landscapes surrounding their fields.
+Furthermore, our work shows that precision yield data can be used to address large-scale agroecological questions, and we recommend its general adoption for the study of agriculture and agroecology.
 
-Our work showed an intermediate increase in wheat yield with distance from shelterbelts and bare ground, but what is the cause of these patterns?
-We hypothesize that the potential yield-increasing mechanisms may be a) microclimate effects, b) soil effects, or c) beneficial animal spillover.
+Our analysis showed decreased yield at the field edge -- this is a very common pattern that growers and agronomists are familiar with -- but we also identified an intermediate increase in wheat yield with distance from shelterbelts and bare ground.
+We hypothesize that the potential causes of this intermediate yield ``boost'' may be microclimate effects, soil effects, or beneficial animal spillover.
 Shelterbelts can act to capture snow and moderate downwind microclimate, reducing water stress on crop plants and increasing yield (Staple \& Lehane 1955; Kowalchuk \& de Jong 1995; Bao \emph{et al.} 2012).
 However, Kowalchuk \& de Jong (1995) found that positive microclimate effects of shelterbelts on wheat yield occur at around 20 m (3-4 times the shelterbelt height), while our work shows a peak at around 50 m.
-This additionally complicated by the fact that water stress varies from year-to-year, and that the impact of the shelterbelt also depends on the orientation to prevailing winds and the physical structure of the shelterbelt (Pelton 1967; Kort 1988; Singh \emph{et al.} 1999).
-Unexpectedly, we also found this intermediate yield maximum with distance away from bare field boundaries, indicating that the patterns we found may be at least somewhat related to water stress.
+This is also complicated by the fact that water stress varies from year-to-year, and that the impact of the shelterbelt also depends on the orientation to prevailing winds and the physical structure of the shelterbelt (Pelton 1967; Kort 1988; Singh \emph{et al.} 1999).
+Unexpectedly, we also found this intermediate yield peak near bare field boundaries, indicating that the patterns we found may be at least somewhat related to water stress.
 While we consider microclimate to be the most likely cause of the intermediate increase, it may also be related to soil effects from shelterbelts, where erosion is reduced and leaf litter from the shelterbelt can cause an increase in soil deposition (de Jong \& Kowalchuk 1995).
@@
 First, field boundaries may not have any influence on beneficial insect abundance if separated from the larger landscape context, or spillover of beneficial organisms may be low.
-Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are low, which may vary among year or location.
+Second, beneficial insect abundance may not influence crop yield if pest pressure or pollination deficits are already low, which may vary among year or location.
 Abiotic ecosystem services, such as reduced wind speed or temperature, also depend on the structure of the boundary and whether wind or temperature are limiting crop production.
-Our work suggests that we need to better examine the context in which a) beneficial organisms spillover from field boundaries and b) under what conditions field boundaries provide tangible ecosystem services (Karp \emph{et al.} 2018).
+Our work suggests that we need to better examine the context in which beneficial organisms spillover from field boundaries, and under what conditions field boundaries provide tangible ecosystem services (Karp \emph{et al.} 2018).
 We consider these reasons below:
@@
 SNL generally has a positive effect on beneficial insect abundance, but field boundaries only represent a fraction of the SNL that surrounds a given field.
-Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales (Nguyen \emph{et al.} 2021) which could be particularly important when beneficial arthropods are disperse very broadly (Öberg \emph{et al.} 2007).
+Our analysis only examined boundary types, without considering the landscape context that fields were embedded in, but some studies show the influence of landscape composition at larger spatial scales (Nguyen \emph{et al.} 2022) which could be particularly important when beneficial arthropods disperse very broadly (Öberg \emph{et al.} 2007).
 There could also be interactions between landscape composition at different scales (Motzke \emph{et al.} 2016; Haslem \emph{et al.} 2021; Robinson \emph{et al.} 2021), but this is not broadly studied.
-Landscape complexity (Fahrig \emph{et al.} 2011) and connectivity are also associated with higher abundance of beneficial arthropods.
+Landscape complexity and connectivity are also associated with higher abundance of beneficial arthropods (Fahrig \emph{et al.} 2011).
 This means that field boundaries such as wetlands or shelterbelts may only be helpful if connected to a larger network of non-crop habitat (but see Fahrig \emph{et al.} 2019).
@@
 Our first scenario (Scenario 1, Figure 1) assumes that beneficial organisms disperse from boundaries into the field, and had a generally positive effect on crop yield.
-However, field boundaries may not have a positive effect if a) ecosystem dis-services -- pest insects or weed seeds -- also disperse along with beneficials, or if b) beneficials do not disperse from the boundary environment.
+However, field boundaries may not have a positive effect if ecosystem dis-services -- pest insects or weed seeds -- also disperse along with beneficials, or if beneficials do not disperse from the boundary environment.
 Pests and weeds can also disperse from field boundaries, which may negate any positive effect of ecosystem services (Zhang \emph{et al.} 2007).
 However, this is still poorly studied, and parsing out services from disservices is impossible using crop yield data \emph{alone}.
-Ecosystem services from field boundaries may also be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales Robinson \emph{et al.} (2021).
+Ecosystem services from field boundaries may also be limited if beneficial organisms stay in field boundaries, or disperse at wider spatial scales (Öberg \emph{et al.} 2007).
 Lastly, pest pressure or pollination requirements may be low enough that ecosystem service provision does not significantly change yield.
@@
 Between-year effects may be important, especially during hot dry years (Kort 1988), but this was not considered in our analysis.
-Future work could examine whether boundaries have differing effects during hot or cool years, as shading or cooling effects may increase yield during hot years but have negative (or no) effect during cooler years.
+Future work could examine whether boundaries have differing effects during hot or cool years, as shading or cooling effects may increase yield during hot years but have no effect (or negative effects) during cooler years.
 Finally, about 50\% of growers in western Canada have collected precision agriculture data, and about 85\% have yield monitoring capability (Steel 2017), which means that our work has only scratched the surface of what could be (collectively) one of the richest agroecological datasets ever created.
@@
 We thank Trent Clark, Dean Hubbard, Alvin French, and Kristina Polziehn for providing yield data for us, and providing insight on their yield data.
-We also thank Autumn Barnes and Keith Gabert of the Canola Council of Canada, and Laurel Thompson and JP Pettyjohn of Lakeland College for connecting us with farmers and agronomists.
+We also thank Autumn Barnes of Lethbridge College, Keith Gabert of the Canola Council of Canada, and Laurel Thompson and JP Pettyjohn of Lakeland College for connecting us with farmers and agronomists.
 Finally, we thank Ben Woodcock and an anonymous second reviewer for their comments on the original manuscript.
@@
 
-\leavevmode\hypertarget{ref-nguyen2021}{}%
-Nguyen, L.H., Robinson, S.V.J.
-\& Galpern, P.
-(2021).
-Medium-resolution multispectral satellite imagery in precision agriculture: Mapping precision canola ({\emph{Brassica napus}} {L.}) Yield using {Sentinel-2} time series.
-\emph{Precision Agriculture}, In Review (Preprint: https://www.essoar.org/doi/10.1002/essoar.10507479.2).
+\leavevmode\hypertarget{ref-nguyen2022}{}%
+Nguyen, L.H., Robinson, S.V.J. \& Galpern, P. (2022). Effects of landscape complexity on crop productivity: An assessment from space. \emph{Agriculture, Ecosystems \& Environment}, 328, 107849.
 

### 2022-03-11 runModels.R

@@ -703,10 +703,8 @@ samp_mean <- vector(mode='list',length=3) %>% set_names(croptype)
 for(i in 1:length(croptype)){
   isCrop <- which(tolower(datSource$crop)==croptype[i] & datSource$use)
   # samp_mean[[i]] <- backTrans(samplePreds(1,useRows=isCrop,samp=FALSE))
-  samp_mean[[i]] <- samplePreds(1,useRows=isCrop,samp=FALSE)
+  samp_mean[[i]][[1]] <- samplePreds(1,useRows=isCrop,samp=FALSE)
 } #Takes about 10 secs
-
-
 
 lapply(samp$wheat,function(x) x$coverDist[c('STANDARD','SHELTERBELT','WETLAND')]) %>% set_names(paste0('samp',1:length(.))) %>% 
   lapply(.,function(x) bind_rows(x,.id='boundary') %>% filter(dist<400)) %>% 
@@ -718,7 +716,8 @@ lapply(samp$wheat,function(x) x$coverDist[c('STANDARD','SHELTERBELT','WETLAND')]
 
 
 #Create field boundary polygon
-qSecDist <- 805/4 #Quarter section = 805 x 805 m
+qSecDist <- 805 #Quarter section = 805 x 805 m
+# qSecDist <- 805/2 #half Quarter section 
 fieldPoly <- st_polygon(x=list(matrix(c(0,0,0,qSecDist,qSecDist,qSecDist,qSecDist,0,0,0),nrow=5,ncol=2,byrow = TRUE))) #Square field
 fieldBoundary <- lapply(1:(length(st_coordinates(fieldPoly)[, 1]) - 1), #4 boundary segments
                         function(i){
@@ -754,8 +753,8 @@ rm(qSecDist,fieldPoly,fieldBoundary)
 # grid1 <- sampGrid %>% mutate(m=NA,sd=NA,y=NA) %>% st_drop_geometry()
 # 
 # #Looks OK
-# samp1 %>% bind_rows(.id='boundary') %>% filter(dist<400) %>% 
-#   pivot_longer(predMean:predLogSD) %>% 
+# samp1 %>% bind_rows(.id='boundary') %>% filter(dist<400) %>%
+#   pivot_longer(predMean:predLogSD) %>%
 #   ggplot(aes(x=dist,y=value))+geom_point()+
 #   facet_grid(name~boundary,scales='free_y')
 
[…]
